### 2021-11-19 Paper/main.tex
Paper latex, simple loss rgb anchor

% EMNLP header. Do not change.
\pdfoutput=1
\documentclass[11pt]{article}
\usepackage[review]{emnlp2021}
\usepackage{times}
\usepackage{latexsym}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{microtype}
% End header.

%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%
% EDITING LINK: 
% https://www.overleaf.com/3575751412sjsdmszkxsyh
%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%

\usepackage{soul}
\usepackage{url}
% \usepackage[hidelinks]{hyperref}
% \usepackage[small]{caption}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{booktabs}
\usepackage{algorithm}
\usepackage{algorithmic}
\urlstyle{same}

% =======================================
% my imported packages

\usepackage{amsmath}
\usepackage{bm}
\usepackage{amssymb}
\usepackage{listings}
% \usepackage[ruled,vlined]{algorithm2e}
\include{pythonlisting}

\newcommand{\pseudosection}[1]{\vspace{1.5ex}\noindent \textit{{#1:~~}}}
% \newcommand{\pseudosection}[1]\indent{{{}}}

\usepackage{todonotes}
\newcommand{\todokdinline}[1]{\todo[color=red!20,inline]{{KD: \small #1}}}
\newcommand{\todokd}[1]{\todo[color=red!20]{{\small #1 -- Kasra}}}
\newcommand{\todocainline}[1]{\todo[color=yellow!20,inline]{{CA: \small #1}}}
\newcommand{\todocm}[1]{\todo[color=green!40]{\small #1 -- Cynthia}}
\newcommand{\todocmi}[1]{\todo[inline,color=green!40]{\small #1 -- Cynthia}}
\newcommand{\todoff}[1]{\todo[color=blue!20]{\small #1 -- Frank}}
\newcommand{\todoed}[1]{\todo[color=cyan!20]{\small #1 -- Ed}}

\newcommand{\CitationNeeded}[1]{{\textbf {\color{red}Cite #1}}}
\newcommand{\ProofRead}[1]{{\textbf {\color{red}Proofread please #1}}}
\newcommand{\Rephrase}[1]{{\textbf {\color{red}Rephrase please #1}}}

% \newcommand{\citet}[1]{\cite{#1}} % This is used because \citet didn't work with their author kit % It does now

% =======================================

\title{MMA: Multi-modal Manifold Alignment}

\author{
First Author$^1$
\and
Second Author$^2$\and
Third Author$^{2,3}$\And
Fourth Author$^4$
\affiliations
$^1$First Affiliation\\
$^2$Second Affiliation\\
$^3$Third Affiliation\\
$^4$Fourth Affiliation
\emails
\{first, second\}@example.com,
third@other.example.com,
fourth@example.com
}

\begin{document}

Todo list

\begin{enumerate}
    \item Lit search 
\end{enumerate}

\newpage

\maketitle
\setcounter{page}{1}

% ***************** The BEGINNING of the paper *****************

\begin{abstract}

\end{abstract}

%===================================================================
\section{Lit Search}

\subsection{Multimodal Machine Learning: A Survey and Taxonomy}

\begin{itemize}
\item \href{https://ieeexplore.ieee.org/document/8269806}{url}
\end{itemize}

\citet{baltrusaitisMultimodalMachineLearning2019} proposes a new taxonomy of multimodal machine learning by introducing five technical challenges in addition to typical early and late fusion categorization. This taxonomy includes the following categories.
\begin{enumerate}
\item \textit{Representation}: represent and summarize multimodal data in a way that exploits the complementarity and redundancy of multiple modalities.
\begin{enumerate}
\item joint: combine the unimodal signals into the same representation space.
\item coordinated: process unimodal signals separately, but enforce certain similarity or structure constraints on them to bring them to a coordinated space.
\end{enumerate}
\item \textit{Translation}: map data from one modality to another.
\begin{enumerate}
\item example-based
\item generative
\end{enumerate}
\item \textit{Alignment}: identify the direct relations between (sub)elements from two or more different modalities.
\begin{enumerate}
\item explicit
\item implicit
\end{enumerate}
\item \textit{Fusion}: join information from two or more modalities to perform a prediction.
\begin{enumerate}
\item model-agnostic
\item model-based
\end{enumerate}
\item \textit{Co-learning}: transfer knowledge between modalities, their representation, and their predictive models
\begin{enumerate}
\item parallel data
\item non-parallel data
\item hybrid data
\end{enumerate}
\end{enumerate}

The authors mention that while joint representations have been used in situations to construct representations of more than two modalities, coordinated spaces have, so far, been mostly limited to two. This means that our research is novel and we can extend similarity measures to more than two modalities.

\subsection{Self-Supervised MultiModal Versatile Networks}

\begin{itemize}
\item \href{https://proceedings.neurips.cc/paper/2020/file/0060ef47b12160b9198302ebdb144dcf-Paper.pdf?utm\_campaign=NLP\%20News\&utm\_medium=email\&utm\_source=Revue\%20newsletter}{url}
\end{itemize}

This work uses self-supervised contrastive learning to learn and combine representations from three modalities of visual, audio, and language. They learn a \textit{multimodal versatile network} that has the following four properties:
\begin{enumerate}
\item Ability to take any of the modalities as input
\item Respecting the specificity of modalities: audio and visual modalities are fine-grained while language modality is coarse-grained.
\item Comparability of different modalities using \textbf{dot product} even if not seen together during training 
\item Efficiently applicable to visual data either in the form of dynamic videos or static images
\end{enumerate}

The authors consider the following three different configurations of the modality spaces which they call \textit{modality embedding graphs}.
\begin{enumerate}
\item Shared: all modalities map to the same space. This respects property 3 but violates property 2.
\item Disjoint: visual-audio and visual-text spaces. This respects property 2 but violates property 3.
\item Fine and coarse: audio and visual domains are map to the same space since they are fine-grained. These embeddings are then mapped to a lower dimensional space where text is also mapped. This respects both properties 2 and 3.
\end{enumerate}

The loss function they define consists of two terms. The first term is to train the fine-grained space of visual and audio embeddings by minimizing the distance between the similar pair (positives) from the same location of a video, and maximizing the distance between negative pairs (negatives) sampled from different videos. The second term is to train the coarse-grained space which consists of visual and text embeddings. Note that they don't consider audio embeddings in this space since they don't want to learn automatic speech recognition (ASR). Text and visual domains are not aligned as well as audio and visual domains (e.g. sound of playing piano versus the text that only says playing piano for a video of playing piano). To cure this, they consider a set of positive pairs instead of a single pair.
If a modality is missed, the corresponding term is dropped from the loss function.

\subsection{Separating Self-Expression and Visual Content in Hashtag Supervision}

\begin{itemize}
\item \href{https://openaccess.thecvf.com/content\_cvpr\_2018/papers/Veit\_Separating\_Self-Expression\_and\_CVPR\_2018\_paper.pdf}{url}
\end{itemize}

The authors train a joint model of images, hashtags, and users to perform image retrieval. This is similar to our task where given a description, we want to find the image in the scene that best matches the description. The idea is to form a three-way tensor product model. They use a ranking loss to train the model where the score of an observed triplet is higher than an unobserved triplet. They sample six negative triplets per positive sample triplet, and use each of them as a negative in the loss.
The downstream retrieval task is then simply done by taking the arg max of the tensor product for a given user.

\subsection{Semi-Heterogeneous Three-Way Joint Embedding Network for Sketch-Based Image Retrieval}

\begin{itemize}
\item \href{https://ieeexplore.ieee.org/document/8809264}{url}
\end{itemize}

\subsection{Deep Multimodal Learning for Affective Analysis and Retrieval}

\begin{itemize}
\item \href{https://ieeexplore.ieee.org/abstract/document/7277066?casa\_token=aBp6BxcszHwAAAAA:3NoMiFrZbn7tXfavF1rgkCiGWbFI2arxn8Xb6iDF79q4zBZHWi7PWWhf6xW-xJwYdFALbmRENo4}{url}
\end{itemize}

\subsection{An Efficient Framework for Zero-Shot Sketch-Based Image Retrieval}

\begin{itemize}
\item \href{https://arxiv.org/pdf/2102.04016.pdf}{url}
\end{itemize}
This paper uses two modalities only; image and sketch. However, it uses a \textit{quadruplet} to compute the loss which can be useful in our research. A quadruplet is composed of a sketch picture as an anchor, a negative example from sketch domain, a negative example from picture domain, and a positive example from picture domain.

\subsection{SMIL: Multimodal Learning with Severely Missing Modality}

\begin{itemize}
\item \href{https://www.aaai.org/AAAI21Papers/AAAI-437.MaM.pdf}{url}
\end{itemize}
Uses Bayesian meta-learning framework to perform multimodal learning with partially missing modalities in training/test data.
One of their experiments is multi-label classification of movie genres with bimodal data including poster of movies and description of the movie from IMDB.

\subsection{Multimodal Language Analysis in the Wild: CMU-MOSEI Dataset and Interpretable Dynamic Fusion Graph}

\begin{itemize}
\item \href{https://aclanthology.org/P18-1208.pdf}{url}
\end{itemize}

\subsection{Multimodal Learning with Incomplete Modalities by Knowledge Distillation}

\begin{itemize}
\item \href{https://dl.acm.org/doi/pdf/10.1145/3394486.3403234}{url}
\end{itemize}
They first train models on each modality independently
\subsection{SWAFN: Sentimental Words Aware Fusion Network for Multimodal Sentiment Analysis}

\begin{itemize}
\item \href{https://aclanthology.org/2020.coling-main.93.pdf}{url}
\end{itemize}

\subsection{Multimodal Learning for Human Action Recognition Via Bimodal/Multimodal Hybrid Centroid Canonical Correlation Analysis}

\begin{itemize}
\item \href{https://ieeexplore.ieee.org/document/8489981}{url}
\end{itemize}

\subsection{Deception Detection Using a Multimodal Approach}

\begin{itemize}
\item \href{https://dl.acm.org/doi/pdf/10.1145/2663204.2663229?casa\_token=KneK7B7xvLwAAAAA:Ajz0N96ygq8ktwkz0IVUqTt8NozCg2wR6n\_x2xntdHqZBh6VXW\_8VbO4GeY4VvMsDJlMwzkhVXQSJA}{url}
\end{itemize}
They use \textit{language}, \textit{physiological response}, and \textit{thermal sensing} to detect deceit.

\subsection{M3ER: Multiplicative Multimodal Emotion Recognition using Facial, Textual, and Speech Cues}

\begin{itemize}
\item \href{https://ojs.aaai.org/index.php/AAAI/article/view/5492}{url}
\end{itemize}

\subsection{Found in Translation: Learning Robust Joint Representations by Cyclic Translations between Modalities}

\begin{itemize}
\item \href{https://ojs.aaai.org/index.php/AAAI/article/view/4666}{url}
\end{itemize}
\subsection{Select-Additive Learning: Improving Generalization In Multimodal Sentiment Analysis}

\begin{itemize}
\item \href{https://ieeexplore.ieee.org/stamp/stamp.jsp?tp=\&arnumber=8019301}{url}
\end{itemize}

\subsection{Attention-Based Multimodal Fusion for Video Description}

\begin{itemize}
\item \href{https://openaccess.thecvf.com/content\_ICCV\_2017/papers/Hori\_Attention-Based\_Multimodal\_Fusion\_ICCV\_2017\_paper.pdf}{url}
\end{itemize}
\subsection{\st{3W-AlignNet: a Feature Alignment Framework for Person Search with Three-Way Decision Theory}}

\begin{itemize}
\item \href{https://link.springer.com/article/10.1007/s12559-021-09898-7}{url}
\end{itemize}

This is not very related since they use three-way decision theory to select bounding boxes as positive, negative, and boundary.
Three-way here refers to something else and not modalities.

\subsection{\st{Unsupervised Domain Adaptation for Face Recognition in Unlabeled Videos}}
\begin{itemize}
\item \href{https://openaccess.thecvf.com/content\_ICCV\_2017/papers/Sohn\_Unsupervised\_Domain\_Adaptation\_ICCV\_2017\_paper.pdf}{url}
\end{itemize}
Might be relevant but not super relevant.

\subsection{Heterogeneous Sensor Data Fusion By Deep Multimodal Encoding}
\begin{itemize}
\item \href{https://ieeexplore.ieee.org/stamp/stamp.jsp?tp=&arnumber=7874158}{url}
\end{itemize}
This one takes two modalities as input and outputs predictions in a third modality.

%===================================================================
\section{Introduction}
\label{intro}

% Grounded language learning~\cite{Mooney2008Grounded,RichardsDarvishMatuszekCategoryFree20,triplet_loss_2021_CVPR} is a cross-modal learning task to connect natural language with the physical environment in which the natural language is used.

% Our contributions are as follows:
% \begin{enumerate}
%     \item 
%     \item 
%     \item 
% \end{enumerate}

%===================================================================

\section{Related Work}
\label{sec:Related-Work}

%===================================================================

\section{Problem Description}
\label{sec:Problem-Description}

%===================================================================

\section{Approach}
\label{sec:Method}

For $M$ modalities, we define a distance based loss function that contains $M+2(M-1)!$ terms. 
We sample two data points from each modality, one positive point and one negative point. The objective is then to (I) minimize the distance between each pair of positive points from heterogeneous modalities which results in $(M-1)!$ terms, (II) maximize the distance between each pair of positive and negative points from heterogeneous modalities which contains $(M-1)!$ terms again, and (III) maximize the distance between each pair of positive and negative points from the homogeneous modalities which contains $M$ terms.
Equation~\ref{eq:objective} shows the objective function we defined above with changing maximization to minimization by changing the sign.

% \begin{equation}\label{eq:objective-detailed}
%     \min \sum_{i=1}^{M-1} \sum_{j=i+1}^{M} ||x_{i}^{+} - x_{j}^{+}|| - \sum_{i=1}^{M-1} \sum_{j=i+1}^{N} ||x_{i}^{+} - x_{j}^{-}|| - 
%     \sum_{i=1}^{M} ||x_{i}^{+} - x_{i}^{-}||
% \end{equation}

\begin{aligned}\label{eq:objective}
    \mathcal{L}  &= \sum_{i=1}^{M-1} \sum_{j=i+1}^{M} ||f_i(x_{i}^{+}) - f_j(x_{j}^{+})||
    - ||f_i(x_{i}^{+}) - f_j(x_{j}^{-})|| \\ 
    & - ||f_i(x_{i}^{-}) - f_j(x_{j}^{+})|| - \sum_{i=1}^{M} ||f_i(x_{i}^{+}) - f_i(x_{i}^{-}) ||
\end{aligned}
where the subscripts $i$ and $j$ represent modalities, and $f_i(.)$ and $f_j(.)$ represent functions that map input data to the shared latent space using a deep neural network.

\todokdinline{I can add another term for minimizing the distance between similar pairs (anchor and positive) from the same domain. This would look like the equation below. However, it breaks the nice procedure of sampling two data points per modality}
$+ \sum_{i=1}^{M} ||f_i(x_i^{a}) - f_i(x_i^{+})||$

Our method is a generalization of well-known similarity-based metric learning and manifold alignment methods such as triplet loss~\CitationNeeded{}.
If $M=2$ which means the number of modalities is 2, then our objective function reduces to quadruplet loss method.
If $M=2$ and we ignore the last term in the objective function, it results in the triplet loss method. 
If $M=1$, only the last term remains in the loss function which is exactly the pairwise distance-based loss function.

\subsection{Network Architecture}
\label{sec:Model}

%===================================================================

\section{Experiments}
\label{sec:Experiments}

\subsection{Data}
\label{sec:Data}

\subsection{Setup}
\label{sec:setup}

\subsection{Quantitative Results}
\label{sec:Quantitative}

\subsection{Qualitative Results}
\label{sec:Qualitative}

\subsection{Baselines}

%===================================================================

\section{Conclusion}
\label{sec:conclusion}

%===================================================================

% ***************** The END of the paper *****************

\clearpage
\section*{Acknowledgments}
%% The file named.bst is a bibliography style file for BibTeX 0.99c
\bibliographystyle{named}
\bibliography{ref}
\clearpage

\appendix
\label{sec:appendix}
\section{Qualitative Analysis}

\section{Model}

\section{Experiments}

\section{Results}

\section{Setup}

\end{document}

### 2021-11-19 Code/ML.py

@@ -1,0 +1,330 @@
+import torch
+import numpy as np
+import random
+import os, argparse, json, pickle
+from copy import deepcopy
+import wandb
+from tqdm import tqdm
+from sklearn.metrics import precision_recall_fscore_support
+from sklearn.metrics.pairwise import euclidean_distances, cosine_similarity
+
+from models import TheModel
+from datasets import load_all_data
+from losses import mma_loss
+from utils import set_seeds, setup_device, initialize_result_keeper, prf_metrics, mrr_acc_metrics
+
+
+def train(config, models, dataset):
+
+    results, outputs, logs, reconstruct, examples = initialize_result_keeper(config)    
+    
+    # Initialize the optimizer
+    optimizers = {}
+    for modality in models.keys():
+        model = models[modality]
+        optimizers[modality] = torch.optim.Adam(model.parameters(), lr=config.learning_rate)
+    
+    for epoch in tqdm(range(config.epochs)):
+        for modality in models.keys():
+            model = models[modality]
+            model.train()
+        logs[epoch] = {}
+        results[epoch] = {}
+        # do the training
+        print('\ntraining for the new epoch')
+        logs[epoch] = training(config, models, dataset, 'train', optimizers, epoch)
+
+        # evaluation on valid and possibly test        
+        for portion in config.portions:
+            if portion != 'train':
+                print('evaluating model on {} set'.format(portion))
+                results[epoch][portion], outputs[portion] = evaluate(config, models, dataset, portion)
+
+        # update best and save outputs of the model if valid f1 is the best
+        if results[epoch]['valid']['f1'] >= results['best']['best-valid']['f1']:
+            for modality in models.keys():
+                model = models[modality]
+                torch.save(model.state_dict(), config.results_dir+modality+'_model_state.pt')
+                # wandb.save(config.results_dir+modality+'_model_state.pt')
+            pickle.dump(outputs, open(config.results_dir+'outputs.pkl', 'wb'))
+            for portion in config.portions:
+                if portion != 'train':
+                    results['best']['best-'+portion].update(dict(zip(results['best']['best-'+portion].keys(), results[epoch][portion].values())))
+
+        wandb.log({**{'epoch': epoch}, **logs[epoch], **results[epoch]})
+        wandb.run.summary.update(results['best'])
[…]

### 2021-11-19 Code/datasets.py

@@ -1,0 +1,275 @@
+import random, os, pickle, json
+import pandas as pd
+import numpy as np
+import scipy 
+import torch
+from torch.utils.data import Dataset, DataLoader
+import torchvision
+from sklearn.model_selection import train_test_split
+import flair
+import re
+from skimage import io
+import skimage
+
+
+class GoLD_Dataset(Dataset):
+
+    def __init__(self, config, device, computed_vision_embeds, train='train', computed_lang_embeds=None):
+    
+        self.a_override = False
+        self.config = config
+        self.device = device
+        self.train = train
+        if config.task == 'gold':
+            root_dir = config.data_dir+'gold/images'
+            # csv_file = '../../../../data/gold/text.tsv'
+            csv_file = '../data/gold_text.tsv'
+        elif config.task in ['RIVR', 'gauss_noise', 'dropout_noise', 'snp_noise']:
+            # root_dir = config.data_dir+'VR_GoLD_structure/images'
+            root_dir = config.data_dir+'simulation/images'
+            csv_file = '../data/rivr_data.tsv'
+
+        full_dataset = pd.read_csv(csv_file, header=0, delimiter='\t', keep_default_na=False)
+        dataset = full_dataset.reset_index()
+        normalize = torchvision.transforms.Normalize(mean=[0.485, 0.456, 0.406], std=[0.229, 0.224, 0.225])
+        resize = torchvision.transforms.Resize((128,128))
+        transform_rgb = torchvision.transforms.Compose([torchvision.transforms.ToPILImage(), resize,
+                                                        torchvision.transforms.ToTensor(), normalize])
+
+        depth2rgb = lambda x: skimage.img_as_ubyte(skimage.color.gray2rgb(x/np.max(x)))
+        transform_depth = torchvision.transforms.Compose([depth2rgb,
+                                                        torchvision.transforms.ToPILImage(), resize,
+                                                        torchvision.transforms.ToTensor(), normalize])
+
+        if config.task == 'gold':
+            self.images = dataset["item_id"]
+            self.descriptions = dataset["text"]
+            self.object_names = [self.getObjectName(image) for image in self.images]
+            self.instance_names = [self.getInstanceName(image) for image in self.images]
+        elif config.task in ['RIVR', 'gauss_noise', 'dropout_noise', 'snp_noise']:
+            self.images = dataset["object_instance"]
+            self.descriptions = dataset["transcription_text"]
+            self.instance_names = dataset["object_instance"]
[…]

### 2021-11-19 Code/losses.py

@@ -1,0 +1,22 @@
+import torch
+import torch.nn.functional as F
+
+
+
+def mma_loss(positive, negative, models, margin=0.4):
+    '''
+    Simple MMA loss using cosine
+    '''
+    # cos_dist(a, p) - cos_dist(a, n) = 1 - cos(a,p) - (1- cos(a, n)) = cos(a, n) - cos(a, p)
+    triplet_loss = lambda a, p, n: torch.clamp(F.cosine_similarity(a, n) - F.cosine_similarity(a, p) + margin, 0.0, 2.0 + margin)
+
+    batch_loss = {'total': 0.0}
+    modalities = ['depth', 'language']
+    anchor = 'rgb'
+    loss = 0.0
+    
+    for modality in modalities:
+        loss = loss + torch.sum(triplet_loss(models[anchor](positive[anchor])['decoded'], models[modality](positive[modality])['decoded'], models[modality](negative[modality])['decoded']))
+
+    batch_loss['total'] = loss / len(positive['instance']) # average loss in this batch
+    return batch_loss

### 2021-11-19 Code/models.py

@@ -1,0 +1,26 @@
+import numpy as np
+import torch
+import torch.nn.functional as F
+import torchvision
+import flair
+
+class TheModel(torch.nn.Module):
+    def __init__(self, config, feature_size, device, embed_dim=1024):
+        super(TheModel, self).__init__()
+
+        self.config = config
+        self.device = device
+        # mapping layers to the shared space    
+        self.fc1 = torch.nn.Linear(feature_size, feature_size)
+        self.fc2 = torch.nn.Linear(feature_size, feature_size)
+        self.fc3 = torch.nn.Linear(feature_size, embed_dim)
+        # End of mapping layers
+        
+    def forward(self, feature):
+        output = {}
+        output['feature'] = feature.to(self.device)
+        output['decoded'] = F.relu(self.fc1(output['feature']))
+        output['decoded'] = F.relu(self.fc2(output['decoded']))
+        output['decoded'] = self.fc3(output['decoded'])
+
+        return output

### 2021-11-19 Code/run_cluster.sh

@@ -1,0 +1,85 @@
+#!/bin/bash
+#SBATCH --job-name=sim2real-gold
+#SBATCH --mem=80000
+#SBATCH --gres=gpu:1
+#SBATCH --cpus-per-task=16
+#SBATCH --time=108:00:00
+#SBATCH --constraint=rtx_8000|rtx_6000|rtx_2080
+#SBATCH --output=results/jobs/slurm-%j.out
+# Make sure to make the results/jobs directory before running this script.
+
+export TMPDIR=/scratch/${SLURM_JOB_ID}
+
+# Sample command for 1 job: sbatch run_cluster.sh my-experiment no_neg_sampling 1024 42 0.50
+
+exp_name=$1 # softmax, baseline, frozen, develop
+negative_sampling=$2 # no_neg_sampling , neg_sampling  , random_neg_sampling 
+embed_dim=$3 # 100 , 150 , 300, 550 This is the latent dimension
+random_seed=$4 # 32 , 64 , 128
+task=$5 # gold , RIVR , gauss_noise , dropout_noise , snp_noise
+# eval_mode=$2 # train , train-test , test
+# activation=$9 # softplus, relu, tanh
+
+eval_mode='train-test' # train , train-test , test
+activation='relu'
+
+# setup directories
+exp_full_name=$exp_name'-'$task'-'$negative_sampling'-'$embed_dim
+results_dir='results/'$exp_name'-'$task'-'$negative_sampling'-'$embed_dim'/seed-'$random_seed'/'
+train_result_dir=$results_dir'train/'
+valid_result_dir=$results_dir'valid/'
+test_result_dir=$results_dir'test/'
+all_results='results/all/'
+
+mkdir -p $results_dir
+mkdir -p $train_result_dir
+mkdir -p $valid_result_dir
+mkdir -p $test_result_dir
+mkdir -p $all_results
+
+# Mostly constant
+# task='gold' # 'gold' or 'RIVR'
+pred=0.50 # 0.40, 0.50, 0.60
+lr=0.001
+track=1 # 0 to turn off wandb and tensorboard tracking
+gpu_num=$(nvidia-smi --query-gpu=memory.free --format=csv,nounits,noheader | nl -v 0 | sort -nrk 2 | cut -f 1 | head -n 1 | xargs) # which one is free?
+
+# Frequently changing
+data_type='rgbd' # 'rgb', 'depth', 'rgbd'
[…]

### 2021-11-19 Code/utils.py

@@ -1,0 +1,175 @@
+import os
+import random
+import torch
+import numpy as np
+import scipy as sp
+import pickle, json
+import matplotlib.pyplot as plt
+from sklearn.metrics import precision_recall_fscore_support
+
+
+def set_seeds(config):
+	"""
+	To be able to perform multiple runs with different seed.
+    Also to report results that can be reproduced
+	"""
+	random.seed(config.random_seed)
+	np.random.seed(config.random_seed)
+	torch.manual_seed(config.random_seed)
+	torch.cuda.manual_seed(config.random_seed)
+	torch.cuda.manual_seed_all(config.random_seed)
+	torch.backends.cudnn.deterministic = True
+	torch.backends.cudnn.enabled = False
+	torch.backends.cudnn.benchmark = False
+
+
+    
+def setup_device(gpu_num=0):
+    device_name = 'cuda:'+str(gpu_num) if torch.cuda.is_available() else 'cpu'
+    device = torch.device(device_name)
+    return device
+
+
+
+def initialize_result_keeper(config):
+    """
+    Initialize empty dicts with the desired format.
+    Also loads results/ouputs/reconstruct to add
+    best model's results/ouputs/reconstructs on test set.
+    """
+    results = {
+        'best': {
+            'best-train': {'mif1': -1.0, 'bif1': -1.0, 'mrr': -1.0, 'acc': -1.0, 'f1': -1.0, 'precision': -1.0, 'recall': -1.0, "TP": -1, "TN": -1, "FP": -1, "FN": -1},
+            'best-valid': {'mif1': -1.0, 'bif1': -1.0, 'mrr': -1.0, 'acc': -1.0, 'f1': -1.0, 'precision': -1.0, 'recall': -1.0, "TP": -1, "TN": -1, "FP": -1, "FN": -1},
+            'best-test': {'mif1': -1.0, 'bif1': -1.0, 'mrr': -1.0, 'acc': -1.0, 'f1': -1.0, 'precision': -1.0, 'recall': -1.0, "TP": -1, "TN": -1, "FP": -1, "FN": -1}
+        },              
+    } # stores the metircs per epoch: results[epoch][portion][metric]
+
+    outputs = {
+        'train': {},
+        'valid': {},
+        'test': {}
+    } # outputs of the best model only: outputs[portion][model_outputs]. Contains the example for best model.
+    
+    reconstruct = {
+        'train': {},
+        'valid': {},
+        'test': {}
[…]

### 2022-01-13 Code/ML.py
speech 16kHz, simple contrastive loss

@@ -10,7 +10,7 @@ from sklearn.metrics.pairwise import euclidean_distances, cosine_similarity
 
 from models import TheModel
 from datasets import load_all_data
-from losses import mma_loss
+from losses import mma_loss, contrastive_loss
 from utils import set_seeds, setup_device, initialize_result_keeper, prf_metrics, mrr_acc_metrics
 
 
@@ -81,7 +81,8 @@ def training(config, models, dataset, portion, optimizers, epoch):
         for optimizer in optimizers.values():
             optimizer.zero_grad()
         
-        batch_loss = mma_loss(data['pos'], data['neg'], models)
+        # batch_loss = mma_loss(data['pos'], data['neg'], models)
+        batch_loss = contrastive_loss(data['pos'], data['neg'], models)
         # saving average loss per epoch. values in batch_loss have backward_fn and requires_grad
         running_loss.update(dict(zip(batch_loss.keys(), [running_loss[key] + batch_loss[key].item() for key in batch_loss.keys()] )))
         # if batch_index % 20 == 0:
@@ -171,10 +172,11 @@ def object_retrieval_task_sampling(config, language, rgb, depth, audio, object_n
         ar_matrix_distance = 1 - cosine_similarity(audio, rgb)
         ad_matrix_distance = 1 - cosine_similarity(audio, depth)
         la_matrix_distance = 1 - cosine_similarity(language, audio)
-        # matrix_distance = (lr_matrix_distance + ld_matrix_distance) / 2
+        matrix_distance = (lr_matrix_distance + ld_matrix_distance) / 2 # when textual language is anchor
+        # matrix_distance = (ar_matrix_distance + ad_matrix_distance) / 2 # when audio is anchor
         # matrix_distance = (lr_matrix_distance + ld_matrix_distance + la_matrix_distance) / 3
         # matrix_distance = (lr_matrix_distance + ld_matrix_distance + ar_matrix_distance + ad_matrix_distance) / 4
-        matrix_distance = (lr_matrix_distance + ld_matrix_distance + ar_matrix_distance + ad_matrix_distance + la_matrix_distance) / 5
+        # matrix_distance = (lr_matrix_distance + ld_matrix_distance + ar_matrix_distance + ad_matrix_distance + la_matrix_distance) / 5
         # TODO Save/return all these matrices
     elif config.distance == 'euclidean':
         lr_matrix_distance = euclidean_distances(language, rgb)

### 2022-01-13 Code/datasets.py

@@ -172,6 +172,7 @@ class GoLD_Dataset(Dataset):
             print('negative sampling in progress ...')
             # caculate distance matrix using language embeddings
             self.dists = scipy.spatial.distance.squareform(scipy.spatial.distance.pdist(self.language_embeddings_data, metric='cosine'))
+            # self.dists = scipy.spatial.distance.squareform(scipy.spatial.distance.pdist(self.audio_embeddings_data, metric='cosine'))
             for d in range(len(self.dists)):
                 self.dists[d] = scipy.stats.rankdata(self.dists[d], method='ordinal')
     
@@ -218,7 +219,7 @@ class GoLD_Dataset(Dataset):
         return sentence.get_embedding()
     
     def proc_speech(self, audio_name):
-        audio_file = self.root_dir + 'speech/' + audio_name + '.wav'
+        audio_file = self.root_dir + 'speech_16/' + audio_name + '.wav'
         speech, _ = sf.read(audio_file)
         input_values = self.auido_processor(speech, return_tensors="pt", sampling_rate=16000).input_values.cuda()
         with torch.no_grad():

### 2022-01-13 Code/losses.py

@@ -11,8 +11,8 @@ def mma_loss(positive, negative, models, margin=0.4):
     triplet_loss = lambda a, p, n: torch.clamp(F.cosine_similarity(a, n) - F.cosine_similarity(a, p) + margin, 0.0, 2.0 + margin)
 
     batch_loss = {'total': 0.0}
-    modalities = ['rgb', 'depth', 'audio']
-    anchor = 'language'
+    modalities = ['rgb', 'depth']
+    anchor = 'audio'
     loss = 0.0
     
     for modality in modalities:
@@ -20,3 +20,39 @@ def mma_loss(positive, negative, models, margin=0.4):
 
     batch_loss['total'] = loss / len(positive['instance']) # average loss in this batch
     return batch_loss
+
+
+
+def contrastive_loss(positive, negative, models):
+    batch_loss = {'total': 0.0}
+    modalities = ['rgb', 'depth']
+    anchor = 'language'
+    loss = 0.0
+    temperature = 0.1
+
+    for modality in modalities:
+        # numerator = torch.matmul(models[anchor](positive[anchor])['decoded'], models[modality](positive[modality])['decoded'].T)
+        # denominator_pos = torch.matmul(models[anchor](positive[anchor])['decoded'],models[modality](positive[modality])['decoded'].T)
+        # denominator_neg = torch.matmul(models[anchor](positive[anchor])['decoded'],models[modality](negative[modality])['decoded'].T)
+        
+        # numerator = torch.div(torch.diagonal(numerator), temperature)
+        # denominator_pos = torch.div(torch.diagonal(denominator_pos), temperature)
+        # denominator_neg = torch.div(torch.diagonal(denominator_neg), temperature)
+
+        numerator = torch.div(F.cosine_similarity(models[anchor](positive[anchor])['decoded'], models[modality](positive[modality])['decoded']), temperature)
+        denominator_pos = torch.div(F.cosine_similarity(models[anchor](positive[anchor])['decoded'],models[modality](positive[modality])['decoded']), temperature)
+        denominator_neg = torch.div(F.cosine_similarity(models[anchor](positive[anchor])['decoded'],models[modality](negative[modality])['decoded']), temperature)
+
+        denominator = torch.log(torch.exp(denominator_pos) + torch.exp(denominator_neg))
+
+        # print(f'numerator: {numerator}, denominator: {denominator}')
+        loss = loss + torch.sum(numerator - denominator)
+
[…]

### 2022-01-13 Code/run_cluster.sh

@@ -68,7 +68,7 @@ echo "copying gold dataset ..."
 mkdir -p $TMPDIR/data/gold
 org_data=../../../../data/gold/images
 rsync -a $org_data $TMPDIR/data/gold
-audio_data=../../../../data/gold/speech
+audio_data=../../../../data/gold/speech_16
 rsync -a $audio_data $TMPDIR/data/gold
 elif [[ " ${sim[*]} " =~ " $task " ]]
 then

### 2022-01-13 Code/run_cluster_loop.sh

@@ -13,7 +13,7 @@ for neg in neg_sampling
 do
 for dim in 1024
 do
-sbatch run_cluster.sh five-mma-simple-cosine-lang-anchor $neg $dim $seed $task
+sbatch run_cluster.sh my-contrastive-loss-cosine-text-anchor $neg $dim $seed $task
 done
 done
 done

### 2022-01-13 Paper/lit-search.tex
paper updated, ICML deadline coming

--- 2021-11-19 Paper/main.tex
@@
-% EMNLP header. Do not change.
-\pdfoutput=1
-\documentclass[11pt]{article}
-\usepackage[review]{emnlp2021}
-\usepackage{times}
-\usepackage{latexsym}
-\usepackage[T1]{fontenc}
-\usepackage[utf8]{inputenc}
-\usepackage{microtype}
-% End header.
-
-%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%
-% EDITING LINK: 
-% https://www.overleaf.com/3575751412sjsdmszkxsyh
-%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%
-
-\usepackage{soul}
-\usepackage{url}
-% \usepackage[hidelinks]{hyperref}
-% \usepackage[small]{caption}
-\usepackage{graphicx}
-\usepackage{amsmath}
-\usepackage{amsthm}
-\usepackage{booktabs}
-\usepackage{algorithm}
-\usepackage{algorithmic}
-\urlstyle{same}
-
-% =======================================
-% my imported packages
-
-\usepackage{amsmath}
-\usepackage{bm}
-\usepackage{amssymb}
-\usepackage{listings}
-% \usepackage[ruled,vlined]{algorithm2e}
-\include{pythonlisting}
-
-\newcommand{\pseudosection}[1]{\vspace{1.5ex}\noindent \textit{{#1:~~}}}
-% \newcommand{\pseudosection}[1]\indent{{{}}}
-
-\usepackage{todonotes}
-\newcommand{\todokdinline}[1]{\todo[color=red!20,inline]{{KD: \small #1}}}
-\newcommand{\todokd}[1]{\todo[color=red!20]{{\small #1 -- Kasra}}}
-\newcommand{\todocainline}[1]{\todo[color=yellow!20,inline]{{CA: \small #1}}}
-\newcommand{\todocm}[1]{\todo[color=green!40]{\small #1 -- Cynthia}}
-\newcommand{\todocmi}[1]{\todo[inline,color=green!40]{\small #1 -- Cynthia}}
-\newcommand{\todoff}[1]{\todo[color=blue!20]{\small #1 -- Frank}}
-\newcommand{\todoed}[1]{\todo[color=cyan!20]{\small #1 -- Ed}}
-
-\newcommand{\CitationNeeded}[1]{{\textbf {\color{red}Cite #1}}}
-\newcommand{\ProofRead}[1]{{\textbf {\color{red}Proofread please #1}}}
-\newcommand{\Rephrase}[1]{{\textbf {\color{red}Rephrase please #1}}}
-
-% \newcommand{\citet}[1]{\cite{#1}} % This is used because \citet didn't work with their author kit % It does now
-
-% =======================================
-
-\title{MMA: Multi-modal Manifold Alignment}
-
-\author{
-First Author$^1$
-\and
-Second Author$^2$\and
-Third Author$^{2,3}$\And
-Fourth Author$^4$
-\affiliations
-$^1$First Affiliation\\
-$^2$Second Affiliation\\
-$^3$Third Affiliation\\
-$^4$Fourth Affiliation
-\emails
-\{first, second\}@example.com,
-third@other.example.com,
-fourth@example.com
-}
-
-\begin{document}
-
-Todo list
-
-\begin{enumerate}
-    \item Lit search 
-\end{enumerate}
-
-\newpage
-
-\maketitle
-\setcounter{page}{1}
-
-% ***************** The BEGINNING of the paper *****************
-
-\begin{abstract}
-
-\end{abstract}
-
-%===================================================================
 \section{Lit Search}
-
 \subsection{Multimodal Machine Learning: A Survey and Taxonomy}
@@
 %===================================================================
-\section{Introduction}
-\label{intro}
-
-% Grounded language learning~\cite{Mooney2008Grounded,RichardsDarvishMatuszekCategoryFree20,triplet_loss_2021_CVPR} is a cross-modal learning task to connect natural language with the physical environment in which the natural language is used.
-
-% Our contributions are as follows:
-% \begin{enumerate}
-%     \item 
-%     \item 
-%     \item 
-% \end{enumerate}
-
-%===================================================================
-
-\section{Related Work}
-\label{sec:Related-Work}
-
-%===================================================================
-
-\section{Problem Description}
-\label{sec:Problem-Description}
-
-%===================================================================
-
-\section{Approach}
-\label{sec:Method}
-
-For $M$ modalities, we define a distance based loss function that contains $M+2(M-1)!$ terms. 
-We sample two data points from each modality, one positive point and one negative point.
-The objective is then to (I) minimize the distance between each pair of positive points from heterogeneous modalities which results in $(M-1)!$ terms, (II) maximize the distance between each pair of positive and negative points from heterogeneous modalities which contains $(M-1)!$ terms again, and (III) maximize the distance between each pair of positive and negative points from the homogeneous modalities which contains $M$ terms.
-Equation~\ref{eq:objective} shows the objective function we defined above with changing maximization to minimization by changing the sign.
-
-% \begin{equation}\label{eq:objective-detailed}
-%     \min \sum_{i=1}^{M-1} \sum_{j=i+1}^{M} ||x_{i}^{+} - x_{j}^{+}|| - \sum_{i=1}^{M-1} \sum_{j=i+1}^{N} ||x_{i}^{+} - x_{j}^{-}|| - 
-%     \sum_{i=1}^{M} ||x_{i}^{+} - x_{i}^{-}||
-% \end{equation}
-
-\begin{aligned}\label{eq:objective}
-    \mathcal{L}  &= \sum_{i=1}^{M-1} \sum_{j=i+1}^{M} ||f_i(x_{i}^{+}) - f_j(x_{j}^{+})||
-    - ||f_i(x_{i}^{+}) - f_j(x_{j}^{-})|| \\ 
-    & - ||f_i(x_{i}^{-}) - f_j(x_{j}^{+})|| - \sum_{i=1}^{M} ||f_i(x_{i}^{+}) - f_i(x_{i}^{-}) ||
-\end{aligned}
-where the subscripts $i$ and $j$ represent modalities, and $f_i(.)$ and $f_j(.)$ represent functions that map input data to the shared latent space using a deep neural network.
-
-\todokdinline{I can add another term for minimizing the distance between similar pairs (anchor and positive) from the same domain.
-This would look like the equation below.
-However, it breaks the nice procedure of sampling two data points per modality}
-$+ \sum_{i=1}^{M} ||f_i(x_i^{a}) - f_i(x_i^{+})||$
-
-Our method is a generalization of well-known similarity-based metric learning and manifold alignment methods such as triplet loss~\CitationNeeded{}.
-If $M=2$ which means the number of modalities is 2, then our objective function reduces to quadruplet loss method.
-If $M=2$ and we ignore the last term in the objective function, it results in the triplet loss method. 
-If $M=1$, only the last term remains in the loss function which is exactly the pairwise distance-based loss function.
-
-\subsection{Network Architecture}
-\label{sec:Model}
-
-%===================================================================
-
-\section{Experiments}
-\label{sec:Experiments}
-
-\subsection{Data}
-\label{sec:Data}
-
-\subsection{Setup}
-\label{sec:setup}
-
-\subsection{Quantitative Results}
-\label{sec:Quantitative}
-
-\subsection{Qualitative Results}
-\label{sec:Qualitative}
-
-\subsection{Baselines}
-
-%===================================================================
-
-\section{Conclusion}
-\label{sec:conclusion}
-
-%===================================================================
-
-% ***************** The END of the paper *****************
-
-\clearpage
-\section*{Acknowledgments}
-%% The file named.bst is a bibliography style file for BibTeX 0.99c
-\bibliographystyle{named}
-\bibliography{ref}
-\clearpage
-
-\appendix
-\label{sec:appendix}
-\section{Qualitative Analysis}
-
-\section{Model}
-
-\section{Experiments}
-
-\section{Results}
-
-\section{Setup}
-
-\end{document}

### 2022-01-13 Paper/main.tex
paper updated, ICML deadline coming

@@
 \usepackage{url}
-% \usepackage[hidelinks]{hyperref}
+% \usepackage[hidelinks]{hyperref}https://www.overleaf.com/project/614b5cc61d38655dce466002
 % \usepackage[small]{caption}
@@
 \newcommand{\ProofRead}[1]{{\textbf {\color{red}Proofread please #1}}}
+\newcommand{\Complete}[0]{{\textbf {\color{red}Complete it }}}
 \newcommand{\Rephrase}[1]{{\textbf {\color{red}Rephrase please #1}}}
@@
 
-\title{MMA: Multi-modal Manifold Alignment}
+\title{MMA: Multi-modal Manifold Alignment \\ MAME: Manifold Alignment of Multimodal Embeddings}
 
@@
 \begin{enumerate}
-    \item Lit search 
+    \item \st{Lit search}
 \end{enumerate}
@@
 \begin{abstract}
-
+    To be written ...
 \end{abstract}
@@
 %===================================================================
-\section{Lit Search}
 
-\subsection{Multimodal Machine Learning: A Survey and Taxonomy}
+\section{Introduction}
+\label{intro}
+Humans use multiple sensory inputs to interact with the world around them.
+Most methods in machine learning use at most two sensory inputs that are very different such as visual data (such as RGB images) and natural language.
+However, using extra similar inputs from different modalities is underexplored. This includes incorporating depth and RGB, written natural language and speech, and a combination of all of them.
+Most of the proposed works cannot be extended to arbitrary number of modalities, and we address this shortcomings by proposing a generalized distance-based loss function.
 
-\begin{itemize}
-\item \href{https://ieeexplore.ieee.org/document/8269806}{url}
-\end{itemize}
+% Grounded language learning~\cite{Mooney2008Grounded,RichardsDarvishMatuszekCategoryFree20,triplet_loss_2021_CVPR} is a cross-modal learning task to connect natural language with the physical environment in which the natural language is used.
 
-\citet{baltrusaitisMultimodalMachineLearning2019} proposes a new taxonomy of multimodal machine learning by introducing five technical challenges in addition to typical early and late fusion categorization.
-This taxonomy includes the following categories.
+\todokdinline{Are we doing contrastive learning? If so, mention it here}
+
+Our contributions are as follows:
 \begin{enumerate}
-\item \textit{Representation}: represent and summarize multimodal data in a way that exploits the complementarity and redundancy of multiple modalities.
-\begin{enumerate}
-\item joint: combine the unimodal signals into the same representation space.
-\item coordinated: process unimodal signals separately, but enforce certain similarity or structure constraints on them to bring them to a coordinated space.
+    \item Proposing a general distance-based loss function formulation for multimodal learning that can be extended to arbitrary number of modalities
+    \item Outperforming state-of-the-art models for grounded language learning by treating depth image as its own separate channel.
+    \item 
 \end{enumerate}
-\item \textit{Translation}: map data from one modality to another.
-\begin{enumerate}
-\item example-based
-\item generative
-\end{enumerate}
-\item \textit{Alignment}: identify the direct relations between (sub)elements from two or more different modalities.
-\begin{enumerate}
-\item explicit
-\item implicit
-\end{enumerate}
-\item \textit{Fusion}: join information from two or more modalities to perform a prediction.
-\begin{enumerate}
-\item model-agnostic
-\item model-based
-\end{enumerate}
-\item \textit{Co-learning}: transfer knowledge between modalities, their representation, and their predictive models
-\begin{enumerate}
-\item parallel data
-\item non-parallel data
-\item hybrid data
-\end{enumerate}
-\end{enumerate}
+%===================================================================
 
-The authors mention that while joint representations have been used in situations to construct representations of more than two modalities, coordinated spaces have, so far, been mostly limited to two.
-This means that our research is novel and we can extend similarity measures to more than two modalities.
+\section{Related Work}
+\label{sec:Related-Work}
+The growing number of multimodal datasets demonstrates the importance and the challenge of this task.
+\citet{bagher-zadeh-etal-2018-multimodal} introduce a large dataset of multimodal opinion sentiment and emotion intensity (CMU-MOSEI) for sentiment analysis and emotion recognition.
+They also introduce a novel multimodal fusion technique called the Dynamic Fusion Graph (DFG).
+\citet{GoLD_UMBC} present a multimodal dataset of household objects containing RGB images, depth images, written language, and spoken language.
 
-\subsection{Self-Supervised MultiModal Versatile Networks}
+\citet{baltrusaitisMultimodalMachineLearning2019} propose a new taxonomy of multimodal machine learning by introducing five technical challenges in addition to typical early and late fusion categorization.
+They cover different approaches to multimodal machine learning such as \textit{representation}, \textit{translation}, \textit{alignment}, \textit{fusion}, and \textit{co-learning}.
+In this project, we focus on alignment.
 
-\begin{itemize}
-\item \href{https://proceedings.neurips.cc/paper/2020/file/0060ef47b12160b9198302ebdb144dcf-Paper.pdf?utm\_campaign=NLP\%20News\&utm\_medium=email\&utm\_source=Revue\%20newsletter}{url}
-\end{itemize}
+We review different methods that have been proposed for alignment learning.
+\citet{alayrac2020self} use a self-supervised contrastive learning method to learn and combine representations from three modalities of visual, audio, and language.
+Audio and visual domains are mapped to the same space because they are fine-grained, containing dense information for each frame of video.
+These embeddings are then mapped to a coarse space where text is also mapped.
+Their network respects specificity and comparability of different modalities.
+The loss function they define consists of two distance-based terms for the corresponding spaces.
 
-This work uses self-supervised contrastive learning to learn and combine representations from three modalities of visual, audio, and language.
-They learn a \textit{multimodal versatile network} that has the following four properties:
-\begin{enumerate}
-\item Ability to take any of the modalities as input
-\item Respecting the specificity of modalities: audio and visual modalities are fine-grained while language modality is coarse-grained.
-\item Comparability of different modalities using \textbf{dot product} even if not seen together during training 
-\item Efficiently applicable to visual data either in the form of dynamic videos or static images
-\end{enumerate}
+\citet{Nguyen-RSS-20} take a similar approach and perform pairwise cosine similarity to align images and natural language text descriptions to perform object retrieval.
 
-The authors consider the following three different configurations of the modality spaces which they call \textit{modality embedding graphs}.
-\begin{enumerate}
-\item Shared: all modalities map to the same space. This respects property 3 but violates property 2.
-\item Disjoint: visual-audio and visual-text spaces. This respects property 2 but violates property 3.
-\item Fine and coarse: audio and visual domains are map to the same space since they are fine-grained.
-These embeddings are then mapped to a lower dimensional space where text is also mapped.
-This respects both properties 2 and 3.
-\end{enumerate}
+\citet{triplet_loss_2021_CVPR} take a cross-modal manifold alignment approach to grounded language learning using triplet loss.
+They perform object retrieval by connecting text descriptions of objects to their corresponding RGB-D images.
 
-The loss function they define consists of two terms.
-The first term is to train the fine-grained space of visual and audio embeddings by minimizing the distance between the similar pair (positives) from the same location of a video, and maximizing the distance between negative pairs (negatives) sampled from different videos.
-The second term is to train the coarse-grained space which consists of visual and text embeddings.
-Note that they don't consider audio embeddings in this space since they don't want to learn automatic speech recognition (ASR).
-Text and visual domains are not aligned as well as audio and visual domains (e.g.
-sound of playing piano versus the text that only says playing piano for a video of playing piano).
-To cure this, they consider a set of positive pairs instead of a single pair.
-If a modality is missed, the corresponding term is dropped from the loss function.
+Another related approach is contrastive learning which is based on the idea of similarity and distance~\cite{NEURIPS2020_supervised_contrastive,chen2020simple}.
+Triplet loss is also a special case of contrastive loss~\cite{NEURIPS2020_supervised_contrastive}.
+Contrastive loss is mostly used in self-supervised learning~\cite{alayrac2020self,chen2020simple} with the exception of the work proposed by \citet{NEURIPS2020_supervised_contrastive} where they use contrastive loss for a supervised problem.
+Most of these methods are implemented for a unimodal dataset (i.e.
+dataset of one type of data such as RGB images only).
 
-\subsection{Separating Self-Expression and Visual Content in Hashtag Supervision}
-
-\begin{itemize}
-\item \href{https://openaccess.thecvf.com/content\_cvpr\_2018/papers/Veit\_Separating\_Self-Expression\_and\_CVPR\_2018\_paper.pdf}{url}
-\end{itemize}
-
-The authors train a joint model of images, hashtags, and users to perform image retrieval.
-This is similar to our task where given a description, we want to find the image in the scene that best matches the description.
-The idea is to form a three-way tensor product model.
+Some of the related works take more than two input modalities.
+\citet{Veit_2018_CVPR} train a joint model of images, hashtags, and users to perform image retrieval.
+Their task is similar to our task where given a description, we want to find the image/object in the scene that best matches the description.
+Their idea to tackle this problem is to form a three-way tensor product model.
 They use a ranking loss to train the model where the score of an observed triplet is higher than an unobserved triplet.
@@
 
-\subsection{Semi-Heterogeneous Three-Way Joint Embedding Network for Sketch-Based Image Retrieval}
+\citet{semihet_three_way_Lei_2020} use inputs from three modalities of image, sketch, and edgemap to perform image retrieval given the sketch.
+They use co-attention mechanism, and their loss function contains alignment loss, cross-entropy loss, and sketch-edgemap contrastive loss.
+\citet{Mittal2020M3ER} use Canonical Correlational Analysis (CCA) to differentiate between ineffective and effective modalities for the emotion recognition task from multiple input modalities including face, speech, and text.
 
-\begin{itemize}
-\item \href{https://ieeexplore.ieee.org/document/8809264}{url}
-\end{itemize}
+Multimodal learning has been used for different applications including but not limited to object retrieval.
+\citet{Deception_ICMI_2014} use \textit{language}, \textit{physiological response}, and \textit{thermal sensing} to detect deceit.
+\citet{het_data_fusion_liu_IEEE_2017} take two modalities as input and output predictions in a third modality.
 
-\subsection{Deep Multimodal Learning for Affective Analysis and Retrieval}
+\todo[inline]{talk about CLIP and DALL-E here}
 
-\begin{itemize}
-\item \href{https://ieeexplore.ieee.org/abstract/document/7277066?casa\_token=aBp6BxcszHwAAAAA:3NoMiFrZbn7tXfavF1rgkCiGWbFI2arxn8Xb6iDF79q4zBZHWi7PWWhf6xW-xJwYdFALbmRENo4}{url}
-\end{itemize}
-
-\subsection{An Efficient Framework for Zero-Shot Sketch-Based Image Retrieval}
-
-\begin{itemize}
-\item \href{https://arxiv.org/pdf/2102.04016.pdf}{url}
-\end{itemize}
-This paper uses two modalities only; image and sketch.
-However, it uses a \textit{quadruplet} to compute the loss which can be useful in our research.
-A quadruplet is composed of a sketch picture as an anchor, a negative example from sketch domain, a negative example from picture domain, and a positive example from picture domain.
-
-\subsection{SMIL: Multimodal Learning with Severely Missing Modality}
-
-\begin{itemize}
-\item \href{https://www.aaai.org/AAAI21Papers/AAAI-437.MaM.pdf}{url}
-\end{itemize}
-Uses Bayesian meta-learning framework to perform multimodal learning with partially missing modalities in training/test data.
-One of their experiments is multi-label classification of movie genres with bimodal data including poster of movies and description of the movie from IMDB.
-
-\subsection{Multimodal Language Analysis in the Wild: CMU-MOSEI Dataset and Interpretable Dynamic Fusion Graph}
-
-\begin{itemize}
-\item \href{https://aclanthology.org/P18-1208.pdf}{url}
-\end{itemize}
-
-\subsection{Multimodal Learning with Incomplete Modalities by Knowledge Distillation}
-
-\begin{itemize}
-\item \href{https://dl.acm.org/doi/pdf/10.1145/3394486.3403234}{url}
-\end{itemize}
-They first train models on each modality independently
-\subsection{SWAFN: Sentimental Words Aware Fusion Network for Multimodal Sentiment Analysis}
-
-\begin{itemize}
-\item \href{https://aclanthology.org/2020.coling-main.93.pdf}{url}
-\end{itemize}
-
-\subsection{Multimodal Learning for Human Action Recognition Via Bimodal/Multimodal Hybrid Centroid Canonical Correlation Analysis}
-
-\begin{itemize}
-\item \href{https://ieeexplore.ieee.org/document/8489981}{url}
-\end{itemize}
-
-\subsection{Deception Detection Using a Multimodal Approach}
-
-\begin{itemize}
-\item \href{https://dl.acm.org/doi/pdf/10.1145/2663204.2663229?casa\_token=KneK7B7xvLwAAAAA:Ajz0N96ygq8ktwkz0IVUqTt8NozCg2wR6n\_x2xntdHqZBh6VXW\_8VbO4GeY4VvMsDJlMwzkhVXQSJA}{url}
-\end{itemize}
-They use \textit{language}, \textit{physiological response}, and \textit{thermal sensing} to detect deceit.
-
-\subsection{M3ER: Multiplicative Multimodal Emotion Recognition using Facial, Textual, and Speech Cues}
-
-\begin{itemize}
-\item \href{https://ojs.aaai.org/index.php/AAAI/article/view/5492}{url}
-\end{itemize}
-
-\subsection{Found in Translation: Learning Robust Joint Representations by Cyclic Translations between Modalities}
-
-\begin{itemize}
-\item \href{https://ojs.aaai.org/index.php/AAAI/article/view/4666}{url}
-\end{itemize}
-\subsection{Select-Additive Learning: Improving Generalization In Multimodal Sentiment Analysis}
-
-\begin{itemize}
-\item \href{https://ieeexplore.ieee.org/stamp/stamp.jsp?tp=\&arnumber=8019301}{url}
-\end{itemize}
-
-\subsection{Attention-Based Multimodal Fusion for Video Description}
-
-\begin{itemize}
-\item \href{https://openaccess.thecvf.com/content\_ICCV\_2017/papers/Hori\_Attention-Based\_Multimodal\_Fusion\_ICCV\_2017\_paper.pdf}{url}
-\end{itemize}
-\subsection{\st{3W-AlignNet: a Feature Alignment Framework for Person Search with Three-Way Decision Theory}}
-
-\begin{itemize}
-\item \href{https://link.springer.com/article/10.1007/s12559-021-09898-7}{url}
-\end{itemize}
-
-This is not very related since they use three-way decision theory to select bounding boxes as positive, negative, and boundary.
-Three-way here refers to something else and not modalities.
-
-\subsection{\st{Unsupervised Domain Adaptation for Face Recognition in Unlabeled Videos}}
-\begin{itemize}
-\item \href{https://openaccess.thecvf.com/content\_ICCV\_2017/papers/Sohn\_Unsupervised\_Domain\_Adaptation\_ICCV\_2017\_paper.pdf}{url}
-\end{itemize}
-Might be relevant but not super relevant.
-
-\subsection{Heterogeneous Sensor Data Fusion By Deep Multimodal Encoding}
-\begin{itemize}
-\item \href{https://ieeexplore.ieee.org/stamp/stamp.jsp?tp=&arnumber=7874158}{url}
-\end{itemize}
-This one takes two modalities as input and outputs predictions in a third modality.
-
-%===================================================================
-\section{Introduction}
-\label{intro}
-
-% Grounded language learning~\cite{Mooney2008Grounded,RichardsDarvishMatuszekCategoryFree20,triplet_loss_2021_CVPR} is a cross-modal learning task to connect natural language with the physical environment in which the natural language is used.
-
-% Our contributions are as follows:
-% \begin{enumerate}
-%     \item 
-%     \item 
-%     \item 
-% \end{enumerate}
-
-%===================================================================
-
-\section{Related Work}
-\label{sec:Related-Work}
+However, to the best of our knowledge, there has not been any work to incorporate more than 3 modalities, and that's the novelty of our work.
 
@@
 
-\begin{aligned}\label{eq:objective}
-    \mathcal{L}  &= \sum_{i=1}^{M-1} \sum_{j=i+1}^{M} ||f_i(x_{i}^{+}) - f_j(x_{j}^{+})||
-    - ||f_i(x_{i}^{+}) - f_j(x_{j}^{-})|| \\ 
-    & - ||f_i(x_{i}^{-}) - f_j(x_{j}^{+})|| - \sum_{i=1}^{M} ||f_i(x_{i}^{+}) - f_i(x_{i}^{-}) ||
-\end{aligned}
+\begin{equation}\label{eq:objective}
+\begin{split}
+    \mathcal{L}  &= \sum_{i=1}^{M-1} \sum_{j=i+1}^{M} ||f_i(x_{i}^{+}) - f_j(x_{j}^{+})|| \\
+    & - ||f_i(x_{i}^{+}) - f_j(x_{j}^{-})||
+    - ||f_i(x_{i}^{-}) - f_j(x_{j}^{+})|| \\
+    & - \sum_{i=1}^{M} ||f_i(x_{i}^{+}) - f_i(x_{i}^{-}) ||
+\end{split}
+\end{equation}
 where the subscripts $i$ and $j$ represent modalities, and $f_i(.)$ and $f_j(.)$ represent functions that map input data to the shared latent space using a deep neural network.
@@
 
-Our method is a generalization of well-known similarity-based metric learning and manifold alignment methods such as triplet loss~\CitationNeeded{}.
+Our method is a generalization of well-known similarity-based metric learning and manifold alignment methods such as triplet loss~\cite{Carvalho-cooking-triplet,triplet_loss_2021_CVPR}.
 If $M=2$ which means the number of modalities is 2, then our objective function reduces to quadruplet loss method.
@@
 
+\subsection{Loss Functions}
+We try three different distance based loss functions.
+In order to use cosine distance, we have to subtract the cosine between two embeddings from 1 which means $1 - \cos(e_1, e_2)$.
+
+\subsubsection{Simple triplet extension}
+\label{sub:simple-mma}
+
+First one is a modification of the triplet loss function idea.
+For each data point (referring to an instance of an object with all of its modalities), we choose a negative data point either based on object class name, or by negative sampling~\cite{Pillai_Matuszek_2018}.
+We now have two data points; we consider the first one as positive, and the second one as negative.
+However, the anchor won't be another data point as it is in the triplet loss terminology.
+Instead we choose one domain as an anchor.
+
+\begin{itemize}
+    \item data point: a row in the csv/tsv file.
+    \item Positive: a set of embeddings of the one data point (rgb image of an apple, depth image of an apple, language description of an apple)
+    \item Negative: a set of embeddings of the another data point of a different object (rgb image of a book, depth image of a book, language description of a book)
+    \item Anchor: A fixed modality to act as anchor in the case of simple MMA loss, where we don't minimize the distance between the pair of embeddings from all modalities of the same data point, and we don't maximize the distance between positive an negative datapoints of the same modality.
+\end{itemize}
+
+Since the downstream task is to predict/retrieve the desired object among multiple objects given a natural language description, it makes sense to train the model using language as anchor. 
+We tried RGB as anchor, and as expected the performance was worse.
+
+Equation~\ref{eq:objective-simple-mma} shows the mathematical formulation of the proposed loss function which can be generalized to arbitrary number of modalities.
+
+\begin{equation}
+\label{eq:objective-simple-mma}
+\begin{split}
+    \mathcal{L}  &= \sum_{m=1}^{M-1} || f_a(x_{a}^{+}) - f_m(x_{m}^{+}) || - \\ 
+    & || f_a(x_{a}^{+}) -f_m(x_{m}^{-}) || 
+    \\
+    &= \sum_{m=1}^{M-1} 1 - \cos(f_a(x_{a}^{+}), f_m(x_{m}^{+})) \\ & - (1 - \cos(f_a(x_{a}^{+}) ,f_m(x_{m}^{-}))) \\
+    &= \sum_{m=1}^{M-1} \cos(f_a(x_{a}^{+}) ,f_m(x_{m}^{-})) \\ 
+    & - \cos(f_a(x_{a}^{+}), f_m(x_{m}^{+}))
+\end{split}
+\end{equation}
+
+where $a$ represents \textit{anchor} modality, $M$ is the number of modalities, $p$ represents the positive objects, and $n$ represents the negative objects.
+
+An illustration of this loss function is provided in figure~\ref{fig:simple-mma-loss}.
+
+This version of MMA loss function can be seen as a contrastive loss usually used in the domain of self-supervised learning~\cite{chen2020simple}.
+However, our proposed loss function has two advantages over the traditional contrastive loss expressed in equation~\ref{eq:contrastive-loss}.
+The first advantage is that our loss function do not loop over multiple positives and negatives in a large batch.
+Instead we sample only two objects (positive and negative) each of which have $M$ modalities which gives us $2M$ datapoints (or embeddings).
+Hence, our model can be trained using smaller batch sizes and reduces the number of negative samples we need.
+Second advantage is that this loss function can be used in a multimodal setting with an arbitrary number of modalities, and is not limited to a single data type (e.g.
+RGB images) which is the most common usage of contrastive loss~\cite{chen2020simple, NEURIPS2020_supervised_contrastive}.
+
+\begin{equation}\label{eq:contrastive-loss}
+    -\sum_{i \in I} \log \frac{\exp (z_i . z_{j(i)} / \tau) }{\sum_{a \in A(i)} \exp (z_i.z_a / \tau)}
+\end{equation}
+
+Therefore, our proposed loss function would modify the equation~\ref{eq:contrastive-loss} to the loss function formulated in equation~\ref{eq:simple-MMA-contrastive}.
+Our formulation is a generalized version of the contrastive loss where $M$ is the number of modalities (e.g.
+RGB image, depth image, speech, text), but it can be the ``augmented views'' instead of different modalities.
+
+\begin{equation}\label{eq:simple-MMA-contrastive}
+    -\sum_{m=1}^{M-1} \log \frac{\exp (z_a . z_{m}^{+}/ \tau) }{ \exp (z_a.z_m^{+} / \tau) + \exp (z_a.z_m^{-} / \tau)}
+\end{equation}
+
+\todokdinline{I should run the traditional contrastive loss to see if my method works better to make this claim.}
+
+\begin{figure*}[tbh]
+\centering
+\includegraphics[width=2.0\columnwidth]{Figures/simple-MMA.pdf}
+\caption{A high-level prototype of the distances used in the simple MMA loss with 3 modalities.
+Orange arrows indicate maximizing the distance, blue arrows indicate minimizing the distance, and green arrows show the margin.}
+\label{fig:simple-mma-loss}
+\end{figure*}
+
+\subsubsection{All in loss with 12 terms}
+\subsubsection{subset of All in loss with 9 terms}
+
 \subsection{Network Architecture}
 \label{sec:Model}
+
+We use BERT \CitationNeeded{} to featurize the textual input, and wave2vec2 \CitationNeeded{} to extract audio embeddings from speech both of which output a 3072-dimensional embedding vector.
+To process images, we use ResNet152 \CitationNeeded{} for both RGB and depth images which gives us a 2048-dimensional embedding vector.
+We then use 3 fully connected layers to map these embeddings to a shared 1024-dimensional space where we can compute the distance between all embeddings.
 
@@
 
+\subsection{4 Modalities}
+We ran an experiment with RGB, depth, language.
+The loss function doesn't require any changes beyond increasing the number of loops in equation~\ref{eq:objective-simple-mma} (or increasing the value of $M$ by 1).
+The non-trivial part is the downstream prediction task. In the case of three modalities, language is the anchor, and we compute the distance of other two modalities from it and then average them.
+With speech coming into the picture, we have three choices.
+First, treat it in a similar way to RGB and depth; which means computing the distance of RGB, depth, and speech from text, and then take an average of three of them.
+Second way is to compute the distance of RGB and depth from language and from speech which gives us 4 distance matrices, and then take average of the 4 of them.
+Third way is similar to the second methods, but we compute the distance between language and speech as well and then take average of 5 distance matrices.
+The problem with second and third method, is that the simple MMA loss function in equation~\ref{eq:objective-simple-mma} is not designed in a way to minimize the distance between speech and visual embeddings explicitly.
+This is supported by our experimental results where the first method results in a better performance compared to the other two methods.
+Third method (averaging 5 matrices) is the worst, then second one (averaging 4 matrices) is bad. First method has the best results.
+
+\todokdinline{try the more complicated MMA that might be able to capture these relationships.}
+
+\subsection{Contrastive Loss}
+We compare our model against contrastive loss as formulated in equation~\ref{eq:contrastive-loss}.
+In order to implement this loss function, we use cosine similarity as suggested in the SimCLR paper~\cite{chen2020simple}.
+The other version mentioned by~\citet{NEURIPS2020_supervised_contrastive} uses inner dot product which can lead to instabilities since the dot product is not bounded.
+\todokdinline{try contrastive loss with cross-entropy, dot product, and the supervised contrastive method itself (without triplets).}
+
+\subsection{Metrics}
+\label{sec:metrics}
+To evaluate our model we measure different performance metrics on a retrieval task where the model has to select 1 object among 5 objects given a language description.
+Only one of the objects corresponds to the description and the rest are from different object classes.
+\todokdinline{Try with different number of objects.}
+
+To evaluate the performance, we compute the cosine distance between a given natural language description and 5 randomly selected images (1 of which corresponds to the description, and others are from different object classes).
+We have to modify the way we compute the distance since we treat depth images as a separate input modality and do not append it to the RGB features as in the triplet loss method~\cite{triplet_loss_2021_CVPR}.
+To handle this, we compute the distance between 1 language embedding and all candidate RGB embeddings, and we compute the distance between the same language embedding and all candidate depth embeddings corresponding to the RGB embeddings.
+We then take average of these two distance matrices.
+
+Moreover, instead of choosing an arbitrary threshold, we choose the closest image embedding (average distance of rgb and depth from language) as the prediction.
+
+The best metric to capture the performance in such a scenario would be mean reciprocal rank (MRR) and accuracy.
+Accuracy, micro F1 score (sklearn), flatten binary f1 score (sklearn), and my f1 score computation are all the same in this task since for each prediction we either have a true positive and no false positives and no false negatives (no FP no FN), or we have no true positives and 1 false positive and 1 false negative.
+However, subset MRR is equal or greater than those because if the model do not rank the correct answer as 1, it still gets some score and is greater than 0 (in accuracy the score is ether 0 or 1).
+
 \subsection{Data}
 \label{sec:Data}
+
+We use a recent multimodal dataset called GoLD~\cite{GoLD_UMBC}.
+The original paper uses raw RGB and raw depth images in which other items are present in the background.
+We use a masked version of the images where the background (including the white turntable) is deleted and one object is present in each image.
+We show that a masked version converges faster, however, they both converge to the same performance after enough epoch as shown  in figure~\ref{fig:mask-vs-raw}.
+
+\begin{figure*}[tbh]
+\centering
+\includegraphics[width=2.0\columnwidth]{Figures/raw-mask-test-mrr.pdf}
+\caption{Mean Reciprocal Rank (MRR) on different versions of GoLD dataset~\cite{GoLD_UMBC}.
+The raw version is depicted in green and masked version is depicted in orange averaged over 3 different random seeds.
+The higher is better.}
+\label{fig:mask-vs-raw}
+\end{figure*}
+
+Since the wave2vec2~\CitationNeeded{} model is trained on 16hz speech data, we need to convert he speech files to 16hz.
 
@@
 
+We use \Complete optimizer with a learning rate of \Complete.
+The models are trained for \Complete epochs on a \Complete GPU.
+
+\subsection{Baselines}
+We compare our MMA model against the triplet loss method proposed by~\citet{GoLD_UMBC} and~\citet{triplet_loss_2021_CVPR}.
+Triplet loss function consists of three data points including anchor, positive, and negative from two modalities.
+The anchor, positive, and negative can be chosen from different modalities in each batch.
+The triplet loss method has two major disadvantages.
+First, it cannot be used for more than three modalities since there are only 3 data points in the loss function.
+Second, a batch size of greater than 1 cannot be easily implemented if the anchor, positive, and negative come from different modalities for each batch.
+To address this issues, we sample more than one positive and one negative data points.
+Essentially, we sample one positive data point and one negative data point from each modality which results in 2(M-1) + 1 data points where 1 is the single anchor data point from anchor modality of the same object class as positive datapoints.
+We would have 2M datapoints if we use the negative data point from the anchor modality too.
+The anchor data point is always from the same modality and corresponds to the same object class sampled for positive data points.
+
 \subsection{Quantitative Results}
@@
 
+We evaluate our proposed simple MMA loss function~\ref{sub:simple-mma} against the triplet loss method~\cite{triplet_loss_2021_CVPR}.
+Our model significantly outperforms the triplet method across all epochs while it converges faster as shown in figure~\ref{fig:simple-MMA-mrr}.
+Table~\ref{table:quantitative} summarizes the performance of all models.
+
+\begin{table*}[tbh]
+\centering
+\begin{tabular}{|c|c|c|}
+\toprule
+\textbf{Method} & \textbf{Accuracy} & \textbf{MRR} \\ %\hline
+\midrule
+Full MMA (Text) w/ neg & ?$\pm$? & ?$\pm$? \\
+Simple MMA (Text) w/ neg & \textbf{0.7949}$\pm$? & \textbf{0.8778}$\pm$? \\
+Contrastive (Text) w/ neg & ?$\pm$? & ?$\pm$? \\
+Triplet (Text) w/ neg & 0.7497$\pm$? & 0.8522$\pm$? \\
+Full MMA (Speech) w/ neg & ?$\pm$? & ?$\pm$? \\
+Simple MMA (Speech) w/ neg & \textbf{0.7949}$\pm$? & \textbf{0.8778}$\pm$? \\
+Contrastive (Speech) w/ neg & ?$\pm$? & ?$\pm$? \\
+Triplet (Speech) w/ neg & 0.7497$\pm$? & 0.8522$\pm$? \\
+Triplet (Speech) w/o neg & 0.7497$\pm$? & 0.8522$\pm$? \\
+\bottomrule
+\end{tabular}
+\caption{\label{table:quantitative}Average and standard deviation of accuracy and MRR over 3 runs with 3 different random seeds on a hel-out test set.
+Both metrics are from 0 to 1, and the higher is better.\todokdinline{maybe adding results on cropped and raw dataset as well.
+Also results using objec class isntead of negative sampling}}
+\end{table*}
+
+\begin{figure*}[tbh]
+\centering
+\includegraphics[width=2.0\columnwidth]{Figures/test.mrr.pdf}
+\caption{Mean Reciprocal Rank (MRR) of our proposed method against triplet loss method averaged over 3 different random seeds for a downstream task of object detection.
+The higher is better.
+Blue represents our proposed simple MMA loss function, and orange represent the triplet loss function~\cite{triplet_loss_2021_CVPR}.}
+\label{fig:simple-MMA-mrr}
+\end{figure*}
+
 \subsection{Qualitative Results}
 \label{sec:Qualitative}
-
-\subsection{Baselines}
 
@@
 \clearpage
+\clearpage
 \section*{Acknowledgments}
@@
 \label{sec:appendix}
+
 \section{Qualitative Analysis}
@@
 
+\input{lit-search}
+
 \end{document}

### 2022-01-15 Code/ML.py
supervised contrastive loss added

@@ -10,21 +10,27 @@ from sklearn.metrics.pairwise import euclidean_distances, cosine_similarity
 
 from models import TheModel
 from datasets import load_all_data
-from losses import mma_loss, contrastive_loss
-from utils import set_seeds, setup_device, initialize_result_keeper, prf_metrics, mrr_acc_metrics
+from losses import mma_loss, contrastive_loss, SupConLoss
+from utils import set_seeds, setup_device, initialize_result_keeper, prf_metrics, mrr_acc_metrics, adjust_learning_rate
 
 
-def train(config, models, dataset):
+def train(config, models, dataset, device):
 
     results, outputs, logs, reconstruct, examples = initialize_result_keeper(config)    
     
+    criterion = SupConLoss().to(device)
     # Initialize the optimizer
     optimizers = {}
     for modality in models.keys():
         model = models[modality]
-        optimizers[modality] = torch.optim.Adam(model.parameters(), lr=config.learning_rate)
+        if config.method == 'MMA':
+            optimizers[modality] = torch.optim.Adam(model.parameters(), lr=config.learning_rate)
+        elif config.method == 'supervised-contrastive':
+            optimizers[modality] = torch.optim.SGD(model.parameters(), lr=0.05, momentum=0.9, weight_decay=1e-4)
     
     for epoch in tqdm(range(config.epochs)):
+        if config.method == 'supervised-contrastive':
+            adjust_learning_rate(config, optimizers, epoch+1)
         for modality in models.keys():
             model = models[modality]
             model.train()
@@ -32,7 +38,7 @@ def train(config, models, dataset):
         results[epoch] = {}
         # do the training
         print('\ntraining for the new epoch')
-        logs[epoch] = training(config, models, dataset, 'train', optimizers, epoch)
+        logs[epoch] = training(config, models, dataset, 'train', optimizers, epoch, criterion)
 
         # evaluation on valid and possibly test        
         for portion in config.portions:
@@ -68,7 +74,7 @@ def train(config, models, dataset):
 
 
 
-def training(config, models, dataset, portion, optimizers, epoch):
+def training(config, models, dataset, portion, optimizers, epoch, criterion):
     """
     performs one epoch training loop over all data
     """
@@ -76,13 +82,19 @@ def training(config, models, dataset, portion, optimizers, epoch):
     running_loss = {
         'total': 0.0,
     }
-        
+
     for batch_index, data in enumerate(dataset[portion]):
         for optimizer in optimizers.values():
[…]

### 2022-01-15 Code/losses.py

@@ -1,5 +1,7 @@
 import torch
 import torch.nn.functional as F
+import torch.nn as nn
+from sklearn import preprocessing
 
 
 
@@ -12,7 +14,7 @@ def mma_loss(positive, negative, models, margin=0.4):
 
     batch_loss = {'total': 0.0}
     modalities = ['rgb', 'depth']
-    anchor = 'audio'
+    anchor = 'language'
     loss = 0.0
     
     for modality in modalities:
@@ -56,3 +58,103 @@ def contrastive_loss(positive, negative, models):
 
 def original_contrastive_loss():
     pass
+
+
+# Code adopted from https://github.com/HobbitLong/SupContrast
+class SupConLoss(nn.Module):
+    """
+    Author: Yonglong Tian (yonglong@mit.edu)
+    Date: May 07, 2020
+    Supervised Contrastive Learning: https://arxiv.org/pdf/2004.11362.pdf.
+    It also supports the unsupervised contrastive loss in SimCLR
+    """
+    def __init__(self, temperature=0.07, contrast_mode='all',
+                 base_temperature=0.07):
+        super(SupConLoss, self).__init__()
+        self.temperature = temperature
+        self.contrast_mode = contrast_mode
+        self.base_temperature = base_temperature
+
+    def forward(self, features, labels=None, mask=None):
+        """Compute loss for model. If both `labels` and `mask` are None,
+        it degenerates to SimCLR unsupervised loss:
+        https://arxiv.org/pdf/2002.05709.pdf
+        Args:
+            features: hidden vector of shape [bsz, n_views, ...].
+            labels: ground truth of shape [bsz].
+            mask: contrastive mask of shape [bsz, bsz], mask_{i,j}=1 if sample j
+                has the same class as sample i. Can be asymmetric.
+        Returns:
+            A loss scalar.
+        """
+        batch_loss = {'total': 0.0}
+        device = (torch.device('cuda')
+                  if features.is_cuda
+                  else torch.device('cpu'))
+
+        if len(features.shape) < 3:
+            raise ValueError('`features` needs to be [bsz, n_views, ...],'
+                             'at least 3 dimensions are required')
+        if len(features.shape) > 3:
+            features = features.view(features.shape[0], features.shape[1], -1)
+
+        batch_size = features.shape[0]
+        if labels is not None and mask is not None:
+            raise ValueError('Cannot define both `labels` and `mask`')
+        elif labels is None and mask is None:
+            mask = torch.eye(batch_size, dtype=torch.float32).to(device)
[…]

### 2022-01-15 Code/utils.py

@@ -6,6 +6,7 @@ import scipy as sp
 import pickle, json
 import matplotlib.pyplot as plt
 from sklearn.metrics import precision_recall_fscore_support
+import math
 
 
 def set_seeds(config):
@@ -173,3 +174,24 @@ def Visualize(pkldata,title,has_x_axis,xlablel,ylabel,xscale,yscale,legend,figsi
         plt.legend(loc='lower right')
     plt.savefig(location+title+'.png')
     plt.close('all')
+
+
+
+def adjust_learning_rate(args, optimizers, epoch):
+    # Code adopted from https://github.com/HobbitLong/SupContrast
+    lr = args.learning_rate
+    args.lr_decay_rate = 0.1
+    args.lr_decay_epochs = [120, 150, 190]
+    args.cosine = True
+    for optimizer in optimizers.values():
+        if args.cosine:
+            eta_min = lr * (args.lr_decay_rate ** 3)
+            lr = eta_min + (lr - eta_min) * (
+                    1 + math.cos(math.pi * epoch / args.epochs)) / 2
+        else:
+            steps = np.sum(epoch > np.asarray(args.lr_decay_epochs))
+            if steps > 0:
+                lr = lr * (args.lr_decay_rate ** steps)
+
+        for param_group in optimizer.param_groups:
+            param_group['lr'] = lr

### 2022-01-25 Code/ML.py
EMMA: explicit anchors loss, schedule lr for Adam

@@ -1,3 +1,4 @@
+from cmath import exp
 import torch
 import numpy as np
 import random
@@ -10,7 +11,7 @@ from sklearn.metrics.pairwise import euclidean_distances, cosine_similarity
 
 from models import TheModel
 from datasets import load_all_data
-from losses import mma_loss, contrastive_loss, SupConLoss
+from losses import mma_loss, contrastive_loss, SupConLoss, explicit_anchor_mma_loss
 from utils import set_seeds, setup_device, initialize_result_keeper, prf_metrics, mrr_acc_metrics, adjust_learning_rate
 
 
@@ -21,16 +22,22 @@ def train(config, models, dataset, device):
     criterion = SupConLoss().to(device)
     # Initialize the optimizer
     optimizers = {}
+    schedulers = {}
     for modality in models.keys():
         model = models[modality]
         if config.optimizer == 'Adam':
             optimizers[modality] = torch.optim.Adam(model.parameters(), lr=config.learning_rate)
+            lr_lambda = lambda epoch : config.learning_rate if epoch < int(config.epochs/3) else (config.learning_rate*0.1 if epoch < int(config.epochs/3) * 2 else config.learning_rate*0.1)
+            schedulers[modality] = torch.optim.lr_scheduler.LambdaLR(optimizers[modality], lr_lambda)
         elif config.optimizer == 'SGD':
             optimizers[modality] = torch.optim.SGD(model.parameters(), lr=0.05, momentum=0.9, weight_decay=1e-4)
     
     for epoch in tqdm(range(config.epochs)):
         if config.optimizer == 'SGD':
             adjust_learning_rate(config, optimizers, epoch+1)
+        elif config.optimizer == 'Adam':
+            for scheduler in schedulers.values():
+                scheduler.step()
         for modality in models.keys():
             model = models[modality]
             model.train()
@@ -92,10 +99,12 @@ def training(config, models, dataset, portion, optimizers, epoch, criterion):
         
         if config.method == 'MMA':
             batch_loss = mma_loss(data['pos'], data['neg'], models)
+        elif config.method == 'eMMA':
+            batch_loss = explicit_anchor_mma_loss(data['pos'], data['neg'], models)
         elif config.method == 'contrastive':
             batch_loss = contrastive_loss(data['pos'], data['neg'], models)
         elif config.method == 'supervised-contrastive':
[…]

### 2022-01-25 Code/losses.py

@@ -25,6 +25,30 @@ def mma_loss(positive, negative, models, margin=0.4):
 
 
 
+def explicit_anchor_mma_loss(positive, negative, models, margin=0.4):
+    '''
+    MMA loss using cosine and explicit maximization of the two anchor points
+    '''
+    # cos_dist(a, p) - cos_dist(a, n) = 1 - cos(a,p) - (1- cos(a, n)) = cos(a, n) - cos(a, p)
+    triplet_loss = lambda a, p, n: torch.clamp(F.cosine_similarity(a, n) - F.cosine_similarity(a, p) + margin, 0.0, 2.0 + margin)
+
+    batch_loss = {'total': 0.0}
+    modalities = ['rgb', 'depth', 'audio']
+    anchor = 'language'
+    loss = 0.0
+    
+    for modality in modalities:
+        loss = loss + torch.sum(triplet_loss(models[anchor](positive[anchor])['decoded'], models[modality](positive[modality])['decoded'], models[modality](negative[modality])['decoded']))
+
+    # loss = loss + torch.sum(torch.clamp(1 - F.cosine_similarity(models[anchor](positive[anchor])['decoded'], models[anchor](negative[anchor])['decoded'])), 0.0 , 2.0 + margin)
+    loss = loss + torch.sum(torch.clamp(F.cosine_similarity(models[anchor](positive[anchor])['decoded'], models[anchor](negative[anchor])['decoded']), 0.0 , 1.0 + margin))
+    # loss = loss + torch.sum(torch.clamp(-1.0 + F.cosine_similarity(models[anchor](positive[anchor])['decoded'], models[anchor](negative[anchor])['decoded'])+ margin, 0.0 , 2.0 + margin))
+    
+    batch_loss['total'] = loss / len(positive['instance']) # average loss in this batch
+    return batch_loss
+
+
+
 def contrastive_loss(positive, negative, models):
     batch_loss = {'total': 0.0}
     modalities = ['rgb', 'depth']

### 2022-01-26 Code/analysis.py
graphs of results, norm of embeddings

@@ -1,0 +1,96 @@
+import os, json, csv
+from re import M
+import numpy as np
+import pandas as pd
+from glob import glob
+import matplotlib.pyplot as plt
+import tikzplotlib
+
+
+
+exp_ugly_names = {
+    '4M-eMMA': '4M-eMMA-cosine-submodalities-text-anchor-gold-no_neg_sampling-1024', 
+    '4M-SupCon': '4M-supervised-contrastive-SGD-cosine-submodalities-text-first-gold-no_neg_sampling-1024',
+    '4M-simple-MMA': '4M-mma-simple-SGD-cosine-submodalities-text-anchor-gold-no_neg_sampling-1024',
+}
+
+experiments = {
+    '4M-eMMA': {},
+    '4M-SupCon': {},
+    '4M-simple-MMA': {},
+}
+
+
+for method in experiments.keys():
+    for seed in [7, 24, 42]:
+        experiments[method]['seed-'+str(seed)] = glob(os.path.join('results', f'*{exp_ugly_names[method]}*/seed-{str(seed)}/'), recursive=True)[0]
+
+# print(experiments)
+
+
+results = {}
+for method in experiments.keys():
+    results[method] = {}
+    for seed in experiments[method].keys():
+        res = json.load(open(experiments[method][seed]+'results.json'))
+        del res['best']
+        results[method][seed] = res
+    results[method]['avg'] = {'best':{}}
+    results[method]['std'] = {'best':{}}
+    # computing average and std over seeds for each metric of each portion of each epoch
+    for epoch in results[method][seed].keys():
+        results[method]['avg'][epoch] = {}
+        results[method]['std'][epoch] = {}
+        for portion in results[method]['seed-42'][epoch].keys():
+            results[method]['avg'][epoch][portion] = {}
+            results[method]['std'][epoch][portion] = {}
+            for metric in results[method]['seed-42'][epoch][portion].keys():
+                results[method]['avg'][epoch][portion][metric] = np.mean([results[method][seed][epoch][portion][metric] for seed in experiments[method].keys()])
+                results[method]['std'][epoch][portion][metric] = np.std([results[method][seed][epoch][portion][metric] for seed in experiments[method].keys()])
+    
+    # computing best for each metric and portion. We find the best average among all epochs but for the std we have to use the std corresponding to that best average.
[…]

### 2022-01-26 Code/analysis.py
4M contrastive method

@@ -12,13 +12,10 @@ exp_ugly_names = {
     '4M-eMMA': '4M-eMMA-cosine-submodalities-text-anchor-gold-no_neg_sampling-1024', 
     '4M-SupCon': '4M-supervised-contrastive-SGD-cosine-submodalities-text-first-gold-no_neg_sampling-1024',
     '4M-simple-MMA': '4M-mma-simple-SGD-cosine-submodalities-text-anchor-gold-no_neg_sampling-1024',
+    '4M-contrastive': '4M-contrastive-cosine-submodalities-text-anchor-gold-no_neg_sampling-1024',
 }
 
-experiments = {
-    '4M-eMMA': {},
-    '4M-SupCon': {},
-    '4M-simple-MMA': {},
-}
+experiments = {name:{} for name in exp_ugly_names.keys()}
 
 
 for method in experiments.keys():
@@ -70,15 +67,26 @@ styles_markers = ['-o', ':+', ':o','--o','-.o', '-+', '--+','-.+', '-v',':v','--
 
 name = 'epochs'
 portion = 'test'
+metric_correct_names = {
+    'mrr_lrd': 'mrr.trd',
+    'mrr_ard': 'mrr.srd',
+    'mrr_lar': 'mrr.tsr',
+    'mrr_lad': 'mrr.tsd',
+    'mrr_lard': 'mrr.tsrd',
+    'mrr_lr': 'mrr.tr',
+    'mrr_ld': 'mrr.td',
+    'mrr_ar': 'mrr.sr',
+    'mrr_ad': 'mrr.sd',
+}
 for metr in ['mrr_lrd', 'mrr_ard', 'mrr_lard']:
 # for metr in ['mrr_ard']:
     fig = plt.figure(figsize=(25,20))
     ax = fig.add_subplot(1,1,1)
     for method_idx, method in enumerate(results.keys()):
-        plt.plot([str(epoch) for epoch in results[method]['avg'].keys()], [results[method]['avg'][epoch][portion][metr] for epoch in results[method]['avg'].keys()], styles_markers[method_idx], label=method)
-        low = np.asarray([results[method]['avg'][epoch][portion][metr] for epoch in results[method]['avg'].keys()]) - np.asarray([results[method]['std'][epoch][portion][metr] for epoch in results[method]['std'].keys()])
[…]

### 2022-01-26 Code/losses.py

@@ -51,7 +51,7 @@ def explicit_anchor_mma_loss(positive, negative, models, margin=0.4):
 
 def contrastive_loss(positive, negative, models):
     batch_loss = {'total': 0.0}
-    modalities = ['rgb', 'depth']
+    modalities = ['rgb', 'depth', 'audio']
     anchor = 'language'
     loss = 0.0
     temperature = 0.1

### 2022-01-26 Code/run_cluster_loop.sh

@@ -14,7 +14,7 @@ for neg in no_neg_sampling
 do
 for dim in 1024
 do
-sbatch run_cluster.sh mma-simple-SGD-cosine-text-anchor $neg $dim $seed $task
+sbatch run_cluster.sh 4M-contrastive-cosine-submodalities-text-anchor $neg $dim $seed $task
 done
 done
 done

### 2022-01-26 Paper/lit-search.tex
Paper update before SIGIR writeup

@@
 
+\subsection{Multimodal Contrastive Training for Visual Representation Learning}
+\href{https://arxiv.org/pdf/2104.12836.pdf}{paper pdf}
+
+\subsection{CrossCLR: Cross-modal Contrastive Learning For Multi-modal Video Representations}
+\href{https://openaccess.thecvf.com/content/ICCV2021/papers/Zolfaghari_CrossCLR_Cross-Modal_Contrastive_Learning_for_Multi-Modal_Video_Representations_ICCV_2021_paper.pdf}{paper pdf}
+
 %===================================================================

### 2022-01-26 Paper/main.tex
Paper update before SIGIR writeup

@@
-% EMNLP header. Do not change.
-\pdfoutput=1
-\documentclass[11pt]{article}
-\usepackage[review]{emnlp2021}
-\usepackage{times}
-\usepackage{latexsym}
-\usepackage[T1]{fontenc}
-\usepackage[utf8]{inputenc}
-\usepackage{microtype}
-% End header.
+\documentclass[sigconf,natbib=true,anonymous=true]{acmart}
+% \documentclass[manuscript,screen,review]{acmart}
+
+% Needed to fix an acmart problem with line numbering. AFAICT line 
+% numbers are not expected in sigconf, so it should be a no-op.
+\makeatletter
+\providecommand{\@LN}[2]{}
+\makeatother
+
+% % Fonts used in the template cannot be substituted; margin 
+% % adjustments are not allowed.
+% %
+% % \BibTeX command to typeset BibTeX logo in the docs
+% \AtBeginDocument{%
+%   \providecommand\BibTeX{{%
+%     \normalfont B\kern-0.5em{\scshape i\kern-0.25em b}\kern-0.8em\TeX}}}
+
+%% Rights management information.  This information is sent to you
+%% when you complete the rights form.  These commands have SAMPLE
+%% values in them; it is your responsibility as an author to replace
+%% the commands and values with those provided to you when you
+%% complete the rights form.
+\setcopyright{acmcopyright}
+\copyrightyear{2022}
+\acmYear{2022}
+\acmDOI{XXXXXXX.XXXXXXX}
+
+%% These commands are for a PROCEEDINGS abstract or paper.
+\acmConference[SIGIR '22]{Make sure to enter the correct
+  conference title from your rights confirmation email}{June 03--05,
+  2018}{Woodstock, NY}
+%
+%  Uncomment \acmBooktitle if the title of the proceedings is different
+%  from ``Proceedings of ...''!
+%
+\acmBooktitle{SIGIR '22: Proceedings of the 45th International ACM SIGIR Conference on Research and Development in Information Retrieval,
+ July 11--15, 2022, Madrid, Spain} 
+% \acmPrice{15.00}
+% \acmISBN{978-1-4503-XXXX-X/18/06}
+
+%
+% Submission ID.
+% Use this when submitting an article to a sponsored event. You'll
+% receive a unique submission ID from the organizers
+% of the event, and this ID should be used as the parameter to this command.
+% \acmSubmissionID{123-A56-BU3}
+
+%
+% The majority of ACM publications use numbered citations and
+% references.  The command \citestyle{authoryear} switches to the
+% "author year" style.
+%
+% If you are preparing content for an event
+% sponsored by ACM SIGGRAPH, you must use the "author year" style of
+% citations and references.
+% Uncommenting
+% the next command will enable that style.
+%\citestyle{acmauthoryear}
+
+%% The "author" command and its associated commands are used to define
+%% the authors and their affiliations.
+%% Of note is the shared affiliation of the first two authors, and the
+%% "authornote" and "authornotemark" commands
+%% used to denote shared contribution to the research.
+\author{Ben Trovato}
+\authornote{Both authors contributed equally to this research.}
+\email{trovato@corporation.com}
+\orcid{1234-5678-9012}
+\author{G.K.M. Tobin}
+\authornotemark[1]
+\email{webmaster@marysville-ohio.com}
+\affiliation{%
+  \institution{Institute for Clarity in Documentation}
+  \streetaddress{P.O. Box 1212}
+  \city{Dublin}
+  \state{Ohio}
+  \country{USA}
+  \postcode{43017-6221}
+}
+
+\author{Lars Th{\o}rv{\"a}ld}
+\affiliation{%
+  \institution{The Th{\o}rv{\"a}ld Group}
+  \streetaddress{1 Th{\o}rv{\"a}ld Circle}
+  \city{Hekla}
+  \country{Iceland}}
+\email{larst@affiliation.org}
+
+\author{Valerie B\'eranger}
+\affiliation{%
+  \institution{Inria Paris-Rocquencourt}
+  \city{Rocquencourt}
+  \country{France}
+}
+
+\author{Aparna Patel}
+\affiliation{%
+ \institution{Rajiv Gandhi University}
+ \streetaddress{Rono-Hills}
+ \city{Doimukh}
+ \state{Arunachal Pradesh}
+ \country{India}}
+
+\author{Huifen Chan}
+\affiliation{%
+  \institution{Tsinghua University}
+  \streetaddress{30 Shuangqing Rd}
+  \city{Haidian Qu}
+  \state{Beijing Shi}
+  \country{China}}
+
+\author{Charles Palmer}
+\affiliation{%
+  \institution{Palmer Research Laboratories}
+  \streetaddress{8600 Datapoint Drive}
+  \city{San Antonio}
+  \state{Texas}
+  \country{USA}
+  \postcode{78229}}
+\email{cpalmer@prl.com}
+
+\author{John Smith}
+\affiliation{%
+  \institution{The Th{\o}rv{\"a}ld Group}
+  \streetaddress{1 Th{\o}rv{\"a}ld Circle}
+  \city{Hekla}
+  \country{Iceland}}
+\email{jsmith@affiliation.org}
+
+\author{Julius P. Kumquat}
+\affiliation{%
+  \institution{The Kumquat Consortium}
+  \city{New York}
+  \country{USA}}
+\email{jpkumquat@consortium.net}
 
@@
 
-\usepackage{soul}
+% \usepackage{soul} %% Soul does not play well with acmart
 \usepackage{url}
-% \usepackage[hidelinks]{hyperref}https://www.overleaf.com/project/614b5cc61d38655dce466002
+% % \usepackage[hidelinks]{hyperref}https://www.overleaf.com/project/614b5cc61d38655dce466002
 % \usepackage[small]{caption}
 \usepackage{graphicx}
-\usepackage{amsmath}
 \usepackage{amsthm}
@@
 
-% =======================================
-% my imported packages
+% % =======================================
+% % my imported packages
 
@@
 \usepackage{bm}
-\usepackage{amssymb}
+% \usepackage{amssymb}          % duplicates something in acmart
 \usepackage{listings}
@@
 \newcommand{\todokd}[1]{\todo[color=red!20]{{\small #1 -- Kasra}}}
-\newcommand{\todocainline}[1]{\todo[color=yellow!20,inline]{{CA: \small #1}}}
+% \newcommand{\todocainline}[1]{\todo[color=yellow!20,inline]{{CA: \small #1}}}
 \newcommand{\todocm}[1]{\todo[color=green!40]{\small #1 -- Cynthia}}
@@
 % =======================================
-
-\title{MMA: Multi-modal Manifold Alignment \\ MAME: Manifold Alignment of Multimodal Embeddings}
-
-\author{
-First Author$^1$
-\and
-Second Author$^2$\and
-Third Author$^{2,3}$\And
-Fourth Author$^4$
-\affiliations
-$^1$First Affiliation\\
-$^2$Second Affiliation\\
-$^3$Third Affiliation\\
-$^4$Fourth Affiliation
-\emails
-\{first, second\}@example.com,
-third@other.example.com,
-fourth@example.com
-}
+\title{Multimodal Contrastive Learning for Physical Object Retrieval}
 
@@
 
-Todo list
-
-\begin{enumerate}
-    \item \st{Lit search}
-\end{enumerate}
-
-\newpage
-
-\maketitle
-\setcounter{page}{1}
+% Todo list
+% \begin{enumerate}
+%     \item \st{Lit search}
+% \end{enumerate}
+% \newpage
 
@@
 \begin{abstract}
-    To be written ...
+    % Grounded language understanding, in which natural language is used as a query against objects in a physical environment, allows a real-world, intuitive mechanism by which users can instruct physical agents to engage in tasks such as object retrieval.
+Visuolinguistic approaches to such object inference tasks typically involve training on large pools of image/text pairs and then using language to subselect elements of the sensed environment.
+However, physical agents such as robots typically have access to sensory and interactive modalities beyond vision, and learning from multiple modalities can improve performance on downstream tasks.
+In order to fully leverage multimodal training data while being robust to missing information, we propose a generalized distance-based loss function that can be extended to learn retrieval models that incorporate an arbitrary number of modalities.
+We demonstrate the usability of our model on a grounded language object retrieval scenario, where an intelligent agent has to select an object given an unconstrained language command.
+We leverage four modalities including vision, depth sensing, text, and speech, and we show that our model can outperform state-of-the-art contrastive models when modalities are ablated.
+    % \todokdinline{Text is easier to learn from compared to speech, and our proposed model allow us to minimize the distance between text and speech and when text is missing during the test time, speech can be used to perform the task, and it outperforms the supervised contrastive learning because such a leverage is missing there.}
+
+    % FRANK:
+
+    We propose a generalized distance-based loss function that can be used to learn retrieval models that incorporate an arbitrary number of views of 
+    a particular piece of data. %
+    We study this in the context of grounded language understanding, in which natural language is used as a query against objects in a physical environment, allowing a real-world, intuitive mechanism by which users can instruct physical agents to engage in tasks such as object retrieval.
+    In our setting, each view is a different modality of data, such as written text, spoken language, and images. 
+    % Visuolinguistic approaches to such object inference tasks typically involve training on large pools of image/text pairs and then using language to subselect elements of the sensed environment. 
+    %Such an approach could improve the ability of physical agents, such as robots typically have access to sensory and interactive modalities beyond vision, and learning from multiple modalities can improve performance on downstream tasks.
+    We demonstrate the usability of our model on a grounded language object retrieval scenario, where an intelligent agent has to select an object given an unconstrained language command.
+We leverage four modalities including vision, depth sensing, text, and speech, and we show that our model can outperform state-of-the-art contrastive models when modalities are ablated.
+
 \end{abstract}
+
+\settopmatter{printfolios=true} % Page numbers please 
+9 pages without references.
+%===================================================================
+
+\maketitle
+% \setcounter{page}{1}
 
@@
 
+% \section{SIGIR Papers}
+% \begin{enumerate}
+%     \item \href{https://dl.acm.org/doi/pdf/10.1145/3397271.3401430}{FashionBERT}
+%     \item \href{https://dl.acm.org/doi/pdf/10.1145/3397271.3401265?casa_token=W8NqNWp0UQkAAAAA:w-iRUnNhmytDKCUiYzaUPG0bI_yyo8Bzxwro_eOC-zl5VDIGRVT2dvixvnCwLJ0iEKqYo1LkQIWXVA}{CrossBERT}
+
+% \end{enumerate}
+
 \section{Introduction}
 \label{intro}
-Humans use multiple sensory inputs to interact with the world around them.
-Most methods in machine learning use at most two sensory inputs that are very different such as visual data (such as RGB images) and natural language.
-However, using extra similar inputs from different modalities is underexplored. This includes incorporating depth and RGB, written natural language and speech, and a combination of all of them.
-Most of the proposed works cannot be extended to arbitrary number of modalities, and we address this shortcomings by proposing a generalized distance-based loss function.
 
-% Grounded language learning~\cite{Mooney2008Grounded,RichardsDarvishMatuszekCategoryFree20,triplet_loss_2021_CVPR} is a cross-modal learning task to connect natural language with the physical environment in which the natural language is used.
+Humans use multiple sensory inputs to understand, interact with, and retrieve information from the world around them.
+It is intuitive that agents learning about the world, upon encountering new concepts and new objects, should form a model that incorporates information from all available sensors and data sources.
 
-\todokdinline{Are we doing contrastive learning? If so, mention it here}
+Grounded language understanding, in which natural language is used as a query against objects in a physical environment, allows a real-world, intuitive mechanism by which users can instruct physical agents to engage in tasks such as object retrieval.
+Visuolinguistic approaches to such object inference tasks typically involve training on large pools of image/text pairs and then using language to subselect elements of the sensed environment.
+Although physical agents such as robots typically have access to sensory and interactive modalities beyond vision, and learning from multiple modalities can improve performance on downstream tasks, most approaches use at most two sensory inputs (e.g., visual data such as RGB images) with single labels, such as those provided by textual natural language.
+Simultaneously using additional inputs from different modalities is an underexplored area, in part due to the domain-specific nature of such \textit{n}-ary learning approaches.
 
+In order to fully leverage multimodal training data while being robust to missing information, we propose a generalized distance-based loss function that can be extended to learn retrieval models that incorporate an arbitrary number of modalities.
+We consider the domain of grounded language-based object retrieval~\cite{Mooney2008Grounded, RichardsDarvishMatuszekCategoryFree20, triplet_loss_2021_CVPR}.\todo{check whether these are still the appropriate citations}
+% , a cross-modal learning task in which the goal is to learn to understand natural language and its connection to the physical environment in which the language occurs. 
+Approaches to acquiring grounded language have explored various combinations of sensor inputs such as depth and RGB with labels provided by textual language or speech.
+This work has significant overlap with robotics, as it pertains to the use of language in the physical domain in which robots operate.
+However, despite the multisensory nature of most robot platforms, much of the existing work has not previously been extended to include an arbitrary number of modalities.
+
+The high level goal of this work is to take arbitrary input modalities about novel objects in the world, including both spoken and written language, and build a model that allows a robot system to correctly retrieve that object given a subset of those input modalities.
+This is a generalization of approaches to grounded language learning in which specific input modalities are labeled with language to allow for future identification, but explicitly seeks to be agnostic about the nature of the individual sensory and linguistic inputs.
 Our contributions are as follows:
 \begin{enumerate}
-    \item Proposing a general distance-based loss function formulation for multimodal learning that can be extended to arbitrary number of modalities
-    \item Outperforming state-of-the-art models for grounded language learning by treating depth image as its own separate channel.
-    \item 
+    \item Proposing a general distance-based loss function formulation for multimodal learning that can be extended to an arbitrary number of modalities.
+    \item Outperforming state-of-the-art models for grounded language learning by differentiating visual input modalities such as RGB and depth, rather than combining them into a single modality.
+    \item Demonstrating the separate utility of speech and text as sources of label information by treating them as sensory input modalities, rather than explicitly as labels.
+    \item Outperforming state-of-the-art contrastive models when one or more modalities are missing during test.
 \end{enumerate}
+
+\todocmi{NOTES: Some things are probably less informative or more ambiguous (e.g., depth for differentiating round things).
+So instead of predicting a single vector output, predict a vector and a covariance that captures uncertainty; could potentially use that to do more accurate search, or to handle really uncertain modalities (e.g., just drop speech when it's uninformative).}
+
+\todocmi{NOTES: What if one of your modalities fails? Does that super-bork the contrastive approach? Possibly Simple MMA will be more robust to dropped modalities.}
+
 %===================================================================
+
+\todo[inline]{Edward Framing idea: Sensors sometimes break. Look at accuracy retrieval using against target object with various sub-modalities available/not? I will run those experiments today! }
 
@@
 \label{sec:Related-Work}
+
 The growing number of multimodal datasets demonstrates the importance and the challenge of this task.
 \citet{bagher-zadeh-etal-2018-multimodal} introduce a large dataset of multimodal opinion sentiment and emotion intensity (CMU-MOSEI) for sentiment analysis and emotion recognition.
-They also introduce a novel multimodal fusion technique called the Dynamic Fusion Graph (DFG).
-\citet{GoLD_UMBC} present a multimodal dataset of household objects containing RGB images, depth images, written language, and spoken language.
+They also introduce a novel multimodal fusion technique called the Dynamic Fusion Graph (DFG).~\citet{GoLD_UMBC} present a multimodal dataset of household objects containing RGB images, depth images, written language, and spoken language, which has been used to support learning grounded language directly from speech given small datasets~\cite{KebeAAAI2022}.
+\todo[inline]{Look at multisensory datasets from robotics}
 
@@
 They cover different approaches to multimodal machine learning such as \textit{representation}, \textit{translation}, \textit{alignment}, \textit{fusion}, and \textit{co-learning}.
-In this project, we focus on alignment.
+In this project, we focus on alignment, in which \textcolor{red}{\dots}.
 
-We review different methods that have been proposed for alignment learning.
+We review different methods that have been proposed for alignment learning in the language domain.
 \citet{alayrac2020self} use a self-supervised contrastive learning method to learn and combine representations from three modalities of visual, audio, and language.
@@
 The loss function they define consists of two distance-based terms for the corresponding spaces.
+This work differs from ours in that \textcolor{red}{\dots}.
+\citet{Nguyen-RSS-20} take a similar approach and perform pairwise cosine similarity to align images and natural language text descriptions to perform object retrieval tasks.
+\citet{triplet_loss_2021_CVPR} take a cross-modal manifold alignment approach to grounded language learning using triplet loss, and perform object retrieval by connecting text descriptions of objects to their corresponding RGB-D images.
 
-\citet{Nguyen-RSS-20} take a similar approach and perform pairwise cosine similarity to align images and natural language text descriptions to perform object retrieval.
+Another related approach is contrastive learning, which is based on the idea of similarity and distance~\cite{NEURIPS2020_supervised_contrastive,chen2020simple}.
+Triplet loss is also a special case of contrastive loss~\cite{NEURIPS2020_supervised_contrastive}.
+Contrastive loss is mostly used in self-supervised learning~\cite{alayrac2020self,chen2020simple}, with the exception of the work proposed by \citet{NEURIPS2020_supervised_contrastive} where they use contrastive loss for a supervised problem.
+Most of these methods are implemented for a non-multimodal dataset (e.g., RGB images only).
+Our work is a specialization of contrastive loss that is focused on the sensor-and-language problem.\todo[inline]{Can we think of something better to say about how this relates to contrastive learning generally?}
 
-\citet{triplet_loss_2021_CVPR} take a cross-modal manifold alignment approach to grounded language learning using triplet loss.
-They perform object retrieval by connecting text descriptions of objects to their corresponding RGB-D images.
-
-Another related approach is contrastive learning which is based on the idea of similarity and distance~\cite{NEURIPS2020_supervised_contrastive,chen2020simple}.
-Triplet loss is also a special case of contrastive loss~\cite{NEURIPS2020_supervised_contrastive}.
-Contrastive loss is mostly used in self-supervised learning~\cite{alayrac2020self,chen2020simple} with the exception of the work proposed by \citet{NEURIPS2020_supervised_contrastive} where they use contrastive loss for a supervised problem.
-Most of these methods are implemented for a unimodal dataset (i.e.
-dataset of one type of data such as RGB images only).
-
-Some of the related works take more than two input modalities.
+Some of the most closely related work focuses on learning based on more than two input modalities.
 \citet{Veit_2018_CVPR} train a joint model of images, hashtags, and users to perform image retrieval.
@@
 The downstream retrieval task is then simply done by taking the arg max of the tensor product for a given user.
+\todo[inline]{This work differs from ours in that---what?
+Does it becomes computationally complex as additional modalities are added?
+The question is, could we not just apply their approach, and why not?}
 
@@
 \citet{Mittal2020M3ER} use Canonical Correlational Analysis (CCA) to differentiate between ineffective and effective modalities for the emotion recognition task from multiple input modalities including face, speech, and text.
-
 Multimodal learning has been used for different applications including but not limited to object retrieval.
@@
 \citet{het_data_fusion_liu_IEEE_2017} take two modalities as input and output predictions in a third modality.
+\citet{tursun2021efficient} uses quadruplet loss for two modalities only; image and sketch.
+A quadruplet is composed of a sketch picture as an anchor, a negative example from sketch domain, a negative example from picture domain,  and a positive example from picture domain.
+However, our method is not restricted to two modalities only.~\citet{chen2017beyond} also uses quadruplet loss, but does not support multiple domains.
+% \todo[inline]{talk about CLIP and DALL-E here}
 
-\todo[inline]{talk about CLIP and DALL-E here}
-
-However, to the best of our knowledge, there has not been any work to incorporate more than 3 modalities, and that's the novelty of our work.
+To the best of our knowledge, there has been limited work on incorporating more than three modalities into a general learning approach.
+Our proposed loss function closely resembles work on quadruplet loss~\cite{chen2017beyond,tursun2021efficient}
+% \todo{Beyond triplet loss: a deep quadruplet network for person re-identification, but no multi-domain -Ed}, 
+but is intended to scale to an arbitrary number of modalities.
 
@@
 
+Given a language command either text or speech that describes an object, we want our model to retrieve the correct object from a set of objects.
+This problem is also known as grounded language learning in the fields of robotics and natural language processing.
+If we have a spoken language command $x_s$, a textual language command $x_t$, a set of RGB images $X_r = \{x_r^{(1)}, x_r^{(2)}, ..., x_r^{(n)}\}$, and a set of depth images $X_r = \{x_d^{(1)}, x_d^{(2)}, ..., x_d^{(n)}\}$ the task is to retrieve the correct object by choosing the index that has the minimum distance from either of the language commands across all modalities.
+We consider (I) ``sr'': the distance between $x_s$ and all RGB images in $X_r$, (II) ``sd'': the distance between $x_s$ and all depth images in $X_d$, (III) ``tr'': the distance between $x_t$ and all RGB images in $X_r$, and (IV) ``td'': the distance between $x_t$ and all depth images in $X_d$.
+We use the average of these distances to select the correct object.
+Depending on the missing/available sensors during test time, we might have any combination of these four distances.
+For example if text modality is missing during the test, we can take average of $sr$ and $sd$.
+If we have all four modalities, we take the average of $sr$, $sd$, $tr$, and $td$.
+This method allows us to extend our model to support arbitrary number of modalities.
+
 %===================================================================
@@
 
-For $M$ modalities, we define a distance based loss function that contains $M+2(M-1)!$ terms. 
+For $M$ modalities, we define a distance based loss function which is a generalization of well-known similarity-based metric learning and manifold alignment methods such as triplet loss~\cite{Carvalho-cooking-triplet,triplet_loss_2021_CVPR}, quadruplet loss~\cite{chen2017beyond}, and is similar to contrastive loss under some settings.
 We sample two data points from each modality, one positive point and one negative point. 
+For each set of positive data points (referring to an instance of an object with all of its modalities), we choose a set of negative data points either based on object class name, or by semantic negative sampling~\cite{Pillai_Matuszek_2018}.
+We now have two data points; we consider the first one as positive, and the second one as negative.
+However, the anchor won't be another data point as it is in the triplet loss terminology.
+Instead we choose one domain as an anchor.
+
+\begin{itemize}
+    % \item data point: a row in the csv/tsv file.
+    \item Positive: a set of embeddings of the one data point (rgb image of an apple, depth image of an apple, language description of an apple)
+    \item Negative: a set of embeddings of the another data point of a different object (rgb image of a book, depth image of a book, language description of a book)
+    \item Anchor: A fixed modality to act as anchor in the case of simple MMA loss, where we don't minimize the distance between the pair of embeddings from all modalities of the same data point, and we don't maximize the distance between positive an negative datapoints of the same modality.
+\end{itemize}
+
 The objective is then to (I) minimize the distance between each pair of positive points from heterogeneous modalities which results in $(M-1)!$ terms, (II) maximize the distance between each pair of positive and negative points from heterogeneous modalities which contains $(M-1)!$ terms again, and (III) maximize the distance between each pair of positive and negative points from the homogeneous modalities which contains $M$ terms.
+Altogether, our proposed loss function contains $M+2(M-1)!$ terms.
 Equation~\ref{eq:objective} shows the objective function we defined above with changing maximization to minimization by changing the sign.
@@
 
+% \begin{equation}\label{eq:objective}
+% \begin{split}
+%     \mathcal{L}  &= \sum_{i=1}^{M-1} \sum_{j=i+1}^{M} ||f_i(x_{i}^{+}) - f_j(x_{j}^{+})|| \\
+%     & - ||f_i(x_{i}^{+}) - f_j(x_{j}^{-})||
+%     - ||f_i(x_{i}^{-}) - f_j(x_{j}^{+})|| \\
+%     & - \sum_{i=1}^{M} ||f_i(x_{i}^{+}) - f_i(x_{i}^{-}) ||
+% \end{split}
+% \end{equation}
+% where $f_i(.)$ and $f_j(.)$ represent functions that map input data to the shared latent space using a deep neural network.
+
+% d(,) notation instead of || - ||
+% \begin{equation}\label{eq:objective}
+% \begin{split}
+%     \mathcal{L}  &= \sum_{i=1}^{M-1} \sum_{j=i+1}^{M} d(z_{i}^{+}, z_{j}^{+}) 
+%      - d(z_{i}^{+}, z_{j}^{-}) \\ 
+%     & - d(z_{i}^{-}, z_{j}^{+}) - \sum_{i=1}^{M} d(z_{i}^{+}, z_{i}^{-})
+% \end{split}
+% \end{equation}
+
 \begin{equation}\label{eq:objective}
 \begin{split}
-    \mathcal{L}  &= \sum_{i=1}^{M-1} \sum_{j=i+1}^{M} ||f_i(x_{i}^{+}) - f_j(x_{j}^{+})|| \\
-    & - ||f_i(x_{i}^{+}) - f_j(x_{j}^{-})||
-    - ||f_i(x_{i}^{-}) - f_j(x_{j}^{+})|| \\
-    & - \sum_{i=1}^{M} ||f_i(x_{i}^{+}) - f_i(x_{i}^{-}) ||
+    \mathcal{L}  &= \sum_{i=1}^{M-1} \sum_{j=i+1}^{M} ||z_{i}^{+} - z_{j}^{+}|| 
+     - ||z_{i}^{+} - z_{j}^{-}|| - ||z_{i}^{-} - z_{j}^{+}|| \\ 
+    &  - \sum_{i=1}^{M} ||z_{i}^{+} - z_{i}^{-} ||
 \end{split}
 \end{equation}
-where the subscripts $i$ and $j$ represent modalities, and $f_i(.)$ and $f_j(.)$ represent functions that map input data to the shared latent space using a deep neural network.
+where the subscripts $i$ and $j$ represent modalities, and $z$ is the embedding we get by applying a mapping function $f$ which in our case is a neural network on our input data.
+In other words $z_m = f_m(x_m)$ where each modality $m$ has a specific model $f_m$ that is different from the models for other modalities and these models do not share their weights.
+To measure the distance, we use cosine distance.
+In order to use cosine \textit{distance}, we have to subtract the cosine of the \textit{angle} between two embeddings (which represents similarity) from 1: $1 - \cos(e_1, e_2)$.
 
@@
 However, it breaks the nice procedure of sampling two data points per modality}
+\todo{Edward wonders if $f_i(x)$ should instead produce $\mu_x$ and $\Sigma_x$, to capture uncertainty of embedding location, when working with so many modalities?
+(e.g., not all modalities share the same ambiguities)}
+\todo{Does Edward mean the actual network produces $\Sigma_x$, e.g., something like a VAE? How would uncertainty be applied to the inference problem?}
 $+ \sum_{i=1}^{M} ||f_i(x_i^{a}) - f_i(x_i^{+})||$
 
-Our method is a generalization of well-known similarity-based metric learning and manifold alignment methods such as triplet loss~\cite{Carvalho-cooking-triplet,triplet_loss_2021_CVPR}.
-If $M=2$ which means the number of modalities is 2, then our objective function reduces to quadruplet loss method.
-If $M=2$ and we ignore the last term in the objective function, it results in the triplet loss method. 
+If $M=2$ which means the number of modalities is 2, then our objective function reduces to the quadruplet loss method~\cite{chen2017beyond} if we ignore the third term, and negative samples for each modality do not belong to the same class.
+If $M=2$ and we ignore the last two terms in the objective function, it results in the triplet loss method. 
 If $M=1$, only the last term remains in the loss function which is exactly the pairwise distance-based loss function.
 
-\subsection{Loss Functions}
-We try three different distance based loss functions.
-In order to use cosine distance, we have to subtract the cosine between two embeddings from 1 which means $1 - \cos(e_1, e_2)$.
+We demonstrate performance using three different distance-based loss functions.
+\todokdinline{Either remove it or experiment with all versions of MMA (12 terms, 9 terms, and explicit anchor maximization}
 
-\subsubsection{Simple triplet extension}
+\subsection{Simple MMA}
 \label{sub:simple-mma}
+Our first proposed loss function is a modification of the triplet loss function idea, and can be derived from equation~\ref{eq:objective} by fixing one modality as anchor, and ignoring the last two terms in equation~\ref{eq:objective}.
 
-First one is a modification of the triplet loss function idea.
-For each data point (referring to an instance of an object with all of its modalities), we choose a negative data point either based on object class name, or by negative sampling~\cite{Pillai_Matuszek_2018}.
+% For each data point (referring to an instance of an object with all of its modalities), we choose a negative data point either based on object class name, or by semantic negative sampling~\cite{Pillai_Matuszek_2018}.
 We now have two data points; we consider the first one as positive, and the second one as negative.
@@
 
-\begin{itemize}
-    \item data point: a row in the csv/tsv file.
-    \item Positive: a set of embeddings of the one data point (rgb image of an apple, depth image of an apple, language description of an apple)
-    \item Negative: a set of embeddings of the another data point of a different object (rgb image of a book, depth image of a book, language description of a book)
-    \item Anchor: A fixed modality to act as anchor in the case of simple MMA loss, where we don't minimize the distance between the pair of embeddings from all modalities of the same data point, and we don't maximize the distance between positive an negative datapoints of the same modality.
-\end{itemize}
+% \begin{itemize}
+%     % \item data point: a row in the csv/tsv file.
+%     \item Positive: a set of embeddings of the one data point (rgb image of an apple, depth image of an apple, language description of an apple)
+%     \item Negative: a set of embeddings of the another data point of a different object (rgb image of a book, depth image of a book, language description of a book)
+%     \item Anchor: A fixed modality to act as anchor in the case of simple MMA loss, where we don't minimize the distance between the pair of embeddings from all modalities of the same data point, and we don't maximize the distance between positive an negative datapoints of the same modality.
+% \end{itemize}
 
-Since the downstream task is to predict/retrieve the desired object among multiple objects given a natural language description, it makes sense to train the model using language as anchor. 
-We tried RGB as anchor, and as expected the performance was worse.
+Since the downstream task in the grounded language learning domain is to predict/retrieve a desired object among multiple objects given a natural language description, it makes sense to train the model using language as an anchor, and in fact anchoring learning around language outperforms anchoring on RGB.
 
 Equation~\ref{eq:objective-simple-mma} shows the mathematical formulation of the proposed loss function which can be generalized to arbitrary number of modalities.
+
+% \begin{equation}
+% \label{eq:objective-simple-mma}
+% \begin{split}
+%     \mathcal{L}  &= \sum_{m=1}^{M-1} || f_a(x_{a}^{+}) - f_m(x_{m}^{+}) || - \\ 
+%     & || f_a(x_{a}^{+}) -f_m(x_{m}^{-}) || 
+%     \\
+%     &= \sum_{m=1}^{M-1} 1 - \cos(f_a(x_{a}^{+}), f_m(x_{m}^{+})) \\ & - (1 - \cos(f_a(x_{a}^{+}) ,f_m(x_{m}^{-}))) \\
+%     &= \sum_{m=1}^{M-1} \cos(f_a(x_{a}^{+}) ,f_m(x_{m}^{-})) \\ 
+%     & - \cos(f_a(x_{a}^{+}), f_m(x_{m}^{+}))
+% \end{split}
+% \end{equation}
 
@@
 \begin{split}
-    \mathcal{L}  &= \sum_{m=1}^{M-1} || f_a(x_{a}^{+}) - f_m(x_{m}^{+}) || - \\ 
-    & || f_a(x_{a}^{+}) -f_m(x_{m}^{-}) || 
-    \\
-    &= \sum_{m=1}^{M-1} 1 - \cos(f_a(x_{a}^{+}), f_m(x_{m}^{+})) \\ & - (1 - \cos(f_a(x_{a}^{+}) ,f_m(x_{m}^{-}))) \\
-    &= \sum_{m=1}^{M-1} \cos(f_a(x_{a}^{+}) ,f_m(x_{m}^{-})) \\ 
-    & - \cos(f_a(x_{a}^{+}), f_m(x_{m}^{+}))
+    \mathcal{L}  &= \sum_{m=1}^{M-1} || z_{a}^{+} - z_{m}^{+} || - || z_{a}^{+} - z_{m}^{-} || \\
+    &= \sum_{m=1}^{M-1} 1 - \cos(z_{a}^{+}, z_{m}^{+}) - (1 - \cos(z_{a}^{+}, z_{m}^{-})) \\
+    &= \sum_{m=1}^{M-1} \cos(z_{a}^{+} ,z_{m}^{-}) - \cos(z_{a}^{+}, z_{m}^{+})
 \end{split}
@@
 
-where $a$ represents \textit{anchor} modality, $M$ is the number of modalities, $p$ represents the positive objects, and $n$ represents the negative objects.
+where subscript $a$ represents \textit{anchor} modality, $M$ is the number of modalities, superscript $+$ represents the positive objects, and superscript $-$ represents the negative objects.
 
 An illustration of this loss function is provided in figure~\ref{fig:simple-mma-loss}.
+
+\begin{figure}[tb]
+\centering
+\includegraphics[width=\columnwidth]{Figures/simple-MMA.pdf}
+\caption{A high-level prototype of the distances used in the simple MMA loss with 3 modalities.
+Orange arrows indicate maximizing the distance, blue arrows indicate minimizing the distance, and green lines show the margin.
+Depth images are brightened to be visible.
+\todo[inline]{We need to show speech as the 4th modality as well.}}
+\label{fig:simple-mma-loss}
+\end{figure}
 
@@
 Hence, our model can be trained using smaller batch sizes and reduces the number of negative samples we need.
-Second advantage is that this loss function can be used in a multimodal setting with an arbitrary number of modalities, and is not limited to a single data type (e.g.
+The second advantage is that this loss function can be used in a multimodal setting with an arbitrary number of modalities, and is not limited to a single data type (e.g.
 RGB images) which is the most common usage of contrastive loss~\cite{chen2020simple, NEURIPS2020_supervised_contrastive}.
-
-\begin{equation}\label{eq:contrastive-loss}
-    -\sum_{i \in I} \log \frac{\exp (z_i . z_{j(i)} / \tau) }{\sum_{a \in A(i)} \exp (z_i.z_a / \tau)}
-\end{equation}
 
@@
 
+% \begin{equation}\label{eq:simple-MMA-contrastive}
+%     -\sum_{m=1}^{M-1} \log \frac{\exp (sim(f_a(x_a^+) , f_m(x_{m}^{+})/ \tau) )}{ \exp (sim(f_a(x_a^+) , f_m(x_{m}^{+}) / \tau)) + \exp (sim(f_a(x_a^+), f_m(x_{m}^{-}) / \tau))}
+% \end{equation}
+
 \begin{equation}\label{eq:simple-MMA-contrastive}
-    -\sum_{m=1}^{M-1} \log \frac{\exp (z_a . z_{m}^{+}/ \tau) }{ \exp (z_a.z_m^{+} / \tau) + \exp (z_a.z_m^{-} / \tau)}
+    -\sum_{m=1}^{M-1} \log \frac{\exp (sim(z_a^+ , z_{m}^{+})/ \tau) }{ \exp (sim(z_a^+ , z_{m}^{+}) / \tau) + \exp (sim(z_a^+, z_{m}^{-}) / \tau)}
 \end{equation}
+where $z = f(x)$.
 
-\todokdinline{I should run the traditional contrastive loss to see if my method works better to make this claim.}
+% \todokdinline{I should run the traditional contrastive loss to see if my method works better to make this claim.}
 
-\begin{figure*}[tbh]
-\centering
-\includegraphics[width=2.0\columnwidth]{Figures/simple-MMA.pdf}
-\caption{A high-level prototype of the distances used in the simple MMA loss with 3 modalities.
-Orange arrows indicate maximizing the distance, blue arrows indicate minimizing the distance, and green arrows show the margin.}
-\label{fig:simple-mma-loss}
-\end{figure*}
+% \subsubsection{All in loss with 12 terms}
+% \subsubsection{subset of All in loss with 9 terms}
 
-\subsubsection{All in loss with 12 terms}
-\subsubsection{subset of All in loss with 9 terms}
+\subsection{EMMA}
+\label{sub:emma}
+In this version, we add the last term in equation~\ref{eq:objective} which maximizes the distance between positive and negative instances from the same modality.
+However, instead of looping over all modalities, we only do it for the anchor modality.
+The reason for that is because a margin between positive and negative points from other modalities (excluding anchor) is enforced.
+We formulate this method in equation~\ref{eq:objective-emma}.
+
+This approaches closes the gap against supervised contrastive learning method when dropping speech, and increases the improvement even more when dropping text.
+
+\begin{equation}
+\label{eq:objective-emma}
+\begin{split}
+    \mathcal{L} = \sum_{m=1}^{M-1} \cos(z_{a}^{+} ,z_{m}^{-}) - \cos(z_{a}^{+}, z_{m}^{+}) - \cos(z_{a}^{+}, z_{a}^{-})
+\end{split}
+\end{equation}
 
@@
 
-We use BERT \CitationNeeded{} to featurize the textual input, and wave2vec2 \CitationNeeded{} to extract audio embeddings from speech both of which output a 3072-dimensional embedding vector.
-To process images, we use ResNet152 \CitationNeeded{} for both RGB and depth images which gives us a 2048-dimensional embedding vector.
-We then use 3 fully connected layers to map these embeddings to a shared 1024-dimensional space where we can compute the distance between all embeddings.
+We use BERT~\cite{devlin-etal-2019-bert} embeddings from flair~\cite{akbik2019flair,akbik-etal-2019-pooled} to featurize the textual input, and wav2vec2~\cite{wav2vec2} to extract audio embeddings from speech both of which output a 3072-dimensional embedding vector.
+To process images, we use ResNet152~\cite{He_resnet_2016} for both RGB and depth images which gives us a 2048-dimensional embedding vector.
+We then use 3 fully connected layers with ReLU activation function~\cite{relu2010} to map these embeddings to a shared 1024-dimensional space where we can compute the distance between all embeddings.
 
@@
 
-\subsection{4 Modalities}
-We ran an experiment with RGB, depth, language.
-The loss function doesn't require any changes beyond increasing the number of loops in equation~\ref{eq:objective-simple-mma} (or increasing the value of $M$ by 1).
-The non-trivial part is the downstream prediction task. In the case of three modalities, language is the anchor, and we compute the distance of other two modalities from it and then average them.
-With speech coming into the picture, we have three choices.
-First, treat it in a similar way to RGB and depth; which means computing the distance of RGB, depth, and speech from text, and then take an average of three of them.
-Second way is to compute the distance of RGB and depth from language and from speech which gives us 4 distance matrices, and then take average of the 4 of them.
-Third way is similar to the second methods, but we compute the distance between language and speech as well and then take average of 5 distance matrices.
-The problem with second and third method, is that the simple MMA loss function in equation~\ref{eq:objective-simple-mma} is not designed in a way to minimize the distance between speech and visual embeddings explicitly.
-This is supported by our experimental results where the first method results in a better performance compared to the other two methods.
-Third method (averaging 5 matrices) is the worst, then second one (averaging 4 matrices) is bad. First method has the best results.
-
-\todokdinline{try the more complicated MMA that might be able to capture these relationships.}
-
-\subsection{Contrastive Loss}
-We compare our model against contrastive loss as formulated in equation~\ref{eq:contrastive-loss}.
-In order to implement this loss function, we use cosine similarity as suggested in the SimCLR paper~\cite{chen2020simple}.
-The other version mentioned by~\citet{NEURIPS2020_supervised_contrastive} uses inner dot product which can lead to instabilities since the dot product is not bounded.
-\todokdinline{try contrastive loss with cross-entropy, dot product, and the supervised contrastive method itself (without triplets).}
-
-\subsection{Metrics}
-\label{sec:metrics}
-To evaluate our model we measure different performance metrics on a retrieval task where the model has to select 1 object among 5 objects given a language description.
-Only one of the objects corresponds to the description and the rest are from different object classes.
-\todokdinline{Try with different number of objects.}
-
-To evaluate the performance, we compute the cosine distance between a given natural language description and 5 randomly selected images (1 of which corresponds to the description, and others are from different object classes).
-We have to modify the way we compute the distance since we treat depth images as a separate input modality and do not append it to the RGB features as in the triplet loss method~\cite{triplet_loss_2021_CVPR}.
-To handle this, we compute the distance between 1 language embedding and all candidate RGB embeddings, and we compute the distance between the same language embedding and all candidate depth embeddings corresponding to the RGB embeddings.
-We then take average of these two distance matrices.
-
-Moreover, instead of choosing an arbitrary threshold, we choose the closest image embedding (average distance of rgb and depth from language) as the prediction.
-
-The best metric to capture the performance in such a scenario would be mean reciprocal rank (MRR) and accuracy.
-Accuracy, micro F1 score (sklearn), flatten binary f1 score (sklearn), and my f1 score computation are all the same in this task since for each prediction we either have a true positive and no false positives and no false negatives (no FP no FN), or we have no true positives and 1 false positive and 1 false negative.
-However, subset MRR is equal or greater than those because if the model do not rank the correct answer as 1, it still gets some score and is greater than 0 (in accuracy the score is ether 0 or 1).
+We test our model against different baselines in different scenarios and settings using multiple metrics.
 
@@
 We use a recent multimodal dataset called GoLD~\cite{GoLD_UMBC}.
-The original paper uses raw RGB and raw depth images in which other items are present in the background.
+The original paper uses raw RGB and raw depth images in which other objects are present in the background.
 We use a masked version of the images where the background (including the white turntable) is deleted and one object is present in each image.
-We show that a masked version converges faster, however, they both converge to the same performance after enough epoch as shown  in figure~\ref{fig:mask-vs-raw}.
+We show that a masked version converges faster, however, they both converge to the same performance as shown in the supplementary material in figure~\ref{fig:mask-vs-raw}.
+Speech from GoLD is converted to 16 Hz in order to match wav2vec2 speech featurization model~\cite{wav2vec2}.
 
-\begin{figure*}[tbh]
-\centering
-\includegraphics[width=2.0\columnwidth]{Figures/raw-mask-test-mrr.pdf}
-\caption{Mean Reciprocal Rank (MRR) on different versions of GoLD dataset~\cite{GoLD_UMBC}.
-The raw version is depicted in green and masked version is depicted in orange averaged over 3 different random seeds.
-The higher is better.}
-\label{fig:mask-vs-raw}
-\end{figure*}
+\subsection{Metrics}
+\label{sec:metrics}
+To evaluate our model we measure different performance metrics on a retrieval task where the model has to select an object among 5 objects given a language description.
+Only one of the objects corresponds to the description and the rest are from different object classes (e.g., one apple among a set including a fork, a mug, a lemon, and a bell pepper).
 
-Since the wave2vec2~\CitationNeeded{} model is trained on 16hz speech data, we need to convert he speech files to 16hz.
+% \todokdinline{Try with different number of objects.}
+
+To evaluate the performance, we compute the cosine distance between the given natural language description and 5 randomly selected images (1 of which corresponds to the description, with the others from different object classes).
+We handle depth images as a separate input modality (as opposed to appending depth to RGB, as others have done~\cite{triplet_loss_2021_CVPR}).
+We compute the distance between one textual language embedding and all candidate RGB embeddings, and we compute the distance between the same language embedding and all candidate depth embeddings corresponding to the RGB embeddings.
+We then take average of these two distance matrices.
+Instead of choosing an empirical threshold beyond which objects are considered to be `referred to,' we choose the closest image embedding (average distance of RGB and depth from language) as the prediction.
+
+The best metric to capture the performance in such a scenario is mean reciprocal rank (MRR), in which the inverse of the rank of the correct object for a set of queries and then averaged.
+Accuracy, micro F1 score (sklearn), flattened binary f1 score (sklearn), and F1 score computation are all the same in this task, since for each prediction we either have a true positive and no false positives and no false negatives (no FP no FN), or we have no true positives and 1 false positive and 1 false negative.
+Subset MRR is equal or greater than those, because if the model does not rank the correct answer as 1, it still gets some score and is greater than 0 (in accuracy the score is ether 0 or 1).
 
@@
 
-We use \Complete optimizer with a learning rate of \Complete.
-The models are trained for \Complete epochs on a \Complete GPU.
+Similar to~\citet{NEURIPS2020_supervised_contrastive}, we also use stochastic gradient descent (SGD) optimizer with momentum~\cite{ruder2016overviewSGD} with a flexible learning rate starting with 0.05.
+The models are trained for 200 epochs with a batch size of 64 on a Quadro RTX 8000 GPU.
+We use the same setup for all our experiments to have a fair comparison between different methods.
+
+\subsection{4 Modalities}
+We consider an experiment in which we incorporate RGB, depth, speech, and written language.
+The loss function requires no changes beyond increasing the number of loops in equation~\ref{eq:objective-simple-mma} (or increasing the value of $M$ by 1).
+Our goal is the non-trivial downstream prediction task: determining what objects are being referred to by arbitrary language from a small number of examples.
+In the case of three modalities, written language is used the anchor, and we compute the distance of RGB and depth modalities from it and then average them.
+However, when speech is incorporated as an additional sensory modality, we have three immediate possible choices.
+\textit{First}, we can treat speech in a similar way to RGB and depth: compute the distance of RGB, depth, and speech from text, and then take an average of three of them.
+\textit{Second}, we can compute the distance of RGB and depth from language and from speech which gives us 4 distance matrices, and then take average of these four.
+\textit{Third}, similar to the second method, we can compute the distance between language and speech as well and then take the average of 5 distance matrices.
+
+Of these, the first method is both more consistent with the goal of making a computationally tractable approach to arbitrary input modalities, and more consistent with existing work on using featurized speech directly as a sensory learning input; the simple MMA loss function in equation~\ref{eq:objective-simple-mma} is not designed to minimize the distance between speech and visual embeddings explicitly.
+This is supported by empirical results, in which the first method results in a better performance compared to the other two methods.
+% Third method (averaging 5 matrices) is the worst, then second one (averaging 4 matrices) is bad. First method has the best results.
+
+\todokdinline{try the more complicated MMA that might be able to capture these relationships.}
 
 \subsection{Baselines}
-We compare our MMA model against the triplet loss method proposed by~\citet{GoLD_UMBC} and~\citet{triplet_loss_2021_CVPR}.
-Triplet loss function consists of three data points including anchor, positive, and negative from two modalities.
+We compare our MMA model against different contrastive learning methods including the triplet loss~\cite{GoLD_UMBC, triplet_loss_2021_CVPR}, the traditional version of contrastive loss usually used in self-supervised settings~\cite{chen2020simple}, and supervised contrastive learning~\cite{NEURIPS2020_supervised_contrastive}.
+The triplet loss function consists of three data points including anchor, positive, and negative from two modalities.
 The anchor, positive, and negative can be chosen from different modalities in each batch.
@@
 Second, a batch size of greater than 1 cannot be easily implemented if the anchor, positive, and negative come from different modalities for each batch.
-To address this issues, we sample more than one positive and one negative data points.
+To address these issues, we sample more than one positive and one negative data points.
 Essentially, we sample one positive data point and one negative data point from each modality which results in 2(M-1) + 1 data points where 1 is the single anchor data point from anchor modality of the same object class as positive datapoints.
@@
 The anchor data point is always from the same modality and corresponds to the same object class sampled for positive data points.
+
+\subsubsection{Contrastive Loss}
+\label{sub:baseline-contrastive}
+
+We compare our model against contrastive loss as formulated in equation~\ref{eq:contrastive-loss}.
+In order to implement this loss function, we use cosine similarity as suggested in the SimCLR paper~\cite{chen2020simple}.
+The other version (mentioned by~\citet{NEURIPS2020_supervised_contrastive}) uses an inner dot product that can lead to instabilities since the dot product is not bounded.
+% \todokdinline{try contrastive loss with cross-entropy, dot product, and the supervised contrastive method itself (without triplets).}
+
+% Original Formula
+\begin{equation}\label{eq:contrastive-loss}
+    -\sum_{i \in I} \log \frac{\exp (sim(z_i , z_{j(i)}) / \tau) }{\sum_{a \in A(i)} \exp (sim(z_i, z_a) / \tau)}
+\end{equation}
+
+% \begin{equation}\label{eq:contrastive-loss}
+%     -\sum_{i \in I} \log \frac{\exp (sim(f(x_i) , z_{j(i)}) / \tau) }{\sum_{a \in A(i)} \exp (sim(f(x_i), f(x_a)) / \tau)}
+% \end{equation}
+where $i$ is anchor, $j(i)$ is the set of positives excluding anchor, $a$ is the set of all positives and negatives excluding anchor, and $z = f(x)$.
+
+\subsubsection{Supervised Contrastive Learning}
+\label{sub:baseline-supcon}
+
+\citet{NEURIPS2020_supervised_contrastive} propose a supervised way of performing contrastive learning which is shown in equation~\ref{eq:supervised-contrastive}.
+
+\begin{equation}\label{eq:supervised-contrastive}
+    \sum_{i \in I} \frac{-1}{|P(i)|} \sum_{p \in P(i)} \log \frac{\exp (z_i \cdot z_p / \tau) }{\sum_{a \in A(i)} \exp (z_i \cdot z_a / \tau)}
+\end{equation}
+
+% \begin{equation}\label{eq:supervised-contrastive-my-notation}
+%     \sum_{i \in I} \frac{-1}{|P(i)|} \sum_{p \in P(i)} \log \frac{\exp (f_i(x_i) \cdot f_p(x_p) / \tau) }{\sum_{a \in A(i)} \exp (f_i(x_i) \cdot f_a(x_a) / \tau)}
+% \end{equation}
+
+While this model is a strong baseline and the performance of our model is very close to this model, if the text modality is dropped, this method does not perform as good as our proposed model as shown in figure~\ref{fig:simple-MMA-mrr.srd}.
+
+\begin{figure}[tbh]
+\centering
+\includegraphics[width=.99\columnwidth]{Figures/average-seeds-epochs-mrr_ard.pdf}
+\caption{Mean Reciprocal Rank (MRR) on the held-out test set when text modality is missing averaged over 3 runs with different random seeds.
+The higher is better.
+Green is Supervised Contrastive Learning~\cite{NEURIPS2020_supervised_contrastive} and pink represents our proposed simple MMA loss function without semantic negative sampling w/ SGD w/ adjusting learning rate.
+% \todokdinline{Add legend to the graph.}
+}
+
+\label{fig:simple-MMA-mrr.srd}
+\end{figure}
 
@@
 \centering
-\begin{tabular}{|c|c|c|}
+\begin{tabular}{|c|c|c|c|c|c|}
 \toprule
-\textbf{Method} & \textbf{Accuracy} & \textbf{MRR} \\ %\hline
+\textbf{Method} & \textbf{Modalities} & \textbf{Accuracy} & \textbf{MRR.srd} & \textbf{MRR.trd} & \textbf{MRR.tsrd} \\ %\hline
 \midrule
-Full MMA (Text) w/ neg & ?$\pm$? & ?$\pm$? \\
-Simple MMA (Text) w/ neg & \textbf{0.7949}$\pm$? & \textbf{0.8778}$\pm$? \\
-Contrastive (Text) w/ neg & ?$\pm$? & ?$\pm$? \\
-Triplet (Text) w/ neg & 0.7497$\pm$? & 0.8522$\pm$? \\
-Full MMA (Speech) w/ neg & ?$\pm$? & ?$\pm$? \\
-Simple MMA (Speech) w/ neg & \textbf{0.7949}$\pm$? & \textbf{0.8778}$\pm$? \\
-Contrastive (Speech) w/ neg & ?$\pm$? & ?$\pm$? \\
-Triplet (Speech) w/ neg & 0.7497$\pm$? & 0.8522$\pm$? \\
-Triplet (Speech) w/o neg & 0.7497$\pm$? & 0.8522$\pm$? \\
+Simple MMA (Text) w/o neg  SGD & 3 & 0.8168$\pm$? & 0.8962$\pm$? & ? & ? \\
+eMMA (Text) w/o neg SGD & 3 & 0.8168$\pm$? & 0.8962$\pm$? & ? & ? \\
+Contrastive Cosine (Text) w/ neg & 3 & 0.7894$\pm$? & 0.8788$\pm$?  & ? & ?\\
+Supervised Contrastive (Text)& 3 & 0.8332$\pm$? & 0.9057$\pm$? & ? & ?\\
+Simple MMA (Text) w/o neg  SGD & 4 & ?$\pm$? & ?$\pm$? & ? & ? \\
+Supervised Contrastive (Text)& 4 & ?$\pm$? & ?$\pm$? & ? & ? \\
+Simple MMA (Speech) w/o neg & 3 & 0.7949$\pm$? & 0.8778$\pm$? & ? & ?  \\
+Triplet (Speech) w/o neg & 3 & 0.7497$\pm$? & 0.8522$\pm$? & ? & -  \\
+% Full MMA (Text) w/ neg & 3 & ?$\pm$? & ?$\pm$? \\
+% Simple MMA (Text) w/ neg Adam & 3  & 0.7949$\pm$? & 0.8778$\pm$? \\
+% Simple MMA (Text) w/ neg SGD & 3 & 0.7905$\pm$? & 0.8821$\pm$? \\
+% Triplet (Text) w/ neg & 3 & 0.7497$\pm$? & 0.8522$\pm$? \\
+% Full MMA (Speech) w/ neg & 3 & ?$\pm$? & ?$\pm$? \\
+% Simple MMA (Speech) w/ neg & 3 & 0.7949$\pm$? & 0.8778$\pm$? \\
+% Contrastive (Speech) w/ neg & 3 & ?$\pm$? & ?$\pm$? \\
+% Triplet (Speech) w/ neg & 3 & 0.7497$\pm$? & 0.8522$\pm$? \\
 \bottomrule
 \end{tabular}
-\caption{\label{table:quantitative}Average and standard deviation of accuracy and MRR over 3 runs with 3 different random seeds on a hel-out test set.
-Both metrics are from 0 to 1, and the higher is better.\todokdinline{maybe adding results on cropped and raw dataset as well.
-Also results using objec class isntead of negative sampling}}
+\caption{\label{table:quantitative}Average and standard deviation of accuracy and MRR over 3 runs with 3 different random seeds on a held-out test set.
+Both metrics are from 0 to 1, and the higher is better.
+% \todokdinline{maybe adding results on cropped and raw dataset as well. Also results using object class instead of negative sampling}
+}
 \end{table*}
 
-\begin{figure*}[tbh]
+\begin{figure}[tbh]
 \centering
-\includegraphics[width=2.0\columnwidth]{Figures/test.mrr.pdf}
-\caption{Mean Reciprocal Rank (MRR) of our proposed method against triplet loss method averaged over 3 different random seeds for a downstream task of object detection.
+\includegraphics[width=.99\columnwidth]{Figures/average-seeds-epochs-mrr_lrd.pdf}
+\caption{Mean Reciprocal Rank (MRR) on the held-out test set when speech modality is dropped averaged over 3 runs with different random seeds for a downstream task of object retrieval.
 The higher is better.
-Blue represents our proposed simple MMA loss function, and orange represent the triplet loss function~\cite{triplet_loss_2021_CVPR}.}
+Orange represents Supervised Contrastive Learning~\cite{NEURIPS2020_supervised_contrastive}, \Complete{} represents the triplet loss function~\cite{triplet_loss_2021_CVPR}.
+\todokdinline{remove simple MMA and probably add triplet loss. Any ideas?}
+}
+% \todocmi{Maybe MRR is wrong and we should be looking at top-ranked result?}
 \label{fig:simple-MMA-mrr}
-\end{figure*}
+\end{figure}
 
@@
 \label{sec:Qualitative}
+
+\begin{enumerate}
+    \item DONE Perform t-sne visualization of embeddings.
+    \item TODO Dropout modalities MRR
+\end{enumerate}
+
+\subsection{Failed Experiments}
+We tried a lot of different approaches and they did not work better than our model or the baselines, but we list them here to save time for readers who want to try them.
+
+\subsubsection{Two Anchors}
+Since speech is similar to text in its usage when it comes to the downstream task of object retrieval (given a single language command, find the correct object among multiple objects), we decided to have text and speech as two anchors instead of having one anchor only.
+When there is one anchor, the distance between positive pairs text and speech is minimized and the distance between negative pairs are minimized.
+However, this is not related to the downstream task since we never want to find the correct speech for a given text or vice versa.
+The second term added to the last function is exactly similar to the loss function when speech is the anchor, and the first term is the same as when language is the anchor.
+Therefore, the performance is somewhere in the middle of those two approaches but more towards the speech when the metrics are computed based and speech matrices.
 
@@
 
+% %% The acknowledgments section is defined using the "acks" environment
+% %% (and NOT an unnumbered section). This ensures the proper
+% %% identification of the section in the article metadata, and the
+% %% consistent spelling of the heading.
+\begin{acks}
+To Robert, for the bagels and explaining CMYK and color spaces.
+\end{acks}
+% \section*{Acknowledgments}
+
 \clearpage
 \clearpage
-\section*{Acknowledgments}
-%% The file named.bst is a bibliography style file for BibTeX 0.99c
-\bibliographystyle{named}
+% %%
+% %% The next two lines define the bibliography style to be used, and
+% %% the bibliography file.
+\bibliographystyle{ACM-Reference-Format}
 \bibliography{ref}
-\clearpage
 
+% %%
+% %% If your work has an appendix, this is the place to put it.
 \appendix
@@
 \section{Qualitative Analysis}
+\section{Model}
 
-\section{Model}
+\section{Data}
+
+\begin{figure*}[tbh]
+\centering
+\includegraphics[width=2.0\columnwidth]{Figures/raw-mask-test-mrr.pdf}
+\caption{Mean Reciprocal Rank (MRR) on different versions of GoLD dataset~\cite{GoLD_UMBC}.
+The raw version is depicted in green and masked version is depicted in orange averaged over 3 different random seeds.
+Higher is better.}
+\label{fig:mask-vs-raw}
+\end{figure*}
 
@@
 
+\begin{figure*}[tbh]
+\centering
+\includegraphics[width=2.1\columnwidth]{Figures/test.mrr.pdf}
+\caption{Mean Reciprocal Rank (MRR) of our proposed method against triplet loss method averaged over 3 different random seeds for a downstream task of object detection.
+The higher is better.
+Green is Supervised Contrastive Learning~\cite{NEURIPS2020_supervised_contrastive}, Pink represents our proposed simple MMA loss function without semantic negative sampling w/ SGD w/ adjusting learning rate, Purple is out method with semantic negative sampling, Blue is our method w/ semantic negative sampling w/ Adam and w/o adjusting learning rate, Red is vanilla contrastive loss with cosine, and orange represent the triplet loss.
+function~\cite{triplet_loss_2021_CVPR}.
+}
+
+\label{fig:simple-MMA-mrr}
+\end{figure*}
+
 \section{Setup}
 
-\input{lit-search}
+% \input{lit-search}
 

### 2022-01-30 Code/ML.py
save text & candidates, example predictions

@@ -105,6 +105,7 @@ def training(config, models, dataset, portion, optimizers, epoch, criterion):
             batch_loss = contrastive_loss(data['pos'], data['neg'], models)
         elif config.method == 'supervised-contrastive':
             features = torch.cat([models['language'](data['pos']['language'])['decoded'].unsqueeze(1), models['rgb'](data['pos']['rgb'])['decoded'].unsqueeze(1), models['depth'](data['pos']['depth'])['decoded'].unsqueeze(1), models['audio'](data['pos']['audio'])['decoded'].unsqueeze(1)], dim=1)
+            # features = torch.cat([models['language'](data['pos']['language'])['decoded'].unsqueeze(1), models['rgb'](data['pos']['rgb'])['decoded'].unsqueeze(1), models['depth'](data['pos']['depth'])['decoded'].unsqueeze(1)], dim=1)
             batch_loss = criterion(features, data['pos']['object'])
 
         # saving average loss per epoch. values in batch_loss have backward_fn and requires_grad
@@ -128,7 +129,7 @@ def evaluate(config, models, dataset, portion):
     This function runs the model over valid and/or test set
     Returns f1, precision, accuracy, and the model outputs
     """
-    outputs = {'rgb': [], 'depth': [], 'language': [], 'audio': [], 'object_names': [], 'instace_names': [], 'image_names': []}
+    outputs = {'rgb': [], 'depth': [], 'language': [], 'audio': [], 'object_names': [], 'instace_names': [], 'image_names': [], 'descriptions': []}
     
     for modality in models.keys():
         model = models[modality]
@@ -141,13 +142,14 @@ def evaluate(config, models, dataset, portion):
             outputs['object_names'].extend(data['object'])
             outputs['instace_names'].extend(data['instance'])
             outputs['image_names'].extend(data['img_name'])
+            outputs['descriptions'].extend(data['description'])
         
         for key in outputs.keys():
-            if key not in ['object_names', 'instace_names', 'image_names']:
+            if key not in ['object_names', 'instace_names', 'image_names', 'descriptions']:
                 outputs[key] = np.concatenate(outputs[key], axis=0)
 
         if config.metric_mode == 'sampling':
[…]

### 2022-01-30 Code/analysis.py

@@ -7,13 +7,26 @@ import matplotlib.pyplot as plt
 import tikzplotlib
 
 
-
 exp_ugly_names = {
-    '4M-eMMA': '4M-eMMA-cosine-submodalities-text-anchor-gold-no_neg_sampling-1024', 
-    '4M-SupCon': '4M-supervised-contrastive-SGD-cosine-submodalities-text-first-gold-no_neg_sampling-1024',
-    '4M-simple-MMA': '4M-mma-simple-SGD-cosine-submodalities-text-anchor-gold-no_neg_sampling-1024',
-    '4M-contrastive': '4M-contrastive-cosine-submodalities-text-anchor-gold-no_neg_sampling-1024',
+    '4M-eMMA-text': '4M-eMMA-cosine-submodalities-text-anchor-gold-no_neg_sampling-1024',
+    '3M-eMMA-text': '3M-eMMA-cosine-submodalities-text-anchor-gold-no_neg_sampling-1024',
+    '4M-SupCon-text': '4M-supervised-contrastive-SGD-cosine-submodalities-text-first-gold-no_neg_sampling-1024',
+    '3M-SupCon-text': '3M-supervised-contrastive-SGD-cosine-submodalities-text-first-gold-no_neg_sampling-1024',
+    '4M-contrastive-text': '4M-contrastive-cosine-submodalities-text-anchor-gold-no_neg_sampling-1024',
+    '3M-contrastive-text': '3M-contrastive-SGD-cosine-submodalities-text-anchor-gold-no_neg_sampling-1024',
+
 }
+
+# exp_ugly_names = {
+    # '3M-triplet-text':  'triplet-cosine-Adam-text-anchor-gold-gold-no_neg_sampling-1024',
+    # '3M-eMMA-audio': '3M-eMMA-cosine-submodalities-audio-anchor-gold-no_neg_sampling-1024',
+    # '3m-SupCon-audio': '3M-supervised-contrastive-SGD-cosine-submodalities-audio-first-gold-no_neg_sampling-1024',
+    # '4M-simple-MMA': '4M-mma-simple-SGD-cosine-submodalities-text-anchor-gold-no_neg_sampling-1024',
+#     '3M-triplet-audio': 'triplet-cosine-Adam-audio16-anchor-gold-gold-no_neg_sampling-1024',
+#     '3M-eMMA-audio': '3M-eMMA-cosine-submodalities-audio-anchor-gold-no_neg_sampling-1024',
[…]

### 2022-01-30 Code/run_cluster.sh

@@ -48,6 +48,7 @@ sim=(snp_noise gauss_noise dropout_noise RIVR clean_normalized)
 real=(gold gold_raw gold_cropped gold_no_crop_old)
 
 # Frequently changing
+method='eMMA'
 data_type='rgbd' # 'rgb', 'depth', 'rgbd'
 per_epoch='best' # 'best' or 'all'
 epoch=200
@@ -82,7 +83,7 @@ scratch_data=$TMPDIR/data/
 
 
 python -u ML.py --wandb_track $track --experiment_name $exp_name --epochs $epoch --task $task --data_dir $scratch_data \
---random_seed $random_seed --embed_dim $embed_dim --data_type $data_type \
+--random_seed $random_seed --embed_dim $embed_dim --data_type $data_type --method $method \
 --prediction_thresh $pred  --results_dir $results_dir  --exp_full_name $exp_full_name \
 --learning_rate $lr --eval_mode $eval_mode  --per_epoch $per_epoch --batch_size $batch_size \
 --activation $activation --gpu_num $gpu_num --negative_sampling $negative_sampling 2>&1 | tee -a $results_dir'out.log'

### 2022-01-31 Paper/appendix.tex
paper submitted to SIGIR

--- 2021-11-19 Paper/main.tex
@@
-% EMNLP header. Do not change.
-\pdfoutput=1
-\documentclass[11pt]{article}
-\usepackage[review]{emnlp2021}
-\usepackage{times}
-\usepackage{latexsym}
-\usepackage[T1]{fontenc}
-\usepackage[utf8]{inputenc}
-\usepackage{microtype}
-% End header.
+\section{Qualitative Analysis}
+\section{Approach}
 
-%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%
-% EDITING LINK: 
-% https://www.overleaf.com/3575751412sjsdmszkxsyh
-%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%
-
-\usepackage{soul}
-\usepackage{url}
-% \usepackage[hidelinks]{hyperref}
-% \usepackage[small]{caption}
-\usepackage{graphicx}
-\usepackage{amsmath}
-\usepackage{amsthm}
-\usepackage{booktabs}
-\usepackage{algorithm}
-\usepackage{algorithmic}
-\urlstyle{same}
-
-% =======================================
-% my imported packages
-
-\usepackage{amsmath}
-\usepackage{bm}
-\usepackage{amssymb}
-\usepackage{listings}
-% \usepackage[ruled,vlined]{algorithm2e}
-\include{pythonlisting}
-
-\newcommand{\pseudosection}[1]{\vspace{1.5ex}\noindent \textit{{#1:~~}}}
-% \newcommand{\pseudosection}[1]\indent{{{}}}
-
-\usepackage{todonotes}
-\newcommand{\todokdinline}[1]{\todo[color=red!20,inline]{{KD: \small #1}}}
-\newcommand{\todokd}[1]{\todo[color=red!20]{{\small #1 -- Kasra}}}
-\newcommand{\todocainline}[1]{\todo[color=yellow!20,inline]{{CA: \small #1}}}
-\newcommand{\todocm}[1]{\todo[color=green!40]{\small #1 -- Cynthia}}
-\newcommand{\todocmi}[1]{\todo[inline,color=green!40]{\small #1 -- Cynthia}}
-\newcommand{\todoff}[1]{\todo[color=blue!20]{\small #1 -- Frank}}
-\newcommand{\todoed}[1]{\todo[color=cyan!20]{\small #1 -- Ed}}
-
-\newcommand{\CitationNeeded}[1]{{\textbf {\color{red}Cite #1}}}
-\newcommand{\ProofRead}[1]{{\textbf {\color{red}Proofread please #1}}}
-\newcommand{\Rephrase}[1]{{\textbf {\color{red}Rephrase please #1}}}
-
-% \newcommand{\citet}[1]{\cite{#1}} % This is used because \citet didn't work with their author kit % It does now
-
-% =======================================
-
-\title{MMA: Multi-modal Manifold Alignment}
-
-\author{
-First Author$^1$
-\and
-Second Author$^2$\and
-Third Author$^{2,3}$\And
-Fourth Author$^4$
-\affiliations
-$^1$First Affiliation\\
-$^2$Second Affiliation\\
-$^3$Third Affiliation\\
-$^4$Fourth Affiliation
-\emails
-\{first, second\}@example.com,
-third@other.example.com,
-fourth@example.com
-}
-
-\begin{document}
-
-Todo list
-
-\begin{enumerate}
-    \item Lit search 
-\end{enumerate}
-
-\newpage
-
-\maketitle
-\setcounter{page}{1}
-
-% ***************** The BEGINNING of the paper *****************
-
-\begin{abstract}
-
-\end{abstract}
-
-%===================================================================
-\section{Lit Search}
-
-\subsection{Multimodal Machine Learning: A Survey and Taxonomy}
+For $M$ modalities, we define a distance based loss function which is a generalization of well-known similarity-based metric learning and manifold alignment methods such as triplet loss~\cite{Carvalho-cooking-triplet,triplet_loss_2021_CVPR}, quadruplet loss~\cite{chen2017beyond}, and is similar to contrastive loss under some settings.
+We sample two data points from each modality, one positive point and one negative point. 
+For each set of positive data points (referring to an instance of an object with all of its modalities), we choose a set of negative data points either based on object class name, or by semantic negative sampling~\cite{Pillai_Matuszek_2018}.
+We now have two data points; we consider the first one as positive, and the second one as negative.
+However, the anchor won't be another data point as it is in the triplet loss terminology.
+Instead we choose one domain as an anchor.
 
 \begin{itemize}
-\item \href{https://ieeexplore.ieee.org/document/8269806}{url}
+    % \item data point: a row in the csv/tsv file.
+    \item Positive: a set of embeddings of the one data point (RGB image of an apple, corresponding depth image, text description, and speech description of the same apple)
+    \item Negative: a set of embeddings of the another data point of a different object (RGB image of a book, corresponding depth image, text description, and speech description of the same book)
+    \item Anchor: A fixed modality to act as anchor in the case of simple MMA loss, where we don't minimize the distance between the pair of embeddings from all modalities of the same data point, and we don't maximize the distance between positive an negative datapoints of the same modality.
 \end{itemize}
 
-\citet{baltrusaitisMultimodalMachineLearning2019} proposes a new taxonomy of multimodal machine learning by introducing five technical challenges in addition to typical early and late fusion categorization.
-This taxonomy includes the following categories.
-\begin{enumerate}
-\item \textit{Representation}: represent and summarize multimodal data in a way that exploits the complementarity and redundancy of multiple modalities.
-\begin{enumerate}
-\item joint: combine the unimodal signals into the same representation space.
-\item coordinated: process unimodal signals separately, but enforce certain similarity or structure constraints on them to bring them to a coordinated space.
-\end{enumerate}
-\item \textit{Translation}: map data from one modality to another.
-\begin{enumerate}
-\item example-based
-\item generative
-\end{enumerate}
-\item \textit{Alignment}: identify the direct relations between (sub)elements from two or more different modalities.
-\begin{enumerate}
-\item explicit
-\item implicit
-\end{enumerate}
-\item \textit{Fusion}: join information from two or more modalities to perform a prediction.
-\begin{enumerate}
-\item model-agnostic
-\item model-based
-\end{enumerate}
-\item \textit{Co-learning}: transfer knowledge between modalities, their representation, and their predictive models
-\begin{enumerate}
-\item parallel data
-\item non-parallel data
-\item hybrid data
-\end{enumerate}
-\end{enumerate}
-
-The authors mention that while joint representations have been used in situations to construct representations of more than two modalities, coordinated spaces have, so far, been mostly limited to two.
-This means that our research is novel and we can extend similarity measures to more than two modalities.
-
-\subsection{Self-Supervised MultiModal Versatile Networks}
-
-\begin{itemize}
-\item \href{https://proceedings.neurips.cc/paper/2020/file/0060ef47b12160b9198302ebdb144dcf-Paper.pdf?utm\_campaign=NLP\%20News\&utm\_medium=email\&utm\_source=Revue\%20newsletter}{url}
-\end{itemize}
-
-This work uses self-supervised contrastive learning to learn and combine representations from three modalities of visual, audio, and language.
-They learn a \textit{multimodal versatile network} that has the following four properties:
-\begin{enumerate}
-\item Ability to take any of the modalities as input
-\item Respecting the specificity of modalities: audio and visual modalities are fine-grained while language modality is coarse-grained.
-\item Comparability of different modalities using \textbf{dot product} even if not seen together during training 
-\item Efficiently applicable to visual data either in the form of dynamic videos or static images
-\end{enumerate}
-
-The authors consider the following three different configurations of the modality spaces which they call \textit{modality embedding graphs}.
-\begin{enumerate}
-\item Shared: all modalities map to the same space. This respects property 3 but violates property 2.
-\item Disjoint: visual-audio and visual-text spaces. This respects property 2 but violates property 3.
-\item Fine and coarse: audio and visual domains are map to the same space since they are fine-grained.
-These embeddings are then mapped to a lower dimensional space where text is also mapped.
-This respects both properties 2 and 3.
-\end{enumerate}
-
-The loss function they define consists of two terms.
-The first term is to train the fine-grained space of visual and audio embeddings by minimizing the distance between the similar pair (positives) from the same location of a video, and maximizing the distance between negative pairs (negatives) sampled from different videos.
-The second term is to train the coarse-grained space which consists of visual and text embeddings.
-Note that they don't consider audio embeddings in this space since they don't want to learn automatic speech recognition (ASR).
-Text and visual domains are not aligned as well as audio and visual domains (e.g.
-sound of playing piano versus the text that only says playing piano for a video of playing piano).
-To cure this, they consider a set of positive pairs instead of a single pair.
-If a modality is missed, the corresponding term is dropped from the loss function.
-
-\subsection{Separating Self-Expression and Visual Content in Hashtag Supervision}
-
-\begin{itemize}
-\item \href{https://openaccess.thecvf.com/content\_cvpr\_2018/papers/Veit\_Separating\_Self-Expression\_and\_CVPR\_2018\_paper.pdf}{url}
-\end{itemize}
-
-The authors train a joint model of images, hashtags, and users to perform image retrieval.
-This is similar to our task where given a description, we want to find the image in the scene that best matches the description.
-The idea is to form a three-way tensor product model.
-They use a ranking loss to train the model where the score of an observed triplet is higher than an unobserved triplet.
-They sample six negative triplets per positive sample triplet, and use each of them as a negative in the loss.
-The downstream retrieval task is then simply done by taking the arg max of the tensor product for a given user.
-
-\subsection{Semi-Heterogeneous Three-Way Joint Embedding Network for Sketch-Based Image Retrieval}
-
-\begin{itemize}
-\item \href{https://ieeexplore.ieee.org/document/8809264}{url}
-\end{itemize}
-
-\subsection{Deep Multimodal Learning for Affective Analysis and Retrieval}
-
-\begin{itemize}
-\item \href{https://ieeexplore.ieee.org/abstract/document/7277066?casa\_token=aBp6BxcszHwAAAAA:3NoMiFrZbn7tXfavF1rgkCiGWbFI2arxn8Xb6iDF79q4zBZHWi7PWWhf6xW-xJwYdFALbmRENo4}{url}
-\end{itemize}
-
-\subsection{An Efficient Framework for Zero-Shot Sketch-Based Image Retrieval}
-
-\begin{itemize}
-\item \href{https://arxiv.org/pdf/2102.04016.pdf}{url}
-\end{itemize}
-This paper uses two modalities only; image and sketch.
-However, it uses a \textit{quadruplet} to compute the loss which can be useful in our research.
-A quadruplet is composed of a sketch picture as an anchor, a negative example from sketch domain, a negative example from picture domain, and a positive example from picture domain.
-
-\subsection{SMIL: Multimodal Learning with Severely Missing Modality}
-
-\begin{itemize}
-\item \href{https://www.aaai.org/AAAI21Papers/AAAI-437.MaM.pdf}{url}
-\end{itemize}
-Uses Bayesian meta-learning framework to perform multimodal learning with partially missing modalities in training/test data.
-One of their experiments is multi-label classification of movie genres with bimodal data including poster of movies and description of the movie from IMDB.
-
-\subsection{Multimodal Language Analysis in the Wild: CMU-MOSEI Dataset and Interpretable Dynamic Fusion Graph}
-
-\begin{itemize}
-\item \href{https://aclanthology.org/P18-1208.pdf}{url}
-\end{itemize}
-
-\subsection{Multimodal Learning with Incomplete Modalities by Knowledge Distillation}
-
-\begin{itemize}
-\item \href{https://dl.acm.org/doi/pdf/10.1145/3394486.3403234}{url}
-\end{itemize}
-They first train models on each modality independently
-\subsection{SWAFN: Sentimental Words Aware Fusion Network for Multimodal Sentiment Analysis}
-
-\begin{itemize}
-\item \href{https://aclanthology.org/2020.coling-main.93.pdf}{url}
-\end{itemize}
-
-\subsection{Multimodal Learning for Human Action Recognition Via Bimodal/Multimodal Hybrid Centroid Canonical Correlation Analysis}
-
-\begin{itemize}
-\item \href{https://ieeexplore.ieee.org/document/8489981}{url}
-\end{itemize}
-
-\subsection{Deception Detection Using a Multimodal Approach}
-
-\begin{itemize}
-\item \href{https://dl.acm.org/doi/pdf/10.1145/2663204.2663229?casa\_token=KneK7B7xvLwAAAAA:Ajz0N96ygq8ktwkz0IVUqTt8NozCg2wR6n\_x2xntdHqZBh6VXW\_8VbO4GeY4VvMsDJlMwzkhVXQSJA}{url}
-\end{itemize}
-They use \textit{language}, \textit{physiological response}, and \textit{thermal sensing} to detect deceit.
-
-\subsection{M3ER: Multiplicative Multimodal Emotion Recognition using Facial, Textual, and Speech Cues}
-
-\begin{itemize}
-\item \href{https://ojs.aaai.org/index.php/AAAI/article/view/5492}{url}
-\end{itemize}
-
-\subsection{Found in Translation: Learning Robust Joint Representations by Cyclic Translations between Modalities}
-
-\begin{itemize}
-\item \href{https://ojs.aaai.org/index.php/AAAI/article/view/4666}{url}
-\end{itemize}
-\subsection{Select-Additive Learning: Improving Generalization In Multimodal Sentiment Analysis}
-
-\begin{itemize}
-\item \href{https://ieeexplore.ieee.org/stamp/stamp.jsp?tp=\&arnumber=8019301}{url}
-\end{itemize}
-
-\subsection{Attention-Based Multimodal Fusion for Video Description}
-
-\begin{itemize}
-\item \href{https://openaccess.thecvf.com/content\_ICCV\_2017/papers/Hori\_Attention-Based\_Multimodal\_Fusion\_ICCV\_2017\_paper.pdf}{url}
-\end{itemize}
-\subsection{\st{3W-AlignNet: a Feature Alignment Framework for Person Search with Three-Way Decision Theory}}
-
-\begin{itemize}
-\item \href{https://link.springer.com/article/10.1007/s12559-021-09898-7}{url}
-\end{itemize}
-
-This is not very related since they use three-way decision theory to select bounding boxes as positive, negative, and boundary.
-Three-way here refers to something else and not modalities.
-
-\subsection{\st{Unsupervised Domain Adaptation for Face Recognition in Unlabeled Videos}}
-\begin{itemize}
-\item \href{https://openaccess.thecvf.com/content\_ICCV\_2017/papers/Sohn\_Unsupervised\_Domain\_Adaptation\_ICCV\_2017\_paper.pdf}{url}
-\end{itemize}
-Might be relevant but not super relevant.
-
-\subsection{Heterogeneous Sensor Data Fusion By Deep Multimodal Encoding}
-\begin{itemize}
-\item \href{https://ieeexplore.ieee.org/stamp/stamp.jsp?tp=&arnumber=7874158}{url}
-\end{itemize}
-This one takes two modalities as input and outputs predictions in a third modality.
-
-%===================================================================
-\section{Introduction}
-\label{intro}
-
-% Grounded language learning~\cite{Mooney2008Grounded,RichardsDarvishMatuszekCategoryFree20,triplet_loss_2021_CVPR} is a cross-modal learning task to connect natural language with the physical environment in which the natural language is used.
-
-% Our contributions are as follows:
-% \begin{enumerate}
-%     \item 
-%     \item 
-%     \item 
-% \end{enumerate}
-
-%===================================================================
-
-\section{Related Work}
-\label{sec:Related-Work}
-
-%===================================================================
-
-\section{Problem Description}
-\label{sec:Problem-Description}
-
-%===================================================================
-
-\section{Approach}
-\label{sec:Method}
-
-For $M$ modalities, we define a distance based loss function that contains $M+2(M-1)!$ terms. 
-We sample two data points from each modality, one positive point and one negative point.
 The objective is then to (I) minimize the distance between each pair of positive points from heterogeneous modalities which results in $(M-1)!$ terms, (II) maximize the distance between each pair of positive and negative points from heterogeneous modalities which contains $(M-1)!$ terms again, and (III) maximize the distance between each pair of positive and negative points from the homogeneous modalities which contains $M$ terms.
-Equation~\ref{eq:objective} shows the objective function we defined above with changing maximization to minimization by changing the sign.
+Altogether, our proposed loss function contains $M+2(M-1)!$ terms.
+\Cref{eq:objective} shows the objective function we defined above with changing maximization to minimization by changing the sign.
 
@@
 
-\begin{aligned}\label{eq:objective}
-    \mathcal{L}  &= \sum_{i=1}^{M-1} \sum_{j=i+1}^{M} ||f_i(x_{i}^{+}) - f_j(x_{j}^{+})||
-    - ||f_i(x_{i}^{+}) - f_j(x_{j}^{-})|| \\ 
-    & - ||f_i(x_{i}^{-}) - f_j(x_{j}^{+})|| - \sum_{i=1}^{M} ||f_i(x_{i}^{+}) - f_i(x_{i}^{-}) ||
-\end{aligned}
-where the subscripts $i$ and $j$ represent modalities, and $f_i(.)$ and $f_j(.)$ represent functions that map input data to the shared latent space using a deep neural network.
+% \begin{equation}\label{eq:objective}
+% \begin{split}
+%     \mathcal{L}  &= \sum_{i=1}^{M-1} \sum_{j=i+1}^{M} ||f_i(x_{i}^{+}) - f_j(x_{j}^{+})|| \\
+%     & - ||f_i(x_{i}^{+}) - f_j(x_{j}^{-})||
+%     - ||f_i(x_{i}^{-}) - f_j(x_{j}^{+})|| \\
+%     & - \sum_{i=1}^{M} ||f_i(x_{i}^{+}) - f_i(x_{i}^{-}) ||
+% \end{split}
+% \end{equation}
+% where $f_i(.)$ and $f_j(.)$ represent functions that map input data to the shared latent space using a deep neural network.
+
+% d(,) notation instead of || - ||
+% \begin{equation}\label{eq:objective}
+% \begin{split}
+%     \mathcal{L}  &= \sum_{i=1}^{M-1} \sum_{j=i+1}^{M} d(z_{i}^{+}, z_{j}^{+}) 
+%      - d(z_{i}^{+}, z_{j}^{-}) \\ 
+%     & - d(z_{i}^{-}, z_{j}^{+}) - \sum_{i=1}^{M} d(z_{i}^{+}, z_{i}^{-})
+% \end{split}
+% \end{equation}
+
+\begin{equation}\label{eq:objective}
+\begin{split}
+    \mathcal{L}  &= \sum_{i=1}^{M-1} \sum_{j=i+1}^{M} ||z_{i}^{+} - z_{j}^{+}|| 
+     - ||z_{i}^{+} - z_{j}^{-}|| - ||z_{i}^{-} - z_{j}^{+}|| \\ 
+    &  - \sum_{i=1}^{M} ||z_{i}^{+} - z_{i}^{-} ||
+\end{split}
+\end{equation}
+where the subscripts $i$ and $j$ represent modalities, and $z$ is the embedding we get by applying a mapping function $f$ which in our case is a neural network on our input data.
+In other words $z_m = f_m(x_m)$ where each modality $m$ has a specific model $f_m$ that is different from the models for other modalities and these models do not share their weights.
+To measure the distance, we use cosine distance.
+In order to use cosine \textit{distance}, we have to subtract the cosine of the \textit{angle} between two embeddings (which represents similarity) from 1: $1 - \cos(e_1, e_2)$.
 
@@
 However, it breaks the nice procedure of sampling two data points per modality}
+\todo{Edward wonders if $f_i(x)$ should instead produce $\mu_x$ and $\Sigma_x$, to capture uncertainty of embedding location, when working with so many modalities?
+(e.g., not all modalities share the same ambiguities)}
+\todo{Does Edward mean the actual network produces $\Sigma_x$, e.g., something like a VAE? How would uncertainty be applied to the inference problem?}
 $+ \sum_{i=1}^{M} ||f_i(x_i^{a}) - f_i(x_i^{+})||$
 
-Our method is a generalization of well-known similarity-based metric learning and manifold alignment methods such as triplet loss~\CitationNeeded{}.
-If $M=2$ which means the number of modalities is 2, then our objective function reduces to quadruplet loss method.
-If $M=2$ and we ignore the last term in the objective function, it results in the triplet loss method. 
+If $M=2$ which means the number of modalities is 2, then our objective function reduces to the quadruplet loss method~\cite{chen2017beyond} if we ignore the third term, and negative samples for each modality do not belong to the same class.
+If $M=2$ and we ignore the last two terms in the objective function, it results in the triplet loss method. 
 If $M=1$, only the last term remains in the loss function which is exactly the pairwise distance-based loss function.
 
-\subsection{Network Architecture}
-\label{sec:Model}
+We demonstrate performance using three different distance-based loss functions.
+\todokdinline{Either remove it or experiment with all versions of MMA (12 terms, 9 terms, and explicit anchor maximization}
 
-%===================================================================
+\subsection{Simple MMA}
+\label{sub:simple-mma}
+Our first proposed loss function is a modification of the triplet loss function idea, and can be derived from \cref{eq:objective} by fixing one modality as anchor, and ignoring the last two terms in \cref{eq:objective}.
 
-\section{Experiments}
-\label{sec:Experiments}
+% For each data point (referring to an instance of an object with all of its modalities), we choose a negative data point either based on object class name, or by semantic negative sampling~\cite{Pillai_Matuszek_2018}.
+We now have two data points; we consider the first one as positive, and the second one as negative.
+However, the anchor won't be another data point as it is in the triplet loss terminology.
+Instead we choose one domain as an anchor.
 
-\subsection{Data}
-\label{sec:Data}
+% \begin{itemize}
+%     % \item data point: a row in the csv/tsv file.
+%     \item Positive: a set of embeddings of the one data point (RGB image of an apple, depth image of an apple, language description of an apple)
+%     \item Negative: a set of embeddings of the another data point of a different object (RGB image of a book, depth image of a book, language description of a book)
+%     \item Anchor: A fixed modality to act as anchor in the case of simple MMA loss, where we don't minimize the distance between the pair of embeddings from all modalities of the same data point, and we don't maximize the distance between positive an negative datapoints of the same modality.
+% \end{itemize}
 
-\subsection{Setup}
-\label{sec:setup}
+Since the downstream task in the grounded language learning domain is to predict/retrieve a desired object among multiple objects given a natural language description, it makes sense to train the model using language as an anchor, and in fact anchoring learning around language outperforms anchoring on RGB.
 
-\subsection{Quantitative Results}
-\label{sec:Quantitative}
+\Cref{eq:objective-simple-mma} shows the mathematical formulation of the proposed loss function which can be generalized to arbitrary number of modalities.
 
-\subsection{Qualitative Results}
-\label{sec:Qualitative}
+% \begin{equation}
+% \label{eq:objective-simple-mma}
+% \begin{split}
+%     \mathcal{L}  &= \sum_{m=1}^{M-1} || f_a(x_{a}^{+}) - f_m(x_{m}^{+}) || - \\ 
+%     & || f_a(x_{a}^{+}) -f_m(x_{m}^{-}) || 
+%     \\
+%     &= \sum_{m=1}^{M-1} 1 - \cos(f_a(x_{a}^{+}), f_m(x_{m}^{+})) \\ & - (1 - \cos(f_a(x_{a}^{+}) ,f_m(x_{m}^{-}))) \\
+%     &= \sum_{m=1}^{M-1} \cos(f_a(x_{a}^{+}) ,f_m(x_{m}^{-})) \\ 
+%     & - \cos(f_a(x_{a}^{+}), f_m(x_{m}^{+}))
+% \end{split}
+% \end{equation}
 
-\subsection{Baselines}
+\begin{equation}
+\label{eq:objective-simple-mma}
+\begin{split}
+    \mathcal{L}  &= \sum_{m=1}^{M-1} || z_{a}^{+} - z_{m}^{+} || - || z_{a}^{+} - z_{m}^{-} || \\
+    &= \sum_{m=1}^{M-1} 1 - \cos(z_{a}^{+}, z_{m}^{+}) - (1 - \cos(z_{a}^{+}, z_{m}^{-})) \\
+    &= \sum_{m=1}^{M-1} \cos(z_{a}^{+} ,z_{m}^{-}) - \cos(z_{a}^{+}, z_{m}^{+})
+\end{split}
+\end{equation}
 
-%===================================================================
+where subscript $a$ represents \textit{anchor} modality, $M$ is the number of modalities, superscript $+$ represents the positive objects, and superscript $-$ represents the negative objects.
 
-\section{Conclusion}
-\label{sec:conclusion}
+An illustration of this loss function is provided in \cref{fig:simple-mma-loss}.
 
-%===================================================================
+\begin{figure}[tb]
+\centering
+\includegraphics[width=\columnwidth]{Figures/simple-MMA.pdf}
+\caption{A high-level prototype of the distances used in the simple MMA loss with 3 modalities.
+Orange arrows indicate maximizing the distance, blue arrows indicate minimizing the distance, and green lines show the margin.
+Depth images are brightened to be visible.
+\todo[inline]{We need to show speech as the 4th modality as well.}}
+\label{fig:simple-mma-loss}
+\end{figure}
 
-% ***************** The END of the paper *****************
+This version of MMA loss function can be seen as a contrastive loss usually used in the domain of self-supervised learning~\cite{chen2020simple}.
+However, our proposed loss function has two advantages over the traditional contrastive loss expressed in \cref{eq:contrastive-loss}.
+The first advantage is that our loss function do not loop over multiple positives and negatives in a large batch.
+Instead we sample only two objects (positive and negative) each of which have $M$ modalities which gives us $2M$ datapoints (or embeddings).
+Hence, our model can be trained using smaller batch sizes and reduces the number of negative samples we need.
+The second advantage is that this loss function can be used in a multimodal setting with an arbitrary number of modalities, and is not limited to a single data type (e.g.
+RGB images) which is the most common usage of contrastive loss~\cite{chen2020simple, NEURIPS2020_supervised_contrastive}.
 
-\clearpage
-\section*{Acknowledgments}
-%% The file named.bst is a bibliography style file for BibTeX 0.99c
-\bibliographystyle{named}
-\bibliography{ref}
-\clearpage
+Therefore, our proposed loss function would modify \cref{eq:contrastive-loss} to the loss function formulated in \cref{eq:simple-MMA-contrastive}.
+Our formulation is a generalized version of the contrastive loss where $M$ is the number of modalities (e.g.
+RGB image, depth image, speech, text), but it can be the ``augmented views'' instead of different modalities.
 
-\appendix
-\label{sec:appendix}
-\section{Qualitative Analysis}
+% \begin{equation}\label{eq:simple-MMA-contrastive}
+%     -\sum_{m=1}^{M-1} \log \frac{\exp (sim(f_a(x_a^+) , f_m(x_{m}^{+})/ \tau) )}{ \exp (sim(f_a(x_a^+) , f_m(x_{m}^{+}) / \tau)) + \exp (sim(f_a(x_a^+), f_m(x_{m}^{-}) / \tau))}
+% \end{equation}
 
-\section{Model}
+\begin{equation}\label{eq:simple-MMA-contrastive}
+    -\sum_{m=1}^{M-1} \log \frac{\exp (sim(z_a^+ , z_{m}^{+})/ \tau) }{ \exp (sim(z_a^+ , z_{m}^{+}) / \tau) + \exp (sim(z_a^+, z_{m}^{-}) / \tau)}
+\end{equation}
+where $z = f(x)$.
+
+% \todokdinline{I should run the traditional contrastive loss to see if my method works better to make this claim.}
+
+% \subsubsection{All in loss with 12 terms}
+% \subsubsection{subset of All in loss with 9 terms}
+
+\section{Data}
+
+\begin{figure*}[tbh]
+\centering
+\includegraphics[width=2.0\columnwidth]{Figures/raw-mask-test-mrr.pdf}
+\caption{Mean Reciprocal Rank (MRR) on different versions of GoLD dataset~\cite{GoLD_UMBC}.
+The raw version is depicted in green and masked version is depicted in orange averaged over 3 different random seeds.
+Higher is better.}
+\label{fig:mask-vs-raw}
+\end{figure*}
 
@@
 
+\begin{figure*}[tbh]
+\centering
+\includegraphics[width=2.1\columnwidth]{Figures/test.mrr.pdf}
+\caption{Mean Reciprocal Rank (MRR) of our proposed method against triplet loss method averaged over 3 different random seeds for a downstream task of object detection.
+The higher is better.
+Green is Supervised Contrastive Learning~\cite{NEURIPS2020_supervised_contrastive}, Pink represents our proposed simple MMA loss function without semantic negative sampling w/ SGD w/ adjusting learning rate, Purple is out method with semantic negative sampling, Blue is our method w/ semantic negative sampling w/ Adam and w/o adjusting learning rate, Red is vanilla contrastive loss with cosine, and orange represent the triplet loss.
+function~\cite{triplet_loss_2021_CVPR}.
+}
+
+\label{fig:simple-MMA-mrr}
+\end{figure*}
+
 \section{Setup}
 
-\end{document}
+% \input{lit-search}

### 2022-01-31 Paper/main.tex
paper submitted to SIGIR

--- 2021-11-19 Paper/main.tex
@@
-% EMNLP header. Do not change.
-\pdfoutput=1
-\documentclass[11pt]{article}
-\usepackage[review]{emnlp2021}
-\usepackage{times}
-\usepackage{latexsym}
-\usepackage[T1]{fontenc}
-\usepackage[utf8]{inputenc}
-\usepackage{microtype}
-% End header.
+\documentclass[sigconf,natbib=true,anonymous=true]{acmart}
+% \documentclass[manuscript,screen,review]{acmart}
+
+% Needed to fix an acmart problem with line numbering. AFAICT line 
+% numbers are not expected in sigconf, so it should be a no-op.
+\makeatletter
+\providecommand{\@LN}[2]{}
+\makeatother
+
+% % Fonts used in the template cannot be substituted; margin 
+% % adjustments are not allowed.
+% %
+% % \BibTeX command to typeset BibTeX logo in the docs
+% \AtBeginDocument{%
+%   \providecommand\BibTeX{{%
+%     \normalfont B\kern-0.5em{\scshape i\kern-0.25em b}\kern-0.8em\TeX}}}
+
+%% Rights management information.  This information is sent to you
+%% when you complete the rights form.  These commands have SAMPLE
+%% values in them; it is your responsibility as an author to replace
+%% the commands and values with those provided to you when you
+%% complete the rights form.
+\setcopyright{acmcopyright}
+\copyrightyear{2022}
+\acmYear{2022}
+\acmDOI{XXXXXXX.XXXXXXX}
+
+%% These commands are for a PROCEEDINGS abstract or paper.
+\acmConference[SIGIR '22]{Make sure to enter the correct
+  conference title from your rights confirmation email}{June 03--05,
+  2022}{Madrid, Spain}
+%
+%  Uncomment \acmBooktitle if the title of the proceedings is different
+%  from ``Proceedings of ...''!
+%
+\acmBooktitle{SIGIR '22: Proceedings of the 45th International ACM SIGIR Conference on Research and Development in Information Retrieval,
+ July 11--15, 2022, Madrid, Spain} 
+% \acmPrice{15.00}
+% \acmISBN{978-1-4503-XXXX-X/18/06}
+
+%
+% Submission ID.
+% Use this when submitting an article to a sponsored event. You'll
+% receive a unique submission ID from the organizers
+% of the event, and this ID should be used as the parameter to this command.
+% \acmSubmissionID{123-A56-BU3}
+
+%
+% The majority of ACM publications use numbered citations and
+% references.  The command \citestyle{authoryear} switches to the
+% "author year" style.
+%
+% If you are preparing content for an event
+% sponsored by ACM SIGGRAPH, you must use the "author year" style of
+% citations and references.
+% Uncommenting
+% the next command will enable that style.
+%\citestyle{acmauthoryear}
+
+%% The "author" command and its associated commands are used to define
+%% the authors and their affiliations.
+%% Of note is the shared affiliation of the first two authors, and the
+%% "authornote" and "authornotemark" commands
+%% used to denote shared contribution to the research.
+\author{Ben Trovato}
+\authornote{Both authors contributed equally to this research.}
+\email{trovato@corporation.com}
+\orcid{1234-5678-9012}
+\author{G.K.M. Tobin}
+\authornotemark[1]
+\email{webmaster@marysville-ohio.com}
+\affiliation{%
+  \institution{Institute for Clarity in Documentation}
+  \streetaddress{P.O. Box 1212}
+  \city{Dublin}
+  \state{Ohio}
+  \country{USA}
+  \postcode{43017-6221}
+}
+
+\author{Lars Th{\o}rv{\"a}ld}
+\affiliation{%
+  \institution{The Th{\o}rv{\"a}ld Group}
+  \streetaddress{1 Th{\o}rv{\"a}ld Circle}
+  \city{Hekla}
+  \country{Iceland}}
+\email{larst@affiliation.org}
+
+\author{Valerie B\'eranger}
+\affiliation{%
+  \institution{Inria Paris-Rocquencourt}
+  \city{Rocquencourt}
+  \country{France}
+}
+
+\author{Aparna Patel}
+\affiliation{%
+ \institution{Rajiv Gandhi University}
+ \streetaddress{Rono-Hills}
+ \city{Doimukh}
+ \state{Arunachal Pradesh}
+ \country{India}}
+
+\author{Huifen Chan}
+\affiliation{%
+  \institution{Tsinghua University}
+  \streetaddress{30 Shuangqing Rd}
+  \city{Haidian Qu}
+  \state{Beijing Shi}
+  \country{China}}
+
+\author{Charles Palmer}
+\affiliation{%
+  \institution{Palmer Research Laboratories}
+  \streetaddress{8600 Datapoint Drive}
+  \city{San Antonio}
+  \state{Texas}
+  \country{USA}
+  \postcode{78229}}
+\email{cpalmer@prl.com}
+
+\author{John Smith}
+\affiliation{%
+  \institution{The Th{\o}rv{\"a}ld Group}
+  \streetaddress{1 Th{\o}rv{\"a}ld Circle}
+  \city{Hekla}
+  \country{Iceland}}
+\email{jsmith@affiliation.org}
+
+\author{Julius P. Kumquat}
+\affiliation{%
+  \institution{The Kumquat Consortium}
+  \city{New York}
+  \country{USA}}
+\email{jpkumquat@consortium.net}
 
@@
 
-\usepackage{soul}
+% \usepackage{soul} %% Soul does not play well with acmart
 \usepackage{url}
-% \usepackage[hidelinks]{hyperref}
+% % \usepackage[hidelinks]{hyperref}https://www.overleaf.com/project/614b5cc61d38655dce466002
 % \usepackage[small]{caption}
 \usepackage{graphicx}
-\usepackage{amsmath}
 \usepackage{amsthm}
@@
 \usepackage{algorithmic}
+\usepackage{cleveref}
+\usepackage{lipsum}
+
 \urlstyle{same}
 
-% =======================================
-% my imported packages
+% % =======================================
+% % my imported packages
 
@@
 % =======================================
-
-\title{MMA: Multi-modal Manifold Alignment}
-
-\author{
-First Author$^1$
-\and
-Second Author$^2$\and
-Third Author$^{2,3}$\And
-Fourth Author$^4$
-\affiliations
-$^1$First Affiliation\\
-$^2$Second Affiliation\\
-$^3$Third Affiliation\\
-$^4$Fourth Affiliation
-\emails
-\{first, second\}@example.com,
-third@other.example.com,
-fourth@example.com
-}
+\title{Multimodal Contrastive Learning for Object Retrieval in the Face of Modality Failures}
 
@@
 
-Todo list
-
-\begin{enumerate}
-    \item Lit search 
-\end{enumerate}
-
-\newpage
-
-\maketitle
-\setcounter{page}{1}
+% Todo list
+% \begin{enumerate}
+%     \item \st{Lit search}
+% \end{enumerate}
+% \newpage
 
@@
 \begin{abstract}
+    % Grounded language understanding, in which natural language is used as a query against objects in a physical environment, allows a real-world, intuitive mechanism by which users can instruct physical agents to engage in tasks such as object retrieval.
+Visuolinguistic approaches to such object inference tasks typically involve training on large pools of image/text pairs and then using language to subselect elements of the sensed environment.
+However, physical agents such as robots typically have access to sensory and interactive modalities beyond vision, and learning from multiple modalities can improve performance on downstream tasks.
+In order to fully leverage multimodal training data while being robust to missing information, we propose a generalized distance-based loss function that can be extended to learn retrieval models that incorporate an arbitrary number of modalities.
+We demonstrate the usability of our model on a grounded language object retrieval scenario, where an intelligent agent has to select an object given an unconstrained language command.
+We leverage four modalities including vision, depth sensing, text, and speech, and we show that our model can outperform state-of-the-art contrastive models when modalities are ablated.
 
+    % shorter version
+
+    We propose a generalized distance-based loss function that can be used to learn retrieval models that incorporate an arbitrary number of views of a particular piece of data, and compounded by the challenge of retrieval when a modality becomes unavailable.
+    Our study is motivated by needs in robotics and computer-human interaction, where an agent has many sensors and thus modalities in which a human may interact: voice, text, images, and depth, both to communicate a desired goal and for the agent to recognize a desired target object.
+    Tying these various modalities together relates to the task of grounded language understanding, in which natural language is used as a query against objects in a physical environment, allowing a real-world, intuitive mechanism by which users can instruct physical agents to engage in tasks such as object retrieval.
+    However, current work has neglected the real-world constraint that sensors fail: hardware can become damaged or shipped with defects, sensors can get blocked or obstructed, and various adverse but not uncommon conditions that remove a modality from use.
+For practical applications, we need mechanisms for accurate object retrieval given that a modality may fail without notice.
+    We demonstrate the usability of our model on a grounded language object retrieval scenario, where an intelligent agent has to select an object given an unconstrained language command.
+We leverage four modalities including vision, depth sensing, text, and speech, and we show that our model can outperform state-of-the-art contrastive models when modalities are ablated.
 \end{abstract}
 
+\settopmatter{printfolios=true} % Page numbers please 
+% 9 pages without references.
+
+% \begin{enumerate}
+%     \item Objects of five things we select from
+%     \item \textcolor{green}{DONE} Adding an overview of the whole process 
+%     \item \textcolor{blue}{PROG} Go into more details in section 4 and give examples 
+%     \item Move images around
+%     \item Quant result section is not complete
+%     \item \textcolor{green}{DONE} put the graphs in modality ablation or merge quant and qual
+%     \item \textcolor{green}{DONE} add dropping R D graphs
+%     \item A paragraph in related work about grounded language 
+%     % \item Don't mess with additional experiments
+% \end{enumerate}
 %===================================================================
-\section{Lit Search}
 
-\subsection{Multimodal Machine Learning: A Survey and Taxonomy}
-
-\begin{itemize}
-\item \href{https://ieeexplore.ieee.org/document/8269806}{url}
-\end{itemize}
-
-\citet{baltrusaitisMultimodalMachineLearning2019} proposes a new taxonomy of multimodal machine learning by introducing five technical challenges in addition to typical early and late fusion categorization.
-This taxonomy includes the following categories.
-\begin{enumerate}
-\item \textit{Representation}: represent and summarize multimodal data in a way that exploits the complementarity and redundancy of multiple modalities.
-\begin{enumerate}
-\item joint: combine the unimodal signals into the same representation space.
-\item coordinated: process unimodal signals separately, but enforce certain similarity or structure constraints on them to bring them to a coordinated space.
-\end{enumerate}
-\item \textit{Translation}: map data from one modality to another.
-\begin{enumerate}
-\item example-based
-\item generative
-\end{enumerate}
-\item \textit{Alignment}: identify the direct relations between (sub)elements from two or more different modalities.
-\begin{enumerate}
-\item explicit
-\item implicit
-\end{enumerate}
-\item \textit{Fusion}: join information from two or more modalities to perform a prediction.
-\begin{enumerate}
-\item model-agnostic
-\item model-based
-\end{enumerate}
-\item \textit{Co-learning}: transfer knowledge between modalities, their representation, and their predictive models
-\begin{enumerate}
-\item parallel data
-\item non-parallel data
-\item hybrid data
-\end{enumerate}
-\end{enumerate}
-
-The authors mention that while joint representations have been used in situations to construct representations of more than two modalities, coordinated spaces have, so far, been mostly limited to two.
-This means that our research is novel and we can extend similarity measures to more than two modalities.
-
-\subsection{Self-Supervised MultiModal Versatile Networks}
-
-\begin{itemize}
-\item \href{https://proceedings.neurips.cc/paper/2020/file/0060ef47b12160b9198302ebdb144dcf-Paper.pdf?utm\_campaign=NLP\%20News\&utm\_medium=email\&utm\_source=Revue\%20newsletter}{url}
-\end{itemize}
-
-This work uses self-supervised contrastive learning to learn and combine representations from three modalities of visual, audio, and language.
-They learn a \textit{multimodal versatile network} that has the following four properties:
-\begin{enumerate}
-\item Ability to take any of the modalities as input
-\item Respecting the specificity of modalities: audio and visual modalities are fine-grained while language modality is coarse-grained.
-\item Comparability of different modalities using \textbf{dot product} even if not seen together during training 
-\item Efficiently applicable to visual data either in the form of dynamic videos or static images
-\end{enumerate}
-
-The authors consider the following three different configurations of the modality spaces which they call \textit{modality embedding graphs}.
-\begin{enumerate}
-\item Shared: all modalities map to the same space. This respects property 3 but violates property 2.
-\item Disjoint: visual-audio and visual-text spaces. This respects property 2 but violates property 3.
-\item Fine and coarse: audio and visual domains are map to the same space since they are fine-grained.
-These embeddings are then mapped to a lower dimensional space where text is also mapped.
-This respects both properties 2 and 3.
-\end{enumerate}
-
-The loss function they define consists of two terms.
-The first term is to train the fine-grained space of visual and audio embeddings by minimizing the distance between the similar pair (positives) from the same location of a video, and maximizing the distance between negative pairs (negatives) sampled from different videos.
-The second term is to train the coarse-grained space which consists of visual and text embeddings.
-Note that they don't consider audio embeddings in this space since they don't want to learn automatic speech recognition (ASR).
-Text and visual domains are not aligned as well as audio and visual domains (e.g.
-sound of playing piano versus the text that only says playing piano for a video of playing piano).
-To cure this, they consider a set of positive pairs instead of a single pair.
-If a modality is missed, the corresponding term is dropped from the loss function.
-
-\subsection{Separating Self-Expression and Visual Content in Hashtag Supervision}
-
-\begin{itemize}
-\item \href{https://openaccess.thecvf.com/content\_cvpr\_2018/papers/Veit\_Separating\_Self-Expression\_and\_CVPR\_2018\_paper.pdf}{url}
-\end{itemize}
-
-The authors train a joint model of images, hashtags, and users to perform image retrieval.
-This is similar to our task where given a description, we want to find the image in the scene that best matches the description.
-The idea is to form a three-way tensor product model.
-They use a ranking loss to train the model where the score of an observed triplet is higher than an unobserved triplet.
-They sample six negative triplets per positive sample triplet, and use each of them as a negative in the loss.
-The downstream retrieval task is then simply done by taking the arg max of the tensor product for a given user.
-
-\subsection{Semi-Heterogeneous Three-Way Joint Embedding Network for Sketch-Based Image Retrieval}
-
-\begin{itemize}
-\item \href{https://ieeexplore.ieee.org/document/8809264}{url}
-\end{itemize}
-
-\subsection{Deep Multimodal Learning for Affective Analysis and Retrieval}
-
-\begin{itemize}
-\item \href{https://ieeexplore.ieee.org/abstract/document/7277066?casa\_token=aBp6BxcszHwAAAAA:3NoMiFrZbn7tXfavF1rgkCiGWbFI2arxn8Xb6iDF79q4zBZHWi7PWWhf6xW-xJwYdFALbmRENo4}{url}
-\end{itemize}
-
-\subsection{An Efficient Framework for Zero-Shot Sketch-Based Image Retrieval}
-
-\begin{itemize}
-\item \href{https://arxiv.org/pdf/2102.04016.pdf}{url}
-\end{itemize}
-This paper uses two modalities only; image and sketch.
-However, it uses a \textit{quadruplet} to compute the loss which can be useful in our research.
-A quadruplet is composed of a sketch picture as an anchor, a negative example from sketch domain, a negative example from picture domain, and a positive example from picture domain.
-
-\subsection{SMIL: Multimodal Learning with Severely Missing Modality}
-
-\begin{itemize}
-\item \href{https://www.aaai.org/AAAI21Papers/AAAI-437.MaM.pdf}{url}
-\end{itemize}
-Uses Bayesian meta-learning framework to perform multimodal learning with partially missing modalities in training/test data.
-One of their experiments is multi-label classification of movie genres with bimodal data including poster of movies and description of the movie from IMDB.
-
-\subsection{Multimodal Language Analysis in the Wild: CMU-MOSEI Dataset and Interpretable Dynamic Fusion Graph}
-
-\begin{itemize}
-\item \href{https://aclanthology.org/P18-1208.pdf}{url}
-\end{itemize}
-
-\subsection{Multimodal Learning with Incomplete Modalities by Knowledge Distillation}
-
-\begin{itemize}
-\item \href{https://dl.acm.org/doi/pdf/10.1145/3394486.3403234}{url}
-\end{itemize}
-They first train models on each modality independently
-\subsection{SWAFN: Sentimental Words Aware Fusion Network for Multimodal Sentiment Analysis}
-
-\begin{itemize}
-\item \href{https://aclanthology.org/2020.coling-main.93.pdf}{url}
-\end{itemize}
-
-\subsection{Multimodal Learning for Human Action Recognition Via Bimodal/Multimodal Hybrid Centroid Canonical Correlation Analysis}
-
-\begin{itemize}
-\item \href{https://ieeexplore.ieee.org/document/8489981}{url}
-\end{itemize}
-
-\subsection{Deception Detection Using a Multimodal Approach}
-
-\begin{itemize}
-\item \href{https://dl.acm.org/doi/pdf/10.1145/2663204.2663229?casa\_token=KneK7B7xvLwAAAAA:Ajz0N96ygq8ktwkz0IVUqTt8NozCg2wR6n\_x2xntdHqZBh6VXW\_8VbO4GeY4VvMsDJlMwzkhVXQSJA}{url}
-\end{itemize}
-They use \textit{language}, \textit{physiological response}, and \textit{thermal sensing} to detect deceit.
-
-\subsection{M3ER: Multiplicative Multimodal Emotion Recognition using Facial, Textual, and Speech Cues}
-
-\begin{itemize}
-\item \href{https://ojs.aaai.org/index.php/AAAI/article/view/5492}{url}
-\end{itemize}
-
-\subsection{Found in Translation: Learning Robust Joint Representations by Cyclic Translations between Modalities}
-
-\begin{itemize}
-\item \href{https://ojs.aaai.org/index.php/AAAI/article/view/4666}{url}
-\end{itemize}
-\subsection{Select-Additive Learning: Improving Generalization In Multimodal Sentiment Analysis}
-
-\begin{itemize}
-\item \href{https://ieeexplore.ieee.org/stamp/stamp.jsp?tp=\&arnumber=8019301}{url}
-\end{itemize}
-
-\subsection{Attention-Based Multimodal Fusion for Video Description}
-
-\begin{itemize}
-\item \href{https://openaccess.thecvf.com/content\_ICCV\_2017/papers/Hori\_Attention-Based\_Multimodal\_Fusion\_ICCV\_2017\_paper.pdf}{url}
-\end{itemize}
-\subsection{\st{3W-AlignNet: a Feature Alignment Framework for Person Search with Three-Way Decision Theory}}
-
-\begin{itemize}
-\item \href{https://link.springer.com/article/10.1007/s12559-021-09898-7}{url}
-\end{itemize}
-
-This is not very related since they use three-way decision theory to select bounding boxes as positive, negative, and boundary.
-Three-way here refers to something else and not modalities.
-
-\subsection{\st{Unsupervised Domain Adaptation for Face Recognition in Unlabeled Videos}}
-\begin{itemize}
-\item \href{https://openaccess.thecvf.com/content\_ICCV\_2017/papers/Sohn\_Unsupervised\_Domain\_Adaptation\_ICCV\_2017\_paper.pdf}{url}
-\end{itemize}
-Might be relevant but not super relevant.
-
-\subsection{Heterogeneous Sensor Data Fusion By Deep Multimodal Encoding}
-\begin{itemize}
-\item \href{https://ieeexplore.ieee.org/stamp/stamp.jsp?tp=&arnumber=7874158}{url}
-\end{itemize}
-This one takes two modalities as input and outputs predictions in a third modality.
+\maketitle
+% \setcounter{page}{1}
 
 %===================================================================
+
+% \section{SIGIR Papers}
+% \begin{enumerate}
+%     \item \href{https://dl.acm.org/doi/pdf/10.1145/3397271.3401430}{FashionBERT}
+%     \item \href{https://dl.acm.org/doi/pdf/10.1145/3397271.3401265?casa_token=W8NqNWp0UQkAAAAA:w-iRUnNhmytDKCUiYzaUPG0bI_yyo8Bzxwro_eOC-zl5VDIGRVT2dvixvnCwLJ0iEKqYo1LkQIWXVA}{CrossBERT}
+
+% \end{enumerate}
+
 \section{Introduction}
@@
 
-% Grounded language learning~\cite{Mooney2008Grounded,RichardsDarvishMatuszekCategoryFree20,triplet_loss_2021_CVPR} is a cross-modal learning task to connect natural language with the physical environment in which the natural language is used.
+Humans use multiple sensory inputs to understand, interact with, and retrieve information from the world around them.
+It is intuitive that agents learning about the world, upon encountering new concepts and new objects, should form a model that incorporates information from all available sensors and data sources.
+% 
+Grounded language understanding, in which natural language is used as a query against objects in a physical environment, allows a real-world, intuitive mechanism by which users can instruct physical agents to engage in tasks such as object retrieval.
+Visuolinguistic approaches to such object inference tasks typically involve training on large pools of image/text pairs and then using language to subselect elements of the sensed environment~\cite{hong2021gilbert}.
+Although physical agents such as robots typically have access to sensory and interactive modalities beyond vision, and learning from multiple modalities can improve performance on downstream tasks, most approaches use at most two sensory inputs (e.g., visual data such as RGB images) with single labels, such as those provided by textual natural language.
+Simultaneously using additional inputs from different modalities is an underexplored area, in part due to the domain-specific nature of such \textit{n}-ary learning approaches.
+With the modern proliferation of audio and text based communication and proliferation of home agents (e.g., iRobot vacuums and Alexa/Google Home), there is a growing need to handle more modalities, and simultaneously their potential failures.
 
-% Our contributions are as follows:
-% \begin{enumerate}
-%     \item 
-%     \item 
-%     \item 
-% \end{enumerate}
+One difficulty with working with complex multimodal data is the increased likelihood that one or more modalities may have missing information.
+Current multimodal approaches are typically not robust to the loss of one or more modalities at test time, as may happen if, for example, a physical agent fails to retrieve data from a particular sensor.
+In order to fully leverage multimodal training data while being robust to missing information, we propose a generalized distance-based loss function that can be extended to learn retrieval models that incorporate an arbitrary number of modalities.
+
+We consider the domain of grounded language-based object retrieval~\cite{hu2016natural, triplet_loss_2021_CVPR}, in which objects in an environment must be identified based on linguistic instructions.
+This can be considered a special case of image retrieval~\cite{huang2017deep, ma2020large, novak2015large, vo2019composing} in which objects are identified using visual inputs in combination with other sensor modalities.
+Approaches to acquiring grounded language have explored various combinations of sensor inputs such as depth and RGB with labels provided by textual language or  speech~\cite{RichardsDarvishMatuszekCategoryFree20}.
+However, despite the multisensory nature of object retrieval, much of the existing work has not previously been extended to include an arbitrary number of modalities.
+
+The high level goal of this work is to take arbitrary input modalities about novel objects, including both spoken and written language, and build a model that allows a system to correctly retrieve that object given a subset of those input modalities.
+This is a generalization of approaches to grounded language learning in which specific input modalities are labeled with language to allow for future identification, but explicitly seeks to be agnostic about the nature of the individual sensory and linguistic inputs.
+Our contributions are as follows:
+\begin{enumerate}
+    \item Proposing a general distance-based loss function formulation for multimodal learning that can be extended to an arbitrary number of modalities.
+    \item Outperforming state-of-the-art contrastive learning models when one or more modalities are ablated at test time, demonstrating robustness in the face of missing information.
+    % \item Outperforming state-of-the-art models for grounded language learning by differentiating visual input modalities such as RGB and depth, rather than combining them into a single modality.
+    \item Demonstrating the separate utility of speech and text as sources of label information by treating them as sensory input modalities, rather than explicitly as labels.
+\end{enumerate}
+
+The remainder of this paper is organized as follows.
+In \cref{sec:Related-Work}, we describe existing work on visuolinguistic retrieval, contrastive learning, and multimodal learning.
+In \cref{sec:Problem-Description}, we formalize the object retrieval problem and give examples of the data under consideration.
+In \cref{sec:Method}, we describe our novel approach to learning models of linguistically-driven queries against a multimodal set of objects.
+Finally, in \cref{sec:Experiments}, we demonstrate the effectiveness of our approach against traditional contrastive learning and supervised contrastive learning baselines, both for the general learning problem and in the case of missing information at test time.
+
+% \todocmi{NOTES: Some things are probably less informative or more ambiguous (e.g., depth for differentiating round things).
+So instead of predicting a single vector output, predict a vector and a covariance that captures uncertainty; could potentially use that to do more accurate search, or to handle really uncertain modalities (e.g., just drop speech when it's uninformative).}
 
@@
 
+There is extensive work on the task of image retrieval, of which physical object retrieval can be considered a special case.
+In this task, language is used to formulate queries against datasets of images, for example in text-and-image matching tasks for fashion data~\cite{gao2020fashionbert, wen2021comprehensive}, sketch retrieval~\cite{huang2017deep}, and general photographs of objects~\cite{ma2020large, novak2015large, hong2021gilbert}.
+Prior work has focused solely on language and visual modalities, using text or a combination of language and vision to perform visual retrieval~\cite{vo2019composing}.
+While recent works have focused on grounding with respect to a knowledge set \cite{10.1145/3357384.3357889,10.1145/3397271.3401097}, our work focuses on robust multimodal learning over to perform grounding over an arbitrary number of input modalities.
+
+The growing number of datasets that contain modalities beyond images and text demonstrates the importance and the challenge of this task.
+\citet{bagher-zadeh-etal-2018-multimodal} introduce a large dataset of multimodal opinion sentiment and emotion intensity (CMU-MOSEI) for sentiment analysis and emotion recognition.
+They also introduce a novel multimodal fusion technique called the Dynamic Fusion Graph (DFG).~\citet{GoLD_UMBC} present a multimodal dataset of household objects containing RGB images, depth images, written language, and spoken language, which has been used to support learning grounded language directly from speech given small datasets~\cite{KebeAAAI2022}.
+\citet{baltrusaitisMultimodalMachineLearning2019} propose a new taxonomy of multimodal machine learning by introducing five technical challenges in addition to typical early and late fusion categorization.
+They cover different approaches to multimodal machine learning such as \textit{representation}, \textit{translation}, \textit{alignment}, \textit{fusion}, and \textit{co-learning}.
+In this project, we focus on alignment, the process of identifying the direct relationship between elements across modalities.
+
+We highlight a key difference with recent work like \citet{10.1145/3397271.3401232} and \citet{10.1145/3331184.3331213} who develop multi-modal retrieval models.
+These retrieval models are class based, where the model is trained to recognize any object of the same retrieval class as being ``same''.
+Our method and data is instance based, where different objects are not considered ``same'' even if they have the same class, due to the grounding goals of our work (i.e., the agent should identify the specified object, not equivalent objects).
+These works also do not consider the failure of a modality that is our target of interest.
+
+A number of different approaches have been proposed for linguistic-based alignment learning.
+\citet{alayrac2020self} use a self-supervised contrastive learning method to learn and combine representations from three modalities of visual, audio, and language.
+Audio and visual domains are mapped to the same space because they are fine-grained, containing dense information for each frame of video.
+These embeddings are then mapped to a coarse space where text is also mapped.
+Their network respects specificity and comparability of different modalities.
+The loss function they define consists of two distance-based terms for the corresponding spaces, differing from ours in that they do not handle arbitrarily many modalities and do not focus on robustness in the face of modality drop-outs.
+\citet{Nguyen-RSS-20} take a similar approach and perform pairwise cosine similarity to align images and natural language text descriptions to perform object retrieval tasks.
+\citet{triplet_loss_2021_CVPR} take a cross-modal manifold alignment approach to grounded language learning using triplet loss \cite{Chechik:2010:LSO:1756006.1756042}, and perform object retrieval by connecting text descriptions of objects to their corresponding RGB-D images.
+
+Another related approach is contrastive learning~\cite{qin2021world}, which is based on the idea of similarity and distance of elements of training data~\cite{NEURIPS2020_supervised_contrastive,chen2020simple}.
+Triplet loss is also a special case of contrastive loss~\cite{NEURIPS2020_supervised_contrastive}.
+Contrastive loss is mostly used in self-supervised learning~\cite{bui2021self, alayrac2020self,chen2020simple}, with the exception of \citet{NEURIPS2020_supervised_contrastive}, who use contrastive loss for a supervised problem.
+Most of these methods are implemented for a non-multimodal dataset (e.g., RGB images only).
+Our work is a specialization of contrastive loss that is focused on the multimodal input and- language problem.
+
+We note in particular that standard triplet learning approaches often require an expensive \textit{mining} step to find harder negative samples to enable effective learning \cite{Hoffer2015,Schroff2015,DBLP:conf/eccv/2018-9,DBLP:conf/eccv/ZhaoJQLH18,Zhai2018}.
+This is problematic from throughput and scalability, as run time becomes quadratic in batch size.
+Our approach does not require mining to obtain good performance, alleviating these issues along with their complicated history in reproducibility and parameter tuning \cite{Musgrave2020,Raff2020c,Raff2019_quantify_repro}.
+
+Some of the most closely related work focuses on learning based on more than two input modalities.
+\citet{Veit_2018_CVPR} train a joint model of images, hashtags, and users to perform image retrieval.
+Their task is similar to ours---given a description, we want to find the image/object in the scene that best matches the description.
+They tackle this problem by forming a three-way tensor product model.
+They use a ranking loss to train the model where the score of an observed triplet is higher than an unobserved triplet.
+They sample six negative triplets per positive sample triplet, and use each of them as a negative in the loss.
+The downstream retrieval task is then simply done by taking the argmax of the tensor product for a given user.
+This work differs from ours in that it becomes computationally complex as more and higher-dimensional modalities are added.
+
+\citet{semihet_three_way_Lei_2020} use inputs from three modalities of image, sketch, and edgemap to perform image retrieval given the sketch.
+They use co-attention mechanism, and their loss function contains alignment loss, cross-entropy loss, and sketch-edgemap contrastive loss.
+\citet{Mittal2020M3ER} use Canonical Correlational Analysis (CCA) to differentiate between ineffective and effective modalities for the emotion recognition task from multiple input modalities including face, speech, and text.
+Multimodal learning has been used for different applications including but not limited to object retrieval.
+\citet{Deception_ICMI_2014} use language, physiological response, and thermal sensing to detect deceit.
+\citet{het_data_fusion_liu_IEEE_2017} take two modalities as input and output predictions in a third modality.
+\citet{tursun2021efficient} uses quadruplet loss for two modalities only; image and sketch.
+A quadruplet is composed of a sketch picture as an anchor, a negative example from sketch domain, a negative example from picture domain,  and a positive example from picture domain.
+\citet{chen2017beyond} also uses quadruplet loss, but does not support multiple domains.
+In contrast, our method is not restricted to two modalities.
+% \todo[inline]{talk about CLIP and DALL-E here}
+
+To the best of our knowledge, there has been limited work on incorporating more than three modalities into a general learning approach.
+Our proposed loss function closely resembles work on quadruplet loss~\cite{chen2017beyond,tursun2021efficient} but is intended to scale to an arbitrary number of modalities.
+
 %===================================================================
@@
 \label{sec:Problem-Description}
+
+Given a language command, for example either text or speech, that describes an object, we want our model to retrieve the correct object from a set of objects.
+This problem is an exemplar of tasks found in the area of grounded language learning in the fields of robotics and natural  language processing.
+Intuitively, the goal is to take unconstrained natural language queries and select the appropriate object based on the complete set of sensor inputs available to the agent.
+We demonstrate on a domain containing four modalities, all of which refer to objects in the environment: spoken language, written text, RGB (image) inputs, and depth camera inputs.
+\Cref{fig:experimental-setup} illustrates a small visualization of our object retrieval task: the spoken query ``A white textbook titled algorithms'' is provided to our contrastive model, which identifies the item (outlined in red, and represented by up to four different modalities) as the most likely item the query is referring to.
+
+\begin{figure}[tb]
+\centering
+\includegraphics[width=.99\columnwidth]{Figures/experiment-setup.png}
+\caption{
+The experimental object retrieval setup, in which objects are represented by four modalities: an RGB image, a depth image, spoken descriptions, and textual descriptions.
+Given a query in some modality, our approach seeks to select the object that is the best fit, per a trained model.
+This approach, detailed in \cref{sec:Method}, is able to rank objects as to appropriateness even when one or more modalities is ablated at test time (e.g., textual descriptions are missing), and outperforms state-of-the-art contrastive learning approaches on this task.
+}
+\label{fig:experimental-setup}
+\end{figure}
+
+More formally, we consider a spoken language command $x_s$, a textual language command $x_t$, a set of RGB images $X_r = \{x_r^{(1..n)}\}$, and a set of depth images $X_d = \{x_d^{(1..n)}\}$, the task is to retrieve the correct object by choosing the index that has the minimum distance from either of the language commands across all modalities.
+Depending on which modalities are or are not ablated, we consider up-to four distances: ``sr,'' a vector of distances between $x_s$ and all RGB images in $X_r$; ``sd,'' a vector of distances between $x_s$ and all depth images in $X_d$; ``tr,'' a vector of distances between $x_t$ and all RGB images in $X_r$; and ``td,'' a vector of distances between $x_t$ and all depth images in $X_d$.
+In order to select the correct object, we first perform a component-wise average of the relevant modality pair distances for the available modalities.
+Then, we select the object which had the minimum distance, i.e., we perform an argmin on this average vector of multiple-modality distances.
+Depending on the missing/available sensors during test time, we might have any combination of these four distances.
+For example, if no written instructions are available at test time,\footnote{This setting is particularly salient.
+While large bodies of text are frequently available at training time, a person interacting directly with a physical agent may well prefer to use spoken instructions.} we compute the component-wise average of $sr$ and $sd$, and then select the object whose coordinate resulted in the lowest average distance.
+This method allows us to extend our model to support arbitrary modalities while remaining robust to test cases in which some modalities are missing or incomplete.
 
@@
 
-For $M$ modalities, we define a distance based loss function that contains $M+2(M-1)!$ terms. 
-We sample two data points from each modality, one positive point and one negative point.
-The objective is then to (I) minimize the distance between each pair of positive points from heterogeneous modalities which results in $(M-1)!$ terms, (II) maximize the distance between each pair of positive and negative points from heterogeneous modalities which contains $(M-1)!$ terms again, and (III) maximize the distance between each pair of positive and negative points from the homogeneous modalities which contains $M$ terms.
-Equation~\ref{eq:objective} shows the objective function we defined above with changing maximization to minimization by changing the sign.
+In keeping with previous work on the closely related problem of image retrieval, we focus on contrastive loss approaches, in which the goal is to learn an embedding of data in which similar samples---in our case, samples belonging to the same class of object---are embedded `close' to one another, while dissimilar samples are farther apart.
+We develop a novel loss function that simultaneously minimizes intra-class distances and maximizes inter-class distances across each pair of modalities, yielding a model that is both effective at the retrieval task defined above and robust to modality dropouts at test time.
 
-% \begin{equation}\label{eq:objective-detailed}
-%     \min \sum_{i=1}^{M-1} \sum_{j=i+1}^{M} ||x_{i}^{+} - x_{j}^{+}|| - \sum_{i=1}^{M-1} \sum_{j=i+1}^{N} ||x_{i}^{+} - x_{j}^{-}|| - 
-%     \sum_{i=1}^{M} ||x_{i}^{+} - x_{i}^{-}||
-% \end{equation}
+\subsection{Distance-based Loss}
+\label{ssec:distanceloss}
 
-\begin{aligned}\label{eq:objective}
-    \mathcal{L}  &= \sum_{i=1}^{M-1} \sum_{j=i+1}^{M} ||f_i(x_{i}^{+}) - f_j(x_{j}^{+})||
-    - ||f_i(x_{i}^{+}) - f_j(x_{j}^{-})|| \\ 
-    & - ||f_i(x_{i}^{-}) - f_j(x_{j}^{+})|| - \sum_{i=1}^{M} ||f_i(x_{i}^{+}) - f_i(x_{i}^{-}) ||
-\end{aligned}
-where the subscripts $i$ and $j$ represent modalities, and $f_i(.)$ and $f_j(.)$ represent functions that map input data to the shared latent space using a deep neural network.
+For $M$ modalities, we define a distance-based loss function which can be used for arbitrary number of modalities.
+Our proposed method is an extension of the well-known similarity-based triplet loss~\cite{Carvalho-cooking-triplet,triplet_loss_2021_CVPR}, and is similar to contrastive loss~\cite{chen2020simple,NEURIPS2020_supervised_contrastive} under some settings.
+Triplet loss-based learning works by forcing similar concepts from different domains `together' in some shared embedding space, while forcing dissimilar concepts `apart.' It is so named because it relies on three data points from the training set: a positive, a negative, and an anchor point.
+In our setting, the anchor and positive instances are from the same class, but in different modalities.
 
-\todokdinline{I can add another term for minimizing the distance between similar pairs (anchor and positive) from the same domain.
-This would look like the equation below.
-However, it breaks the nice procedure of sampling two data points per modality}
-$+ \sum_{i=1}^{M} ||f_i(x_i^{a}) - f_i(x_i^{+})||$
+However, standard triplet loss cannot be used for more than two modalities.
+Therefore, we extend this concept as follows.
+During training, we sample two sets of data points from each modality---one positive set (referring to a specific object) and one negative set (referring to some different object).
+Unlike some prior triplet loss methods~\cite{GoLD_UMBC,triplet_loss_2021_CVPR}, the anchor is not randomly chosen from different modalities in each batch.
+Instead, we choose one \textit{domain} as an anchor (in our experiments, language).
+We consider the following formulation:
 
-Our method is a generalization of well-known similarity-based metric learning and manifold alignment methods such as triplet loss~\CitationNeeded{}.
-If $M=2$ which means the number of modalities is 2, then our objective function reduces to quadruplet loss method.
-If $M=2$ and we ignore the last term in the objective function, it results in the triplet loss method. 
-If $M=1$, only the last term remains in the loss function which is exactly the pairwise distance-based loss function.
+\begin{itemize}
+    % \item data point: a row in the csv/tsv file.
+    \item Positive (Instance): A set of embeddings of one data point (e.g., an RGB image of an apple, corresponding depth image, text description, and speech description of the same apple)
+    \item Negative (Instance): A set of embeddings of another data point of a different object (e.g., an RGB image of a book, corresponding depth image, text description, and speech description of the same book)
+    \item Anchor (Modality): A fixed modality of the positive set is chosen as the learning anchor.
+For our experiments, text is used.
+The anchor is used as the basis for learning distances between positive and negative samples.
+    % variance of text is lower and clusters of different objects are more distinct compared to other modalities, therefore, it can train a better model.
+    % where we don't minimize the distance between the pair of embeddings from all modalities of the same data point, and we don't maximize the distance between positive and negative datapoints of the same modality.
+    % \todo{explain this better}
+    % \todokd{I rephrased it. Good?}
+\end{itemize}
+
+The objective is then to first minimize the distance between the anchor and each of the positive points from heterogeneous modalities; second, maximize the distance between the anchor and each of the negative points from heterogeneous modalities; and finally, maximize the distance between positive and negative anchor points from the same modality.
+Altogether, our proposed loss function contains $2M-1$ terms: $(M-1)$ anchor-to-positive distance minimizations, $(M-1)$ anchor-to-negative distance maximizations, and 1 positive-to-negative distance maximization.
+\Cref{eq:objective-emma} shows this objective function.
+
+To measure distance in embedded space, we use cosine similarity between pairs of embeddings, i.e.
+we measure the cosine of the angle between embeddings.
+Cosine similarity is a good choice for high-dimensional data as it is bounded between -1 and 1.
+Other distances, such as Euclidean, grow in value with respect to their dimensionality, resulting in a very large distances for data points.
+Cosine similarity is the opposite of distance, and we need to reverse the logic for maximization and minimization.
+
+\subsection{EMMA: Extended Multimodal Alignment}
+\label{sub:emma}
+Our proposed loss function is a modification of the triplet loss function idea where we fix one modality as anchor and do not sample a positive instance from the same modality. 
+Na\"ively, to extend triplet loss to work with arbitrary number of modalities, we can use two anchors (e.g., text and speech), which results in $2(M-2)$ triplet losses as formulated in~\cref{eq:objective-two-anchors}.
+\begin{equation}
+\label{eq:objective-two-anchors}
+\begin{split}
+    \mathcal{L}  &= \sum_{m=1}^{M-2} || z_{t}^{+} - z_{m}^{+} || - || z_{t}^{+} - z_{m}^{-} || + || z_{s}^{+} - z_{m}^{+} || - || z_{s}^{+} - z_{m}^{-} || \\
+    &= \sum_{m=1}^{M-2} \cos(z_{t}^{+} ,z_{m}^{-}) - \cos(z_{t}^{+}, z_{m}^{+}) + \cos(z_{s}^{+} ,z_{m}^{-}) - \cos(z_{s}^{+}, z_{m}^{+})
+\end{split}
+\end{equation}
+However, this approach performs poorly when one of the anchor modalities is ablated; since there is no explicit minimization between anchors, they do not necessarily map closely to each other, such that only one of the two anchors is actually learned.
+
+An alternative is to apply triplet loss $M-1$ times with a single anchor and $M-1$ `target' modalities.
+In this case, the negative anchor point is disregarded.
+The positive anchor is used as an anchor for all $M-1$ triplet losses, and for each of those triplet losses the positive and negative points are simply selected from the corresponding modalities.
+\Cref{eq:objective-simple-mma} formulates this idea.
+\begin{equation}
+\label{eq:objective-simple-mma}
+\begin{split}
+    \mathcal{L}  &= \sum_{m=1}^{M-1} || z_{a}^{+} - z_{m}^{+} || - || z_{a}^{+} - z_{m}^{-} || \\
+    &= \sum_{m=1}^{M-1} \cos(z_{a}^{+} ,z_{m}^{-}) - \cos(z_{a}^{+}, z_{m}^{+})
+\end{split}
+\end{equation}
+However, this approach fails to take into account the dissimilarity of the positive and negative anchor points.
+This is an important piece of information that \cref{eq:objective-simple-mma} ignores.
+Although this information is implicitly captured when we maximize the distance between positive anchor and other negative modalities, when the negative data point becomes a positive example itself, the anchor and those points are forced to be closer to each other which can happen in three ways: either moving anchor closer to those points, moving those points closer to anchor, or move both somewhere in between.
+Therefore, in the last two cases we do not have explicit direct control over the distance between these points and previous positive points.
+Our experimental results confirm this hypothesis.
+
+Therefore, we add an extra term which is responsible for maximizing the distance between positive and negative instances of the anchor modality, because a margin between positive and negative points from other modalities (excluding anchor) is enforced.
+The first dashed line from the top in \cref{fig:emma-loss} is this extra term that captures the dissimilarity between positive and negative text datapoints.
+We formulate this method in \cref{eq:objective-emma}, which is generalized to an arbitrary number of modalities. We refer to this approach as EMMA, for extended multimodal alignment.
+\begin{equation}
+\label{eq:objective-emma}
+\begin{split}
+    \mathcal{L}  %&= %\sum_{m=1}^{M-1} || z_{a}^{+} - z_{m}^{+} || - || z_{a}^{+} - z_{m}^{-} ||  - || z_{a}^{+} - z_{a}^{-} || \\
+    &= \cos(z_{a}^{+}, z_{a}^{-}) + \sum_{m=1}^{M-1} \left(\cos(z_{a}^{+} ,z_{m}^{-}) - \cos(z_{a}^{+}, z_{m}^{+}) + \alpha_m\right) 
+\end{split}
+\end{equation}
+In \cref{eq:objective-emma}, $a$ represents the \textit{anchor} modality, $M$ is the number of modalities, the superscripts $+$ and $-$ represent positive and negative objects, $\alpha_m$ represents enforced margin for each modality which we set to 0.4 for all modalities without tuning, and $z$ is the embedding we get by applying a mapping function $f$, which in our case is a neural network on our input data.
+In other words, $z_m = f_m(x_m)$, where each modality $m$ has a specific model $f_m$ that is different from the models for other modalities.
+These models do not share their weights.
+The $\cos(\cdot)$ function is a measure of similarity, not distance, and that is why the signs are reversed in the second line.
+An illustration of this loss function is provided in \cref{fig:emma-loss}.
+
+Since the downstream task in the grounded language learning domain is to predict/retrieve a desired object among multiple objects given a natural language description, it makes sense to train the model using language as an anchor, and in fact anchoring learning around language outperforms anchoring on RGB.
+Triplet loss cannot be used for more than 2 modalities.
+Some previous work has concatenated RGB and depth embeddings to create a single ``vision'' embedding for learning~\cite{triplet_loss_2021_CVPR}, but they cannot handle RGB or depth sensor ablation during test.
+Since our method can handle any number of modalities, we can handle such a case by treating depth as a separate modality.
+% Moreover, when using triplet loss with different anchors per batch, the batch size has to be 1 which makes training very slow.
+This approach is most similar to that of supervised contrastive learning~\cite{NEURIPS2020_supervised_contrastive}, but outperforms both that and regular contrastive learning~\cite{chen2020simple} methods when modalities are ablated, most notably written language.
+
+It is noteworthy that our training procedure does not perform any stochastic dropout of modalities to obtain test-time robustness to missing modalities.
+We find that the use of instance level losses grounded to one key modality (text) is sufficient to encourage a representation that satisfies our goals.
+
+\begin{figure*}[tb]
+\centering
+% \includegraphics[width=.9\columnwidth]{Figures/4way-MMA.png}
+\includegraphics[width=2.1\columnwidth]{Figures/e-MMA.png}
+\caption{A high-level prototype of our approach and the distances used in the MMA loss with four modalities.
+Gray arrows indicate pre-trained models that are frozen (i.e.
+the parameters are fixed and are not trained).
+The black arrows show 3 fully connected layers with a ReLU activitation function~\cite{relu2010} after the first two layers.
+These networks are trained.
+Orange vectors are negative data points, green vectors are positive data points, and the blue vector is the anchor (which is the text modality of the positive data points).
+Dashed lines indicate distances to be maximized, dotted lines indicate distances to be minimized, and solid lines show the enforced margin.
+}
+\label{fig:emma-loss}
+\end{figure*}
 
@@
 \label{sec:Model}
+
+Transformers are de facto architectures in natural langauge processing community and have shown great success across different tasks.
+Similar to~\citet{GoLD_UMBC}, we use BERT~\cite{devlin-etal-2019-bert} embeddings contained in the FLAIR library ~\cite{akbik2019flair,akbik-etal-2019-pooled} to featurize the textual input, and wav2vec2~\cite{wav2vec2} to extract audio embeddings from speech.
+Both of these encoders output a 3072-dimensional embedding vector which is generated by concatenating the last four hidden layers of their corresponding networks.
+Both BERT~\cite{devlin-etal-2019-bert} and wav2vec2~\cite{wav2vec2} are self-supervised language models using transformers~\cite{vaswani2017attention}.
+% \todo[inline]{A couple of sentences about FLAIR and wav2vec2 and what they do}
+
+To process images, we use ResNet152~\cite{He_resnet_2016} for both RGB and depth images which gives us a 2048-dimensional embedding vector. Depth images are colorized before passing to the ResNet152.
+We then use 3 fully connected layers with ReLU activation~\cite{relu2010} to map these embeddings to a shared 1024-dimensional space where we can compute the distances between all embeddings.
 
@@
 
+In this section we evaluate the quality of object retrieval models learned using the EMMA loss function.
+We first describe the dataset we use, give the metrics by which we evaluate the performance, the setup of the experiments, the baselines we compare against, and we end the section by a presenting results and analyzing them.
+
 \subsection{Data}
@@
 
-\subsection{Setup}
+We demonstrate the effectiveness of our approach on a recent multimodal dataset called GoLD~\cite{GoLD_UMBC}, which contains RGB image, depth image, written text descriptions, speech descriptions, and transcribed speech descriptions for 207 object instances across 47 object classes (see \cref{fig:emma-loss}).
+There are a total of 16,500 speech and 16,500 textual descriptions.
+The original GoLD paper uses raw RGB and raw depth images in which other objects are present in the background.
+We use a masked version of the images where the background is deleted (we observe that a masked version converges faster, however, they both converge to the same performance).
+Speech is converted to 16 Hz to match the wav2vec2 speech model~\cite{wav2vec2}.
+
+\subsection{Problem Setup and Metrics}
+\label{sec:metrics}
+To evaluate our model we measure different performance metrics on a retrieval task where the model has to select an object among 5 objects given a language description.
+Only one of the objects corresponds to the description and the rest are from different object classes (e.g., one apple among a set including a fork, a mug, a lemon, and a bell pepper).
+
+\subsubsection{Setup}
 \label{sec:setup}
+Similar to~\citet{NEURIPS2020_supervised_contrastive}, we use a stochastic gradient descent (SGD) optimizer with momentum~\cite{ruder2016overviewSGD} with a flexible learning rate starting at 0.05. 
+For the triplet loss experiments we use the Adam~\cite{kingma_adam_2015} optimizer with scheduling learning rate as suggested by the papers.
+The models are trained for 200 epochs with a batch size of 64 on a Quadro RTX 8000 GPU.
 
-\subsection{Quantitative Results}
-\label{sec:Quantitative}
+To evaluate the performance, we compute the distance between the given natural language description and 5 randomly selected images (1 of which corresponds to the description, with the others from different object classes).
+We compute the distance between one textual language embedding and all candidate RGB embeddings, and we compute the distance between the same language embedding and all candidate depth embeddings corresponding to the RGB embeddings.
+We then take average of these two distance matrices.
+Instead of choosing an empirical threshold beyond which objects are considered to be `referred to,' we choose the closest image embedding (average distance of RGB and depth from language) as the prediction.
+In order to use cosine \textit{distance}, we have to subtract the cosine of the \textit{angle} between two embeddings (which represents similarity) from 1: that is, we compute $1 - \cos(e_1, e_2)$.
 
-\subsection{Qualitative Results}
-\label{sec:Qualitative}
+\subsubsection{Metrics}
+
+The best metric to capture the performance in such a scenario is mean reciprocal rank (MRR, \cref{eq:mrr} for $Q$ queries).
+For each query we predict the rank of all objects based on their distance from the language command, and then the inverse rank of the desired objects in all queries are averaged.
+For example, if the model predicts the desired object as the first rank, then MRR = $\frac{1}{1} = 1$ which means a perfect score, and if it predicts the correct object as the last rank among five objects, then MRR = $\frac{1}{5}=0.2$.
+
+\begin{equation} \label{eq:mrr}
+    \mathrm{MRR}=\frac{1}{|Q|} \sum_{i=1}^{|Q|} \frac{1}{\operatorname{rank}_{i}}
+\end{equation}
+
+Accuracy and micro F1 score
+% (sklearn), flattened binary f1 score (sklearn), and F1 score 
+are the same in this task, since for each prediction we either have a true positive and no false positives and no false negatives, or we have no true positives, one false positive and one false negative.
+MRR is a more informative metric because it captures the idea that having the correct object as the second choice should be considered better than having it as a last choice, while in accuracy the score is ``all or nothing''---either 0 or 1.
+Because our approach is designed to be robust to missing information across modalities, we also report MRR for different combinations of modality dropouts.
 
 \subsection{Baselines}
+We compare our EMMA model against different contrastive learning methods including triplet loss, the traditional version of contrastive loss usually used in self-supervised settings~\cite{chen2020simple}, and supervised contrastive learning~\cite{NEURIPS2020_supervised_contrastive}.
+
+\begin{table*}[]
+\centering
+\begin{tabular}{lc|ll|lll}
+\multicolumn{1}{c}{} &  & \multicolumn{2}{c|}{\textbf{Accuracy}} & \multicolumn{3}{c}{\textbf{Mean Reciprocal Rank}} \\ \hline
+\multicolumn{1}{c|}{\textbf{Method}} & \textbf{Modalities} & \multicolumn{1}{c|}{\textbf{\begin{tabular}[c]{@{}c@{}}Speech/RGB/\\ Depth\end{tabular}}} & \multicolumn{1}{c|}{\textbf{\begin{tabular}[c]{@{}c@{}}Text/RGB/\\ Depth\end{tabular}}} & \multicolumn{1}{c|}{\textbf{\begin{tabular}[c]{@{}c@{}}Speech/RGB/\\ Depth\end{tabular}}} & \multicolumn{1}{c|}{\textbf{\begin{tabular}[c]{@{}c@{}}Text/RGB/\\ Depth\end{tabular}}} & \multicolumn{1}{c}{\textbf{\begin{tabular}[c]{@{}c@{}}Text/Speech/\\ RGB/Depth\end{tabular}}} \\ \hline
+\multicolumn{1}{l|}{EMMA} & 4 & \multicolumn{1}{l|}{\textbf{0.65}$\pm$0.01} & 0.83$\pm$0.006 & \multicolumn{1}{l|}{\textbf{0.79}$\pm$0.002} & \multicolumn{1}{l|}{0.90$\pm$0.003} & 0.90$\pm$0.004 \\
+\multicolumn{1}{l|}{EMMA} & 3 & \multicolumn{1}{l|}{--} & 0.84$\pm$0.008 & \multicolumn{1}{l|}{--} & \multicolumn{1}{l|}{0.90$\pm$0.004} & -- \\
+\multicolumn{1}{l|}{Supervised Contrastive} & 4 & \multicolumn{1}{l|}{0.59$\pm$0.007} & 0.84$\pm$0.008 & \multicolumn{1}{l|}{0.75$\pm$0.004} & \multicolumn{1}{l|}{0.91$\pm$0.006} & 0.90$\pm$0.006 \\
+\multicolumn{1}{l|}{Supervised Contrastive} & 3 & \multicolumn{1}{l|}{--} & 0.84$\pm$0.006 & \multicolumn{1}{l|}{--} & \multicolumn{1}{l|}{0.91$\pm$0.004} & -- \\
+\multicolumn{1}{l|}{Contrastive Cosine} & 4 & \multicolumn{1}{l|}{0.40$\pm$0.153} & 0.83$\pm$0.006 & \multicolumn{1}{l|}{0.62$\pm$0.034} & \multicolumn{1}{l|}{0.90$\pm$0.004} & 0.89$\pm$0.005 \\
+\multicolumn{1}{l|}{Contrastive Cosine} & 3 & \multicolumn{1}{l|}{--} & 0.84$\pm$0.005 & \multicolumn{1}{l|}{--} & \multicolumn{1}{l|}{0.91$\pm$0.003} & -- \\
+\multicolumn{1}{l|}{Triplet} & 3 & \multicolumn{1}{l|}{--} & 0.83$\pm$0.008 & \multicolumn{1}{l|}{--} & \multicolumn{1}{l|}{0.90$\pm$0.005} & -- \\ \hline
+\end{tabular}
+\caption{Average and standard deviation of accuracy and MRR over 3 runs with 3 different random seeds on a held-out test set.
+These results represent performance when no modalities are ablated, in which case supervised contrastive loss, contrastive loss, and EMMA perform similarly (``--'' is an ablated modality and can't be evaluated).
+All metrics are from 0 to 1, and higher is better.
+For 5 objects, a random guess would have MRR of 0.33, and the worst case performance would be always ranking the correct item last which gives us 0.2.
+All models are trained using text as anchor.
+The batch size is 64 and optimizer is SGD for all experiments except the triplet loss, where batch size is 1 and optimizer is Adam.}
+\label{table:quantitative}
+\end{table*}
+% %Edward moved table above where refernced b/c LaTeX plays games with double-wide objects
+% \begin{table*}[bth]
+% \centering
+% \begin{tabular}{c|c|c|c|c|c|c}
+% \toprule
+% \textbf{Method} & \textbf{Modalities} & \textbf{acc.srd} & \textbf{acc.trd} & \textbf{MRR.srd} & \textbf{MRR.trd} & \textbf{MRR.tsrd} \\ %\hline
+% \midrule
+% EMMA (Text) & 4 & \textbf{0.65}$\pm$0.01 & 0.83$\pm$0.006 & \textbf{0.79}$\pm$0.002 & 0.90$\pm$0.003 & 0.90$\pm$0.004 \\
+% EMMA (Text) & 3 & - & 0.84$\pm$0.008 & - & 0.90$\pm$0.004 & - \\
+% % Simple MMA (Text)  & 3 & ?$\pm$? & ?$\pm$? & ? & ? \\
+% % Simple MMA (Text) & 4 & ?$\pm$? & ?$\pm$? & ? & ? \\
+% % Simple MMA (Speech) & 3 & ?$\pm$? & ?$\pm$? & ? & ?  \\
+% DONE Supervised Contrastive (Text)& 4 & 0.59$\pm$0.007 & 0.84$\pm$0.008 & 0.75$\pm$0.004 & 0.91$\pm$0.006 & 0.90$\pm$0.006  \\
+% TODO Supervised Contrastive (Text)& 3 & - & 0.84$\pm$0.006 & - & 0.91$\pm$0.004 & -\\
+% DONE Contrastive Cosine (Text) & 4 & 0.40$\pm$0.153 & 0.83$\pm$0.006 & 0.62$\pm$0.034  & 0.90$\pm$0.004 & 0.89$\pm$0.005\\
+% TODO Contrastive Cosine (Text) & 3 & - & 0.84$\pm$0.005 & -  & 0.91$\pm$0.003 & -\\
+% PROJ Triplet (Text) & 3 & - & 0.83$\pm$0.008 & - & 0.90$\pm$0.005 & -  \\
+% % DONE EMMA (Speech) & 3 & ?$\pm$? & ?$\pm$? & - & - \\
+% % DONE EMMA (Speech) & 4 & ?$\pm$? & ?$\pm$? & ? & ? \\
+% % DONE Supervised Contrastive (Speech)& 3 & ?$\pm$? & ?$\pm$? & - & -\\
+% % DONE Supervised Contrastive (Speech)& 4 & ?$\pm$? & ?$\pm$? & ? & ? \\
+% % TODO Contrastive Cosine (Speech) & 3 & ?$\pm$? & ?$\pm$?  & - & -\\
+% % TODO Contrastive Cosine (Speech) & 4 & ?$\pm$? & ?$\pm$?  & ? & ?\\
+% % DONE Triplet (Speech) & 3 & ?$\pm$? & ?$\pm$? & - & -  \\
+
+% % Full MMA (Text) w/ neg & 3 & ?$\pm$? & ?$\pm$? \\
+% % Simple MMA (Text) w/ neg Adam & 3  & 0.7949$\pm$? & 0.8778$\pm$? \\
+% % Simple MMA (Text) w/ neg SGD & 3 & 0.7905$\pm$? & 0.8821$\pm$? \\
+% % Triplet (Text) w/ neg & 3 & 0.7497$\pm$? & 0.8522$\pm$? \\
+% % Full MMA (Speech) w/ neg & 3 & ?$\pm$? & ?$\pm$? \\
+% % Simple MMA (Speech) w/ neg & 3 & 0.7949$\pm$? & 0.8778$\pm$? \\
+% % Contrastive Cosine (Text) w/ neg & 3 & 0.7894$\pm$? & 0.8788$\pm$?  & ? & ?\\
+% % Contrastive (Speech) w/ neg & 3 & ?$\pm$? & ?$\pm$? \\
+% % Triplet (Speech) w/ neg & 3 & 0.7497$\pm$? & 0.8522$\pm$? \\
+% \bottomrule
+% \end{tabular}
+% \caption{\label{table:quantitative}Average and standard deviation of accuracy and MRR over 3 runs with 3 different random seeds on a held-out test set.
+These results represent performance when no modalities are ablated, in which case supervised contrastive loss, contrastive loss, and EMMA perform similarly.
+All metrics are from 0 to 1, and higher is better.
+For 5 objects, a random guess would have MRR of 0.33, and the worst case performance would be always ranking the correct item last which gives us 0.2.
+All models are trained using text as anchor.
+The batch size is 64 and optimizer is SGD for all experiments except the triplet loss, where batch size is 1 and optimizer is Adam.
+% % \todokdinline{maybe adding results on cropped and raw dataset as well. Also results using object class instead of negative sampling}
+% }
+% \end{table*}
+
+\subsubsection{Triplet Loss}
+\label{sub:baseline-triplet}
+The triplet loss function consists of three data points including anchor, positive, and negative from two modalities.
+The anchor, positive, and negative can be chosen from different modalities in each batch.
+The triplet loss method has two major disadvantages.
+First, it cannot be used for more than two modalities.
+Second, a batch size of greater than 1 is hard to implement if the anchor, positive, and negative come from different modalities for each batch.
+% To address these issues in EMMA, we sample more than one positive and negative data points.
+To address these issues in EMMA, we choose one modality as anchor.
+Intuitively, we sample one positive data point and one negative data point from each modality, resulting in 2M data points.
+The anchor data point is always from the same modality and corresponds to the same object class sampled for positive data points.
+\Cref{table:quantitative} shows that our MMA method perform as good as this baseline while out model can be trained 10x faster since we can use batch sizes larger than 1.
+Also, our method can be used for any number of modalities, while previous works~\cite{GoLD_UMBC,triplet_loss_2021_CVPR} concatenate rgb and depth to form a single vector to handle more than two modalites.
+When we used a fixed anchor modality, a batch size of 64, and applied semantic negative sampling~\cite{Pillai_Matuszek_2018}, our method outperformed this baseline.
+However, since this method cannot be applied to four modalities, we do not include it in the main analysis.
+Moreover, our method has the advantage that does not require triplet mining or providing hard examples.
+
+\subsubsection{Contrastive Loss}
+\label{sub:baseline-contrastive}
+
+We compare our model against contrastive loss.
+In order to implement this loss function, we use cosine similarity as suggested in the SimCLR paper~\cite{chen2020simple}.
+Another possibility is to use an inner dot product~\cite{NEURIPS2020_supervised_contrastive}; this can lead to instabilities since the dot product is not bounded.
+This contrastive loss function is formulated in \cref{eq:contrastive-loss}:
+% \todokdinline{try contrastive loss with cross-entropy, dot product, and the supervised contrastive method itself (without triplets).}
+
+% Original Formula
+\begin{equation}\label{eq:contrastive-loss}
+    -\sum_{i \in I} \log \frac{\exp (sim(z_i , z_{j(i)}) / \tau) }{\sum_{a \in A(i)} \exp (sim(z_i, z_a) / \tau)}
+\end{equation}
+% \begin{equation}\label{eq:contrastive-loss}
+%     -\sum_{i \in I} \log \frac{\exp (sim(f(x_i) , z_{j(i)}) / \tau) }{\sum_{a \in A(i)} \exp (sim(f(x_i), f(x_a)) / \tau)}
+% \end{equation}
+where $i$ is anchor, $j(i)$ is the set of positives excluding anchor, $a$ is the set of all positives and negatives excluding anchor, and $z = f(x)$.
+
+\Cref{eq:contrastive-loss} can be extended to the loss function formulated in \cref{eq:mma-contrastive}.
+Our formulation is a generalized version of the contrastive loss where $M$ is the number of modalities (e.g.
+RGB image, depth image, speech, text), but it can be ``augmented views'' instead of different modalities.
+
+\begin{equation}\label{eq:mma-contrastive}
+    -\sum_{m=1}^{M-1} \log \frac{\exp (sim(z_a^+ , z_{m}^{+})/ \tau) }{ \exp (sim(z_a^+ , z_{m}^{+}) / \tau) + \exp (sim(z_a^+, z_{m}^{-}) / \tau)}
+\end{equation}
+where $z = f(x)$.
+
+\subsubsection{Supervised Contrastive Learning}
+\label{sub:baseline-supcon}
+
+\citet{NEURIPS2020_supervised_contrastive} propose a supervised way of performing contrastive learning which is shown in \cref{eq:supervised-contrastive}.
+
+\begin{equation}\label{eq:supervised-contrastive}
+    \sum_{i \in I} \frac{-1}{|P(i)|} \sum_{p \in P(i)} \log \frac{\exp (z_i \cdot z_p / \tau) }{\sum_{a \in A(i)} \exp (z_i \cdot z_a / \tau)}
+\end{equation}
+
+% \begin{equation}\label{eq:supervised-contrastive-my-notation}
+%     \sum_{i \in I} \frac{-1}{|P(i)|} \sum_{p \in P(i)} \log \frac{\exp (f_i(x_i) \cdot f_p(x_p) / \tau) }{\sum_{a \in A(i)} \exp (f_i(x_i) \cdot f_a(x_a) / \tau)}
+% \end{equation}
+
+While this model is a strong baseline, if modalities are ablated, this method does not perform as well as our proposed model (we show and discuss this later, in \cref{fig:epochs-mrr.srd}).
+This is especially true when the text modality is dropped, as is likely in physical agent scenarios, where speech is a natural interaction mechanism.
+
+\subsection{Modality Ablation}
+We consider an experiment in which we incorporate RGB, depth, speech, and written language to train the model.
+The loss function requires no changes beyond increasing the value of $M$ by 1 in \cref{eq:objective-emma}.
+Our goal is the non-trivial downstream prediction task: determining what objects are being referred to by arbitrary language from a small number of examples.
+When we consider only text, RGB, and depth, written language is used as the anchor modality, and we compute the distance of RGB and depth modalities from it and then average them.
+However, when speech is incorporated as an additional fourth sensory modality, we have three possible choices.
+\textit{First}, we can compute the distance of RGB and depth from text and from speech which gives us 4 distance matrices, and then take average of these four.
+\textit{Second}, we can treat speech in a similar way to RGB and depth: compute the distance of RGB, depth, and speech from text, and then take an average of three of them.
+\textit{Third}, similar to the first method, but adding the distance between language and speech as well and then take the average of 5 distance matrices.
+
+Of these, the first method is the most appropriate choice for a robust multimodal alignment approach.
+The second and third options are possible during training, but in real-world object retrieval scenarios, having only one form of language instructions is a reasonable scenario---people are not likely to \textit{both} speak about \textit{and} type in instructions for an agent.
+At test time, depending on which modalities are available to the model, we can use speech, text, or both to compute the distance of RGB and depth embeddings from the linguistic query, and then take the average.
+\Cref{eq:distance-matrix} shows all possible cases of modality dropout and the corresponding distance computations, where ``t'' represents \textit{text}, ``s'' represents \textit{speech}, ``r'' represents \textit{RGB}, ``d'' represents \textit{depth}, $K_{ij}$ for each pair of modalities is the distance of one instance in modality $i$ from all instances in modality $j$, and $K$ is the final distance---a matrix if there are multiple queries and a vector if there is one query.
+\begin{equation}
+\label{eq:distance-matrix}
+K = 
+    \begin{cases} 
+        K_{tr} & \text{speech and depth are missing} \\
+        K_{sr} & \text{text and depth are missing} \\
+        K_{td} & \text{speech and RGB are missing} \\
+        K_{sd} & \text{text and RGB are missing} \\
+        K_{trd} = \frac{K_{tr} + K_{td} }{2} & \text{speech is missing} \\
+        K_{srd} = \frac{K_{sr} + K_{sd} }{2} & \text{text is missing} \\
+        K_{tsr} = \frac{K_{tr} + K_{sr} }{2} & \text{depth is missing} \\
+        K_{tsd} = \frac{K_{td} + K_{sd} }{2} & \text{RGB is missing} \\
+        K_{tsrd} = \frac{K_{tr} + K_{td} + K_{sr} + K_{sd} }{4} & \text{o.w.} \\
+    \end{cases}
+\end{equation}
+
+\Cref{fig:epochs-mrr.srd,fig:epochs-mrr.trd,fig:epochs-mrr.sd,fig:epochs-mrr.sr} show the relative performance of EMMA against baselines when different modalities are ablated.
+
+\subsection{Results}
+\label{sec:results}
+In this section we provide quantitative and qualitative results by comparing our method against supervised contrastive learning~\cite{NEURIPS2020_supervised_contrastive} and na\"ive contrastive loss~\cite{chen2020simple}.
+% We evaluate our EMMA loss function against the na\"ive triplet loss method.
+Our model significantly outperforms the triplet method across all epochs while it converges faster as shown in \cref{fig:epochs-mrr.srd}.
+\Cref{table:quantitative} summarizes the performance of all models with different ablation metrics.
+% \todo[inline]{Why do we call out triplet loss specifically? We evaluate against several baselines? I don't understand this section particularly}
+To provide a better sense of the performance measure, we can think of a model that always ranks the correct object in the second place which is not very bad.
+Such a model would have MRR of $1/2 = 0.5$.
+
+Comparing \cref{fig:epochs-mrr.trd} and \cref{fig:epochs-mrr.srd}, we can see that when the text modality is dropped the performance decreases from about 0.90 to about 0.79, showing that speech cannot completely replace the text.
+We also observe that speech has a higher variance compared to text, and a na\"ive contrastive learning method cannot take advantage of text modality to reduce this variance.
+While the other two methods leverage text modality to align speech better.
+
+There is very little gap in performance when rgb is dropped in \cref{fig:epochs-mrr.sd} compared to when depth is dropped in \cref{fig:epochs-mrr.sr} which shows that our model is robust enough when rgb or depth sensors fail in real-world deployment.
+
+\begin{figure}[tbh]
+\centering
+\includegraphics[width=.99\columnwidth]{Figures/average-seeds-epochs-mrr_ard.pdf}
+\caption{Mean Reciprocal Rank (MRR) on the held-out test set when the text modality is ablated, averaged over 3 runs.
+Green is self-supervised contrastive learning, orange is supervised contrastive learning, and blue represents our proposed EMMA loss function.
+Higher is better.
+% \todo[inline]{Fix description of colors.}
+}
+\label{fig:epochs-mrr.srd}
+\end{figure}
+
+\begin{figure}[tbh]
+\centering
+\includegraphics[width=.99\columnwidth]{Figures/average-seeds-epochs-mrr_lrd.pdf}
+\caption{Mean Reciprocal Rank (MRR) on the held-out test set when the speech modality is ablated, averaged over 3 runs, for the downstream task of object retrieval.
+Green is self-supervised contrastive learning, orange is supervised contrastive learning, and blue represents our proposed EMMA loss function.
+Higher is better.
+When speech as a modality is ablated, all tested methods perform similarly.
+}
+\label{fig:epochs-mrr.trd}
+\end{figure}
+
+\begin{figure}[tbh]
+\centering
+\includegraphics[width=.99\columnwidth]{Figures/average-seeds-epochs-mrr_ar.pdf}
+\caption{Mean Reciprocal Rank (MRR) on the held-out test set when text and depth modalities are ablated, averaged over 3 runs for the downstream task of object retrieval.
+Green is self-supervised contrastive learning, orange is supervised contrastive learning, and blue represents our proposed EMMA loss function.
+Higher is better.
+% Orange represents Supervised Contrastive Learning~\cite{NEURIPS2020_supervised_contrastive}, \Complete{} represents the triplet loss function~\cite{triplet_loss_2021_CVPR}.
+}
+% \todocmi{Maybe MRR is wrong and we should be looking at top-ranked result?}
+\label{fig:epochs-mrr.sr}
+\end{figure}
+
+\begin{figure}[tbh]
+\centering
+\includegraphics[width=.99\columnwidth]{Figures/average-seeds-epochs-mrr_ad.pdf}
+\caption{Mean Reciprocal Rank (MRR) on the held-out test set when text and RGB modalities are dropped averaged over 3 runs for the downstream task of object retrieval.
+Green is self-supervised contrastive learning, orange is supervised contrastive learning, and blue represents our proposed EMMA loss function.
+Higher is better.
+}
+\label{fig:epochs-mrr.sd}
+\end{figure}
+
+\begin{figure*}[tbh]
+\centering
+\includegraphics[width=1.99\columnwidth]{Figures/3D-tsne-4M-eMMA-cosine-submodalities-text-anchor-gold-no_neg_sampling-1024-sbs.png}
+\caption{3D T-SNE~\cite{van2008tsne} projection of test embeddings of 10 randomly selected classes of objects using our proposed method EMMA.
+Each modality is separately projected into a three-dimensional space.
+In a perfect embedding, all instances of a class would be clustered in identical areas of the embedding space across all modalities.
+We can see that EMMA successfully encourages all four modalities to live in a common manifold, allowing accurate retrieval even when modalities are missing.
+}
+\label{fig:3d-tsne}
+\end{figure*}
+
+% \subsection{Qualitative Results}
+% \label{sec:Qualitative}
+We look into the projection of the high-dimensional learned embeddings into a 2-dimensional and 3-dimensional spaces using t-SNE~\cite{van2008tsne}, which is a dimensionality reduction technique to visualize high-dimensional data.
+T-SNE creates a probability distribution over pairs high-dimensional data where similar pairs have a higher probability and dissimilar pairs have a lower probability.
+A similar probability distribution is also defined over pairs of data in the lower dimension (either 2D or 3D), and T-SNE minimizes the KL divergence between these two probability distributions.
+% \textcolor{red}{\dots}\todo{explanation}. 
+% \Cref{fig:2d-tsne} shows the projection of all test embeddings from all different modalities in a 2D space.
+As shown in the graph, instances of the same class but different modality are mapped to almost the same locations and are closer to each other.
+\Cref{fig:3d-tsne} shows the projection onto 3D space to give a better view of the location of embeddings.
+Although these projections are not perfect, combined with the quantitative results, they demonstrate that our model is learning to map instances of the same class closer to each other regardless of their modalities.
+
+% \begin{figure}[tbh]
+% \centering
+% \includegraphics[width=.99\columnwidth]{Figures/2D-tsne-4M-eMMA-cosine-submodalities-text-anchor-gold-no_neg_sampling-1024.pdf}
+% \caption{2D T-SNE~\cite{van2008tsne} projection of test embeddings of 10 randomly selected classes of objects using our proposed method.
+Each modality is separately projected into a two-dimensional space.
+\Rephrase{} \textcolor{red}{We may only want 3D}
+% }
+% \label{fig:2d-tsne}
+% \end{figure}
+
+% \todokdinline{Add the T-SNE projection for supervised contrastive loss as well.}
+
+% \subsection{Failed Experiments}
+% \todokdinline{I suppose I should remove this subsection, right?}
+% We tried a lot of different approaches and they did not work better than our model or the baselines, but we list them here to save time for readers who want to try them.
+
+% \subsubsection{Simple MMA}
+% Copy text from \ref{sub:simple-mma} if we decide to keep this section.
+
+% \subsubsection{Two Anchors}
+% Since speech is similar to text in its usage when it comes to the downstream task of object retrieval (given a single language command, find the correct object among multiple objects), we decided to have text and speech as two anchors instead of having one anchor only.
+When there is one anchor, the distance between positive pairs text and speech is minimized and the distance between negative pairs are minimized.
+However, this is not related to the downstream task since we never want to find the correct speech for a given text or vice versa.
+The second term added to the last function is exactly similar to the loss function when speech is the anchor, and the first term is the same as when language is the anchor.
+Therefore, the performance is somewhere in the middle of those two approaches but more towards the speech when the metrics are computed based and speech matrices.
+
+% \subsection{Example Prediction}
+
+\subsection{Discussion}
+Our proposed model performs well, has been demonstrated to handle four modalities of shared information effectively, and is robust to test-time situations where information from one or more modalities is missing.
+However, there is still room for improvement.
+Specifically, the speech modality is harder to handle.
+\Cref{fig:3d-tsne} shows that although the relative position of instances are correct in the speech space, the distinction and clustering of different objects are not as good as the other three modalities.
+
+Text seems to be the best clustered modality, and that makes sense because the variation in written text is much smaller than the other three modalities.
+Variation in speech is higher because of different accents, native language, gender, age, and etc., variation in RGB and depth is higher than text because of lightning conditions, object's texture and shape, the angle of camera, and other factors.
+
+\begin{figure}[h!]
+\centering
+\includegraphics[width=.8\columnwidth]{Figures/example-rankings.png}
+\caption{EMMA more accurately ranks objects and handles ambiguities with respect to a retrieval query.}
+\label{fig:rankings}
+\end{figure}
+
+An example of the need to consider multiple modalities jointly is shown in \cref{fig:rankings}, showing how EMMA is able to correctly select an object instance from several similarly shaped and describable objects.
 
@@
 
+In this work, we have demonstrated the effectiveness of a novel approach to learning from high-dimensional multimodal information.
+Our approach performs well on an object retrieval task from a testbed that contains four separate modalities, consistent with information that might be available to a physical agent, and outperforms state of the art contrastive learning approaches when one or more modalities is unavailable at test time.
+Our proposed method is general enough to be applied to a variety of multimodal retrieval problems, and is not limited to purely language-based image retrieval.
+
+In future, this work will be extended to solve less clearly delineated problems; for example, differentiating among members of a class, as well as across classes.
+However, this work represents a significant step towards handling such retrieval problems, while not arbitrarily limiting the number of sensor and other modalities that can be incorporated.
+
+% Another future work is to improve the speech modality so it can contribute more to the training.
+Right now when text modality is dropped, the performance decreases by about 0.12.
+Ideally a better model should have the same performance in case of dropout when it is trained using all modalities.
+
 %===================================================================
@@
 
+% %% The acknowledgments section is defined using the "acks" environment
+% %% (and NOT an unnumbered section). This ensures the proper
+% %% identification of the section in the article metadata, and the
+% %% consistent spelling of the heading.
+\begin{acks}
+To Robert, for the bagels and explaining CMYK and color spaces.
+\end{acks}
+% \section*{Acknowledgments}
+
 \clearpage
-\section*{Acknowledgments}
-%% The file named.bst is a bibliography style file for BibTeX 0.99c
-\bibliographystyle{named}
+\clearpage
+% %%
+% %% The next two lines define the bibliography style to be used, and
+% %% the bibliography file.
+\bibliographystyle{ACM-Reference-Format}
 \bibliography{ref}
-\clearpage
 
+% %%
+% %% If your work has an appendix, this is the place to put it.
 \appendix
 \label{sec:appendix}
-\section{Qualitative Analysis}
-
-\section{Model}
-
-\section{Experiments}
-
-\section{Results}
-
-\section{Setup}
 

### 2022-08-17 Paper/main.tex
first draft of paper for TMLR

@@
-\documentclass[sigconf,natbib=true,anonymous=true]{acmart}
-% \documentclass[manuscript,screen,review]{acmart}
+\documentclass[10pt]{article} % For LaTeX2e
+\usepackage{tmlr}
+% If accepted, instead use the following line for the camera-ready submission:
+%\usepackage[accepted]{tmlr}
+% To de-anonymize and remove mentions to TMLR (for example for posting to preprint servers), instead use the following:
+%\usepackage[preprint]{tmlr}
 
-% Needed to fix an acmart problem with line numbering. AFAICT line 
-% numbers are not expected in sigconf, so it should be a no-op.
-\makeatletter
-\providecommand{\@LN}[2]{}
-\makeatother
+% Optional math commands from https://github.com/goodfeli/dlbook_notation.
+\input{math_commands.tex}
 
-% % Fonts used in the template cannot be substituted; margin 
-% % adjustments are not allowed.
-% %
-% % \BibTeX command to typeset BibTeX logo in the docs
-% \AtBeginDocument{%
-%   \providecommand\BibTeX{{%
-%     \normalfont B\kern-0.5em{\scshape i\kern-0.25em b}\kern-0.8em\TeX}}}
+\usepackage{hyperref}
+\usepackage{url}
 
-%% Rights management information.  This information is sent to you
-%% when you complete the rights form.  These commands have SAMPLE
-%% values in them; it is your responsibility as an author to replace
-%% the commands and values with those provided to you when you
-%% complete the rights form.
-\setcopyright{acmcopyright}
-\copyrightyear{2022}
-\acmYear{2022}
-\acmDOI{XXXXXXX.XXXXXXX}
-
-%% These commands are for a PROCEEDINGS abstract or paper.
-\acmConference[SIGIR '22]{Make sure to enter the correct
-  conference title from your rights confirmation email}{June 03--05,
-  2022}{Madrid, Spain}
-%
-%  Uncomment \acmBooktitle if the title of the proceedings is different
-%  from ``Proceedings of ...''!
-%
-\acmBooktitle{SIGIR '22: Proceedings of the 45th International ACM SIGIR Conference on Research and Development in Information Retrieval,
- July 11--15, 2022, Madrid, Spain} 
-% \acmPrice{15.00}
-% \acmISBN{978-1-4503-XXXX-X/18/06}
-
-%
-% Submission ID.
-% Use this when submitting an article to a sponsored event. You'll
-% receive a unique submission ID from the organizers
-% of the event, and this ID should be used as the parameter to this command.
-% \acmSubmissionID{123-A56-BU3}
-
-%
-% The majority of ACM publications use numbered citations and
-% references.  The command \citestyle{authoryear} switches to the
-% "author year" style.
-%
-% If you are preparing content for an event
-% sponsored by ACM SIGGRAPH, you must use the "author year" style of
-% citations and references.
-% Uncommenting
-% the next command will enable that style.
-%\citestyle{acmauthoryear}
-
-%% The "author" command and its associated commands are used to define
-%% the authors and their affiliations.
-%% Of note is the shared affiliation of the first two authors, and the
-%% "authornote" and "authornotemark" commands
-%% used to denote shared contribution to the research.
-\author{Ben Trovato}
-\authornote{Both authors contributed equally to this research.}
-\email{trovato@corporation.com}
-\orcid{1234-5678-9012}
-\author{G.K.M. Tobin}
-\authornotemark[1]
-\email{webmaster@marysville-ohio.com}
-\affiliation{%
-  \institution{Institute for Clarity in Documentation}
-  \streetaddress{P.O. Box 1212}
-  \city{Dublin}
-  \state{Ohio}
-  \country{USA}
-  \postcode{43017-6221}
-}
-
-\author{Lars Th{\o}rv{\"a}ld}
-\affiliation{%
-  \institution{The Th{\o}rv{\"a}ld Group}
-  \streetaddress{1 Th{\o}rv{\"a}ld Circle}
-  \city{Hekla}
-  \country{Iceland}}
-\email{larst@affiliation.org}
-
-\author{Valerie B\'eranger}
-\affiliation{%
-  \institution{Inria Paris-Rocquencourt}
-  \city{Rocquencourt}
-  \country{France}
-}
-
-\author{Aparna Patel}
-\affiliation{%
- \institution{Rajiv Gandhi University}
- \streetaddress{Rono-Hills}
- \city{Doimukh}
- \state{Arunachal Pradesh}
- \country{India}}
-
-\author{Huifen Chan}
-\affiliation{%
-  \institution{Tsinghua University}
-  \streetaddress{30 Shuangqing Rd}
-  \city{Haidian Qu}
-  \state{Beijing Shi}
-  \country{China}}
-
-\author{Charles Palmer}
-\affiliation{%
-  \institution{Palmer Research Laboratories}
-  \streetaddress{8600 Datapoint Drive}
-  \city{San Antonio}
-  \state{Texas}
-  \country{USA}
-  \postcode{78229}}
-\email{cpalmer@prl.com}
-
-\author{John Smith}
-\affiliation{%
-  \institution{The Th{\o}rv{\"a}ld Group}
-  \streetaddress{1 Th{\o}rv{\"a}ld Circle}
-  \city{Hekla}
-  \country{Iceland}}
-\email{jsmith@affiliation.org}
-
-\author{Julius P. Kumquat}
-\affiliation{%
-  \institution{The Kumquat Consortium}
-  \city{New York}
-  \country{USA}}
-\email{jpkumquat@consortium.net}
-
-%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%
-% EDITING LINK: 
-% https://www.overleaf.com/3575751412sjsdmszkxsyh
-%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%
-
-% \usepackage{soul} %% Soul does not play well with acmart
-\usepackage{url}
-% % \usepackage[hidelinks]{hyperref}https://www.overleaf.com/project/614b5cc61d38655dce466002
-% \usepackage[small]{caption}
-\usepackage{graphicx}
-\usepackage{amsthm}
-\usepackage{booktabs}
-\usepackage{algorithm}
-\usepackage{algorithmic}
+% My Packages
 \usepackage{cleveref}
-\usepackage{lipsum}
-
-\urlstyle{same}
-
-% % =======================================
-% % my imported packages
-
-\usepackage{amsmath}
-\usepackage{bm}
-% \usepackage{amssymb}          % duplicates something in acmart
-\usepackage{listings}
-% \usepackage[ruled,vlined]{algorithm2e}
-\include{pythonlisting}
-
-\newcommand{\pseudosection}[1]{\vspace{1.5ex}\noindent \textit{{#1:~~}}}
-% \newcommand{\pseudosection}[1]\indent{{{}}}
-
+\usepackage{comment}
 \usepackage{todonotes}
@@
 \newcommand{\todoff}[1]{\todo[color=blue!20]{\small #1 -- Frank}}
+\newcommand{\todoffinline}[1]{\todo[inline,color=blue!20]{\small #1 -- Frank}}
 \newcommand{\todoed}[1]{\todo[color=cyan!20]{\small #1 -- Ed}}
@@
 \newcommand{\Rephrase}[1]{{\textbf {\color{red}Rephrase please #1}}}
+\newcommand{\TD}[1]{{\color{red}{\textbf{TODO}}} #1}
+\newcommand{\OK}[1]{{\color{green}{\textbf{DONE}}} #1}
+\newcommand{\ours}{\textsc{EMMA}}
+% \newcommand{\ours}{\textsc{EMMA}}
+\newcommand{\geom}{\textsc{Geometric Alignment}}
+\newcommand{\supcon}{\textsc{SupCon}}
 
-% \newcommand{\citet}[1]{\cite{#1}} % This is used because \citet didn't work with their author kit % It does now
+%%%%%%%%%%%%%%%%%%%%
+% EDITING LINK: 
+% https://www.overleaf.com/3575751412sjsdmszkxsyh
+%%%%%%%%%%%%%%%%%%%%
 
-% =======================================
-\title{Multimodal Contrastive Learning for Object Retrieval in the Face of Modality Failures}
+\title{Multimodal Contrastive Learning for Object Retrieval in the Face of Modality Failures \\ Extended Multimodal Alignment \\ Combining Geometric and classification ... }
+
+% Authors must not appear in the submitted version. They should be hidden
+% as long as the tmlr package is used without the [accepted] or [preprint] options.
+% Non-anonymous submissions will be rejected without review.
+
+\author{\name Kyunghyun Cho \email kyunghyun.cho@nyu.edu \\
+      \addr Department of Computer Science\\
+      University of New York
+      \AND
+      \name Raia Hadsell \email raia@google.com \\
+      \addr DeepMind
+      \AND
+      \name Hugo Larochelle \email hugolarochelle@google.com\\
+      \addr Mila, Universit\'e de Montr\'eal \\
+      Google Research\\
+      CIFAR Fellow}
+
+% The \author macro works with any number of authors. Use \AND 
+% to separate the names and addresses of multiple authors.
+
+\newcommand{\fix}{\marginpar{FIX}}
+\newcommand{\new}{\marginpar{NEW}}
+
+\def\month{MM}  % Insert correct month for camera-ready version
+\def\year{YYYY} % Insert correct year for camera-ready version
+\def\openreview{\url{https://openreview.net/forum?id=XXXX}} % Insert correct link to OpenReview for camera-ready version
 
@@
 
-% Todo list
-% \begin{enumerate}
-%     \item \st{Lit search}
-% \end{enumerate}
-% \newpage
-
-% ***************** The BEGINNING of the paper *****************
+\maketitle
 
@@
 
-    We propose a generalized distance-based loss function that can be used to learn retrieval models that incorporate an arbitrary number of views of a particular piece of data, and compounded by the challenge of retrieval when a modality becomes unavailable.
-    Our study is motivated by needs in robotics and computer-human interaction, where an agent has many sensors and thus modalities in which a human may interact: voice, text, images, and depth, both to communicate a desired goal and for the agent to recognize a desired target object.
-    Tying these various modalities together relates to the task of grounded language understanding, in which natural language is used as a query against objects in a physical environment, allowing a real-world, intuitive mechanism by which users can instruct physical agents to engage in tasks such as object retrieval.
-    However, current work has neglected the real-world constraint that sensors fail: hardware can become damaged or shipped with defects, sensors can get blocked or obstructed, and various adverse but not uncommon conditions that remove a modality from use.
-For practical applications, we need mechanisms for accurate object retrieval given that a modality may fail without notice.
+    We propose a generalized geometric combined with cross-entropy loss function that can be used to learn retrieval models that incorporate an arbitrary number of views of a particular piece of data, and compounded by the challenge of retrieval when a modality becomes unavailable.
+    \todoff{this red text can (should) be shortened}
+    \todokd{I shortened it. Please remove both notes if it's resolved}
+    {\color{red}Our study is motivated by needs in robotics and computer-human interaction, where an agent has many sensors and thus modalities in which a human may interact both to communicate a desired goal and for the agent to recognize a desired target object.
+    Tying these various modalities together relates to the task of grounded language understanding. However, there has been little research on integrating more than two modalities.
+    % \todokdinline{Duplicate, it's already mentioned in intro}in which natural language is used as a query against objects in a physical environment, allowing a real-world, intuitive mechanism by which users can instruct physical agents to engage in tasks such as object retrieval.
+    % However, current work has neglected the real-world constraint that sensors fail. For practical applications,
+    %  we need mechanisms for accurate object retrieval given that a modality may fail without notice.
+    }
+    % \todoff{There's very little in here about actual method or conclusions}
+    There has been widely popular works on self-supervised contrastive learning based on cross-entropy, but there is an entire other approach based on explicit geometric alignment.
+However, there has been no work on combining the two approaches for multimodal learning to the best of our knowledge.
+We propose to combine the two approaches and argue that the two are complementary.
     We demonstrate the usability of our model on a grounded language object retrieval scenario, where an intelligent agent has to select an object given an unconstrained language command. 
-We leverage four modalities including vision, depth sensing, text, and speech, and we show that our model can outperform state-of-the-art contrastive models when modalities are ablated.
+    We leverage four modalities including vision, depth sensing, text, and speech, and we show that our model can outperform state-of-the-art contrastive learning models by $2\%$ while converging/learning/training $5$ times faster (takes $80\%$ less time or is $20\%$ of the baseline).
+    %when modalities are ablated. 
+    The code will be publicly available on GitHub and will be included for the camera-ready version (it is redacted for anonymity). 
 \end{abstract}
-
-\settopmatter{printfolios=true} % Page numbers please 
-% 9 pages without references.
-
-% \begin{enumerate}
-%     \item Objects of five things we select from
-%     \item \textcolor{green}{DONE} Adding an overview of the whole process 
-%     \item \textcolor{blue}{PROG} Go into more details in section 4 and give examples 
-%     \item Move images around
-%     \item Quant result section is not complete
-%     \item \textcolor{green}{DONE} put the graphs in modality ablation or merge quant and qual
-%     \item \textcolor{green}{DONE} add dropping R D graphs
-%     \item A paragraph in related work about grounded language 
-%     % \item Don't mess with additional experiments
-% \end{enumerate}
-%===================================================================
-
-\maketitle
-% \setcounter{page}{1}
-
-%===================================================================
-
-% \section{SIGIR Papers}
-% \begin{enumerate}
-%     \item \href{https://dl.acm.org/doi/pdf/10.1145/3397271.3401430}{FashionBERT}
-%     \item \href{https://dl.acm.org/doi/pdf/10.1145/3397271.3401265?casa_token=W8NqNWp0UQkAAAAA:w-iRUnNhmytDKCUiYzaUPG0bI_yyo8Bzxwro_eOC-zl5VDIGRVT2dvixvnCwLJ0iEKqYo1LkQIWXVA}{CrossBERT}
-
-% \end{enumerate}
 
@@
 
-Humans use multiple sensory inputs to understand, interact with, and retrieve information from the world around them.
-It is intuitive that agents learning about the world, upon encountering new concepts and new objects, should form a model that incorporates information from all available sensors and data sources.
+\todoffinline{nix this first sentence and revise the start of the second.
+We're not doing any cogsci-y stuff, or even human studies tied to modality deprivation.
+You could take most of the red text in the abstract I mentioned should be compressed and rework it here}
+\todokdinline{I took a shot at nixing it.
+Please remove both notes if it's resolved.
+I don't get the vibe of cogsci-y stuff or human studies from the first paragraph, would you please refer to those sentences?}
+Humans successfully use multiple sensory inputs to understand, interact with, and retrieve information from the world around them.
+Inspired by the multimodal nature of human interactions with the world, it is intuitive that agents learning about the world, upon encountering new concepts and new objects, should form a model that incorporates information from all available sensors and data sources.
+Benefits of integrating multiple modalities is twofold; first, additional modalities can help in the cases when other modalities fail, and second, complementary information can be extracted from different modalities that can help with understanding the world.
 % 
 Grounded language understanding, in which natural language is used as a query against objects in a physical environment, allows a real-world, intuitive mechanism by which users can instruct physical agents to engage in tasks such as object retrieval.
-Visuolinguistic approaches to such object inference tasks typically involve training on large pools of image/text pairs and then using language to subselect elements of the sensed environment~\cite{hong2021gilbert}.
-Although physical agents such as robots typically have access to sensory and interactive modalities beyond vision, and learning from multiple modalities can improve performance on downstream tasks, most approaches use at most two sensory inputs (e.g., visual data such as RGB images) with single labels, such as those provided by textual natural language.
+Visuolinguistic approaches to such object inference tasks typically involve training on large pools of image/text pairs and then using language to subselect elements of the sensed environment~\cite{hong2021gilbert,Zhuge_2021_CVPR_VLP}.
+\todoffinline{Unless Hong et al is a review paper, add at least one more citation}
+\todokdinline{I believe this sentence was added to cite some relevant work from SIGIR, so we can remove it. However, I added one more citation anyway.}
+Although physical agents such as robots typically have access to sensory and interactive modalities beyond vision, and learning from multiple modalities can improve performance on downstream tasks, most approaches use at most two sensory inputs (e.g., visual data such as RGB plus depth images) with single labels, such as those provided by textual natural language.
 Simultaneously using additional inputs from different modalities is an underexplored area, in part due to the domain-specific nature of such \textit{n}-ary learning approaches.
-With the modern proliferation of audio and text based communication and proliferation of home agents (e.g., iRobot vacuums and Alexa/Google Home), there is a growing need to handle more modalities, and simultaneously their potential failures.
+With the modern proliferation of audio and text based communication and home agents (e.g., Alexa/Google Home), there is a growing need to handle more modalities, and simultaneously their potential failures.
 
-One difficulty with working with complex multimodal data is the increased likelihood that one or more modalities may have missing information.
+One difficulty with working with complex multimodal data is the increased likelihood that one or more modalities may have missing information; hardware can become damaged or shipped with defects, sensors can get blocked or obstructed, and various adverse but not uncommon conditions that remove a modality from use.
 Current multimodal approaches are typically not robust to the loss of one or more modalities at test time, as may happen if, for example, a physical agent fails to retrieve data from a particular sensor.
@@
 
-The high level goal of this work is to take arbitrary input modalities about novel objects, including both spoken and written language, and build a model that allows a system to correctly retrieve that object given a subset of those input modalities.
+\todoff{revise this goal. It sets up the reviewers to take the easy way out and say results are too incremental}
+The high-level goal of this work is to take arbitrary input modalities about novel objects, including both spoken and written language, and build a model that allows a system to correctly retrieve that object given a subset of those input modalities.
 This is a generalization of approaches to grounded language learning in which specific input modalities are labeled with language to allow for future identification, but explicitly seeks to be agnostic about the nature of the individual sensory and linguistic inputs.
@@
 \begin{enumerate}
-    \item Proposing a general distance-based loss function formulation for multimodal learning that can be extended to an arbitrary number of modalities.
-    \item Outperforming state-of-the-art contrastive learning models when one or more modalities are ablated at test time, demonstrating robustness in the face of missing information.
+    \item Proposing a new approach to multimodal learning by combining geometric and cross-entropy based methods and showing that it outperforms and converges faster compared to each method separately.
     % \item Outperforming state-of-the-art models for grounded language learning by differentiating visual input modalities such as RGB and depth, rather than combining them into a single modality.
+
+    \item Our method can be easily extended to an arbitrary number of modalities.\todoffinline{potential counter: isn't the method quadratic?}
+    \todokdinline{Yes, but nonetheless it can be extended.}
+
+    \item Outperforming state-of-the-art contrastive learning~\cite{chen2020simple} and supervised contrastive learning~\cite{NEURIPS2020_supervised_contrastive} models by $2\%$ when all modalities are available, and achieving the peak performance $5$ times faster than both methods.
+    % Ours takes 8 epochs, baselines take 40 epochs.
+    \todokdinline{Should we break this contribution into two contributions, one for faster learning and the other for performance compared to both supcon and vanilla contrastive learning?}
+
+    \item Demonstrating robustness in the face of missing information, when one or more modalities are ablated at test time.
+
     \item Demonstrating the separate utility of speech and text as sources of label information by treating them as sensory input modalities, rather than explicitly as labels.
+
+    % \item Stable model that outperforms state-of-the-art contrastive learning models when batch size decreases.\todoff{add motivation in the intro about why robustness to a smaller batch size is important}
+
 \end{enumerate}
 
-The remainder of this paper is organized as follows.
+% \todoffinline{Super pet peeve of mine: except for books or theses, don't take up space with this roadmap.
+Instead, in the above contributions, add pointers to the sections (or even subsections) where we provide evidence for each contribution}
+% \todokdinline{I think Cynthia added this part for the last submission since we were short a few sentences to reach the page limit. We can remove this.}
+% The remainder of this paper is organized as follows.
 In \cref{sec:Related-Work}, we describe existing work on visuolinguistic retrieval, contrastive learning, and multimodal learning.
@@
 
+\todoffinline{use paragraphs to add some light structure to this block-o-text}
+
+\todoffinline{I'm not convinced this section shouldn't appear toward the end of the paper. Maybe adding that narrative structure will help?}
+
 There is extensive work on the task of image retrieval, of which physical object retrieval can be considered a special case.
@@
 Prior work has focused solely on language and visual modalities, using text or a combination of language and vision to perform visual retrieval~\cite{vo2019composing}.
-While recent works have focused on grounding with respect to a knowledge set \cite{10.1145/3357384.3357889,10.1145/3397271.3401097}, our work focuses on robust multimodal learning over to perform grounding over an arbitrary number of input modalities.
+While recent works have focused on grounding with respect to a knowledge set \cite{10.1145/3357384.3357889,10.1145/3397271.3401097}, our work focuses on robust multimodal learning to perform grounding over an arbitrary number of input modalities.
 
@@
 They cover different approaches to multimodal machine learning such as \textit{representation}, \textit{translation}, \textit{alignment}, \textit{fusion}, and \textit{co-learning}.
-In this project, we focus on alignment, the process of identifying the direct relationship between elements across modalities.
+In this work, we focus on alignment, the process of identifying the direct relationship between elements across modalities.
 
@@
 
-Another related approach is contrastive learning~\cite{qin2021world}, which is based on the idea of similarity and distance of elements of training data~\cite{NEURIPS2020_supervised_contrastive,chen2020simple}.
+In general there are two different family of approaches to multimodal learning in the literature.
+One family is based on classification and cross-entropy losses such as~\citet{NEURIPS2020_supervised_contrastive, chen2020simple} and the other family is geometric based such as~\citet{ICML22GeometricMultimodal,Carvalho-cooking-triplet,salvador2017cooking,triplet_loss_2021_CVPR}.
+In this paper we propose to marry the two families and show that combining geometric approach with cross-entropy method is superior. \todokdinline{Pun intended :-)}
+\todokdinline{Should we move the following two paragraphs to the Approach section?}
+For the classification part of our loss function we employ the supervised contrastive loss proposed by~\citet{NEURIPS2020_supervised_contrastive}.
+The way we formulate our geometric part of the loss function is similar to lifted structured loss~\cite{songCVPR16LiftedStructured}, where they take a geometric approach in a unimodal scenario and consider pairwise distances among all items in a batch, but in this paper we apply our idea to four modalities, and our formulation can be extended to any number of modalities.
+This extended geometric formulation results in a loss function that is similar to \citet{ICML22GeometricMultimodal} where they propose a geometric contrastive learning approach in which they %use the same instance of an MLP network (same weights) after extracting embeddings from each modality, while we use different instances of the same MLP network (different weights).
+They also
+fuse all embeddings as a central embedding and then minimize the distance between each embedding and that central embedding.
+However, their method does not take into account any notion of classification.
+
+The geometric approach is also known as contrastive learning~\cite{qin2021world}, which is based on the idea of similarity and distance of elements of training data~\cite{Carvalho-cooking-triplet,triplet_loss_2021_CVPR,salvador2017cooking}.
 Triplet loss is also a special case of contrastive loss~\cite{NEURIPS2020_supervised_contrastive}.
-Contrastive loss is mostly used in self-supervised learning~\cite{bui2021self, alayrac2020self,chen2020simple}, with the exception of \citet{NEURIPS2020_supervised_contrastive}, who use contrastive loss for a supervised problem.
+Contrastive loss is mostly used in self-supervised learning~\cite{bui2021self, alayrac2020self,chen2020simple}, with some exceptions such as \citet{NEURIPS2020_supervised_contrastive}, who use contrastive loss for a supervised problem.
 Most of these methods are implemented for a non-multimodal dataset (e.g., RGB images only).
@@
 This work differs from ours in that it becomes computationally complex as more and higher-dimensional modalities are added.
+Our proposed geometric loss function resembles work on quadruplet loss~\cite{chen2017beyond,tursun2021efficient} but is intended to scale to an arbitrary number of modalities.
 
@@
 % \todo[inline]{talk about CLIP and DALL-E here}
-
 To the best of our knowledge, there has been limited work on incorporating more than three modalities into a general learning approach.
-Our proposed loss function closely resembles work on quadruplet loss~\cite{chen2017beyond,tursun2021efficient} but is intended to scale to an arbitrary number of modalities.
 
@@
 \centering
-\includegraphics[width=.99\columnwidth]{Figures/experiment-setup.png}
+\includegraphics[width=.5\columnwidth]{Figures/experiment-setup.png}
 \caption{
@@
 Given a query in some modality, our approach seeks to select the object that is the best fit, per a trained model.
-This approach, detailed in \cref{sec:Method}, is able to rank objects as to appropriateness even when one or more modalities is ablated at test time (e.g., textual descriptions are missing), and outperforms state-of-the-art contrastive learning approaches on this task.
+This approach, detailed in \cref{sec:Method}, is able to rank objects as to appropriateness even when one or more modalities is ablated at test time (e.g., depth inputs are missing), and outperforms state-of-the-art contrastive learning approaches on this task.
 }
@@
 
-In keeping with previous work on the closely related problem of image retrieval, we focus on contrastive loss approaches, in which the goal is to learn an embedding of data in which similar samples---in our case, samples belonging to the same class of object---are embedded `close' to one another, while dissimilar samples are farther apart.
-We develop a novel loss function that simultaneously minimizes intra-class distances and maximizes inter-class distances across each pair of modalities, yielding a model that is both effective at the retrieval task defined above and robust to modality dropouts at test time.
+In keeping with previous work on the closely related problem of image retrieval, we focus on contrastive loss approaches, in which the goal is to learn an embedding of data where similar samples---in our case, samples belonging to the same class of object---are embedded `close' to one another, while dissimilar samples are farther apart.
+We develop a novel geometric loss function that simultaneously minimizes intra-class distances and maximizes inter-class distances across each pair of modalities yielding a model that is both effective at the retrieval task defined above and robust to modality dropouts at test time.
+We take an additional step and combine this geometric loss with a classification-based (cross-entropy) loss function which results in a superior model compared to either of geometric or cross-entropy losses alone.
 
-\subsection{Distance-based Loss}
+\subsection{Geometric Alignment Loss}
 \label{ssec:distanceloss}
 
-For $M$ modalities, we define a distance-based loss function which can be used for arbitrary number of modalities.
-Our proposed method is an extension of the well-known similarity-based triplet loss~\cite{Carvalho-cooking-triplet,triplet_loss_2021_CVPR}, and is similar to contrastive loss~\cite{chen2020simple,NEURIPS2020_supervised_contrastive} under some settings.
+% For $M$ modalities, we 
+We define a distance-based loss function which can be used for arbitrary number of modalities.
+Our proposed method is inspired by the well-known similarity-based triplet loss~\cite{Carvalho-cooking-triplet,triplet_loss_2021_CVPR}, and is similar to contrastive loss~\cite{chen2020simple,NEURIPS2020_supervised_contrastive} under some settings.
 Triplet loss-based learning works by forcing similar concepts from different domains `together' in some shared embedding space, while forcing dissimilar concepts `apart.' It is so named because it relies on three data points from the training set: a positive, a negative, and an anchor point.
-In our setting, the anchor and positive instances are from the same class, but in different modalities.
-
 However, standard triplet loss cannot be used for more than two modalities.
-Therefore, we extend this concept as follows.
-During training, we sample two sets of data points from each modality---one positive set (referring to a specific object) and one negative set (referring to some different object).
+Therefore, we modify this concept as follows.
+During training, we sample two different instances and their corresponding representations from all modalities into two sets---one positive set (referring to a specific object) and one negative set (referring to some different object) as shown in \cref{fig:emma-loss}.
 Unlike some prior triplet loss methods~\cite{GoLD_UMBC,triplet_loss_2021_CVPR}, the anchor is not randomly chosen from different modalities in each batch.
-Instead, we choose one \textit{domain} as an anchor (in our experiments, language).
-We consider the following formulation:
-
+Instead, in our setting, every item in the positive set becomes anchor once, and we minimize the distance between that item and other items in the positive set, while minimizing the distance between that item and all items in the negative set.
+It can be seen as an one-to-many relationship instead of an one-to-two relationship in the triplet loss formulation.
+% the anchor and positive items are from the same object class, but in different modalities and each of the positives modalities becomes anchor once. 
+We consider the following terminology:
+% \todokdinline{I feel like using the terminology of triplet loss especially \textbf{anchor} is confusing here or maybe it needs to be rephrased.
+We can alternatively say that we consider all pairwise distances.}
+% \todoffinline{I don't see the problem with anchor, but if you'd rather use a different term, what about query?
+As a note, while you do consider all pairwise distances but you treat some differently than others.}
 \begin{itemize}
@@
     \item Negative (Instance): A set of embeddings of another data point of a different object (e.g., an RGB image of a book, corresponding depth image, text description, and speech description of the same book)
-    \item Anchor (Modality): A fixed modality of the positive set is chosen as the learning anchor.
-For our experiments, text is used.
+    \item Anchor (Modality): Every modality of the positive set is chosen as the learning anchor once.
+In our formulation, anchor is a ``concept'' referring to the points in the positive sets rather than a single instance in itself.
+    %For our experiments, text is used. 
     The anchor is used as the basis for learning distances between positive and negative samples.
@@
 
-The objective is then to first minimize the distance between the anchor and each of the positive points from heterogeneous modalities; second, maximize the distance between the anchor and each of the negative points from heterogeneous modalities; and finally, maximize the distance between positive and negative anchor points from the same modality.
-Altogether, our proposed loss function contains $2M-1$ terms: $(M-1)$ anchor-to-positive distance minimizations, $(M-1)$ anchor-to-negative distance maximizations, and 1 positive-to-negative distance maximization.
-\Cref{eq:objective-emma} shows this objective function.
+The objective is then to first minimize the distance between each pair of positive points from heterogeneous modalities, and second, maximize the distance between each positive and negative points from all modalities.
+%and finally minimize the distance between each pair of negative points from heterogeneous modalities.
+% The objective is then to first minimize the distance between the anchor and each of the other positive points from heterogeneous modalities, and second, maximize the distance between the anchor and each of the negative points from all modalities.
+%and finally, maximize the distance between positive and negative anchor points from the same modality. 
+% When we minimize the distance between each pair of positives, and maximize the distance between each pair of modalities one from the positive set and one from the negative set, the model converges faster than \supcon{} while maintaining the same performance when we have all modalities and a little bit less than Supcon when we drop text.
+We refer to this approach as \geom{}, for extended multimodal alignment which is formulated in \cref{eq:full-emma}. An illustration of this loss function is provided in \cref{fig:emma-loss}.
+
+% \begin{equation}\label{eq:full-emma}
+%     L = \sum_{m_1=1}^{M} \left[ \sum_{m_2=1}^{M} \left[ \max (\cos(z_{m_1}^{+}, z_{m_2}^{-}),0) \right] + \sum_{m_3=m_1+1}^{M} \left[ 1 - \max(\cos(z_{m_1}^{+},z_{m_3}^{+}),0) \right] \right]
+% \end{equation}
+\begin{equation}\label{eq:full-emma}
+    L = \sum_{m_1=1}^{M} \left[ \sum_{m_2=1}^{M} \left[ \max (\cos(z_{m_1}^{+}, z_{m_2}^{-}) -1 + \alpha,0) \right] + \sum_{m_3=m_1+1}^{M} \left[ \max(1 - \cos(z_{m_1}^{+},z_{m_3}^{+}),0) \right] \right]
+\end{equation}
+
+In \cref{eq:full-emma}, $M$ is the number of modalities, the superscripts $+$ and $-$ represent positive and negative objects, $\alpha$ represents enforced margin between each positive and negative points which we set to 0.4 for all modalities without tuning, and $z$ is the embedding we get by applying a mapping function $f$, which in our case is a neural network on our input data.
+In other words, $z_m = f_m(x_m)$, where each modality $m$ has a specific model $f_m$ that is different from the models for other modalities.
+These models do not share their weights.
+The $\cos(\cdot)$ function is a measure of similarity, not distance, and that is why the signs are reversed.
 
@@
 Cosine similarity is a good choice for high-dimensional data as it is bounded between -1 and 1.
-Other distances, such as Euclidean, grow in value with respect to their dimensionality, resulting in a very large distances for data points.
+Other distance metrics, such as Euclidean distance, grow in value with respect to their dimensionality, resulting in a very large distances for data points.
 Cosine similarity is the opposite of distance, and we need to reverse the logic for maximization and minimization.
 
-\subsection{EMMA: Extended Multimodal Alignment}
-\label{sub:emma}
+Altogether, our proposed \geom{} function contains 
+$3M^2-M/2$ terms: 
+$M(M-1)/2$ anchor-to-positive distance minimizations and $M^2$ anchor-to-negative distance maximizations.
+%, and $M(M-1)/2$ \textcolor{red}{optional} negative-to-negative distance minimizations.
+%1 positive-to-negative anchor distance maximization.
+It is noteworthy that our training procedure does not perform any stochastic dropout of modalities to obtain test-time robustness to missing modalities.
+Moreover, our approach does not need to compute the distance between all items in the batch as opposed to \supcon{}~\cite{NEURIPS2020_supervised_contrastive}.
+
+\begin{figure*}[tb]
+\centering
+% \includegraphics[width=.9\columnwidth]{Figures/4way-MMA.png}
+% \includegraphics[width=2.1\columnwidth]{Figures/e-MMA.png}
+\includegraphics[width=1\columnwidth]{Figures/emma-loss.pdf}
+\caption{A high-level prototype of our approach and the distances used in the \geom{} loss with four modalities.
+Gray arrows indicate pre-trained models that are frozen (i.e.
+the parameters are fixed and are not trained).
+The black arrows show 3 fully connected layers with a ReLU activitation function~\cite{relu2010} after the first two layers.
+These networks are trained.
+Orange rectangles are negative embeddings and green rectangles are positive embeddings.
+Dashed lines indicate distances to be maximized while dotted lines indicate distances to be minimized.%, and solid lines show the enforced margin.
+}
+\label{fig:emma-loss}
+\end{figure*}
+
+\subsection{Loss Development Story}
+\todokdinline{Should I move this whole section to supplementary or should I talk about the high-level takeaways in the main paper?}
+
 Our proposed loss function is a modification of the triplet loss function idea where we fix one modality as anchor and do not sample a positive instance from the same modality. 
 Na\"ively, to extend triplet loss to work with arbitrary number of modalities, we can use two anchors (e.g., text and speech), which results in $2(M-2)$ triplet losses as formulated in~\cref{eq:objective-two-anchors}.
+\todoffinline{in this equation block, you're writing things like $|| z_{t}^{+} - z_{m}^{+} ||$. Do you actually mean norm, or would any old scoring function work?}
+\todokdinline{This whole section will be moved to supplementary, but anyway, would it be better if I just get rid of the $||.||$ and use $\cos(x)$ instead?}
+
 \begin{equation}
@@
 \end{equation}
+
 However, this approach performs poorly when one of the anchor modalities is ablated; since there is no explicit minimization between anchors, they do not necessarily map closely to each other, such that only one of the two anchors is actually learned.
@@
 \Cref{eq:objective-simple-mma} formulates this idea.
+
 \begin{equation}
@@
 \end{equation}
+
 However, this approach fails to take into account the dissimilarity of the positive and negative anchor points.
@@
 Therefore, we add an extra term which is responsible for maximizing the distance between positive and negative instances of the anchor modality, because a margin between positive and negative points from other modalities (excluding anchor) is enforced.
-The first dashed line from the top in \cref{fig:emma-loss} is this extra term that captures the dissimilarity between positive and negative text datapoints.
-We formulate this method in \cref{eq:objective-emma}, which is generalized to an arbitrary number of modalities. We refer to this approach as EMMA, for extended multimodal alignment.
+% The first dashed line from the top in \cref{fig:emma-loss} is this extra term that captures the dissimilarity between positive and negative text datapoints.
+We formulate this method in \cref{eq:objective-explicit-fixed-anchor}, which is generalized to an arbitrary number of modalities.
+
 \begin{equation}
-\label{eq:objective-emma}
+\label{eq:objective-explicit-fixed-anchor}
 \begin{split}
     \mathcal{L}  %&= %\sum_{m=1}^{M-1} || z_{a}^{+} - z_{m}^{+} || - || z_{a}^{+} - z_{m}^{-} ||  - || z_{a}^{+} - z_{a}^{-} || \\
-    &= \cos(z_{a}^{+}, z_{a}^{-}) + \sum_{m=1}^{M-1} \left(\cos(z_{a}^{+} ,z_{m}^{-}) - \cos(z_{a}^{+}, z_{m}^{+}) + \alpha_m\right) 
+    &= \max(\cos(z_{a}^{+}, z_{a}^{-}) + \alpha_a, 0) + \sum_{m=1}^{M-1} \max \left(\cos(z_{a}^{+} ,z_{m}^{-}) - \cos(z_{a}^{+}, z_{m}^{+}) + \alpha_m, 0 \right) 
 \end{split}
 \end{equation}
-In \cref{eq:objective-emma}, $a$ represents the \textit{anchor} modality, $M$ is the number of modalities, the superscripts $+$ and $-$ represent positive and negative objects, $\alpha_m$ represents enforced margin for each modality which we set to 0.4 for all modalities without tuning, and $z$ is the embedding we get by applying a mapping function $f$, which in our case is a neural network on our input data.
+
+In \cref{eq:objective-explicit-fixed-anchor}, $a$ represents the \textit{anchor} modality, $M$ is the number of modalities, the superscripts $+$ and $-$ represent positive and negative objects, $\alpha_m$ represents enforced margin for each modality which we set to 0.4 for all modalities without tuning, and $z$ is the embedding we get by applying a mapping function $f$, which in our case is a neural network on our input data.
 In other words, $z_m = f_m(x_m)$, where each modality $m$ has a specific model $f_m$ that is different from the models for other modalities.
 These models do not share their weights.
-The $\cos(\cdot)$ function is a measure of similarity, not distance, and that is why the signs are reversed in the second line.
-An illustration of this loss function is provided in \cref{fig:emma-loss}.
+The $\cos(\cdot)$ function is a measure of similarity, not distance, and that is why the signs are reversed.
 
-Since the downstream task in the grounded language learning domain is to predict/retrieve a desired object among multiple objects given a natural language description, it makes sense to train the model using language as an anchor, and in fact anchoring learning around language outperforms anchoring on RGB.
-Triplet loss cannot be used for more than 2 modalities.
+Since the downstream task we are considering in the grounded language learning domain is to predict/retrieve a desired object among multiple objects given a natural language description, it makes sense to train the model using language as an anchor, and in fact anchoring learning around language outperforms anchoring on RGB.
+However, if use all modalities as anchor once (looping over modalities and compute distances between each pair), it results in a better performance.
+
+% Triplet loss cannot be used for more than 2 modalities.
 Some previous work has concatenated RGB and depth embeddings to create a single ``vision'' embedding for learning~\cite{triplet_loss_2021_CVPR}, but they cannot handle RGB or depth sensor ablation during test.
@@
 % Moreover, when using triplet loss with different anchors per batch, the batch size has to be 1 which makes training very slow.
-This approach is most similar to that of supervised contrastive learning~\cite{NEURIPS2020_supervised_contrastive}, but outperforms both that and regular contrastive learning~\cite{chen2020simple} methods when modalities are ablated, most notably written language.
 
-It is noteworthy that our training procedure does not perform any stochastic dropout of modalities to obtain test-time robustness to missing modalities.
-We find that the use of instance level losses grounded to one key modality (text) is sufficient to encourage a representation that satisfies our goals.
+% This approach is most similar to that of supervised contrastive learning~\cite{NEURIPS2020_supervised_contrastive}, but outperforms both that and regular contrastive learning~\cite{chen2020simple} methods when modalities are ablated, most notably written language.
+
+% We find that the use of instance level losses grounded to one key modality (text) is sufficient to encourage a representation that satisfies our goals.
+
+\begin{comment}
+\subsubsection{Extended Triplet}
+Using an extended version of triplet loss to add all connections, which is formulated in \cref{eq:extended-triplet}, results in a worse performance.
+
+\begin{equation}\label{eq:extended-triplet}
+    L = \sum_{m_1=1}^{M} \left[ \sum_{m_2=m_1+1}^{M} \left[ \text{triplet}(z_{m_1}^{+}, z_{m_2}^{+}, z_{m_2}^{-}) + \text{triplet}(z_{m_1}^{-}, z_{m_2}^{-}, z_{m_2}^{+}) \right] + \cos(z_{m_1}^{+}, z_{m_1}^{-}) -1 + margin \right]
+\end{equation}
+Didn't change much when we drop text: Full eMMA when we have similar objects
+
+Did not work: Run the version with all connections (including among negatives) without triplet and just by using cosine distance between pair of embeddings
+\end{comment}
+
+\begin{comment}
+\subsubsection{Smaller Batch Sizes}
+
+Reducing the batch size from 64 to 16 (for both eMMA and supcon) increases the gap between our model and the baseline in cases where our model beats the baseline (all modalities, dropping speech), but baseline is still better than ours when we drop the text modality.
+
+Batch size of 8 (for both eMMA and supcon) had a similar effect.
+
+In batch size of 2 the trend is still the same, but when we drop the text modality, the MRR graph is flat on 0.4 which is still better than worse case and random case, but it seems model doesn't learn anything with regards to speech.
+To remedy this, we add pulling connections among the modalities of the negative instance as formulated in \cref{eq:full-emma-pull-negatives} and as shown in \cref{fig:full-emma-pull-negatives}, and that solves the flatline problem when we drop the text modality, however, when we have all modalities it performs worse than formulation in \cref{eq:full-emma}.
+
+\begin{equation}\label{eq:full-emma-pull-negatives}
+\begin{split}
+    L = & \sum_{m_1=1}^{M} \left[ \sum_{m_2=1}^{M} \left[ \max (\cos(z_{m_1}^{+}, z_{m_2}^{-}) -1 + margin,0) \right] + \sum_{m_3=m_1+1}^{M} \left[ \max(1 - \cos(z_{m_1}^{+},z_{m_3}^{+}),0) \right] \right.
+\\
+    & \left. + \sum_{m_3=m_1+1}^{M} \left[ \max(1 - \cos(z_{m_1}^{-},z_{m_3}^{-}),0) \right] \right]
+\end{split}
+\end{equation}
 
@@
 \centering
-% \includegraphics[width=.9\columnwidth]{Figures/4way-MMA.png}
-\includegraphics[width=2.1\columnwidth]{Figures/e-MMA.png}
-\caption{A high-level prototype of our approach and the distances used in the MMA loss with four modalities.
-Gray arrows indicate pre-trained models that are frozen (i.e.
-the parameters are fixed and are not trained).
-The black arrows show 3 fully connected layers with a ReLU activitation function~\cite{relu2010} after the first two layers.
-These networks are trained.
-Orange vectors are negative data points, green vectors are positive data points, and the blue vector is the anchor (which is the text modality of the positive data points).
-Dashed lines indicate distances to be maximized, dotted lines indicate distances to be minimized, and solid lines show the enforced margin.
-}
-\label{fig:emma-loss}
+\includegraphics[width=1\columnwidth]{Figures/full-emma-pull-negatives.pdf}
+\caption{A high-level prototype of our approach and the distances used when we have full EMMA and pulling among negatives}
+\label{fig:full-emma-pull-negatives}
 \end{figure*}
+\end{comment}
+
+\begin{comment}
+\subsubsection{Margin}
+% \todokdinline{Try different margins of 0.6, 1.0, and if it improved the results we can try to learn it as a parameter}
+\todokdinline{Should we remove this section?}
+I also tried two different margins so far. A margin of 0.6 and a margin of 1.0. Margin of 0.6 does not change the performance much. However, it converges even faster.
+Margin of 1.0 (this is essentially just cosine similarity because we have: cos -1 +margin) results in a worse performance when we have all modalities.
+The MRR is 0.8995 when margin is 1, while with a margin of 0.6 the MRR is 0.9176, and with a margin of 0.4 the MRR is 0.9198.
+If we drop the text modality, the MRR is 0.7923, 0.7913, and 0.6918 for margins of 0.4, 0.6, and 1.0 respectively.
+\end{comment}
+
+\subsection{EMMA: Extended Multimodal Alignment}
+\label{sub:emma}
+
+\todokdinline{We introduce \supcon{} in Baselines section which is after this section while we use SupCon as a part of our loss function here.
+Should we move the introduction of SupCon to here?
+Or maybe we can discuss the baselines first, and then Approach section follows it?}
+\todocmi{I suggest a section 4.3, Baseline Approach: Supervised Contrastive Learning (pushing EMMA down)}
+
+The main difference between \geom{} and \supcon{} is that in \geom{} we focus on the geometric interpretation of similarity using cosine distances, while in \supcon{}, cosine distances are used to compute a classification-based loss function similar to cross-entropy.
+Both methods are valid and have some advantages that the other method does not offer.
+In \geom{}, the advantages are intuitive learning objective in terms of distance, interpretability of the learned embedding space, and faster convergence.
+The advantage of \supcon{} is that it uses classification objective which is aligned with the downstream task.
+
+We propose to combine the \geom{} and \supcon{} and we refer to this approach as \ours{}, for extended multimodal alignment. \ours{} loss function is formulated in \cref{eq:emma}.
+
+\begin{equation}\label{eq:emma}
+    % \mathcal{L} = \geom{} + \supcon{}
+    \ours{} = \geom{} + \supcon{}
+\end{equation}
+
+Combining these two losses results in faster convergence and better performance when we have all modalities available, and maintaining a good performance when modalities are missing.
+
+\subsection{High-Level Findings/Observations}
+We find that using each modality as anchor results in a better performance compared to when we fix one modality as anchor. This difference is more significant when we test with all modalities.
+
+While the performance of SupCon deteriorates as the batch size is reduced, our \geom{} stays the same when we have all modalities or when we drop speech.
+However, when we drop text, EMMA deteriorate when we decrease batch size, and with a batch size of 2 without pulling negatives, it flatlines.
+In this case (when batch size is 2 and text is dropped), pulling among negatives is important and results in a better MRR.
+
+We discuss the geometric interpretation of our method versus the classification interpretation of SupCon method and advantages of each.
+Finally, we show that combining \geom{} with \supcon{} is superior to either the \geom{} approach or \supcon{} alone.
+
+\begin{comment}
+\subsubsection{EMMA + Binary Classification}
+\todokdinline{Should we remove this section?}
+
+When batch size is 64, it performs worse than baseline (SupCon) and EMMA itself. The experiment for the batch size of 2 also performs worse than other versions of EMMA, except for when we drop text.
+
+There is a slight chance that I have made a mistake in my implementation of binary cross-entropy loss. My guesses (in decreasing order of likelihood) are:
+
+1.
+The inputs to the BCE loss are $sigmoid(\cos_{sim}(z_i,z_j))$ for all $i \neq j$ where i is over the positive modalities only, while j is over both positive and negative modalities.
+Maybe I should've used $1-\cos(pos_i, pos_j)$ for similar embeddings and $cos(pos_i, neg_j) -1$ for dissimilar embeddings instead; similar to how EMMA's geometric loss is structured.
+
+2.
+Let's assume $\text{predicts} = \text{sigmoid}(\cos_{sim}(z_i,z_j))$.
+Their corresponding labels/targets are 1 if i and j are from positive modalities, and 0 if j is from the negative modalities.
+The way I used PyTorch's BCE is that I passed the predicts whose targets are 1 to BCE first, passed the predicts whose targets are 0 to BCE in the next step, and finally added the two together.
+In other words:
+
+\begin{equation}
+\begin{aligned}    
+loss &= 0.0 \\
+loss &= loss + BCE(\text{similar predicts, targets}=1) \\
+loss &= loss + BCE(\text{dissimilar predicts, targets}=0)
+\end{aligned}
+\end{equation}
+
+Adding temperature gives mixed results; when we have all the modalities during the test, adding temperature improves performance over not having a temperature, but is still slightly worse than other methods.
+When we drop the text modality, however, the performance goes down.
+\end{comment}
 
@@
 
-Transformers are de facto architectures in natural langauge processing community and have shown great success across different tasks.
+Transformers are de facto architectures in natural language processing community and have shown great success across different tasks.
 Similar to~\citet{GoLD_UMBC}, we use BERT~\cite{devlin-etal-2019-bert} embeddings contained in the FLAIR library ~\cite{akbik2019flair,akbik-etal-2019-pooled} to featurize the textual input, and wav2vec2~\cite{wav2vec2} to extract audio embeddings from speech.
@@
 To process images, we use ResNet152~\cite{He_resnet_2016} for both RGB and depth images which gives us a 2048-dimensional embedding vector. Depth images are colorized before passing to the ResNet152.
-We then use 3 fully connected layers with ReLU activation~\cite{relu2010} to map these embeddings to a shared 1024-dimensional space where we can compute the distances between all embeddings.
+
+We then use different instantiations of the same multi-layer perceptron (MLP) consisting of 3 fully connected layers with ReLU activation~\cite{relu2010} to map each of these embeddings to a shared 1024-dimensional space where we can compute the distances between all embeddings.
+We note that these MLP networks are different and do not share any weights.
 
@@
 
-We demonstrate the effectiveness of our approach on a recent multimodal dataset called GoLD~\cite{GoLD_UMBC}, which contains RGB image, depth image, written text descriptions, speech descriptions, and transcribed speech descriptions for 207 object instances across 47 object classes (see \cref{fig:emma-loss}).
+We demonstrate the effectiveness of our approach on a recent publicly available multimodal dataset called GoLD~\cite{GoLD_UMBC}, which contains RGB image, depth image, written text descriptions, speech descriptions, and transcribed speech descriptions for 207 object instances across 47 object classes (see \cref{fig:emma-loss}).
 There are a total of 16,500 speech and 16,500 textual descriptions.
@@
 
-\subsection{Problem Setup and Metrics}
-\label{sec:metrics}
+\subsection{Setup and Metrics}
+\label{sec:setupmetrics}
 To evaluate our model we measure different performance metrics on a retrieval task where the model has to select an object among 5 objects given a language description.
@@
 Similar to~\citet{NEURIPS2020_supervised_contrastive}, we use a stochastic gradient descent (SGD) optimizer with momentum~\cite{ruder2016overviewSGD} with a flexible learning rate starting at 0.05. 
-For the triplet loss experiments we use the Adam~\cite{kingma_adam_2015} optimizer with scheduling learning rate as suggested by the papers.
-The models are trained for 200 epochs with a batch size of 64 on a Quadro RTX 8000 GPU.
+% For the triplet loss experiments we use the Adam~\cite{kingma_adam_2015} optimizer with scheduling learning rate as suggested by the papers.
+All models are trained for 200 epochs with a batch size of 64 on a Quadro RTX 8000 GPU.
 
 To evaluate the performance, we compute the distance between the given natural language description and 5 randomly selected images (1 of which corresponds to the description, with the others from different object classes).
-We compute the distance between one textual language embedding and all candidate RGB embeddings, and we compute the distance between the same language embedding and all candidate depth embeddings corresponding to the RGB embeddings.
+We compute the distance between one language embedding and all candidate RGB embeddings, and we compute the distance between the same language embedding and all candidate depth embeddings corresponding to the RGB embeddings.
 We then take average of these two distance matrices.
-Instead of choosing an empirical threshold beyond which objects are considered to be `referred to,' we choose the closest image embedding (average distance of RGB and depth from language) as the prediction.
+Instead of choosing an empirical threshold beyond which objects are considered to be `referred to,' we choose the closest image embedding (average distance of RGB and/or depth from language) as the prediction.
 In order to use cosine \textit{distance}, we have to subtract the cosine of the \textit{angle} between two embeddings (which represents similarity) from 1: that is, we compute $1 - \cos(e_1, e_2)$.
@@
 \subsubsection{Metrics}
-
+\label{sec:metrics}
 The best metric to capture the performance in such a scenario is mean reciprocal rank (MRR, \cref{eq:mrr} for $Q$ queries).
@@
 \subsection{Baselines}
-We compare our EMMA model against different contrastive learning methods including triplet loss, the traditional version of contrastive loss usually used in self-supervised settings~\cite{chen2020simple}, and supervised contrastive learning~\cite{NEURIPS2020_supervised_contrastive}.
+We compare both \ours{} and \geom{} against contrastive learning~\cite{chen2020simple} and supervised contrastive learning~\cite{NEURIPS2020_supervised_contrastive} which we refer to as \supcon{}.
+We consider \supcon{} as the main baseline since it is a general version of multiple contrastive loss functions including triplet loss, the traditional version of contrastive loss usually used in self-supervised settings~\cite{chen2020simple}, and N-pair loss~\cite{NIPS2016N-PairLoss}.
+% and do not include N-pair and triplet loss in our comparisons.
+%contrastive learning methods including triplet loss, the traditional version of contrastive loss usually used in self-supervised settings~\cite{chen2020simple}, and supervised contrastive learning~\cite{NEURIPS2020_supervised_contrastive}.
 
@@
 \centering
+\resizebox{.97\textwidth}{!}{
 \begin{tabular}{lc|ll|lll}
@@
 \multicolumn{1}{c|}{\textbf{Method}} & \textbf{Modalities} & \multicolumn{1}{c|}{\textbf{\begin{tabular}[c]{@{}c@{}}Speech/RGB/\\ Depth\end{tabular}}} & \multicolumn{1}{c|}{\textbf{\begin{tabular}[c]{@{}c@{}}Text/RGB/\\ Depth\end{tabular}}} & \multicolumn{1}{c|}{\textbf{\begin{tabular}[c]{@{}c@{}}Speech/RGB/\\ Depth\end{tabular}}} & \multicolumn{1}{c|}{\textbf{\begin{tabular}[c]{@{}c@{}}Text/RGB/\\ Depth\end{tabular}}} & \multicolumn{1}{c}{\textbf{\begin{tabular}[c]{@{}c@{}}Text/Speech/\\ RGB/Depth\end{tabular}}} \\ \hline
-\multicolumn{1}{l|}{EMMA} & 4 & \multicolumn{1}{l|}{\textbf{0.65}$\pm$0.01} & 0.83$\pm$0.006 & \multicolumn{1}{l|}{\textbf{0.79}$\pm$0.002} & \multicolumn{1}{l|}{0.90$\pm$0.003} & 0.90$\pm$0.004 \\
+\multicolumn{1}{l|}{EMMA} & TSRD & \multicolumn{1}{l|}{\textbf{0.65}$\pm$0.01} & 0.83$\pm$0.006 & \multicolumn{1}{l|}{\textbf{0.79}$\pm$0.002} & \multicolumn{1}{l|}{0.90$\pm$0.003} & 0.90$\pm$0.004 \\
 \multicolumn{1}{l|}{EMMA} & 3 & \multicolumn{1}{l|}{--} & 0.84$\pm$0.008 & \multicolumn{1}{l|}{--} & \multicolumn{1}{l|}{0.90$\pm$0.004} & -- \\
-\multicolumn{1}{l|}{Supervised Contrastive} & 4 & \multicolumn{1}{l|}{0.59$\pm$0.007} & 0.84$\pm$0.008 & \multicolumn{1}{l|}{0.75$\pm$0.004} & \multicolumn{1}{l|}{0.91$\pm$0.006} & 0.90$\pm$0.006 \\
-\multicolumn{1}{l|}{Supervised Contrastive} & 3 & \multicolumn{1}{l|}{--} & 0.84$\pm$0.006 & \multicolumn{1}{l|}{--} & \multicolumn{1}{l|}{0.91$\pm$0.004} & -- \\
-\multicolumn{1}{l|}{Contrastive Cosine} & 4 & \multicolumn{1}{l|}{0.40$\pm$0.153} & 0.83$\pm$0.006 & \multicolumn{1}{l|}{0.62$\pm$0.034} & \multicolumn{1}{l|}{0.90$\pm$0.004} & 0.89$\pm$0.005 \\
+\multicolumn{1}{l|}{\supcon{}} & TSRD & \multicolumn{1}{l|}{0.59$\pm$0.007} & 0.84$\pm$0.008 & \multicolumn{1}{l|}{0.75$\pm$0.004} & \multicolumn{1}{l|}{0.91$\pm$0.006} & 0.90$\pm$0.006 \\
+\multicolumn{1}{l|}{\supcon{}} & 3 & \multicolumn{1}{l|}{--} & 0.84$\pm$0.006 & \multicolumn{1}{l|}{--} & \multicolumn{1}{l|}{0.91$\pm$0.004} & -- \\
+\multicolumn{1}{l|}{Contrastive Cosine} & TSRD & \multicolumn{1}{l|}{0.40$\pm$0.153} & 0.83$\pm$0.006 & \multicolumn{1}{l|}{0.62$\pm$0.034} & \multicolumn{1}{l|}{0.90$\pm$0.004} & 0.89$\pm$0.005 \\
 \multicolumn{1}{l|}{Contrastive Cosine} & 3 & \multicolumn{1}{l|}{--} & 0.84$\pm$0.005 & \multicolumn{1}{l|}{--} & \multicolumn{1}{l|}{0.91$\pm$0.003} & -- \\
-\multicolumn{1}{l|}{Triplet} & 3 & \multicolumn{1}{l|}{--} & 0.83$\pm$0.008 & \multicolumn{1}{l|}{--} & \multicolumn{1}{l|}{0.90$\pm$0.005} & -- \\ \hline
+\multicolumn{1}{l|}{\ours{}} & TSRD & \multicolumn{1}{l|}{--} & 0.83$\pm$0.008 & \multicolumn{1}{l|}{--} & \multicolumn{1}{l|}{0.90$\pm$0.005} & -- \\ \hline
 \end{tabular}
+}
 \caption{Average and standard deviation of accuracy and MRR over 3 runs with 3 different random seeds on a held-out test set.
@@
 All models are trained using text as anchor.
-The batch size is 64 and optimizer is SGD for all experiments except the triplet loss, where batch size is 1 and optimizer is Adam.}
+The batch size is 64 and optimizer is SGD for all experiments except the triplet loss, where batch size is 1 and optimizer is Adam.
+%In the Triplet experiment, RGB and Depth embeddings are concatenated to form a single vision embedding since triplet loss cannot be used for more than two modalities.
+\todokdinline{Should we keep Acc or is MRR enough since table is out of bounds?
+Also, should we include other training modalities in the table?
+I inserted the results in a file called results-table.tex for now so that we can decide later.}
+}
+\todocmi{Let's include everything, and play tricks to shrink the table horizontally. If necessary it can be split into two tables with the same y headers and different x headers.}
+\todoffinline{I've added a resizebox in order to get this to fit.
+This is not the desired solution (though currently the text isn't that small, so it could be okay) but I needed to see the entire table easily.}
 \label{table:quantitative}
@@
 \label{sub:baseline-triplet}
+\todokdinline{Since we do not include triplet in our analysis, it might make more sense to move this section into the introduction of Approach to motivate why we came up with EMMA, thoughts?}
+
 The triplet loss function consists of three data points including anchor, positive, and negative from two modalities.
 The anchor, positive, and negative can be chosen from different modalities in each batch.
-The triplet loss method has two major disadvantages.
-First, it cannot be used for more than two modalities.
-Second, a batch size of greater than 1 is hard to implement if the anchor, positive, and negative come from different modalities for each batch.
+The major disadvantage of the triplet loss method is that it cannot be used for more than two modalities. 
+%Second, a batch size of greater than 1 is hard to implement if the anchor, positive, and negative come from different modalities for each batch.
 % To address these issues in EMMA, we sample more than one positive and negative data points.
-To address these issues in EMMA, we choose one modality as anchor.
-Intuitively, we sample one positive data point and one negative data point from each modality, resulting in 2M data points.
-The anchor data point is always from the same modality and corresponds to the same object class sampled for positive data points.
-\Cref{table:quantitative} shows that our MMA method perform as good as this baseline while out model can be trained 10x faster since we can use batch sizes larger than 1.
-Also, our method can be used for any number of modalities, while previous works~\cite{GoLD_UMBC,triplet_loss_2021_CVPR} concatenate rgb and depth to form a single vector to handle more than two modalites.
-When we used a fixed anchor modality, a batch size of 64, and applied semantic negative sampling~\cite{Pillai_Matuszek_2018}, our method outperformed this baseline.
-However, since this method cannot be applied to four modalities, we do not include it in the main analysis.
-Moreover, our method has the advantage that does not require triplet mining or providing hard examples.
+\todokd{If moved to approach section, build on this to introduce our geometric loss function.}
+To address this issue, in \geom{} we use pairwise distance optimization for all data points. 
+%Intuitively, we sample one positive data point and one negative data point, each of which includes $M$ modalities, resulting in 2M data points.
+Each of the $M$ positives modalities become an anchor once.
+% \Cref{table:quantitative} shows that our MMA method perform as good as this baseline while our model can be trained 10x faster since we can use batch sizes larger than 1. Also, 
+Our method can be used for any number of modalities, while previous works that use triplet loss~\cite{GoLD_UMBC,triplet_loss_2021_CVPR} concatenate RGB and depth to form a single ``vision'' vector to handle three modalities, but they cannot handle RGB or depth sensor ablation during test.
+Since our method can handle any number of modalities, we can handle such a case by treating depth as a separate modality.
+Moreover, our method has the advantage that does not require triplet mining or providing hard negative examples.
+% When we used a fixed anchor modality, a batch size of 64, and applied semantic negative sampling~\cite{Pillai_Matuszek_2018}, our method outperformed this baseline. However, 
+Since this method cannot be applied to four modalities, we do not include it in the main analysis.
 
@@
 
-We compare our model against contrastive loss.
+We compare our model against contrastive learning method presented by~\citet{chen2020simple} where they use the normalized
+temperature-scaled cross entropy loss (NT-Xent).
 In order to implement this loss function, we use cosine similarity as suggested in the SimCLR paper~\cite{chen2020simple}.
-Another possibility is to use an inner dot product~\cite{NEURIPS2020_supervised_contrastive}; this can lead to instabilities since the dot product is not bounded.
-This contrastive loss function is formulated in \cref{eq:contrastive-loss}:
+Another possibility is to use an inner dot product~\cite{NEURIPS2020_supervised_contrastive}; if not normalized, this can lead to numerical instabilities and overflow/underflow since the dot product is not bounded, but the result is the same whether we use normalized inner dot product or cosine similarity.
+Contrastive loss function is formulated in \cref{eq:contrastive-loss}:
 % \todokdinline{try contrastive loss with cross-entropy, dot product, and the supervised contrastive method itself (without triplets).}
@@
 
-\Cref{eq:contrastive-loss} can be extended to the loss function formulated in \cref{eq:mma-contrastive}.
-Our formulation is a generalized version of the contrastive loss where $M$ is the number of modalities (e.g.
-RGB image, depth image, speech, text), but it can be ``augmented views'' instead of different modalities.
+We can treat different modalities of the same instance as ``augmentations'' of that instance, and consider them as the positive points for the anchor.
+\Cref{eq:contrastive-loss} can be rewritten with the sum over positives explicitly shown as formulated in \cref{eq:explicit-contrastive}.
 
-\begin{equation}\label{eq:mma-contrastive}
-    -\sum_{m=1}^{M-1} \log \frac{\exp (sim(z_a^+ , z_{m}^{+})/ \tau) }{ \exp (sim(z_a^+ , z_{m}^{+}) / \tau) + \exp (sim(z_a^+, z_{m}^{-}) / \tau)}
+%, which is equivalent to ``augmented views'' in the contrastive loss terminology.
+
+\todokdinline{Notation for M - i, it should be something like other modalities of the anchor but it shouldn't be over number of modalities.}
+\begin{equation}\label{eq:explicit-contrastive}
+    - \sum_{i \in I} \sum_{p=1}^{M \setminus \{i\} } \log \frac{\exp (sim(z_i , z_{p})/ \tau) }{
+    \sum_{a \in A(i)} \exp (sim(z_i, z_a) / \tau)}
+    %{ \exp (sim(z_i^+ , z_{m}^{+}) / \tau) + \exp (sim(z_i^+, z_{m}^{-}) / \tau)}
 \end{equation}
-where $z = f(x)$.
+where $M$ is the number of modalities (e.g. RGB image, depth image, speech, text) and $z = f(x)$.
 
@@
 
-\citet{NEURIPS2020_supervised_contrastive} propose a supervised way of performing contrastive learning which is shown in \cref{eq:supervised-contrastive}.
+\citet{NEURIPS2020_supervised_contrastive} extend the contrastive learning method (NT-Xent) and propose a supervised way of performing contrastive learning to treat not only augmentations of the anchor, but also every item that shares the same label with anchor as positives.
+This loss function is shown in \cref{eq:supervised-contrastive}.
 
@@
 
-While this model is a strong baseline, if modalities are ablated, this method does not perform as well as our proposed model (we show and discuss this later, in \cref{fig:epochs-mrr.srd}).
+Although, this loss function do not use cosine similarity, they normalize the embeddings before performing dot product which is equivalent to cosine similarity.
+
+The main difference between the contrastive loss baseline in \cref{sub:baseline-contrastive} and \supcon{} is that there is no notion of meaningful negative points in contrastive loss, and everything in the batch that is not the anchor or one of the positive views is considered to be negative.
+In \supcon{}, however, all elements in the batch that have same label as the anchor, are also considered positives in addition to different views of the same instance.
+While the denominators of \cref{eq:explicit-contrastive,eq:supervised-contrastive} stay the same, this subtle difference affects the numerator and includes more positive examples.
+
+While this model is a strong baseline, the authors applied it to a unimodal dataset.
+In this paper we extend this baseline to work with a multimodal dataset and show that it is slower than \ours{} to learn and its performance is also lower when all modalities are available during test time.
+% if modalities are ablated, this method does not perform as well as our proposed model (we show and discuss this later, in \cref{fig:epochs-mrr.srd}).
 This is especially true when the text modality is dropped, as is likely in physical agent scenarios, where speech is a natural interaction mechanism.
@@
 \subsection{Modality Ablation}
+% \todokdinline{Should I remove this section altogether or should I just mention that our method is equal or better in all cases except when we drop text?
+Or should I spend more time to come up with something so that we can perform better than baseline even in this case?}
 We consider an experiment in which we incorporate RGB, depth, speech, and written language to train the model.
-The loss function requires no changes beyond increasing the value of $M$ by 1 in \cref{eq:objective-emma}.
+The loss function requires no changes beyond adjusting the value of $M$ in \cref{eq:full-emma} according to the number of modalities available during training.
 Our goal is the non-trivial downstream prediction task: determining what objects are being referred to by arbitrary language from a small number of examples.
-When we consider only text, RGB, and depth, written language is used as the anchor modality, and we compute the distance of RGB and depth modalities from it and then average them.
+When we consider only text, RGB, and depth, written language is used as the query modality, and we compute the distance of RGB and depth modalities from it and then average them.
 However, when speech is incorporated as an additional fourth sensory modality, we have three possible choices.
@@
 At test time, depending on which modalities are available to the model, we can use speech, text, or both to compute the distance of RGB and depth embeddings from the linguistic query, and then take the average.
-\Cref{eq:distance-matrix} shows all possible cases of modality dropout and the corresponding distance computations, where ``t'' represents \textit{text}, ``s'' represents \textit{speech}, ``r'' represents \textit{RGB}, ``d'' represents \textit{depth}, $K_{ij}$ for each pair of modalities is the distance of one instance in modality $i$ from all instances in modality $j$, and $K$ is the final distance---a matrix if there are multiple queries and a vector if there is one query.
+\Cref{eq:distance-matrix} shows all possible cases of modality dropout and the corresponding distance computations, where ``t'' represents \textit{text}, ``s'' represents \textit{speech}, ``r'' represents \textit{RGB}, ``d'' represents \textit{depth}, $K_{ij}$ for each pair of modalities is the distance of one instance in modality $i$ from all candidate instances in modality $j$, and $K$ is the final distance---a matrix if there are multiple queries and a vector if there is one query.
 \begin{equation}
@@
 
-\Cref{fig:epochs-mrr.srd,fig:epochs-mrr.trd,fig:epochs-mrr.sd,fig:epochs-mrr.sr} show the relative performance of EMMA against baselines when different modalities are ablated.
+\Cref{fig:epochs-mrr.lard,fig:epochs-mrr.srd,fig:epochs-mrr.ld,fig:epochs-mrr.lr} show the relative performance of \ours{} and \geom{} against baselines when different modalities are ablated.
 
@@
 \label{sec:results}
+
 In this section we provide quantitative and qualitative results by comparing our method against supervised contrastive learning~\cite{NEURIPS2020_supervised_contrastive} and na\"ive contrastive loss~\cite{chen2020simple}.
@@
 
-Comparing \cref{fig:epochs-mrr.trd} and \cref{fig:epochs-mrr.srd}, we can see that when the text modality is dropped the performance decreases from about 0.90 to about 0.79, showing that speech cannot completely replace the text.
-We also observe that speech has a higher variance compared to text, and a na\"ive contrastive learning method cannot take advantage of text modality to reduce this variance.
+\Cref{fig:epochs-mrr.lard} Shows that \ours{} learns faster and results in a better performance compared to both \supcon{} and contrastive learning~\cite{chen2020simple} when trained using all modalities and all modalities are available during test.
+We observe that not only does contrastive loss learn more slowly, but that it is prone to overfitting; while this can be addressed with careful tuning of the learning process, it is better not to have to.
+
+\begin{figure}[tbh]
+\centering
+\includegraphics[width=.99\columnwidth]{Figures/average-seeds-epochs-mrr_lard}
+\caption{Mean Reciprocal Rank (MRR) on the held-out test set when all modalities are available, averaged over 5 runs.
+Green is \ours{}, orange is \supcon{}, blue represents our \geom{} loss function, and red is the contrastive loss baseline.
+Higher is better.
+}
+\label{fig:epochs-mrr.lard}
+\end{figure}
+
+% Comparing \cref{fig:epochs-mrr.trd} and 
+When we drop the text modality in \cref{fig:epochs-mrr.srd}, we can see that the performance decreases from about 0.93 to about 0.82, showing that speech cannot completely replace the text.
+\Cref{fig:3d-tsne} depicts that speech modality is not aligned as good as the text modality and that is why when we drop text and use speech as the main query, the performance decreases.
+This supports our hypothesis that a geometric alignment of the latent space is crucial to a good performance for object retrieval and multimodal understanding.
+In \cref{fig:epochs-mrr.srd}, we observe that when speech is used as the query and text modality is ablated, \supcon{} baseline works slightly better than \ours{}, although \ours{} still learns faster.
+The reason is that \supcon{} optimizes for the classification task, and since the speech modality is not well aligned based on \cref{fig:3d-tsne}, using \geom{} makes the downstream task more difficult by trying to pull and push similar and dissimilar data points, respectively.
+More research needs to be done to come up with strategies to align more chaotic modalities.
+% We also observe that speech has a higher variance compared to text, and a na\"ive contrastive learning method cannot take advantage of text modality to reduce this variance.
 While the other two methods leverage text modality to align speech better.
 
-There is very little gap in performance when rgb is dropped in \cref{fig:epochs-mrr.sd} compared to when depth is dropped in \cref{fig:epochs-mrr.sr} which shows that our model is robust enough when rgb or depth sensors fail in real-world deployment.
+There is very little gap in performance when depth or RGB are dropped in \cref{fig:epochs-mrr.lr,fig:epochs-mrr.ld} compared to when we have all modalities in \cref{fig:epochs-mrr.lard} which shows that our model is robust enough when rgb or depth sensors fail in real-world deployment.
+Also, when depth is dropped in \cref{fig:epochs-mrr.lr} performance decreases less compared to when rgb is dropped in \cref{fig:epochs-mrr.ld} which suggests that depth is harder to work with compared to RGB.
 
@@
 \includegraphics[width=.99\columnwidth]{Figures/average-seeds-epochs-mrr_ard.pdf}
-\caption{Mean Reciprocal Rank (MRR) on the held-out test set when the text modality is ablated, averaged over 3 runs.
+\caption{Mean Reciprocal Rank (MRR) on the held-out test set when the text modality is ablated, averaged over 5 runs.
 Green is self-supervised contrastive learning, orange is supervised contrastive learning, and blue represents our proposed EMMA loss function.
 Higher is better.
-% \todo[inline]{Fix description of colors.}
 }
@@
 
+\begin{comment}
 \begin{figure}[tbh]
@@
 \includegraphics[width=.99\columnwidth]{Figures/average-seeds-epochs-mrr_lrd.pdf}
-\caption{Mean Reciprocal Rank (MRR) on the held-out test set when the speech modality is ablated, averaged over 3 runs, for the downstream task of object retrieval.
+\caption{Mean Reciprocal Rank (MRR) on the held-out test set when the speech modality is ablated, averaged over 5 runs, for the downstream task of object retrieval when all objects are unique.
 Green is self-supervised contrastive learning, orange is supervised contrastive learning, and blue represents our proposed EMMA loss function.
@@
 \end{figure}
+\end{comment}
 
@@
 \centering
-\includegraphics[width=.99\columnwidth]{Figures/average-seeds-epochs-mrr_ar.pdf}
-\caption{Mean Reciprocal Rank (MRR) on the held-out test set when text and depth modalities are ablated, averaged over 3 runs for the downstream task of object retrieval.
-Green is self-supervised contrastive learning, orange is supervised contrastive learning, and blue represents our proposed EMMA loss function.
+\includegraphics[width=.99\columnwidth]{Figures/average-seeds-epochs-mrr_lr.pdf}
+\caption{Mean Reciprocal Rank (MRR) on the held-out test set when speech and depth modalities are ablated, averaged over 5 runs for the downstream task of object retrieval when all objects are unique.
+Green is \ours{}, orange is \supcon{}, and blue represents our proposed \geom{} loss function.
 Higher is better.
-% Orange represents Supervised Contrastive Learning~\cite{NEURIPS2020_supervised_contrastive}, \Complete{} represents the triplet loss function~\cite{triplet_loss_2021_CVPR}.
 }
 % \todocmi{Maybe MRR is wrong and we should be looking at top-ranked result?}
-\label{fig:epochs-mrr.sr}
+\label{fig:epochs-mrr.lr}
 \end{figure}
@@
 \centering
-\includegraphics[width=.99\columnwidth]{Figures/average-seeds-epochs-mrr_ad.pdf}
-\caption{Mean Reciprocal Rank (MRR) on the held-out test set when text and RGB modalities are dropped averaged over 3 runs for the downstream task of object retrieval.
-Green is self-supervised contrastive learning, orange is supervised contrastive learning, and blue represents our proposed EMMA loss function.
+\includegraphics[width=.99\columnwidth]{Figures/average-seeds-epochs-mrr_ld.pdf}
+\caption{Mean Reciprocal Rank (MRR) on the held-out test set when speech and RGB modalities are dropped averaged over 5 runs for the downstream task of object retrieval when all objects are unique.
+Green is \ours{}, orange is \supcon{}, and blue represents our proposed \geom{} loss function.
 Higher is better.
 }
-\label{fig:epochs-mrr.sd}
+\label{fig:epochs-mrr.ld}
 \end{figure}
@@
 \centering
-\includegraphics[width=1.99\columnwidth]{Figures/3D-tsne-4M-eMMA-cosine-submodalities-text-anchor-gold-no_neg_sampling-1024-sbs.png}
-\caption{3D T-SNE~\cite{van2008tsne} projection of test embeddings of 10 randomly selected classes of objects using our proposed method EMMA.
+% \includegraphics[width=1.99\columnwidth]{Figures/3D-tsne-4M-eMMA-cosine-submodalities-text-anchor-gold-no_neg_sampling-1024-sbs.png}
+% \includegraphics[width=1\columnwidth]{Figures/3D-tsne-4M-eMMA-cosine-submodalities-text-anchor-gold-no_neg_sampling-1024-sbs.png}
+\includegraphics[width=1\columnwidth]{Figures/3D-tsne-exp-supcon-emma-lard-64-relu-SGD-0.001-unique_objects-gold-no_neg_sampling-1024-sbs.png}
+\caption{3D T-SNE~\cite{van2008tsne} projection of test embeddings of 10 randomly selected classes of objects using our proposed \ours{}.
 Each modality is separately projected into a three-dimensional space.
 In a perfect embedding, all instances of a class would be clustered in identical areas of the embedding space across all modalities.
-We can see that EMMA successfully encourages all four modalities to live in a common manifold, allowing accurate retrieval even when modalities are missing.
+We can see that \ours{} successfully encourages all four modalities to live in a common manifold, allowing accurate retrieval even when modalities are missing.
 }
@@
 % \label{sec:Qualitative}
-We look into the projection of the high-dimensional learned embeddings into a 2-dimensional and 3-dimensional spaces using t-SNE~\cite{van2008tsne}, which is a dimensionality reduction technique to visualize high-dimensional data.
-T-SNE creates a probability distribution over pairs high-dimensional data where similar pairs have a higher probability and dissimilar pairs have a lower probability.
+We look into the projection of the high-dimensional learned embeddings into a 
+%2-dimensional and 
+3-dimensional space using t-SNE~\cite{van2008tsne}, which is a dimensionality reduction technique to visualize high-dimensional data.
+T-SNE creates a probability distribution over pairs of high-dimensional data where similar pairs have a higher probability and dissimilar pairs have a lower probability.
 A similar probability distribution is also defined over pairs of data in the lower dimension (either 2D or 3D), and T-SNE minimizes the KL divergence between these two probability distributions.
@@
 \subsection{Discussion}
-Our proposed model performs well, has been demonstrated to handle four modalities of shared information effectively, and is robust to test-time situations where information from one or more modalities is missing.
+Our proposed model performs well and learns fast, has been demonstrated to handle four modalities of shared information effectively, and is robust to test-time situations where information from one or more modalities is missing.
+
 However, there is still room for improvement.
@@
 \centering
-\includegraphics[width=.8\columnwidth]{Figures/example-rankings.png}
-\caption{EMMA more accurately ranks objects and handles ambiguities with respect to a retrieval query.}
+\includegraphics[width=1\columnwidth]{Figures/example-rankings.pdf}
+\caption{\ours{} more accurately ranks objects and handles ambiguities with respect to a retrieval query compared to \supcon{}.
+Note that the phrase ``light blue'' in the query is very similar to ``light bulb'' and while \supcon{} confuses this and predicts light bulb as the correct object, \ours{} correctly identifies can opener as the intended object, and light bulb is ranked as the second object.
+}
 \label{fig:rankings}
@@
 
-In this work, we have demonstrated the effectiveness of a novel approach to learning from high-dimensional multimodal information.
-Our approach performs well on an object retrieval task from a testbed that contains four separate modalities, consistent with information that might be available to a physical agent, and outperforms state of the art contrastive learning approaches when one or more modalities is unavailable at test time.
+In this work, we have demonstrated the effectiveness of a novel approach to learning from high-dimensional multimodal information even when one or more modalities is unavailable at test time.
+Our approach performs well on an object retrieval task from a testbed that contains four separate modalities, consistent with information that might be available to a physical agent, and outperforms state of the art contrastive learning approaches.
 Our proposed method is general enough to be applied to a variety of multimodal retrieval problems, and is not limited to purely language-based image retrieval.
+\todokdinline{Don't we need evidence for that?}
 
@@
 
-% %% The acknowledgments section is defined using the "acks" environment
-% %% (and NOT an unnumbered section). This ensures the proper
-% %% identification of the section in the article metadata, and the
-% %% consistent spelling of the heading.
-\begin{acks}
-To Robert, for the bagels and explaining CMYK and color spaces.
-\end{acks}
-% \section*{Acknowledgments}
+\bibliography{tmlr}
+\bibliographystyle{tmlr}
 
-\clearpage
-\clearpage
-% %%
-% %% The next two lines define the bibliography style to be used, and
-% %% the bibliography file.
-\bibliographystyle{ACM-Reference-Format}
-\bibliography{ref}
-
-% %%
-% %% If your work has an appendix, this is the place to put it.
 \appendix
+\section{Appendix}
 \label{sec:appendix}

### 2022-08-17 Paper/results-table.tex
first draft of paper for TMLR

method: Geometric, best mrr_ad ** avg: 0.7682305107526883 std: 0.0033724197709617586
method: Geometric, best mrr_ar ** avg: 0.783393817204301 std: 0.002870980798921243
method: Geometric, best mrr_ld ** avg: 0.8963622311827957 std: 0.0038165384708184025
method: Geometric, best mrr_lr ** avg: 0.9112627688172041 std: 0.0073097134335746074
method: Geometric, best mrr_lad ** avg: 0.8920712365591397 std: 0.004478468907650514
method: Geometric, best mrr_lar ** avg: 0.9095463709677419 std: 0.008254095620938519
method: Geometric, best mrr_ard ** avg: 0.7937331989247312 std: 0.002892081483060137
method: Geometric, best mrr_lrd ** avg: 0.9229038978494624 std: 0.005072886474275548
method: Geometric, best mrr_lard ** avg: 0.9214092741935485 std: 0.004504231901153245
method: Geometric, best acc_ad ** avg: 0.6194758064516129 std: 0.005500153720300029
method: Geometric, best acc_ar ** avg: 0.6434274193548386 std: 0.005332041933055135
method: Geometric, best acc_ld ** avg: 0.8202822580645162 std: 0.0057175656281111785
method: Geometric, best acc_lr ** avg: 0.8459677419354839 std: 0.010990399982684721
method: Geometric, best acc_lad ** avg: 0.8108064516129032 std: 0.008065120945062159
method: Geometric, best acc_lar ** avg: 0.8400000000000001 std: 0.014046271275059837
method: Geometric, best acc_ard ** avg: 0.6583870967741936 std: 0.006289030057447286
method: Geometric, best acc_lrd ** avg: 0.8641129032258064 std: 0.008277431297605348
method: Geometric, best acc_lard ** avg: 0.8593548387096774 std: 0.007356633552926914
method: SupCon, best mrr_ad ** avg: 0.7818461021505375 std: 0.0058056222339268986
method: SupCon, best mrr_ar ** avg: 0.7968904569892473 std: 0.005438750540616555
method: SupCon, best mrr_ld ** avg: 0.8903924731182796 std: 0.008784518487631803
method: SupCon, best mrr_lr ** avg: 0.9056451612903225 std: 0.007443727195959732
method: SupCon, best mrr_lad ** avg: 0.8875026881720431 std: 0.0065906156485880715
method: SupCon, best mrr_lar ** avg: 0.9049993279569893 std: 0.006934618496482506
method: SupCon, best mrr_ard ** avg: 0.8120443548387095 std: 0.0038831681475468686
method: SupCon, best mrr_lrd ** avg: 0.919635752688172 std: 0.00421253081123992
method: SupCon, best mrr_lard ** avg: 0.9202607526881721 std: 0.007036465744801571
method: SupCon, best acc_ad ** avg: 0.6417338709677419 std: 0.009174605555378065
method: SupCon, best acc_ar ** avg: 0.6652016129032258 std: 0.010682516325605402
method: SupCon, best acc_ld ** avg: 0.8105241935483869 std: 0.012173413141473797
method: SupCon, best acc_lr ** avg: 0.8364516129032259 std: 0.014036892078731987
method: SupCon, best acc_lad ** avg: 0.8058467741935484 std: 0.011178885534336718
method: SupCon, best acc_lar ** avg: 0.8354435483870969 std: 0.012324733809520844
method: SupCon, best acc_ard ** avg: 0.6869758064516129 std: 0.006569735184573384
method: SupCon, best acc_lrd ** avg: 0.8605645161290323 std: 0.012074307093137743
method: SupCon, best acc_lard ** avg: 0.8581854838709677 std: 0.01293003766775218
method: EMMA, best mrr_ad ** avg: 0.7763420698924731 std: 0.0029124808921936005
method: EMMA, best mrr_ar ** avg: 0.786581317204301 std: 0.006418196740535436
method: EMMA, best mrr_ld ** avg: 0.8986505376344086 std: 0.005024686589459191
method: EMMA, best mrr_lr ** avg: 0.9125739247311827 std: 0.008583791522680621
method: EMMA, best mrr_lad ** avg: 0.8966283602150538 std: 0.0036157579057873954
method: EMMA, best mrr_lar ** avg: 0.9097157258064514 std: 0.00655691760402764
method: EMMA, best mrr_ard ** avg: 0.8032096774193549 std: 0.004465339908648459
method: EMMA, best mrr_lrd ** avg: 0.9270524193548388 std: 0.004950684287669308
method: EMMA, best mrr_lard ** avg: 0.9272177419354838 std: 0.0046582728067974
method: EMMA, best acc_ad ** avg: 0.6353629032258065 std: 0.00531983066789421
method: EMMA, best acc_ar ** avg: 0.6506854838709677 std: 0.010086772331454559
method: EMMA, best acc_ld ** avg: 0.8277822580645161 std: 0.009658415480510912
method: EMMA, best acc_lr ** avg: 0.850725806451613 std: 0.014181056636432354
method: EMMA, best acc_lad ** avg: 0.8216129032258064 std: 0.006363306949864157
method: EMMA, best acc_lar ** avg: 0.8437096774193549 std: 0.012343847760054961
method: EMMA, best acc_ard ** avg: 0.6768951612903227 std: 0.008115363892853094
method: EMMA, best acc_lrd ** avg: 0.8737903225806452 std: 0.007140626974573769
method: EMMA, best acc_lard ** avg: 0.8715322580645161 std: 0.007248425499130212
method: Contrastive, best mrr_ad ** avg: 0.7174408602150538 std: 0.007255402799973243
method: Contrastive, best mrr_ar ** avg: 0.7337029569892473 std: 0.0039224209303724234
method: Contrastive, best mrr_ld ** avg: 0.8972479838709677 std: 0.005357753897962892
method: Contrastive, best mrr_lr ** avg: 0.9082345430107527 std: 0.003749222854686774
method: Contrastive, best mrr_lad ** avg: 0.8912668010752688 std: 0.006122357196224834
method: Contrastive, best mrr_lar ** avg: 0.9026169354838709 std: 0.005789581070670862
method: Contrastive, best mrr_ard ** avg: 0.7496478494623655 std: 0.004373982216482813
method: Contrastive, best mrr_lrd ** avg: 0.9192137096774193 std: 0.00411868564896933
method: Contrastive, best mrr_lard ** avg: 0.9171935483870965 std: 0.005344248282948013
method: Contrastive, best acc_ad ** avg: 0.5482258064516129 std: 0.014024724525739987
method: Contrastive, best acc_ar ** avg: 0.5727016129032259 std: 0.006437231955968572
method: Contrastive, best acc_ld ** avg: 0.82875 std: 0.008802521869277298
method: Contrastive, best acc_lr ** avg: 0.8434677419354838 std: 0.01012939823665928
method: Contrastive, best acc_lad ** avg: 0.8155241935483872 std: 0.00933622313545278
method: Contrastive, best acc_lar ** avg: 0.8326209677419355 std: 0.010162090342290321
method: Contrastive, best acc_ard ** avg: 0.5938306451612902 std: 0.005993840056794196
method: Contrastive, best acc_lrd ** avg: 0.8631048387096774 std: 0.006685527646846855
method: Contrastive, best acc_lard ** avg: 0.8574999999999999 std: 0.008746747574507578

### 2022-10-06 Code/result-analysis/tableACC.tex
Results for TMLR

\begin{tabular}{|p{{2cm}}|p{{2cm}}|p{{2cm}}|p{{2cm}}|p{{2cm}}|p{{2cm}}|p{{2cm}}|p{{2cm}}|p{{2cm}}|p{{2cm}}|} 
 \hline 
Methods & speech/depth & speech/RGB & text/depth & text/RGB & text/speech/ \newline depth & text/speech/ \newline RGB & speech/RGB/ \newline depth & text/RGB/ \newline depth & all \\ 
 \hline 
Geometric & 61.95$\pm$0.55 & 64.34$\pm$0.53 & 82.03$\pm$0.57 & 84.6$\pm$1.1 & 81.08$\pm$0.81 & 84.0$\pm$1.4 & 65.84$\pm$0.63 & 86.41$\pm$0.83 & 85.94$\pm$0.74 \\ 
SupCon & \textbf{64.17}$\pm$0.92 & \textbf{66.52}$\pm$1.07 & 81.05$\pm$1.22 & 83.65$\pm$1.4 & 80.58$\pm$1.12 & 83.54$\pm$1.23 & \textbf{68.7}$\pm$0.66 & 86.06$\pm$1.21 & 85.82$\pm$1.29 \\ 
EMMA & 63.54$\pm$0.53 & 65.07$\pm$1.01 & 82.78$\pm$0.97 & \textbf{85.07}$\pm$1.42 & \textbf{82.16}$\pm$0.64 & \textbf{84.37}$\pm$1.23 & 67.69$\pm$0.81 & \textbf{87.38}$\pm$0.71 & \textbf{87.15}$\pm$0.72 \\ 
Contrastive & 54.82$\pm$1.4 & 57.27$\pm$0.64 & \textbf{82.88}$\pm$0.88 & 84.35$\pm$1.01 & 81.55$\pm$0.93 & 83.26$\pm$1.02 & 59.38$\pm$0.6 & 86.31$\pm$0.67 & 85.75$\pm$0.87 \\ 
\hline \end{tabular}

### 2022-10-06 Code/analysis.py

@@ -86,9 +86,9 @@ metrics = ['mrr_ad', 'mrr_ar', 'mrr_ld', 'mrr_lr', 'mrr_lad', 'mrr_lar', 'mrr_ar
 # col = ['Method'] + [metr.replace('_','.') for metr in metrics] # underline _ mess with latex.
 metrics_mrr = ['mrr_ad', 'mrr_ar', 'mrr_ld', 'mrr_lr', 'mrr_lad', 'mrr_lar', 'mrr_ard', 'mrr_lrd', 'mrr_lard']
 metrics_acc = ['acc_ad', 'acc_ar', 'acc_ld', 'acc_lr', 'acc_lad', 'acc_lar', 'acc_ard', 'acc_lrd', 'acc_lard']
-col = ['Method'] + [metr.replace('_','.').replace('l', 't').replace('a', 's') for metr in metrics_mrr] # underline _ mess with latex, (l)anguage --> (t)ext, and (a)udio --> (s)peech
+col = ['Method'] + metrics_mrr 
 result_mrr = pd.DataFrame(columns=col)
-col = ['Method'] + [metr.replace('_','.').replace('l', 't').replace('acc', 'cc').replace('a', 's').replace('cc', 'acc') for metr in metrics_acc] # underline _ mess with latex.
+col = ['Method'] + metrics_acc
 result_acc = pd.DataFrame(columns=col) 
 best_bold = {}
 for metr in metrics:
@@ -100,14 +100,31 @@ for idx, method in enumerate(list(results.keys())):
     result_acc.at[idx, 'Method'] = method
     for metr in metrics_mrr:
         if method == best_bold[metr]:
-            result_mrr.at[idx, metr.replace('_','.').replace('l', 't').replace('a', 's')] = f"\\textbf{{{round(results[method]['avg']['best']['test'][metr]*100, 2)}}}$\pm${round(results[method]['std']['best']['test'][metr]*100, 2)}"
+            result_mrr.at[idx, metr] = f"\\textbf{{{round(results[method]['avg']['best']['test'][metr]*100, 2)}}}$\pm${round(results[method]['std']['best']['test'][metr]*100, 2)}"
         else:
-            result_mrr.at[idx, metr.replace('_','.').replace('l', 't').replace('a', 's')] = f"{round(results[method]['avg']['best']['test'][metr]*100, 2)}$\pm${round(results[method]['std']['best']['test'][metr]*100, 2)}"
[…]

### 2022-10-06 Code/explicit-latex-table.py

@@ -1,0 +1,118 @@
+import os, json, csv
+from re import M
+from unittest import result
+import numpy as np
+import pandas as pd
+from glob import glob
+import matplotlib.pyplot as plt
+import tikzplotlib
+
+
+exp_ugly_names = {
+    'Geometric': 'exp-full-emma-lard-64-relu-SGD-0.001-unique_objects-gold-no_neg_sampling-1024',
+    'SupCon': 'exp-supcon-lard-64-relu-SGD-0.001-unique_objects-gold-no_neg_sampling-1024',
+    'EMMA': 'exp-supcon-emma-lard-64-relu-SGD-0.001-unique_objects-gold-no_neg_sampling-1024',
+    'Contrastive': 'exp-contrastive-org-lard-64-relu-SGD-0.001-unique_objects-gold-no_neg_sampling-1024',
+}
+
+
+experiments = {name:{} for name in exp_ugly_names.keys()}
+
+
+for method in experiments.keys():
+    for seed in [7, 24, 42, 123, 3407]:
+        experiments[method]['seed-'+str(seed)] = glob(os.path.join('results', f'{exp_ugly_names[method]}/seed-{str(seed)}/'), recursive=True)[0]
+        # experiments[method]['seed-'+str(seed)] = glob(os.path.join('results', f'*{exp_ugly_names[method]}*/seed-{str(seed)}/'), recursive=True)[0]
+
+print(experiments)
+
+
+results = {}
+for method in experiments.keys():
+    results[method] = {}
+    for seed in experiments[method].keys():
+        res = json.load(open(experiments[method][seed]+'results.json'))
+        del res['best']
+        results[method][seed] = res
+    results[method]['avg'] = {'best':{}}
+    results[method]['std'] = {'best':{}}
+    # computing average and std over seeds for each metric of each portion of each epoch
+    for epoch in results[method][seed].keys():
+        results[method]['avg'][epoch] = {}
+        results[method]['std'][epoch] = {}
+        for portion in results[method]['seed-42'][epoch].keys():
+            results[method]['avg'][epoch][portion] = {}
+            results[method]['std'][epoch][portion] = {}
+            for metric in results[method]['seed-42'][epoch][portion].keys():
[…]

### 2022-10-06 Paper/Figures/unused/results-table-old.tex
Paper submitted to TMLR

--- 2022-01-26 Paper/main.tex
@@
-\documentclass[sigconf,natbib=true,anonymous=true]{acmart}
-% \documentclass[manuscript,screen,review]{acmart}
+\begin{table*}[tb]
+    \centering
+    \begin{subtable}[t]{1.00\textwidth}
+    \centering
+    \resizebox{.99\textwidth}{!}{
+    \csvautotabular{resultsMRR.csv}
+    }
+    \end{subtable}
+    \vskip\baselineskip
+    % \hfill
+    \begin{subtable}[t]{1.00\textwidth}
+    \centering
+    \resizebox{.99\textwidth}{!}{
+    \csvautotabular{resultsACC.csv}
+    }
+    \end{subtable}
+\end{table*}
+% \todocm{For Table 1 - change all mrr.xx to MRR \textit{xx} for consistency with eqn 8, and all acc.xx to Acc \textit{xx}---e.g., column headers should be 'MRR \textit{td}'.
+Also make sure the caption edits are correct.}
+% \todo[inline]{Make it two subtables, and remove MRR and Acc from the columns, and spell out sr, tsrd, and ...}
 
-% Needed to fix an acmart problem with line numbering. AFAICT line 
-% numbers are not expected in sigconf, so it should be a no-op.
-\makeatletter
-\providecommand{\@LN}[2]{}
-\makeatother
+% Automatic DataTool Method
+\DTLloaddb{resultsMRR}{resultsMRR.csv}
+\DTLloaddb{resultsACC}{resultsACC.csv}
+\begin{table}[]
+    \centering
+    \resizebox{1.0\textwidth}{!}{
+    \DTLdisplaydb{resultsMRR}
+    }
+    \resizebox{1.0\textwidth}{!}{
+    \DTLdisplaydb{resultsACC}
+    }
+    \caption{Caption}
+    \label{table:quantitative}
+\end{table}
 
-% % Fonts used in the template cannot be substituted; margin 
-% % adjustments are not allowed.
-% %
-% % \BibTeX command to typeset BibTeX logo in the docs
-% \AtBeginDocument{%
-%   \providecommand\BibTeX{{%
-%     \normalfont B\kern-0.5em{\scshape i\kern-0.25em b}\kern-0.8em\TeX}}}
+% Manually constructed table
+\begin{table*}[]
+\centering
+\resizebox{.97\textwidth}{!}{
+\begin{tabular}{lc|ll|lll}
+\multicolumn{1}{c}{} &  & \multicolumn{2}{c|}{\textbf{Accuracy}} & \multicolumn{3}{c}{\textbf{Mean Reciprocal Rank}} \\ \hline
+\multicolumn{1}{c|}{\textbf{Method}} & \textbf{Modalities} & \multicolumn{1}{c|}{\textbf{\begin{tabular}[c]{@{}c@{}}Speech/RGB/\\ Depth\end{tabular}}} & \multicolumn{1}{c|}{\textbf{\begin{tabular}[c]{@{}c@{}}Text/RGB/\\ Depth\end{tabular}}} & \multicolumn{1}{c|}{\textbf{\begin{tabular}[c]{@{}c@{}}Speech/RGB/\\ Depth\end{tabular}}} & \multicolumn{1}{c|}{\textbf{\begin{tabular}[c]{@{}c@{}}Text/RGB/\\ Depth\end{tabular}}} & \multicolumn{1}{c}{\textbf{\begin{tabular}[c]{@{}c@{}}Text/Speech/\\ RGB/Depth\end{tabular}}} \\ \hline
+\multicolumn{1}{l|}{EMMA} & TSRD & \multicolumn{1}{l|}{\textbf{0.65}$\pm$0.01} & 0.83$\pm$0.006 & \multicolumn{1}{l|}{\textbf{0.79}$\pm$0.002} & \multicolumn{1}{l|}{0.90$\pm$0.003} & 0.90$\pm$0.004 \\
+\multicolumn{1}{l|}{EMMA} & 3 & \multicolumn{1}{l|}{--} & 0.84$\pm$0.008 & \multicolumn{1}{l|}{--} & \multicolumn{1}{l|}{0.90$\pm$0.004} & -- \\
+\multicolumn{1}{l|}{\supcon{}} & TSRD & \multicolumn{1}{l|}{0.59$\pm$0.007} & 0.84$\pm$0.008 & \multicolumn{1}{l|}{0.75$\pm$0.004} & \multicolumn{1}{l|}{0.91$\pm$0.006} & 0.90$\pm$0.006 \\
+\multicolumn{1}{l|}{\supcon{}} & 3 & \multicolumn{1}{l|}{--} & 0.84$\pm$0.006 & \multicolumn{1}{l|}{--} & \multicolumn{1}{l|}{0.91$\pm$0.004} & -- \\
+\multicolumn{1}{l|}{Contrastive Cosine} & TSRD & \multicolumn{1}{l|}{0.40$\pm$0.153} & 0.83$\pm$0.006 & \multicolumn{1}{l|}{0.62$\pm$0.034} & \multicolumn{1}{l|}{0.90$\pm$0.004} & 0.89$\pm$0.005 \\
+\multicolumn{1}{l|}{Contrastive Cosine} & 3 & \multicolumn{1}{l|}{--} & 0.84$\pm$0.005 & \multicolumn{1}{l|}{--} & \multicolumn{1}{l|}{0.91$\pm$0.003} & -- \\
+\multicolumn{1}{l|}{\ours{}} & TSRD & \multicolumn{1}{l|}{--} & 0.83$\pm$0.008 & \multicolumn{1}{l|}{--} & \multicolumn{1}{l|}{0.90$\pm$0.005} & -- \\ \hline
+\end{tabular}
+}
+\caption{Average and standard deviation of accuracy and MRR over 3 runs with 3 different random seeds on a held-out test set.
+These results represent performance when no modalities are ablated, in which case supervised contrastive loss, contrastive loss, and EMMA perform similarly (``--'' is an ablated modality and can't be evaluated).
+All metrics are from 0 to 1, and higher is better.
+For 5 objects, a random guess would have MRR of 0.33, and the worst case performance would be always ranking the correct item last which gives us 0.2.
+All models are trained using text as anchor.
+The batch size is 64 and optimizer is SGD for all experiments except the triplet loss, where batch size is 1 and optimizer is Adam.
+%In the Triplet experiment, RGB and Depth embeddings are concatenated to form a single vision embedding since triplet loss cannot be used for more than two modalities.
+} 
+\label{table:quantitative}
+\end{table*}
+% \todo[inline]{This table is not referred to until sec:results, which is waaay down from where it's introduced, and it takes some parsing to figure out what's going on.
+Either add a paragraph under it that starts with something like "Table 1 shows high-level results, in which..." or just move the table to the results section.}
 
-%% Rights management information.  This information is sent to you
-%% when you complete the rights form.  These commands have SAMPLE
-%% values in them; it is your responsibility as an author to replace
-%% the commands and values with those provided to you when you
-%% complete the rights form.
-\setcopyright{acmcopyright}
-\copyrightyear{2022}
-\acmYear{2022}
-\acmDOI{XXXXXXX.XXXXXXX}
+% %Edward moved table above where refernced b/c LaTeX plays games with double-wide objects
+% \begin{table*}[bth]
+% \centering
+% \begin{tabular}{c|c|c|c|c|c|c}
+% \toprule
+% \textbf{Method} & \textbf{Modalities} & \textbf{acc.srd} & \textbf{acc.trd} & \textbf{MRR.srd} & \textbf{MRR.trd} & \textbf{MRR.tsrd} \\ %\hline
+% \midrule
+% EMMA (Text) & 4 & \textbf{0.65}$\pm$0.01 & 0.83$\pm$0.006 & \textbf{0.79}$\pm$0.002 & 0.90$\pm$0.003 & 0.90$\pm$0.004 \\
+% EMMA (Text) & 3 & - & 0.84$\pm$0.008 & - & 0.90$\pm$0.004 & - \\
+% % Simple MMA (Text)  & 3 & ?$\pm$? & ?$\pm$? & ? & ? \\
+% % Simple MMA (Text) & 4 & ?$\pm$? & ?$\pm$? & ? & ? \\
+% % Simple MMA (Speech) & 3 & ?$\pm$? & ?$\pm$? & ? & ?  \\
+% DONE Supervised Contrastive (Text)& 4 & 0.59$\pm$0.007 & 0.84$\pm$0.008 & 0.75$\pm$0.004 & 0.91$\pm$0.006 & 0.90$\pm$0.006  \\
+% TODO Supervised Contrastive (Text)& 3 & - & 0.84$\pm$0.006 & - & 0.91$\pm$0.004 & -\\
+% DONE Contrastive Cosine (Text) & 4 & 0.40$\pm$0.153 & 0.83$\pm$0.006 & 0.62$\pm$0.034  & 0.90$\pm$0.004 & 0.89$\pm$0.005\\
+% TODO Contrastive Cosine (Text) & 3 & - & 0.84$\pm$0.005 & -  & 0.91$\pm$0.003 & -\\
+% PROJ Triplet (Text) & 3 & - & 0.83$\pm$0.008 & - & 0.90$\pm$0.005 & -  \\
+% % DONE EMMA (Speech) & 3 & ?$\pm$? & ?$\pm$? & - & - \\
+% % DONE EMMA (Speech) & 4 & ?$\pm$? & ?$\pm$? & ? & ? \\
+% % DONE Supervised Contrastive (Speech)& 3 & ?$\pm$? & ?$\pm$? & - & -\\
+% % DONE Supervised Contrastive (Speech)& 4 & ?$\pm$? & ?$\pm$? & ? & ? \\
+% % TODO Contrastive Cosine (Speech) & 3 & ?$\pm$? & ?$\pm$?  & - & -\\
+% % TODO Contrastive Cosine (Speech) & 4 & ?$\pm$? & ?$\pm$?  & ? & ?\\
+% % DONE Triplet (Speech) & 3 & ?$\pm$? & ?$\pm$? & - & -  \\
 
-%% These commands are for a PROCEEDINGS abstract or paper.
-\acmConference[SIGIR '22]{Make sure to enter the correct
-  conference title from your rights confirmation email}{June 03--05,
-  2018}{Woodstock, NY}
-%
-%  Uncomment \acmBooktitle if the title of the proceedings is different
-%  from ``Proceedings of ...''!
-%
-\acmBooktitle{SIGIR '22: Proceedings of the 45th International ACM SIGIR Conference on Research and Development in Information Retrieval,
- July 11--15, 2022, Madrid, Spain} 
-% \acmPrice{15.00}
-% \acmISBN{978-1-4503-XXXX-X/18/06}
+% % Full MMA (Text) w/ neg & 3 & ?$\pm$? & ?$\pm$? \\
+% % Simple MMA (Text) w/ neg Adam & 3  & 0.7949$\pm$? & 0.8778$\pm$? \\
+% % Simple MMA (Text) w/ neg SGD & 3 & 0.7905$\pm$? & 0.8821$\pm$? \\
+% % Triplet (Text) w/ neg & 3 & 0.7497$\pm$? & 0.8522$\pm$? \\
+% % Full MMA (Speech) w/ neg & 3 & ?$\pm$? & ?$\pm$? \\
+% % Simple MMA (Speech) w/ neg & 3 & 0.7949$\pm$? & 0.8778$\pm$? \\
+% % Contrastive Cosine (Text) w/ neg & 3 & 0.7894$\pm$? & 0.8788$\pm$?  & ? & ?\\
+% % Contrastive (Speech) w/ neg & 3 & ?$\pm$? & ?$\pm$? \\
+% % Triplet (Speech) w/ neg & 3 & 0.7497$\pm$? & 0.8522$\pm$? \\
+% \bottomrule
+% \end{tabular}
+% \caption{\label{table:quantitative}Average and standard deviation of accuracy and MRR over 3 runs with 3 different random seeds on a held-out test set.
+These results represent performance when no modalities are ablated, in which case supervised contrastive loss, contrastive loss, and EMMA perform similarly.
+All metrics are from 0 to 1, and higher is better.
+For 5 objects, a random guess would have MRR of 0.33, and the worst case performance would be always ranking the correct item last which gives us 0.2.
+All models are trained using text as anchor.
+The batch size is 64 and optimizer is SGD for all experiments except the triplet loss, where batch size is 1 and optimizer is Adam.
+% % \todokdinline{maybe adding results on cropped and raw dataset as well. Also results using object class instead of negative sampling}
+% }
+% \end{table*}
 
-%
-% Submission ID.
-% Use this when submitting an article to a sponsored event. You'll
-% receive a unique submission ID from the organizers
-% of the event, and this ID should be used as the parameter to this command.
-% \acmSubmissionID{123-A56-BU3}
-
-%
-% The majority of ACM publications use numbered citations and
-% references.  The command \citestyle{authoryear} switches to the
-% "author year" style.
-%
-% If you are preparing content for an event
-% sponsored by ACM SIGGRAPH, you must use the "author year" style of
-% citations and references.
-% Uncommenting
-% the next command will enable that style.
-%\citestyle{acmauthoryear}
-
-%% The "author" command and its associated commands are used to define
-%% the authors and their affiliations.
-%% Of note is the shared affiliation of the first two authors, and the
-%% "authornote" and "authornotemark" commands
-%% used to denote shared contribution to the research.
-\author{Ben Trovato}
-\authornote{Both authors contributed equally to this research.}
-\email{trovato@corporation.com}
-\orcid{1234-5678-9012}
-\author{G.K.M. Tobin}
-\authornotemark[1]
-\email{webmaster@marysville-ohio.com}
-\affiliation{%
-  \institution{Institute for Clarity in Documentation}
-  \streetaddress{P.O. Box 1212}
-  \city{Dublin}
-  \state{Ohio}
-  \country{USA}
-  \postcode{43017-6221}
-}
-
-\author{Lars Th{\o}rv{\"a}ld}
-\affiliation{%
-  \institution{The Th{\o}rv{\"a}ld Group}
-  \streetaddress{1 Th{\o}rv{\"a}ld Circle}
-  \city{Hekla}
-  \country{Iceland}}
-\email{larst@affiliation.org}
-
-\author{Valerie B\'eranger}
-\affiliation{%
-  \institution{Inria Paris-Rocquencourt}
-  \city{Rocquencourt}
-  \country{France}
-}
-
-\author{Aparna Patel}
-\affiliation{%
- \institution{Rajiv Gandhi University}
- \streetaddress{Rono-Hills}
- \city{Doimukh}
- \state{Arunachal Pradesh}
- \country{India}}
-
-\author{Huifen Chan}
-\affiliation{%
-  \institution{Tsinghua University}
-  \streetaddress{30 Shuangqing Rd}
-  \city{Haidian Qu}
-  \state{Beijing Shi}
-  \country{China}}
-
-\author{Charles Palmer}
-\affiliation{%
-  \institution{Palmer Research Laboratories}
-  \streetaddress{8600 Datapoint Drive}
-  \city{San Antonio}
-  \state{Texas}
-  \country{USA}
-  \postcode{78229}}
-\email{cpalmer@prl.com}
-
-\author{John Smith}
-\affiliation{%
-  \institution{The Th{\o}rv{\"a}ld Group}
-  \streetaddress{1 Th{\o}rv{\"a}ld Circle}
-  \city{Hekla}
-  \country{Iceland}}
-\email{jsmith@affiliation.org}
-
-\author{Julius P. Kumquat}
-\affiliation{%
-  \institution{The Kumquat Consortium}
-  \city{New York}
-  \country{USA}}
-\email{jpkumquat@consortium.net}
-
-%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%
-% EDITING LINK: 
-% https://www.overleaf.com/3575751412sjsdmszkxsyh
-%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%
-
-% \usepackage{soul} %% Soul does not play well with acmart
-\usepackage{url}
-% % \usepackage[hidelinks]{hyperref}https://www.overleaf.com/project/614b5cc61d38655dce466002
-% \usepackage[small]{caption}
-\usepackage{graphicx}
-\usepackage{amsthm}
-\usepackage{booktabs}
-\usepackage{algorithm}
-\usepackage{algorithmic}
-\urlstyle{same}
-
-% % =======================================
-% % my imported packages
-
-\usepackage{amsmath}
-\usepackage{bm}
-% \usepackage{amssymb}          % duplicates something in acmart
-\usepackage{listings}
-% \usepackage[ruled,vlined]{algorithm2e}
-\include{pythonlisting}
-
-\newcommand{\pseudosection}[1]{\vspace{1.5ex}\noindent \textit{{#1:~~}}}
-% \newcommand{\pseudosection}[1]\indent{{{}}}
-
-\usepackage{todonotes}
-\newcommand{\todokdinline}[1]{\todo[color=red!20,inline]{{KD: \small #1}}}
-\newcommand{\todokd}[1]{\todo[color=red!20]{{\small #1 -- Kasra}}}
-% \newcommand{\todocainline}[1]{\todo[color=yellow!20,inline]{{CA: \small #1}}}
-\newcommand{\todocm}[1]{\todo[color=green!40]{\small #1 -- Cynthia}}
-\newcommand{\todocmi}[1]{\todo[inline,color=green!40]{\small #1 -- Cynthia}}
-\newcommand{\todoff}[1]{\todo[color=blue!20]{\small #1 -- Frank}}
-\newcommand{\todoed}[1]{\todo[color=cyan!20]{\small #1 -- Ed}}
-
-\newcommand{\CitationNeeded}[1]{{\textbf {\color{red}Cite #1}}}
-\newcommand{\ProofRead}[1]{{\textbf {\color{red}Proofread please #1}}}
-\newcommand{\Complete}[0]{{\textbf {\color{red}Complete it }}}
-\newcommand{\Rephrase}[1]{{\textbf {\color{red}Rephrase please #1}}}
-
-% \newcommand{\citet}[1]{\cite{#1}} % This is used because \citet didn't work with their author kit % It does now
-
-% =======================================
-\title{Multimodal Contrastive Learning for Physical Object Retrieval}
-
-\begin{document}
-
-% Todo list
-% \begin{enumerate}
-%     \item \st{Lit search}
-% \end{enumerate}
-% \newpage
-
-% ***************** The BEGINNING of the paper *****************
-
-\begin{abstract}
-    % Grounded language understanding, in which natural language is used as a query against objects in a physical environment, allows a real-world, intuitive mechanism by which users can instruct physical agents to engage in tasks such as object retrieval.
-Visuolinguistic approaches to such object inference tasks typically involve training on large pools of image/text pairs and then using language to subselect elements of the sensed environment.
-However, physical agents such as robots typically have access to sensory and interactive modalities beyond vision, and learning from multiple modalities can improve performance on downstream tasks.
-In order to fully leverage multimodal training data while being robust to missing information, we propose a generalized distance-based loss function that can be extended to learn retrieval models that incorporate an arbitrary number of modalities.
-We demonstrate the usability of our model on a grounded language object retrieval scenario, where an intelligent agent has to select an object given an unconstrained language command.
-We leverage four modalities including vision, depth sensing, text, and speech, and we show that our model can outperform state-of-the-art contrastive models when modalities are ablated.
-    % \todokdinline{Text is easier to learn from compared to speech, and our proposed model allow us to minimize the distance between text and speech and when text is missing during the test time, speech can be used to perform the task, and it outperforms the supervised contrastive learning because such a leverage is missing there.}
-
-    % FRANK:
-
-    We propose a generalized distance-based loss function that can be used to learn retrieval models that incorporate an arbitrary number of views of 
-    a particular piece of data. %
-    We study this in the context of grounded language understanding, in which natural language is used as a query against objects in a physical environment, allowing a real-world, intuitive mechanism by which users can instruct physical agents to engage in tasks such as object retrieval.
-    In our setting, each view is a different modality of data, such as written text, spoken language, and images. 
-    % Visuolinguistic approaches to such object inference tasks typically involve training on large pools of image/text pairs and then using language to subselect elements of the sensed environment. 
-    %Such an approach could improve the ability of physical agents, such as robots typically have access to sensory and interactive modalities beyond vision, and learning from multiple modalities can improve performance on downstream tasks.
-    We demonstrate the usability of our model on a grounded language object retrieval scenario, where an intelligent agent has to select an object given an unconstrained language command.
-We leverage four modalities including vision, depth sensing, text, and speech, and we show that our model can outperform state-of-the-art contrastive models when modalities are ablated.
-
-\end{abstract}
-
-\settopmatter{printfolios=true} % Page numbers please 
-9 pages without references.
-%===================================================================
-
-\maketitle
-% \setcounter{page}{1}
-
-%===================================================================
-
-% \section{SIGIR Papers}
-% \begin{enumerate}
-%     \item \href{https://dl.acm.org/doi/pdf/10.1145/3397271.3401430}{FashionBERT}
-%     \item \href{https://dl.acm.org/doi/pdf/10.1145/3397271.3401265?casa_token=W8NqNWp0UQkAAAAA:w-iRUnNhmytDKCUiYzaUPG0bI_yyo8Bzxwro_eOC-zl5VDIGRVT2dvixvnCwLJ0iEKqYo1LkQIWXVA}{CrossBERT}
-
-% \end{enumerate}
-
-\section{Introduction}
-\label{intro}
-
-Humans use multiple sensory inputs to understand, interact with, and retrieve information from the world around them.
-It is intuitive that agents learning about the world, upon encountering new concepts and new objects, should form a model that incorporates information from all available sensors and data sources.
-
-Grounded language understanding, in which natural language is used as a query against objects in a physical environment, allows a real-world, intuitive mechanism by which users can instruct physical agents to engage in tasks such as object retrieval.
-Visuolinguistic approaches to such object inference tasks typically involve training on large pools of image/text pairs and then using language to subselect elements of the sensed environment.
-Although physical agents such as robots typically have access to sensory and interactive modalities beyond vision, and learning from multiple modalities can improve performance on downstream tasks, most approaches use at most two sensory inputs (e.g., visual data such as RGB images) with single labels, such as those provided by textual natural language.
-Simultaneously using additional inputs from different modalities is an underexplored area, in part due to the domain-specific nature of such \textit{n}-ary learning approaches.
-
-In order to fully leverage multimodal training data while being robust to missing information, we propose a generalized distance-based loss function that can be extended to learn retrieval models that incorporate an arbitrary number of modalities.
-We consider the domain of grounded language-based object retrieval~\cite{Mooney2008Grounded, RichardsDarvishMatuszekCategoryFree20, triplet_loss_2021_CVPR}.\todo{check whether these are still the appropriate citations}
-% , a cross-modal learning task in which the goal is to learn to understand natural language and its connection to the physical environment in which the language occurs. 
-Approaches to acquiring grounded language have explored various combinations of sensor inputs such as depth and RGB with labels provided by textual language or speech.
-This work has significant overlap with robotics, as it pertains to the use of language in the physical domain in which robots operate.
-However, despite the multisensory nature of most robot platforms, much of the existing work has not previously been extended to include an arbitrary number of modalities.
-
-The high level goal of this work is to take arbitrary input modalities about novel objects in the world, including both spoken and written language, and build a model that allows a robot system to correctly retrieve that object given a subset of those input modalities.
-This is a generalization of approaches to grounded language learning in which specific input modalities are labeled with language to allow for future identification, but explicitly seeks to be agnostic about the nature of the individual sensory and linguistic inputs.
-Our contributions are as follows:
-\begin{enumerate}
-    \item Proposing a general distance-based loss function formulation for multimodal learning that can be extended to an arbitrary number of modalities.
-    \item Outperforming state-of-the-art models for grounded language learning by differentiating visual input modalities such as RGB and depth, rather than combining them into a single modality.
-    \item Demonstrating the separate utility of speech and text as sources of label information by treating them as sensory input modalities, rather than explicitly as labels.
-    \item Outperforming state-of-the-art contrastive models when one or more modalities are missing during test.
-\end{enumerate}
-
-\todocmi{NOTES: Some things are probably less informative or more ambiguous (e.g., depth for differentiating round things).
-So instead of predicting a single vector output, predict a vector and a covariance that captures uncertainty; could potentially use that to do more accurate search, or to handle really uncertain modalities (e.g., just drop speech when it's uninformative).}
-
-\todocmi{NOTES: What if one of your modalities fails? Does that super-bork the contrastive approach? Possibly Simple MMA will be more robust to dropped modalities.}
-
-%===================================================================
-
-\todo[inline]{Edward Framing idea: Sensors sometimes break. Look at accuracy retrieval using against target object with various sub-modalities available/not? I will run those experiments today! }
-
-\section{Related Work}
-\label{sec:Related-Work}
-
-The growing number of multimodal datasets demonstrates the importance and the challenge of this task.
-\citet{bagher-zadeh-etal-2018-multimodal} introduce a large dataset of multimodal opinion sentiment and emotion intensity (CMU-MOSEI) for sentiment analysis and emotion recognition.
-They also introduce a novel multimodal fusion technique called the Dynamic Fusion Graph (DFG).~\citet{GoLD_UMBC} present a multimodal dataset of household objects containing RGB images, depth images, written language, and spoken language, which has been used to support learning grounded language directly from speech given small datasets~\cite{KebeAAAI2022}.
-\todo[inline]{Look at multisensory datasets from robotics}
-
-\citet{baltrusaitisMultimodalMachineLearning2019} propose a new taxonomy of multimodal machine learning by introducing five technical challenges in addition to typical early and late fusion categorization.
-They cover different approaches to multimodal machine learning such as \textit{representation}, \textit{translation}, \textit{alignment}, \textit{fusion}, and \textit{co-learning}.
-In this project, we focus on alignment, in which \textcolor{red}{\dots}.
-
-We review different methods that have been proposed for alignment learning in the language domain.
-\citet{alayrac2020self} use a self-supervised contrastive learning method to learn and combine representations from three modalities of visual, audio, and language.
-Audio and visual domains are mapped to the same space because they are fine-grained, containing dense information for each frame of video.
-These embeddings are then mapped to a coarse space where text is also mapped.
-Their network respects specificity and comparability of different modalities.
-The loss function they define consists of two distance-based terms for the corresponding spaces.
-This work differs from ours in that \textcolor{red}{\dots}.
-\citet{Nguyen-RSS-20} take a similar approach and perform pairwise cosine similarity to align images and natural language text descriptions to perform object retrieval tasks.
-\citet{triplet_loss_2021_CVPR} take a cross-modal manifold alignment approach to grounded language learning using triplet loss, and perform object retrieval by connecting text descriptions of objects to their corresponding RGB-D images.
-
-Another related approach is contrastive learning, which is based on the idea of similarity and distance~\cite{NEURIPS2020_supervised_contrastive,chen2020simple}.
-Triplet loss is also a special case of contrastive loss~\cite{NEURIPS2020_supervised_contrastive}.
-Contrastive loss is mostly used in self-supervised learning~\cite{alayrac2020self,chen2020simple}, with the exception of the work proposed by \citet{NEURIPS2020_supervised_contrastive} where they use contrastive loss for a supervised problem.
-Most of these methods are implemented for a non-multimodal dataset (e.g., RGB images only).
-Our work is a specialization of contrastive loss that is focused on the sensor-and-language problem.\todo[inline]{Can we think of something better to say about how this relates to contrastive learning generally?}
-
-Some of the most closely related work focuses on learning based on more than two input modalities.
-\citet{Veit_2018_CVPR} train a joint model of images, hashtags, and users to perform image retrieval.
-Their task is similar to our task where given a description, we want to find the image/object in the scene that best matches the description.
-Their idea to tackle this problem is to form a three-way tensor product model.
-They use a ranking loss to train the model where the score of an observed triplet is higher than an unobserved triplet.
-They sample six negative triplets per positive sample triplet, and use each of them as a negative in the loss.
-The downstream retrieval task is then simply done by taking the arg max of the tensor product for a given user.
-\todo[inline]{This work differs from ours in that---what?
-Does it becomes computationally complex as additional modalities are added?
-The question is, could we not just apply their approach, and why not?}
-
-\citet{semihet_three_way_Lei_2020} use inputs from three modalities of image, sketch, and edgemap to perform image retrieval given the sketch.
-They use co-attention mechanism, and their loss function contains alignment loss, cross-entropy loss, and sketch-edgemap contrastive loss.
-\citet{Mittal2020M3ER} use Canonical Correlational Analysis (CCA) to differentiate between ineffective and effective modalities for the emotion recognition task from multiple input modalities including face, speech, and text.
-Multimodal learning has been used for different applications including but not limited to object retrieval.
-\citet{Deception_ICMI_2014} use \textit{language}, \textit{physiological response}, and \textit{thermal sensing} to detect deceit.
-\citet{het_data_fusion_liu_IEEE_2017} take two modalities as input and output predictions in a third modality.
-\citet{tursun2021efficient} uses quadruplet loss for two modalities only; image and sketch.
-A quadruplet is composed of a sketch picture as an anchor, a negative example from sketch domain, a negative example from picture domain,  and a positive example from picture domain.
-However, our method is not restricted to two modalities only.~\citet{chen2017beyond} also uses quadruplet loss, but does not support multiple domains.
-% \todo[inline]{talk about CLIP and DALL-E here}
-
-To the best of our knowledge, there has been limited work on incorporating more than three modalities into a general learning approach.
-Our proposed loss function closely resembles work on quadruplet loss~\cite{chen2017beyond,tursun2021efficient}
-% \todo{Beyond triplet loss: a deep quadruplet network for person re-identification, but no multi-domain -Ed}, 
-but is intended to scale to an arbitrary number of modalities.
-
-%===================================================================
-
-\section{Problem Description}
-\label{sec:Problem-Description}
-
-Given a language command either text or speech that describes an object, we want our model to retrieve the correct object from a set of objects.
-This problem is also known as grounded language learning in the fields of robotics and natural language processing.
-If we have a spoken language command $x_s$, a textual language command $x_t$, a set of RGB images $X_r = \{x_r^{(1)}, x_r^{(2)}, ..., x_r^{(n)}\}$, and a set of depth images $X_r = \{x_d^{(1)}, x_d^{(2)}, ..., x_d^{(n)}\}$ the task is to retrieve the correct object by choosing the index that has the minimum distance from either of the language commands across all modalities.
-We consider (I) ``sr'': the distance between $x_s$ and all RGB images in $X_r$, (II) ``sd'': the distance between $x_s$ and all depth images in $X_d$, (III) ``tr'': the distance between $x_t$ and all RGB images in $X_r$, and (IV) ``td'': the distance between $x_t$ and all depth images in $X_d$.
-We use the average of these distances to select the correct object.
-Depending on the missing/available sensors during test time, we might have any combination of these four distances.
-For example if text modality is missing during the test, we can take average of $sr$ and $sd$.
-If we have all four modalities, we take the average of $sr$, $sd$, $tr$, and $td$.
-This method allows us to extend our model to support arbitrary number of modalities.
-
-%===================================================================
-
-\section{Approach}
-\label{sec:Method}
-
-For $M$ modalities, we define a distance based loss function which is a generalization of well-known similarity-based metric learning and manifold alignment methods such as triplet loss~\cite{Carvalho-cooking-triplet,triplet_loss_2021_CVPR}, quadruplet loss~\cite{chen2017beyond}, and is similar to contrastive loss under some settings.
-We sample two data points from each modality, one positive point and one negative point. 
-For each set of positive data points (referring to an instance of an object with all of its modalities), we choose a set of negative data points either based on object class name, or by semantic negative sampling~\cite{Pillai_Matuszek_2018}.
-We now have two data points; we consider the first one as positive, and the second one as negative.
-However, the anchor won't be another data point as it is in the triplet loss terminology.
-Instead we choose one domain as an anchor.
-
-\begin{itemize}
-    % \item data point: a row in the csv/tsv file.
-    \item Positive: a set of embeddings of the one data point (rgb image of an apple, depth image of an apple, language description of an apple)
-    \item Negative: a set of embeddings of the another data point of a different object (rgb image of a book, depth image of a book, language description of a book)
-    \item Anchor: A fixed modality to act as anchor in the case of simple MMA loss, where we don't minimize the distance between the pair of embeddings from all modalities of the same data point, and we don't maximize the distance between positive an negative datapoints of the same modality.
-\end{itemize}
-
-The objective is then to (I) minimize the distance between each pair of positive points from heterogeneous modalities which results in $(M-1)!$ terms, (II) maximize the distance between each pair of positive and negative points from heterogeneous modalities which contains $(M-1)!$ terms again, and (III) maximize the distance between each pair of positive and negative points from the homogeneous modalities which contains $M$ terms.
-Altogether, our proposed loss function contains $M+2(M-1)!$ terms.
-Equation~\ref{eq:objective} shows the objective function we defined above with changing maximization to minimization by changing the sign.
-
-% \begin{equation}\label{eq:objective-detailed}
-%     \min \sum_{i=1}^{M-1} \sum_{j=i+1}^{M} ||x_{i}^{+} - x_{j}^{+}|| - \sum_{i=1}^{M-1} \sum_{j=i+1}^{N} ||x_{i}^{+} - x_{j}^{-}|| - 
-%     \sum_{i=1}^{M} ||x_{i}^{+} - x_{i}^{-}||
-% \end{equation}
-
-% \begin{equation}\label{eq:objective}
-% \begin{split}
-%     \mathcal{L}  &= \sum_{i=1}^{M-1} \sum_{j=i+1}^{M} ||f_i(x_{i}^{+}) - f_j(x_{j}^{+})|| \\
-%     & - ||f_i(x_{i}^{+}) - f_j(x_{j}^{-})||
-%     - ||f_i(x_{i}^{-}) - f_j(x_{j}^{+})|| \\
-%     & - \sum_{i=1}^{M} ||f_i(x_{i}^{+}) - f_i(x_{i}^{-}) ||
-% \end{split}
-% \end{equation}
-% where $f_i(.)$ and $f_j(.)$ represent functions that map input data to the shared latent space using a deep neural network.
-
-% d(,) notation instead of || - ||
-% \begin{equation}\label{eq:objective}
-% \begin{split}
-%     \mathcal{L}  &= \sum_{i=1}^{M-1} \sum_{j=i+1}^{M} d(z_{i}^{+}, z_{j}^{+}) 
-%      - d(z_{i}^{+}, z_{j}^{-}) \\ 
-%     & - d(z_{i}^{-}, z_{j}^{+}) - \sum_{i=1}^{M} d(z_{i}^{+}, z_{i}^{-})
-% \end{split}
-% \end{equation}
-
-\begin{equation}\label{eq:objective}
-\begin{split}
-    \mathcal{L}  &= \sum_{i=1}^{M-1} \sum_{j=i+1}^{M} ||z_{i}^{+} - z_{j}^{+}|| 
-     - ||z_{i}^{+} - z_{j}^{-}|| - ||z_{i}^{-} - z_{j}^{+}|| \\ 
-    &  - \sum_{i=1}^{M} ||z_{i}^{+} - z_{i}^{-} ||
-\end{split}
-\end{equation}
-where the subscripts $i$ and $j$ represent modalities, and $z$ is the embedding we get by applying a mapping function $f$ which in our case is a neural network on our input data.
-In other words $z_m = f_m(x_m)$ where each modality $m$ has a specific model $f_m$ that is different from the models for other modalities and these models do not share their weights.
-To measure the distance, we use cosine distance.
-In order to use cosine \textit{distance}, we have to subtract the cosine of the \textit{angle} between two embeddings (which represents similarity) from 1: $1 - \cos(e_1, e_2)$.
-
-\todokdinline{I can add another term for minimizing the distance between similar pairs (anchor and positive) from the same domain.
-This would look like the equation below.
-However, it breaks the nice procedure of sampling two data points per modality}
-\todo{Edward wonders if $f_i(x)$ should instead produce $\mu_x$ and $\Sigma_x$, to capture uncertainty of embedding location, when working with so many modalities?
-(e.g., not all modalities share the same ambiguities)}
-\todo{Does Edward mean the actual network produces $\Sigma_x$, e.g., something like a VAE? How would uncertainty be applied to the inference problem?}
-$+ \sum_{i=1}^{M} ||f_i(x_i^{a}) - f_i(x_i^{+})||$
-
-If $M=2$ which means the number of modalities is 2, then our objective function reduces to the quadruplet loss method~\cite{chen2017beyond} if we ignore the third term, and negative samples for each modality do not belong to the same class.
-If $M=2$ and we ignore the last two terms in the objective function, it results in the triplet loss method. 
-If $M=1$, only the last term remains in the loss function which is exactly the pairwise distance-based loss function.
-
-We demonstrate performance using three different distance-based loss functions.
-\todokdinline{Either remove it or experiment with all versions of MMA (12 terms, 9 terms, and explicit anchor maximization}
-
-\subsection{Simple MMA}
-\label{sub:simple-mma}
-Our first proposed loss function is a modification of the triplet loss function idea, and can be derived from equation~\ref{eq:objective} by fixing one modality as anchor, and ignoring the last two terms in equation~\ref{eq:objective}.
-
-% For each data point (referring to an instance of an object with all of its modalities), we choose a negative data point either based on object class name, or by semantic negative sampling~\cite{Pillai_Matuszek_2018}.
-We now have two data points; we consider the first one as positive, and the second one as negative.
-However, the anchor won't be another data point as it is in the triplet loss terminology.
-Instead we choose one domain as an anchor.
-
-% \begin{itemize}
-%     % \item data point: a row in the csv/tsv file.
-%     \item Positive: a set of embeddings of the one data point (rgb image of an apple, depth image of an apple, language description of an apple)
-%     \item Negative: a set of embeddings of the another data point of a different object (rgb image of a book, depth image of a book, language description of a book)
-%     \item Anchor: A fixed modality to act as anchor in the case of simple MMA loss, where we don't minimize the distance between the pair of embeddings from all modalities of the same data point, and we don't maximize the distance between positive an negative datapoints of the same modality.
-% \end{itemize}
-
-Since the downstream task in the grounded language learning domain is to predict/retrieve a desired object among multiple objects given a natural language description, it makes sense to train the model using language as an anchor, and in fact anchoring learning around language outperforms anchoring on RGB.
-
-Equation~\ref{eq:objective-simple-mma} shows the mathematical formulation of the proposed loss function which can be generalized to arbitrary number of modalities.
-
-% \begin{equation}
-% \label{eq:objective-simple-mma}
-% \begin{split}
-%     \mathcal{L}  &= \sum_{m=1}^{M-1} || f_a(x_{a}^{+}) - f_m(x_{m}^{+}) || - \\ 
-%     & || f_a(x_{a}^{+}) -f_m(x_{m}^{-}) || 
-%     \\
-%     &= \sum_{m=1}^{M-1} 1 - \cos(f_a(x_{a}^{+}), f_m(x_{m}^{+})) \\ & - (1 - \cos(f_a(x_{a}^{+}) ,f_m(x_{m}^{-}))) \\
-%     &= \sum_{m=1}^{M-1} \cos(f_a(x_{a}^{+}) ,f_m(x_{m}^{-})) \\ 
-%     & - \cos(f_a(x_{a}^{+}), f_m(x_{m}^{+}))
-% \end{split}
-% \end{equation}
-
-\begin{equation}
-\label{eq:objective-simple-mma}
-\begin{split}
-    \mathcal{L}  &= \sum_{m=1}^{M-1} || z_{a}^{+} - z_{m}^{+} || - || z_{a}^{+} - z_{m}^{-} || \\
-    &= \sum_{m=1}^{M-1} 1 - \cos(z_{a}^{+}, z_{m}^{+}) - (1 - \cos(z_{a}^{+}, z_{m}^{-})) \\
-    &= \sum_{m=1}^{M-1} \cos(z_{a}^{+} ,z_{m}^{-}) - \cos(z_{a}^{+}, z_{m}^{+})
-\end{split}
-\end{equation}
-
-where subscript $a$ represents \textit{anchor} modality, $M$ is the number of modalities, superscript $+$ represents the positive objects, and superscript $-$ represents the negative objects.
-
-An illustration of this loss function is provided in figure~\ref{fig:simple-mma-loss}.
-
-\begin{figure}[tb]
-\centering
-\includegraphics[width=\columnwidth]{Figures/simple-MMA.pdf}
-\caption{A high-level prototype of the distances used in the simple MMA loss with 3 modalities.
-Orange arrows indicate maximizing the distance, blue arrows indicate minimizing the distance, and green lines show the margin.
-Depth images are brightened to be visible.
-\todo[inline]{We need to show speech as the 4th modality as well.}}
-\label{fig:simple-mma-loss}
-\end{figure}
-
-This version of MMA loss function can be seen as a contrastive loss usually used in the domain of self-supervised learning~\cite{chen2020simple}.
-However, our proposed loss function has two advantages over the traditional contrastive loss expressed in equation~\ref{eq:contrastive-loss}.
-The first advantage is that our loss function do not loop over multiple positives and negatives in a large batch.
-Instead we sample only two objects (positive and negative) each of which have $M$ modalities which gives us $2M$ datapoints (or embeddings).
-Hence, our model can be trained using smaller batch sizes and reduces the number of negative samples we need.
-The second advantage is that this loss function can be used in a multimodal setting with an arbitrary number of modalities, and is not limited to a single data type (e.g.
-RGB images) which is the most common usage of contrastive loss~\cite{chen2020simple, NEURIPS2020_supervised_contrastive}.
-
-Therefore, our proposed loss function would modify the equation~\ref{eq:contrastive-loss} to the loss function formulated in equation~\ref{eq:simple-MMA-contrastive}.
-Our formulation is a generalized version of the contrastive loss where $M$ is the number of modalities (e.g.
-RGB image, depth image, speech, text), but it can be the ``augmented views'' instead of different modalities.
-
-% \begin{equation}\label{eq:simple-MMA-contrastive}
-%     -\sum_{m=1}^{M-1} \log \frac{\exp (sim(f_a(x_a^+) , f_m(x_{m}^{+})/ \tau) )}{ \exp (sim(f_a(x_a^+) , f_m(x_{m}^{+}) / \tau)) + \exp (sim(f_a(x_a^+), f_m(x_{m}^{-}) / \tau))}
-% \end{equation}
-
-\begin{equation}\label{eq:simple-MMA-contrastive}
-    -\sum_{m=1}^{M-1} \log \frac{\exp (sim(z_a^+ , z_{m}^{+})/ \tau) }{ \exp (sim(z_a^+ , z_{m}^{+}) / \tau) + \exp (sim(z_a^+, z_{m}^{-}) / \tau)}
-\end{equation}
-where $z = f(x)$.
-
-% \todokdinline{I should run the traditional contrastive loss to see if my method works better to make this claim.}
-
-% \subsubsection{All in loss with 12 terms}
-% \subsubsection{subset of All in loss with 9 terms}
-
-\subsection{EMMA}
-\label{sub:emma}
-In this version, we add the last term in equation~\ref{eq:objective} which maximizes the distance between positive and negative instances from the same modality.
-However, instead of looping over all modalities, we only do it for the anchor modality.
-The reason for that is because a margin between positive and negative points from other modalities (excluding anchor) is enforced.
-We formulate this method in equation~\ref{eq:objective-emma}.
-
-This approaches closes the gap against supervised contrastive learning method when dropping speech, and increases the improvement even more when dropping text.
-
-\begin{equation}
-\label{eq:objective-emma}
-\begin{split}
-    \mathcal{L} = \sum_{m=1}^{M-1} \cos(z_{a}^{+} ,z_{m}^{-}) - \cos(z_{a}^{+}, z_{m}^{+}) - \cos(z_{a}^{+}, z_{a}^{-})
-\end{split}
-\end{equation}
-
-\subsection{Network Architecture}
-\label{sec:Model}
-
-We use BERT~\cite{devlin-etal-2019-bert} embeddings from flair~\cite{akbik2019flair,akbik-etal-2019-pooled} to featurize the textual input, and wav2vec2~\cite{wav2vec2} to extract audio embeddings from speech both of which output a 3072-dimensional embedding vector.
-To process images, we use ResNet152~\cite{He_resnet_2016} for both RGB and depth images which gives us a 2048-dimensional embedding vector.
-We then use 3 fully connected layers with ReLU activation function~\cite{relu2010} to map these embeddings to a shared 1024-dimensional space where we can compute the distance between all embeddings.
-
-%===================================================================
-
-\section{Experiments}
-\label{sec:Experiments}
-
-We test our model against different baselines in different scenarios and settings using multiple metrics.
-
-\subsection{Data}
-\label{sec:Data}
-
-We use a recent multimodal dataset called GoLD~\cite{GoLD_UMBC}.
-The original paper uses raw RGB and raw depth images in which other objects are present in the background.
-We use a masked version of the images where the background (including the white turntable) is deleted and one object is present in each image.
-We show that a masked version converges faster, however, they both converge to the same performance as shown in the supplementary material in figure~\ref{fig:mask-vs-raw}.
-Speech from GoLD is converted to 16 Hz in order to match wav2vec2 speech featurization model~\cite{wav2vec2}.
-
-\subsection{Metrics}
-\label{sec:metrics}
-To evaluate our model we measure different performance metrics on a retrieval task where the model has to select an object among 5 objects given a language description.
-Only one of the objects corresponds to the description and the rest are from different object classes (e.g., one apple among a set including a fork, a mug, a lemon, and a bell pepper).
-
-% \todokdinline{Try with different number of objects.}
-
-To evaluate the performance, we compute the cosine distance between the given natural language description and 5 randomly selected images (1 of which corresponds to the description, with the others from different object classes).
-We handle depth images as a separate input modality (as opposed to appending depth to RGB, as others have done~\cite{triplet_loss_2021_CVPR}).
-We compute the distance between one textual language embedding and all candidate RGB embeddings, and we compute the distance between the same language embedding and all candidate depth embeddings corresponding to the RGB embeddings.
-We then take average of these two distance matrices.
-Instead of choosing an empirical threshold beyond which objects are considered to be `referred to,' we choose the closest image embedding (average distance of RGB and depth from language) as the prediction.
-
-The best metric to capture the performance in such a scenario is mean reciprocal rank (MRR), in which the inverse of the rank of the correct object for a set of queries and then averaged.
-Accuracy, micro F1 score (sklearn), flattened binary f1 score (sklearn), and F1 score computation are all the same in this task, since for each prediction we either have a true positive and no false positives and no false negatives (no FP no FN), or we have no true positives and 1 false positive and 1 false negative.
-Subset MRR is equal or greater than those, because if the model does not rank the correct answer as 1, it still gets some score and is greater than 0 (in accuracy the score is ether 0 or 1).
-
-\subsection{Setup}
-\label{sec:setup}
-
-Similar to~\citet{NEURIPS2020_supervised_contrastive}, we also use stochastic gradient descent (SGD) optimizer with momentum~\cite{ruder2016overviewSGD} with a flexible learning rate starting with 0.05.
-The models are trained for 200 epochs with a batch size of 64 on a Quadro RTX 8000 GPU.
-We use the same setup for all our experiments to have a fair comparison between different methods.
-
-\subsection{4 Modalities}
-We consider an experiment in which we incorporate RGB, depth, speech, and written language.
-The loss function requires no changes beyond increasing the number of loops in equation~\ref{eq:objective-simple-mma} (or increasing the value of $M$ by 1).
-Our goal is the non-trivial downstream prediction task: determining what objects are being referred to by arbitrary language from a small number of examples.
-In the case of three modalities, written language is used the anchor, and we compute the distance of RGB and depth modalities from it and then average them.
-However, when speech is incorporated as an additional sensory modality, we have three immediate possible choices.
-\textit{First}, we can treat speech in a similar way to RGB and depth: compute the distance of RGB, depth, and speech from text, and then take an average of three of them.
-\textit{Second}, we can compute the distance of RGB and depth from language and from speech which gives us 4 distance matrices, and then take average of these four.
-\textit{Third}, similar to the second method, we can compute the distance between language and speech as well and then take the average of 5 distance matrices.
-
-Of these, the first method is both more consistent with the goal of making a computationally tractable approach to arbitrary input modalities, and more consistent with existing work on using featurized speech directly as a sensory learning input; the simple MMA loss function in equation~\ref{eq:objective-simple-mma} is not designed to minimize the distance between speech and visual embeddings explicitly.
-This is supported by empirical results, in which the first method results in a better performance compared to the other two methods.
-% Third method (averaging 5 matrices) is the worst, then second one (averaging 4 matrices) is bad. First method has the best results.
-
-\todokdinline{try the more complicated MMA that might be able to capture these relationships.}
-
-\subsection{Baselines}
-We compare our MMA model against different contrastive learning methods including the triplet loss~\cite{GoLD_UMBC, triplet_loss_2021_CVPR}, the traditional version of contrastive loss usually used in self-supervised settings~\cite{chen2020simple}, and supervised contrastive learning~\cite{NEURIPS2020_supervised_contrastive}.
-The triplet loss function consists of three data points including anchor, positive, and negative from two modalities.
-The anchor, positive, and negative can be chosen from different modalities in each batch.
-The triplet loss method has two major disadvantages.
-First, it cannot be used for more than three modalities since there are only 3 data points in the loss function.
-Second, a batch size of greater than 1 cannot be easily implemented if the anchor, positive, and negative come from different modalities for each batch.
-To address these issues, we sample more than one positive and one negative data points.
-Essentially, we sample one positive data point and one negative data point from each modality which results in 2(M-1) + 1 data points where 1 is the single anchor data point from anchor modality of the same object class as positive datapoints.
-We would have 2M datapoints if we use the negative data point from the anchor modality too.
-The anchor data point is always from the same modality and corresponds to the same object class sampled for positive data points.
-
-\subsubsection{Contrastive Loss}
-\label{sub:baseline-contrastive}
-
-We compare our model against contrastive loss as formulated in equation~\ref{eq:contrastive-loss}.
-In order to implement this loss function, we use cosine similarity as suggested in the SimCLR paper~\cite{chen2020simple}.
-The other version (mentioned by~\citet{NEURIPS2020_supervised_contrastive}) uses an inner dot product that can lead to instabilities since the dot product is not bounded.
-% \todokdinline{try contrastive loss with cross-entropy, dot product, and the supervised contrastive method itself (without triplets).}
-
-% Original Formula
-\begin{equation}\label{eq:contrastive-loss}
-    -\sum_{i \in I} \log \frac{\exp (sim(z_i , z_{j(i)}) / \tau) }{\sum_{a \in A(i)} \exp (sim(z_i, z_a) / \tau)}
-\end{equation}
-
-% \begin{equation}\label{eq:contrastive-loss}
-%     -\sum_{i \in I} \log \frac{\exp (sim(f(x_i) , z_{j(i)}) / \tau) }{\sum_{a \in A(i)} \exp (sim(f(x_i), f(x_a)) / \tau)}
-% \end{equation}
-where $i$ is anchor, $j(i)$ is the set of positives excluding anchor, $a$ is the set of all positives and negatives excluding anchor, and $z = f(x)$.
-
-\subsubsection{Supervised Contrastive Learning}
-\label{sub:baseline-supcon}
-
-\citet{NEURIPS2020_supervised_contrastive} propose a supervised way of performing contrastive learning which is shown in equation~\ref{eq:supervised-contrastive}.
-
-\begin{equation}\label{eq:supervised-contrastive}
-    \sum_{i \in I} \frac{-1}{|P(i)|} \sum_{p \in P(i)} \log \frac{\exp (z_i \cdot z_p / \tau) }{\sum_{a \in A(i)} \exp (z_i \cdot z_a / \tau)}
-\end{equation}
-
-% \begin{equation}\label{eq:supervised-contrastive-my-notation}
-%     \sum_{i \in I} \frac{-1}{|P(i)|} \sum_{p \in P(i)} \log \frac{\exp (f_i(x_i) \cdot f_p(x_p) / \tau) }{\sum_{a \in A(i)} \exp (f_i(x_i) \cdot f_a(x_a) / \tau)}
-% \end{equation}
-
-While this model is a strong baseline and the performance of our model is very close to this model, if the text modality is dropped, this method does not perform as good as our proposed model as shown in figure~\ref{fig:simple-MMA-mrr.srd}.
-
-\begin{figure}[tbh]
-\centering
-\includegraphics[width=.99\columnwidth]{Figures/average-seeds-epochs-mrr_ard.pdf}
-\caption{Mean Reciprocal Rank (MRR) on the held-out test set when text modality is missing averaged over 3 runs with different random seeds.
-The higher is better.
-Green is Supervised Contrastive Learning~\cite{NEURIPS2020_supervised_contrastive} and pink represents our proposed simple MMA loss function without semantic negative sampling w/ SGD w/ adjusting learning rate.
-% \todokdinline{Add legend to the graph.}
-}
-
-\label{fig:simple-MMA-mrr.srd}
-\end{figure}
-
-\subsection{Quantitative Results}
-\label{sec:Quantitative}
-
-We evaluate our proposed simple MMA loss function~\ref{sub:simple-mma} against the triplet loss method~\cite{triplet_loss_2021_CVPR}.
-Our model significantly outperforms the triplet method across all epochs while it converges faster as shown in figure~\ref{fig:simple-MMA-mrr}.
-Table~\ref{table:quantitative} summarizes the performance of all models.
-
-\begin{table*}[tbh]
-\centering
-\begin{tabular}{|c|c|c|c|c|c|}
-\toprule
-\textbf{Method} & \textbf{Modalities} & \textbf{Accuracy} & \textbf{MRR.srd} & \textbf{MRR.trd} & \textbf{MRR.tsrd} \\ %\hline
-\midrule
-Simple MMA (Text) w/o neg  SGD & 3 & 0.8168$\pm$? & 0.8962$\pm$? & ? & ? \\
-eMMA (Text) w/o neg SGD & 3 & 0.8168$\pm$? & 0.8962$\pm$? & ? & ? \\
-Contrastive Cosine (Text) w/ neg & 3 & 0.7894$\pm$? & 0.8788$\pm$?  & ? & ?\\
-Supervised Contrastive (Text)& 3 & 0.8332$\pm$? & 0.9057$\pm$? & ? & ?\\
-Simple MMA (Text) w/o neg  SGD & 4 & ?$\pm$? & ?$\pm$? & ? & ? \\
-Supervised Contrastive (Text)& 4 & ?$\pm$? & ?$\pm$? & ? & ? \\
-Simple MMA (Speech) w/o neg & 3 & 0.7949$\pm$? & 0.8778$\pm$? & ? & ?  \\
-Triplet (Speech) w/o neg & 3 & 0.7497$\pm$? & 0.8522$\pm$? & ? & -  \\
-% Full MMA (Text) w/ neg & 3 & ?$\pm$? & ?$\pm$? \\
-% Simple MMA (Text) w/ neg Adam & 3  & 0.7949$\pm$? & 0.8778$\pm$? \\
-% Simple MMA (Text) w/ neg SGD & 3 & 0.7905$\pm$? & 0.8821$\pm$? \\
-% Triplet (Text) w/ neg & 3 & 0.7497$\pm$? & 0.8522$\pm$? \\
-% Full MMA (Speech) w/ neg & 3 & ?$\pm$? & ?$\pm$? \\
-% Simple MMA (Speech) w/ neg & 3 & 0.7949$\pm$? & 0.8778$\pm$? \\
-% Contrastive (Speech) w/ neg & 3 & ?$\pm$? & ?$\pm$? \\
-% Triplet (Speech) w/ neg & 3 & 0.7497$\pm$? & 0.8522$\pm$? \\
-\bottomrule
-\end{tabular}
-\caption{\label{table:quantitative}Average and standard deviation of accuracy and MRR over 3 runs with 3 different random seeds on a held-out test set.
-Both metrics are from 0 to 1, and the higher is better.
-% \todokdinline{maybe adding results on cropped and raw dataset as well. Also results using object class instead of negative sampling}
-}
-\end{table*}
-
-\begin{figure}[tbh]
-\centering
-\includegraphics[width=.99\columnwidth]{Figures/average-seeds-epochs-mrr_lrd.pdf}
-\caption{Mean Reciprocal Rank (MRR) on the held-out test set when speech modality is dropped averaged over 3 runs with different random seeds for a downstream task of object retrieval.
-The higher is better.
-Orange represents Supervised Contrastive Learning~\cite{NEURIPS2020_supervised_contrastive}, \Complete{} represents the triplet loss function~\cite{triplet_loss_2021_CVPR}.
-\todokdinline{remove simple MMA and probably add triplet loss. Any ideas?}
-}
-% \todocmi{Maybe MRR is wrong and we should be looking at top-ranked result?}
-\label{fig:simple-MMA-mrr}
-\end{figure}
-
-\subsection{Qualitative Results}
-\label{sec:Qualitative}
-
-\begin{enumerate}
-    \item DONE Perform t-sne visualization of embeddings.
-    \item TODO Dropout modalities MRR
-\end{enumerate}
-
-\subsection{Failed Experiments}
-We tried a lot of different approaches and they did not work better than our model or the baselines, but we list them here to save time for readers who want to try them.
-
-\subsubsection{Two Anchors}
-Since speech is similar to text in its usage when it comes to the downstream task of object retrieval (given a single language command, find the correct object among multiple objects), we decided to have text and speech as two anchors instead of having one anchor only.
-When there is one anchor, the distance between positive pairs text and speech is minimized and the distance between negative pairs are minimized.
-However, this is not related to the downstream task since we never want to find the correct speech for a given text or vice versa.
-The second term added to the last function is exactly similar to the loss function when speech is the anchor, and the first term is the same as when language is the anchor.
-Therefore, the performance is somewhere in the middle of those two approaches but more towards the speech when the metrics are computed based and speech matrices.
-
-%===================================================================
-
-\section{Conclusion}
-\label{sec:conclusion}
-
-%===================================================================
-
-% ***************** The END of the paper *****************
-
-% %% The acknowledgments section is defined using the "acks" environment
-% %% (and NOT an unnumbered section). This ensures the proper
-% %% identification of the section in the article metadata, and the
-% %% consistent spelling of the heading.
-\begin{acks}
-To Robert, for the bagels and explaining CMYK and color spaces.
-\end{acks}
-% \section*{Acknowledgments}
-
-\clearpage
-\clearpage
-% %%
-% %% The next two lines define the bibliography style to be used, and
-% %% the bibliography file.
-\bibliographystyle{ACM-Reference-Format}
-\bibliography{ref}
-
-% %%
-% %% If your work has an appendix, this is the place to put it.
-\appendix
-\label{sec:appendix}
-
-\section{Qualitative Analysis}
-\section{Model}
-
-\section{Data}
-
-\begin{figure*}[tbh]
-\centering
-\includegraphics[width=2.0\columnwidth]{Figures/raw-mask-test-mrr.pdf}
-\caption{Mean Reciprocal Rank (MRR) on different versions of GoLD dataset~\cite{GoLD_UMBC}.
-The raw version is depicted in green and masked version is depicted in orange averaged over 3 different random seeds.
-Higher is better.}
-\label{fig:mask-vs-raw}
-\end{figure*}
-
-\section{Experiments}
-
-\section{Results}
-
-\begin{figure*}[tbh]
-\centering
-\includegraphics[width=2.1\columnwidth]{Figures/test.mrr.pdf}
-\caption{Mean Reciprocal Rank (MRR) of our proposed method against triplet loss method averaged over 3 different random seeds for a downstream task of object detection.
-The higher is better.
-Green is Supervised Contrastive Learning~\cite{NEURIPS2020_supervised_contrastive}, Pink represents our proposed simple MMA loss function without semantic negative sampling w/ SGD w/ adjusting learning rate, Purple is out method with semantic negative sampling, Blue is our method w/ semantic negative sampling w/ Adam and w/o adjusting learning rate, Red is vanilla contrastive loss with cosine, and orange represent the triplet loss.
-function~\cite{triplet_loss_2021_CVPR}.
-}
-
-\label{fig:simple-MMA-mrr}
-\end{figure*}
-
-\section{Setup}
-
-% \input{lit-search}
-
-\end{document}
+method: Geometric, best mrr_ad ** avg: 0.7682305107526883 std: 0.0033724197709617586
+method: Geometric, best mrr_ar ** avg: 0.783393817204301 std: 0.002870980798921243
+method: Geometric, best mrr_ld ** avg: 0.8963622311827957 std: 0.0038165384708184025
+method: Geometric, best mrr_lr ** avg: 0.9112627688172041 std: 0.0073097134335746074
+method: Geometric, best mrr_lad ** avg: 0.8920712365591397 std: 0.004478468907650514
+method: Geometric, best mrr_lar ** avg: 0.9095463709677419 std: 0.008254095620938519
+method: Geometric, best mrr_ard ** avg: 0.7937331989247312 std: 0.002892081483060137
+method: Geometric, best mrr_lrd ** avg: 0.9229038978494624 std: 0.005072886474275548
+method: Geometric, best mrr_lard ** avg: 0.9214092741935485 std: 0.004504231901153245
+method: Geometric, best acc_ad ** avg: 0.6194758064516129 std: 0.005500153720300029
+method: Geometric, best acc_ar ** avg: 0.6434274193548386 std: 0.005332041933055135
+method: Geometric, best acc_ld ** avg: 0.8202822580645162 std: 0.0057175656281111785
+method: Geometric, best acc_lr ** avg: 0.8459677419354839 std: 0.010990399982684721
+method: Geometric, best acc_lad ** avg: 0.8108064516129032 std: 0.008065120945062159
+method: Geometric, best acc_lar ** avg: 0.8400000000000001 std: 0.014046271275059837
+method: Geometric, best acc_ard ** avg: 0.6583870967741936 std: 0.006289030057447286
+method: Geometric, best acc_lrd ** avg: 0.8641129032258064 std: 0.008277431297605348
+method: Geometric, best acc_lard ** avg: 0.8593548387096774 std: 0.007356633552926914
+method: SupCon, best mrr_ad ** avg: 0.7818461021505375 std: 0.0058056222339268986
+method: SupCon, best mrr_ar ** avg: 0.7968904569892473 std: 0.005438750540616555
+method: SupCon, best mrr_ld ** avg: 0.8903924731182796 std: 0.008784518487631803
+method: SupCon, best mrr_lr ** avg: 0.9056451612903225 std: 0.007443727195959732
+method: SupCon, best mrr_lad ** avg: 0.8875026881720431 std: 0.0065906156485880715
+method: SupCon, best mrr_lar ** avg: 0.9049993279569893 std: 0.006934618496482506
+method: SupCon, best mrr_ard ** avg: 0.8120443548387095 std: 0.0038831681475468686
+method: SupCon, best mrr_lrd ** avg: 0.919635752688172 std: 0.00421253081123992
+method: SupCon, best mrr_lard ** avg: 0.9202607526881721 std: 0.007036465744801571
+method: SupCon, best acc_ad ** avg: 0.6417338709677419 std: 0.009174605555378065
+method: SupCon, best acc_ar ** avg: 0.6652016129032258 std: 0.010682516325605402
+method: SupCon, best acc_ld ** avg: 0.8105241935483869 std: 0.012173413141473797
+method: SupCon, best acc_lr ** avg: 0.8364516129032259 std: 0.014036892078731987
+method: SupCon, best acc_lad ** avg: 0.8058467741935484 std: 0.011178885534336718
+method: SupCon, best acc_lar ** avg: 0.8354435483870969 std: 0.012324733809520844
+method: SupCon, best acc_ard ** avg: 0.6869758064516129 std: 0.006569735184573384
+method: SupCon, best acc_lrd ** avg: 0.8605645161290323 std: 0.012074307093137743
+method: SupCon, best acc_lard ** avg: 0.8581854838709677 std: 0.01293003766775218
+method: EMMA, best mrr_ad ** avg: 0.7763420698924731 std: 0.0029124808921936005
+method: EMMA, best mrr_ar ** avg: 0.786581317204301 std: 0.006418196740535436
+method: EMMA, best mrr_ld ** avg: 0.8986505376344086 std: 0.005024686589459191
+method: EMMA, best mrr_lr ** avg: 0.9125739247311827 std: 0.008583791522680621
+method: EMMA, best mrr_lad ** avg: 0.8966283602150538 std: 0.0036157579057873954
+method: EMMA, best mrr_lar ** avg: 0.9097157258064514 std: 0.00655691760402764
+method: EMMA, best mrr_ard ** avg: 0.8032096774193549 std: 0.004465339908648459
+method: EMMA, best mrr_lrd ** avg: 0.9270524193548388 std: 0.004950684287669308
+method: EMMA, best mrr_lard ** avg: 0.9272177419354838 std: 0.0046582728067974
+method: EMMA, best acc_ad ** avg: 0.6353629032258065 std: 0.00531983066789421
+method: EMMA, best acc_ar ** avg: 0.6506854838709677 std: 0.010086772331454559
+method: EMMA, best acc_ld ** avg: 0.8277822580645161 std: 0.009658415480510912
+method: EMMA, best acc_lr ** avg: 0.850725806451613 std: 0.014181056636432354
+method: EMMA, best acc_lad ** avg: 0.8216129032258064 std: 0.006363306949864157
+method: EMMA, best acc_lar ** avg: 0.8437096774193549 std: 0.012343847760054961
+method: EMMA, best acc_ard ** avg: 0.6768951612903227 std: 0.008115363892853094
+method: EMMA, best acc_lrd ** avg: 0.8737903225806452 std: 0.007140626974573769
+method: EMMA, best acc_lard ** avg: 0.8715322580645161 std: 0.007248425499130212
+method: Contrastive, best mrr_ad ** avg: 0.7174408602150538 std: 0.007255402799973243
+method: Contrastive, best mrr_ar ** avg: 0.7337029569892473 std: 0.0039224209303724234
+method: Contrastive, best mrr_ld ** avg: 0.8972479838709677 std: 0.005357753897962892
+method: Contrastive, best mrr_lr ** avg: 0.9082345430107527 std: 0.003749222854686774
+method: Contrastive, best mrr_lad ** avg: 0.8912668010752688 std: 0.006122357196224834
+method: Contrastive, best mrr_lar ** avg: 0.9026169354838709 std: 0.005789581070670862
+method: Contrastive, best mrr_ard ** avg: 0.7496478494623655 std: 0.004373982216482813
+method: Contrastive, best mrr_lrd ** avg: 0.9192137096774193 std: 0.00411868564896933
+method: Contrastive, best mrr_lard ** avg: 0.9171935483870965 std: 0.005344248282948013
+method: Contrastive, best acc_ad ** avg: 0.5482258064516129 std: 0.014024724525739987
+method: Contrastive, best acc_ar ** avg: 0.5727016129032259 std: 0.006437231955968572
+method: Contrastive, best acc_ld ** avg: 0.82875 std: 0.008802521869277298
+method: Contrastive, best acc_lr ** avg: 0.8434677419354838 std: 0.01012939823665928
+method: Contrastive, best acc_lad ** avg: 0.8155241935483872 std: 0.00933622313545278
+method: Contrastive, best acc_lar ** avg: 0.8326209677419355 std: 0.010162090342290321
+method: Contrastive, best acc_ard ** avg: 0.5938306451612902 std: 0.005993840056794196
+method: Contrastive, best acc_lrd ** avg: 0.8631048387096774 std: 0.006685527646846855
+method: Contrastive, best acc_lard ** avg: 0.8574999999999999 std: 0.008746747574507578

### 2022-10-06 Paper/appendix.tex
Paper submitted to TMLR

@@
-\section{Qualitative Analysis}
-\section{Approach}
+\section{Loss Function Development Story}
 
-For $M$ modalities, we define a distance based loss function which is a generalization of well-known similarity-based metric learning and manifold alignment methods such as triplet loss~\cite{Carvalho-cooking-triplet,triplet_loss_2021_CVPR}, quadruplet loss~\cite{chen2017beyond}, and is similar to contrastive loss under some settings.
-We sample two data points from each modality, one positive point and one negative point. 
-For each set of positive data points (referring to an instance of an object with all of its modalities), we choose a set of negative data points either based on object class name, or by semantic negative sampling~\cite{Pillai_Matuszek_2018}.
-We now have two data points; we consider the first one as positive, and the second one as negative.
-However, the anchor won't be another data point as it is in the triplet loss terminology.
-Instead we choose one domain as an anchor.
-
-\begin{itemize}
-    % \item data point: a row in the csv/tsv file.
-    \item Positive: a set of embeddings of the one data point (RGB image of an apple, corresponding depth image, text description, and speech description of the same apple)
-    \item Negative: a set of embeddings of the another data point of a different object (RGB image of a book, corresponding depth image, text description, and speech description of the same book)
-    \item Anchor: A fixed modality to act as anchor in the case of simple MMA loss, where we don't minimize the distance between the pair of embeddings from all modalities of the same data point, and we don't maximize the distance between positive an negative datapoints of the same modality.
-\end{itemize}
-
-The objective is then to (I) minimize the distance between each pair of positive points from heterogeneous modalities which results in $(M-1)!$ terms, (II) maximize the distance between each pair of positive and negative points from heterogeneous modalities which contains $(M-1)!$ terms again, and (III) maximize the distance between each pair of positive and negative points from the homogeneous modalities which contains $M$ terms.
-Altogether, our proposed loss function contains $M+2(M-1)!$ terms.
-\Cref{eq:objective} shows the objective function we defined above with changing maximization to minimization by changing the sign.
-
-% \begin{equation}\label{eq:objective-detailed}
-%     \min \sum_{i=1}^{M-1} \sum_{j=i+1}^{M} ||x_{i}^{+} - x_{j}^{+}|| - \sum_{i=1}^{M-1} \sum_{j=i+1}^{N} ||x_{i}^{+} - x_{j}^{-}|| - 
-%     \sum_{i=1}^{M} ||x_{i}^{+} - x_{i}^{-}||
-% \end{equation}
-
-% \begin{equation}\label{eq:objective}
-% \begin{split}
-%     \mathcal{L}  &= \sum_{i=1}^{M-1} \sum_{j=i+1}^{M} ||f_i(x_{i}^{+}) - f_j(x_{j}^{+})|| \\
-%     & - ||f_i(x_{i}^{+}) - f_j(x_{j}^{-})||
-%     - ||f_i(x_{i}^{-}) - f_j(x_{j}^{+})|| \\
-%     & - \sum_{i=1}^{M} ||f_i(x_{i}^{+}) - f_i(x_{i}^{-}) ||
-% \end{split}
-% \end{equation}
-% where $f_i(.)$ and $f_j(.)$ represent functions that map input data to the shared latent space using a deep neural network.
-
-% d(,) notation instead of || - ||
-% \begin{equation}\label{eq:objective}
-% \begin{split}
-%     \mathcal{L}  &= \sum_{i=1}^{M-1} \sum_{j=i+1}^{M} d(z_{i}^{+}, z_{j}^{+}) 
-%      - d(z_{i}^{+}, z_{j}^{-}) \\ 
-%     & - d(z_{i}^{-}, z_{j}^{+}) - \sum_{i=1}^{M} d(z_{i}^{+}, z_{i}^{-})
-% \end{split}
-% \end{equation}
-
+Our proposed \geom{} loss function can be rewritten as shown in \cref{eq:objective}.
 \begin{equation}\label{eq:objective}
 \begin{split}
-    \mathcal{L}  &= \sum_{i=1}^{M-1} \sum_{j=i+1}^{M} ||z_{i}^{+} - z_{j}^{+}|| 
-     - ||z_{i}^{+} - z_{j}^{-}|| - ||z_{i}^{-} - z_{j}^{+}|| \\ 
-    &  - \sum_{i=1}^{M} ||z_{i}^{+} - z_{i}^{-} ||
+    \mathcal{L}  &= \sum_{i=1}^{M-1} \sum_{j=i+1}^{M} dist( z_{i}^{+} , z_{j}^{+}) 
+     - dist( z_{i}^{+} , z_{j}^{-}) - dist( z_{i}^{-} , z_{j}^{+}) - \sum_{i=1}^{M} dist( z_{i}^{+} , z_{i}^{-} )
 \end{split}
 \end{equation}
-where the subscripts $i$ and $j$ represent modalities, and $z$ is the embedding we get by applying a mapping function $f$ which in our case is a neural network on our input data.
-In other words $z_m = f_m(x_m)$ where each modality $m$ has a specific model $f_m$ that is different from the models for other modalities and these models do not share their weights.
-To measure the distance, we use cosine distance.
+
+We designed and experimented with alternative versions of our \geom{} loss function to analyze if they offer better results.
+We start with a modification of the triplet loss function idea where we fix one modality as anchor and do not sample a positive instance from the same modality. 
+Na\"ively, to extend triplet loss to work with arbitrary number of modalities, we can use two anchors (e.g., text and speech), which results in $2(M-2)$ triplet losses as formulated in~\cref{eq:objective-two-anchors}.
+
+% \todoffinline{in this equation block, you're writing things like $|| z_{t}^{+} - z_{m}^{+} ||$. Do you actually mean norm, or would any old scoring function work?}
+
+\begin{equation}
+\label{eq:objective-two-anchors}
+\begin{split}
+    \mathcal{L}  &= \sum_{m=1}^{M-2} dist( z_{t}^{+} , z_{m}^{+} ) - dist( z_{t}^{+} , z_{m}^{-} ) + dist( z_{s}^{+} , z_{m}^{+} ) - dist( z_{s}^{+} , z_{m}^{-} ) \\
+    &= \sum_{m=1}^{M-2} \cos(z_{t}^{+} ,z_{m}^{-}) - \cos(z_{t}^{+}, z_{m}^{+}) + \cos(z_{s}^{+} ,z_{m}^{-}) - \cos(z_{s}^{+}, z_{m}^{+})
+\end{split}
+\end{equation}
+where $M$ is the number of modalities, $t$ represents text modality, $s$ represents speech modality, and the superscripts $+$ and $-$ represent positive and negative objects, respectively.
+The $\cos(\cdot)$ function is a measure of similarity, not distance, and that is why the signs are reversed. To measure the distance, we use cosine distance.
 In order to use cosine \textit{distance}, we have to subtract the cosine of the \textit{angle} between two embeddings (which represents similarity) from 1: $1 - \cos(e_1, e_2)$.
 
-\todokdinline{I can add another term for minimizing the distance between similar pairs (anchor and positive) from the same domain.
-This would look like the equation below.
-However, it breaks the nice procedure of sampling two data points per modality}
-\todo{Edward wonders if $f_i(x)$ should instead produce $\mu_x$ and $\Sigma_x$, to capture uncertainty of embedding location, when working with so many modalities?
-(e.g., not all modalities share the same ambiguities)}
-\todo{Does Edward mean the actual network produces $\Sigma_x$, e.g., something like a VAE? How would uncertainty be applied to the inference problem?}
-$+ \sum_{i=1}^{M} ||f_i(x_i^{a}) - f_i(x_i^{+})||$
+However, this approach performs poorly when one of the anchor modalities is ablated; since there is no explicit minimization between the two anchors, text (t) and speech (s) do not necessarily map closely to each other, such that only one of the two anchors is actually learned.
 
-If $M=2$ which means the number of modalities is 2, then our objective function reduces to the quadruplet loss method~\cite{chen2017beyond} if we ignore the third term, and negative samples for each modality do not belong to the same class.
-If $M=2$ and we ignore the last two terms in the objective function, it results in the triplet loss method. 
-If $M=1$, only the last term remains in the loss function which is exactly the pairwise distance-based loss function.
-
-We demonstrate performance using three different distance-based loss functions.
-\todokdinline{Either remove it or experiment with all versions of MMA (12 terms, 9 terms, and explicit anchor maximization}
-
-\subsection{Simple MMA}
-\label{sub:simple-mma}
-Our first proposed loss function is a modification of the triplet loss function idea, and can be derived from \cref{eq:objective} by fixing one modality as anchor, and ignoring the last two terms in \cref{eq:objective}.
-
-% For each data point (referring to an instance of an object with all of its modalities), we choose a negative data point either based on object class name, or by semantic negative sampling~\cite{Pillai_Matuszek_2018}.
-We now have two data points; we consider the first one as positive, and the second one as negative.
-However, the anchor won't be another data point as it is in the triplet loss terminology.
-Instead we choose one domain as an anchor.
-
-% \begin{itemize}
-%     % \item data point: a row in the csv/tsv file.
-%     \item Positive: a set of embeddings of the one data point (RGB image of an apple, depth image of an apple, language description of an apple)
-%     \item Negative: a set of embeddings of the another data point of a different object (RGB image of a book, depth image of a book, language description of a book)
-%     \item Anchor: A fixed modality to act as anchor in the case of simple MMA loss, where we don't minimize the distance between the pair of embeddings from all modalities of the same data point, and we don't maximize the distance between positive an negative datapoints of the same modality.
-% \end{itemize}
-
-Since the downstream task in the grounded language learning domain is to predict/retrieve a desired object among multiple objects given a natural language description, it makes sense to train the model using language as an anchor, and in fact anchoring learning around language outperforms anchoring on RGB.
-
-\Cref{eq:objective-simple-mma} shows the mathematical formulation of the proposed loss function which can be generalized to arbitrary number of modalities.
-
-% \begin{equation}
-% \label{eq:objective-simple-mma}
-% \begin{split}
-%     \mathcal{L}  &= \sum_{m=1}^{M-1} || f_a(x_{a}^{+}) - f_m(x_{m}^{+}) || - \\ 
-%     & || f_a(x_{a}^{+}) -f_m(x_{m}^{-}) || 
-%     \\
-%     &= \sum_{m=1}^{M-1} 1 - \cos(f_a(x_{a}^{+}), f_m(x_{m}^{+})) \\ & - (1 - \cos(f_a(x_{a}^{+}) ,f_m(x_{m}^{-}))) \\
-%     &= \sum_{m=1}^{M-1} \cos(f_a(x_{a}^{+}) ,f_m(x_{m}^{-})) \\ 
-%     & - \cos(f_a(x_{a}^{+}), f_m(x_{m}^{+}))
-% \end{split}
-% \end{equation}
+An alternative is to apply triplet loss $M-1$ times with a single fixed anchor (text in our case) and $M-1$ `target' modalities.
+In this case, the negative anchor point is disregarded.
+The positive anchor is used as an anchor for all $M-1$ triplet losses, and for each of those triplet losses the positive and negative points are simply selected from the corresponding modalities.
+This can be derived from \cref{eq:full-emma} by fixing one modality as anchor, and excluding the distance between the positive and negative anchors or $\cos(z_{a}^{+}, z_{a}^{-})$.
+\Cref{eq:objective-simple-mma} formulates this idea.
 
@@
 \begin{split}
-    \mathcal{L}  &= \sum_{m=1}^{M-1} || z_{a}^{+} - z_{m}^{+} || - || z_{a}^{+} - z_{m}^{-} || \\
+    \mathcal{L}  &= \sum_{m=1}^{M-1} dist( z_{a}^{+} , z_{m}^{+} ) - dist( z_{a}^{+} , z_{m}^{-} ) \\
     &= \sum_{m=1}^{M-1} 1 - \cos(z_{a}^{+}, z_{m}^{+}) - (1 - \cos(z_{a}^{+}, z_{m}^{-})) \\
@@
 \end{equation}
+Here, subscript $a$ represents the \textit{anchor} modality, $M$ is the number of modalities, superscript $+$ represents the positive objects, and superscript $-$ represents the negative objects.
 
-where subscript $a$ represents \textit{anchor} modality, $M$ is the number of modalities, superscript $+$ represents the positive objects, and superscript $-$ represents the negative objects.
+However, this approach fails to take into account the dissimilarity of the positive and negative anchor points.
+This is an important piece of information that \cref{eq:objective-simple-mma} ignores.
+Although this information is implicitly captured when we maximize the distance between positive anchor and other negative modalities, when the negative data point becomes a positive example itself, the anchor and those points are forced to be closer to each other which can happen in three ways: either moving anchor closer to those points, moving those points closer to anchor, or move both somewhere in between.
+Therefore, in the last two cases we do not have explicit direct control over the distance between these points and previous positive points.
+Our experimental results confirm this hypothesis.
 
-An illustration of this loss function is provided in \cref{fig:simple-mma-loss}.
+Therefore, we add an extra term which is responsible for maximizing the distance between positive and negative instances of the anchor modality, because a margin between positive and negative points from other modalities (excluding anchor) is enforced.
+This can be derived from \cref{eq:full-emma} by fixing one modality as anchor.
+% The first dashed line from the top in \cref{fig:emma-loss} is this extra term that captures the dissimilarity between positive and negative text datapoints.
+We formulate this method in \cref{eq:objective-explicit-fixed-anchor}, which is generalized to an arbitrary number of modalities.
 
-\begin{figure}[tb]
+\begin{equation}
+\label{eq:objective-explicit-fixed-anchor}
+\begin{split}
+    \mathcal{L}  %&= %\sum_{m=1}^{M-1} || z_{a}^{+} - z_{m}^{+} || - || z_{a}^{+} - z_{m}^{-} ||  - || z_{a}^{+} - z_{a}^{-} || \\
+    &= \max(\cos(z_{a}^{+}, z_{a}^{-}) + \alpha_a, 0) + \sum_{m=1}^{M-1} \max \left(\cos(z_{a}^{+} ,z_{m}^{-}) - \cos(z_{a}^{+}, z_{m}^{+}) + \alpha_m, 0 \right) \\  
+    &= \sum_{m=1}^{M}  \cos(z_{a}^{+} ,z_{m}^{-}) - \sum_{m=1}^{M-1} \cos(z_{a}^{+}, z_{m}^{+})
+\end{split}
+\end{equation}
+
+In \cref{eq:objective-explicit-fixed-anchor}, $a$ represents the \textit{anchor} modality, $M$ is the number of modalities, the superscripts $+$ and $-$ represent positive and negative objects, $\alpha_m$ represents enforced margin for each modality which we set to 0.4 for all modalities without tuning, and $z$ is the embedding we get by applying a mapping function $f$, which in our case is a neural network on our input data.
+In other words, $z_m = f_m(x_m)$, where each modality $m$ has a specific model $f_m$ that is different from the models for other modalities. These models do not share their weights.
+
+Since the downstream task we are considering in the grounded language learning domain is to predict/retrieve a desired object among multiple objects given a natural language description, it makes sense to train the model using language as an anchor, and in fact anchoring learning around language outperforms anchoring on RGB.
+However, our proposed \geom{} loss function formulated in \cref{eq:full-emma} where we treat all modalities as anchor once still results in a better performance.
+
+% Triplet loss cannot be used for more than 2 modalities.
+Some previous work has concatenated RGB and depth embeddings to create a single ``vision'' embedding for learning~\cite{triplet_loss_2021_CVPR}, but they cannot handle RGB or depth sensor ablation during test.
+Since our method can handle any number of modalities, we can handle such a case by treating depth as a separate modality.
+% Moreover, when using triplet loss with different anchors per batch, the batch size has to be 1 which makes training very slow.
+
+% This approach is most similar to that of supervised contrastive learning~\cite{NEURIPS2020_supervised_contrastive}, but outperforms both that and regular contrastive learning~\cite{chen2020simple} methods.
+% Especially when modalities are ablated, most notably written language.
+
+% We find that the use of instance level losses grounded to one key modality (text) is sufficient to encourage a representation that satisfies our goals.
+
+\subsection{Extended Triplet}
+Using an extended version of triplet loss to consider all pairwise connections which is formulated in \cref{eq:extended-triplet}, results in a worse performance.
+
+\begin{equation}\label{eq:extended-triplet}
+    L = \sum_{m_1=1}^{M} \left[ \sum_{m_2=m_1+1}^{M} \left[ \text{triplet}(z_{m_1}^{+}, z_{m_2}^{+}, z_{m_2}^{-}) + \text{triplet}(z_{m_1}^{-}, z_{m_2}^{-}, z_{m_2}^{+}) \right] + \cos(z_{m_1}^{+}, z_{m_1}^{-}) -1 + margin \right]
+\end{equation}
+
+% Didn't change much when we drop text: Full eMMA when we have similar objects
+
+% Did not work: Run the version with all connections (including among negatives) without triplet and just by using cosine distance between pair of embeddings
+
+\subsection{Smaller Batch Sizes And Pulling Among Negatives}
+
+Reducing the batch size from 64 to 16 and 8 (for both \ours{} and \supcon{}) increases the gap between our model and the baseline in cases where our model beats the baseline (all modalities, dropping speech), but baseline is still better than ours when we drop the text modality.
+
+In batch size of 2 the trend is still the same, but when we drop the text modality, the MRR graph is flat at 0.4.
+This is still better than the worst and random cases, but it implies that the model does not learn anything with regards to speech.
+To remedy this, we add connections among the modalities of the negative instance as formulated in \cref{eq:full-emma-pull-negatives} and as shown in \cref{fig:full-emma-pull-negatives}, and that solves the flatline problem when we drop the text modality; however, when we have all modalities it performs worse than \ours{}.
+
+\begin{equation}\label{eq:full-emma-pull-negatives}
+\begin{split}
+    L = & \sum_{m_1=1}^{M} \left[ \sum_{m_2=1}^{M} \left[ \max (\cos(z_{m_1}^{+}, z_{m_2}^{-}) -1 + margin,0) \right] + \sum_{m_3=m_1+1}^{M} \left[ \max(1 - \cos(z_{m_1}^{+},z_{m_3}^{+}),0) \right] \right.
+\\
+    & \left. + \sum_{m_3=m_1+1}^{M} \left[ \max(1 - \cos(z_{m_1}^{-},z_{m_3}^{-}),0) \right] \right]
+\end{split}
+\end{equation}
+
+\begin{figure*}[tb]
 \centering
-\includegraphics[width=\columnwidth]{Figures/simple-MMA.pdf}
-\caption{A high-level prototype of the distances used in the simple MMA loss with 3 modalities.
-Orange arrows indicate maximizing the distance, blue arrows indicate minimizing the distance, and green lines show the margin.
-Depth images are brightened to be visible.
-\todo[inline]{We need to show speech as the 4th modality as well.}}
-\label{fig:simple-mma-loss}
-\end{figure}
-
-This version of MMA loss function can be seen as a contrastive loss usually used in the domain of self-supervised learning~\cite{chen2020simple}.
-However, our proposed loss function has two advantages over the traditional contrastive loss expressed in \cref{eq:contrastive-loss}.
-The first advantage is that our loss function do not loop over multiple positives and negatives in a large batch.
-Instead we sample only two objects (positive and negative) each of which have $M$ modalities which gives us $2M$ datapoints (or embeddings).
-Hence, our model can be trained using smaller batch sizes and reduces the number of negative samples we need.
-The second advantage is that this loss function can be used in a multimodal setting with an arbitrary number of modalities, and is not limited to a single data type (e.g.
-RGB images) which is the most common usage of contrastive loss~\cite{chen2020simple, NEURIPS2020_supervised_contrastive}.
-
-Therefore, our proposed loss function would modify \cref{eq:contrastive-loss} to the loss function formulated in \cref{eq:simple-MMA-contrastive}.
-Our formulation is a generalized version of the contrastive loss where $M$ is the number of modalities (e.g.
-RGB image, depth image, speech, text), but it can be the ``augmented views'' instead of different modalities.
-
-% \begin{equation}\label{eq:simple-MMA-contrastive}
-%     -\sum_{m=1}^{M-1} \log \frac{\exp (sim(f_a(x_a^+) , f_m(x_{m}^{+})/ \tau) )}{ \exp (sim(f_a(x_a^+) , f_m(x_{m}^{+}) / \tau)) + \exp (sim(f_a(x_a^+), f_m(x_{m}^{-}) / \tau))}
-% \end{equation}
-
-\begin{equation}\label{eq:simple-MMA-contrastive}
-    -\sum_{m=1}^{M-1} \log \frac{\exp (sim(z_a^+ , z_{m}^{+})/ \tau) }{ \exp (sim(z_a^+ , z_{m}^{+}) / \tau) + \exp (sim(z_a^+, z_{m}^{-}) / \tau)}
-\end{equation}
-where $z = f(x)$.
-
-% \todokdinline{I should run the traditional contrastive loss to see if my method works better to make this claim.}
-
-% \subsubsection{All in loss with 12 terms}
-% \subsubsection{subset of All in loss with 9 terms}
-
-\section{Data}
-
-\begin{figure*}[tbh]
-\centering
-\includegraphics[width=2.0\columnwidth]{Figures/raw-mask-test-mrr.pdf}
-\caption{Mean Reciprocal Rank (MRR) on different versions of GoLD dataset~\cite{GoLD_UMBC}.
-The raw version is depicted in green and masked version is depicted in orange averaged over 3 different random seeds.
-Higher is better.}
-\label{fig:mask-vs-raw}
+\includegraphics[width=1\columnwidth]{Figures/unused/full-emma-pull-negatives.pdf}
+\caption{A high-level prototype of our approach and the distances used when we have add pulling among negative modalities.}
+\label{fig:full-emma-pull-negatives}
 \end{figure*}
 
-\section{Experiments}
+\subsection{Margin}
+% \todokdinline{Try different margins of 0.6, 1.0, and if it improved the results we can try to learn it as a parameter}
+% \todokdinline{Should we remove this section?}
+We experimented with two different margins values: a margin of 0.6 and a margin of 1.0.
+Using a margin of 0.6 does not change the performance much.
+However, it converges even faster.
+Using a margin of 1.0 (this is essentially just cosine similarity because we have $cos -1 + \text{margin}$) results in a worse performance when we have all modalities.
+% The MRR is 0.8995 when margin is 1, while with a margin of 0.6 the MRR is 0.9176, and with a margin of 0.4 the MRR is 0.9198.
+If we drop the text modality, the MRR is 0.7923, 0.7913, and 0.6918 for margins of 0.4, 0.6, and 1.0 respectively.
 
 \section{Results}
-
-\begin{figure*}[tbh]
-\centering
-\includegraphics[width=2.1\columnwidth]{Figures/test.mrr.pdf}
-\caption{Mean Reciprocal Rank (MRR) of our proposed method against triplet loss method averaged over 3 different random seeds for a downstream task of object detection.
-The higher is better.
-Green is Supervised Contrastive Learning~\cite{NEURIPS2020_supervised_contrastive}, Pink represents our proposed simple MMA loss function without semantic negative sampling w/ SGD w/ adjusting learning rate, Purple is out method with semantic negative sampling, Blue is our method w/ semantic negative sampling w/ Adam and w/o adjusting learning rate, Red is vanilla contrastive loss with cosine, and orange represent the triplet loss.
-function~\cite{triplet_loss_2021_CVPR}.
-}
-
-\label{fig:simple-MMA-mrr}
-\end{figure*}
-
-\section{Setup}
-
-% \input{lit-search}
+We find that using each modality as anchor results in a better performance compared to when we fix one modality as anchor. This difference is more significant when we test with all modalities. 
+% 
+While the performance of \supcon{} deteriorates as the batch size is reduced, our \geom{} stays the same when we have all modalities or when we drop speech.
+However, when we drop text, EMMA deteriorates with decreasing batch size, and with a batch size of 2 without pulling negatives, it flatlines.
+In this case (when batch size is 2 and text is dropped), pulling among negatives is important and results in a better performance.

### 2022-10-06 Paper/main.tex
Paper submitted to TMLR

@@
 \usepackage{tmlr}
+%%\usepackage{showframe}
 % If accepted, instead use the following line for the camera-ready submission:
-%\usepackage[accepted]{tmlr}
+% \usepackage[accepted]{tmlr}
 % To de-anonymize and remove mentions to TMLR (for example for posting to preprint servers), instead use the following:
-%\usepackage[preprint]{tmlr}
+% \usepackage[preprint]{tmlr}
 
@@
 \usepackage{comment}
-\usepackage{todonotes}
+\usepackage{csvsimple}
+\usepackage{datatool}
+\usepackage{todonotes}[disable]
+\usepackage{graphicx}
+\usepackage{subcaption}
 \newcommand{\todokdinline}[1]{\todo[color=red!20,inline]{{KD: \small #1}}}
@@
 
-\title{Multimodal Contrastive Learning for Object Retrieval in the Face of Modality Failures \\ Extended Multimodal Alignment \\ Combining Geometric and classification ... }
+\title{Multimodal Language Learning in the Face of Missing Modalities}
+% \\ Extended Multimodal Alignment \\ Combining Geometric and classification ... }
 
@@
 
-\author{\name Kyunghyun Cho \email kyunghyun.cho@nyu.edu \\
+\author{\name Kasra Darvish \email kasradarvish@umbc.edu \\
       \addr Department of Computer Science\\
-      University of New York
+      University of Maryland Baltimore County
       \AND
-      \name Raia Hadsell \email raia@google.com \\
-      \addr DeepMind
+      \name Edward Raff \email Raff\_Edward@bah.com \\
+      \addr University of Maryland Baltimore County \\
+      Booz Allen Hamilton
       \AND
-      \name Hugo Larochelle \email hugolarochelle@google.com\\
-      \addr Mila, Universit\'e de Montr\'eal \\
-      Google Research\\
+      \name Francis Ferraro \email ferraro@umbc.edu \\
+      \addr University of Maryland Baltimore County
+      \AND
+      \name Cynthia Matuszek \email cmat@umbc.edu\\
+      \addr University of Maryland Baltimore County \\
       CIFAR Fellow}
@@
 
-    We propose a generalized geometric combined with cross-entropy loss function that can be used to learn retrieval models that incorporate an arbitrary number of views of a particular piece of data, and compounded by the challenge of retrieval when a modality becomes unavailable.
-    \todoff{this red text can (should) be shortened}
-    \todokd{I shortened it. Please remove both notes if it's resolved}
-    {\color{red}Our study is motivated by needs in robotics and computer-human interaction, where an agent has many sensors and thus modalities in which a human may interact both to communicate a desired goal and for the agent to recognize a desired target object.
-    Tying these various modalities together relates to the task of grounded language understanding. However, there has been little research on integrating more than two modalities.
-    % \todokdinline{Duplicate, it's already mentioned in intro}in which natural language is used as a query against objects in a physical environment, allowing a real-world, intuitive mechanism by which users can instruct physical agents to engage in tasks such as object retrieval.
-    % However, current work has neglected the real-world constraint that sensors fail. For practical applications,
-    %  we need mechanisms for accurate object retrieval given that a modality may fail without notice.
-    }
-    % \todoff{There's very little in here about actual method or conclusions}
-    There has been widely popular works on self-supervised contrastive learning based on cross-entropy, but there is an entire other approach based on explicit geometric alignment.
-However, there has been no work on combining the two approaches for multimodal learning to the best of our knowledge.
-We propose to combine the two approaches and argue that the two are complementary.
+    We propose \textit{extended multimodal alignment (EMMA)}, a generalized geometric method combined with a cross-entropy loss function that can be used to learn retrieval models that incorporate an arbitrary number of views of a particular piece of data, compounded by the challenge of retrieval when a modality becomes unavailable.
+Our study is motivated by needs in robotics and computer-human interaction, where an agent has many sensors and thus modalities with which a human may interact, both to communicate a desired goal and for the agent to recognize a desired target object.
+For such problems, there has been little research on integrating more than two modalities.
+While there have been widely popular works on self-supervised contrastive learning based on cross-entropy, there is an entirely separate family of approaches based on explicit geometric alignment.
+However, to the best of our knowledge there has been no work on combining the two approaches for multimodal learning.
+We propose to combine both families of approaches and argue that the two are complementary.
 We demonstrate the usability of our model on a grounded language object retrieval scenario, where an intelligent agent has to select an object given an unconstrained language command.
-    We leverage four modalities including vision, depth sensing, text, and speech, and we show that our model can outperform state-of-the-art contrastive learning models by $2\%$ while converging/learning/training $5$ times faster (takes $80\%$ less time or is $20\%$ of the baseline).
+We leverage four modalities including vision, depth sensing, text, and speech, and we show that our model converges approximately 5 times faster than previous strong baselines, and out-performs or is strongly competitive with state-of-the-art contrastive learning.
+% (in 80\% less time).
     %when modalities are ablated. 
-    The code will be publicly available on GitHub and will be included for the camera-ready version (it is redacted for anonymity). 
+    The code is publicly available on GitHub and will be included for the camera-ready version (it is redacted for anonymity). 
 \end{abstract}
@@
 
-\todoffinline{nix this first sentence and revise the start of the second.
-We're not doing any cogsci-y stuff, or even human studies tied to modality deprivation.
-You could take most of the red text in the abstract I mentioned should be compressed and rework it here}
-\todokdinline{I took a shot at nixing it.
-Please remove both notes if it's resolved.
-I don't get the vibe of cogsci-y stuff or human studies from the first paragraph, would you please refer to those sentences?}
-Humans successfully use multiple sensory inputs to understand, interact with, and retrieve information from the world around them.
-Inspired by the multimodal nature of human interactions with the world, it is intuitive that agents learning about the world, upon encountering new concepts and new objects, should form a model that incorporates information from all available sensors and data sources.
-Benefits of integrating multiple modalities is twofold; first, additional modalities can help in the cases when other modalities fail, and second, complementary information can be extracted from different modalities that can help with understanding the world.
+Inspired by the multimodal nature of human interaction with the world, it is intuitive that agents learning about the world, upon encountering new concepts and new objects, should form a model that incorporates information from all available sensors and data sources.
+The benefits of integrating multiple modalities are twofold; first, complementary information can be extracted from different modalities that can help with understanding the world, and second, additional modalities can help in the cases when one or more sources of data about the world become unavailable.
 % 
 Grounded language understanding, in which natural language is used as a query against objects in a physical environment, allows a real-world, intuitive mechanism by which users can instruct physical agents to engage in tasks such as object retrieval.
-Visuolinguistic approaches to such object inference tasks typically involve training on large pools of image/text pairs and then using language to subselect elements of the sensed environment~\cite{hong2021gilbert,Zhuge_2021_CVPR_VLP}.
-\todoffinline{Unless Hong et al is a review paper, add at least one more citation}
-\todokdinline{I believe this sentence was added to cite some relevant work from SIGIR, so we can remove it. However, I added one more citation anyway.}
+Visuolinguistic approaches to such object inference tasks typically involve training on large pools of image/text pairs and then using language to subselect elements of the sensed environment~\citep{hong2021gilbert,Zhuge_2021_CVPR_VLP}.
+
 Although physical agents such as robots typically have access to sensory and interactive modalities beyond vision, and learning from multiple modalities can improve performance on downstream tasks, most approaches use at most two sensory inputs (e.g., visual data such as RGB plus depth images) with single labels, such as those provided by textual natural language.
@@
 
-One difficulty with working with complex multimodal data is the increased likelihood that one or more modalities may have missing information; hardware can become damaged or shipped with defects, sensors can get blocked or obstructed, and various adverse but not uncommon conditions that remove a modality from use.
+One difficulty with working with complex multimodal data is the increased likelihood that one or more modalities may have missing information.
+Hardware can become damaged or defective, sensors can get blocked or obstructed, and various adverse but not uncommon conditions can remove a modality from use.
 Current multimodal approaches are typically not robust to the loss of one or more modalities at test time, as may happen if, for example, a physical agent fails to retrieve data from a particular sensor.
@@
 
-We consider the domain of grounded language-based object retrieval~\cite{hu2016natural, triplet_loss_2021_CVPR}, in which objects in an environment must be identified based on linguistic instructions.
-This can be considered a special case of image retrieval~\cite{huang2017deep, ma2020large, novak2015large, vo2019composing} in which objects are identified using visual inputs in combination with other sensor modalities.
-Approaches to acquiring grounded language have explored various combinations of sensor inputs such as depth and RGB with labels provided by textual language or  speech~\cite{RichardsDarvishMatuszekCategoryFree20}.
+We consider the domain of grounded language-based object retrieval~\citep{hu2016natural, triplet_loss_2021_CVPR}, in which objects in an environment must be identified based on linguistic instructions.
+This can be considered a special case of image retrieval~\citep{huang2017deep, ma2020large, novak2015large, vo2019composing} in which objects are identified using visual inputs in combination with other sensor modalities.
+Approaches to acquiring grounded language have explored various combinations of sensor inputs such as depth and RGB with labels provided by textual language or  speech~\citep{RichardsDarvishMatuszekCategoryFree20}.
 However, despite the multisensory nature of object retrieval, much of the existing work has not previously been extended to include an arbitrary number of modalities.
 
-\todoff{revise this goal. It sets up the reviewers to take the easy way out and say results are too incremental}
-The high-level goal of this work is to take arbitrary input modalities about novel objects, including both spoken and written language, and build a model that allows a system to correctly retrieve that object given a subset of those input modalities.
+% \todo[inline]{Whats the name of our method?
+When I skim the paper its hard to tell what ``the new thing is'', and then I have to read.
+I think being explicit here, with something like ``\textit{we propose our new method BLA}'' will help -Ed}
+
+The ultimate aim of this work is to take arbitrary input modalities about novel objects, including both spoken and written language, and build a model that allows a system to correctly retrieve that object given a subset of those input modalities.
+In this work, we propose our new method \ours{}, a generalized geometric method combined with a cross-entropy loss function that can be used to learn retrieval models that incorporate an arbitrary number of views of a particular piece of data.
+We demonstrate steps towards this ultimate goal with a mechanism that learns quickly from a variety of input signals and is robust to ablation of certain inputs.
 This is a generalization of approaches to grounded language learning in which specific input modalities are labeled with language to allow for future identification, but explicitly seeks to be agnostic about the nature of the individual sensory and linguistic inputs.
+
 Our contributions are as follows:
 \begin{enumerate}
-    \item Proposing a new approach to multimodal learning by combining geometric and cross-entropy based methods and showing that it outperforms and converges faster compared to each method separately.
+    \item Proposing a new approach to multimodal learning by combining geometric and cross-entropy methods and showing that it outperforms and converges faster compared to each method separately.
     % \item Outperforming state-of-the-art models for grounded language learning by differentiating visual input modalities such as RGB and depth, rather than combining them into a single modality.
 
-    \item Our method can be easily extended to an arbitrary number of modalities.\todoffinline{potential counter: isn't the method quadratic?}
-    \todokdinline{Yes, but nonetheless it can be extended.}
+    \item Our method is designed to be applicable to additional modalities as they become available.
 
-    \item Outperforming state-of-the-art contrastive learning~\cite{chen2020simple} and supervised contrastive learning~\cite{NEURIPS2020_supervised_contrastive} models by $2\%$ when all modalities are available, and achieving the peak performance $5$ times faster than both methods.
-    % Ours takes 8 epochs, baselines take 40 epochs.
-    \todokdinline{Should we break this contribution into two contributions, one for faster learning and the other for performance compared to both supcon and vanilla contrastive learning?}
+    \item Our method converges approximately 5 times faster than previous strong baselines, and out-performs or is strongly competitive with state-of-the-art contrastive learning~\citep{chen2020simple} and supervised contrastive learning~\citep{NEURIPS2020_supervised_contrastive} models when all modalities are available.
 
-    \item Demonstrating robustness in the face of missing information, when one or more modalities are ablated at test time.
+    \item Demonstrating robustness in the face of missing input signals when one or more modalities are ablated at test time, and proposing a simple averaging method to quantify performance for multimodal object retrieval scenarios.
 
@@
 
-\todoffinline{use paragraphs to add some light structure to this block-o-text}
+% \todoffinline{I'm not convinced this section shouldn't appear toward the end of the paper. Maybe adding that narrative structure will help?}
 
-\todoffinline{I'm not convinced this section shouldn't appear toward the end of the paper. Maybe adding that narrative structure will help?}
+% \todocmi{I'm not sure what narrative structure is being referred to here. I have a preference for having related work sections towards the beginning.}
 
 There is extensive work on the task of image retrieval, of which physical object retrieval can be considered a special case.
-In this task, language is used to formulate queries against datasets of images, for example in text-and-image matching tasks for fashion data~\cite{gao2020fashionbert, wen2021comprehensive}, sketch retrieval~\cite{huang2017deep}, and general photographs of objects~\cite{ma2020large, novak2015large, hong2021gilbert}.
-Prior work has focused solely on language and visual modalities, using text or a combination of language and vision to perform visual retrieval~\cite{vo2019composing}.
-While recent works have focused on grounding with respect to a knowledge set \cite{10.1145/3357384.3357889,10.1145/3397271.3401097}, our work focuses on robust multimodal learning to perform grounding over an arbitrary number of input modalities.
+In this task, language is used to formulate queries against datasets of images, for example in text-and-image matching tasks for fashion data~\citep{gao2020fashionbert, wen2021comprehensive}, sketch retrieval~\citep{huang2017deep}, and general photographs of objects~\citep{ma2020large, novak2015large, hong2021gilbert}.
+Prior work has focused solely on language and visual modalities, using text or a combination of language and vision to perform visual retrieval~\citep{vo2019composing}.
+While recent works have focused on grounding with respect to a knowledge set \citep{10.1145/3357384.3357889,10.1145/3397271.3401097}, our work focuses on robust multimodal learning to perform grounding over an arbitrary number of input modalities.
 
@@
 \citet{bagher-zadeh-etal-2018-multimodal} introduce a large dataset of multimodal opinion sentiment and emotion intensity (CMU-MOSEI) for sentiment analysis and emotion recognition.
-They also introduce a novel multimodal fusion technique called the Dynamic Fusion Graph (DFG).~\citet{GoLD_UMBC} present a multimodal dataset of household objects containing RGB images, depth images, written language, and spoken language, which has been used to support learning grounded language directly from speech given small datasets~\cite{KebeAAAI2022}.
+They also introduce a novel multimodal fusion technique called the Dynamic Fusion Graph (DFG).~\citet{GoLD_UMBC} present a multimodal dataset of household objects containing RGB images, depth images, written language, and spoken language, which has been used to support learning grounded language directly from speech given small datasets~\citep{KebeAAAI2022}.
 \citet{baltrusaitisMultimodalMachineLearning2019} propose a new taxonomy of multimodal machine learning by introducing five technical challenges in addition to typical early and late fusion categorization.
@@
 We highlight a key difference with recent work like \citet{10.1145/3397271.3401232} and \citet{10.1145/3331184.3331213} who develop multi-modal retrieval models.
-These retrieval models are class based, where the model is trained to recognize any object of the same retrieval class as being ``same''.
-Our method and data is instance based, where different objects are not considered ``same'' even if they have the same class, due to the grounding goals of our work (i.e., the agent should identify the specified object, not equivalent objects).
-These works also do not consider the failure of a modality that is our target of interest.
+These retrieval models are class based, where the model is trained to recognize any object of the same retrieval class as being the `same.' Our method and data is instance based, where different objects are not considered the same even if they have the same class, due to the grounding goals of our work (i.e., the agent should identify the specified object, not equivalent objects).
+These works also do not consider performance in the case of failure of an input modality.
 
@@
 \citet{alayrac2020self} use a self-supervised contrastive learning method to learn and combine representations from three modalities of visual, audio, and language.
-Audio and visual domains are mapped to the same space because they are fine-grained, containing dense information for each frame of video.
-These embeddings are then mapped to a coarse space where text is also mapped.
+Audio and visual domains are mapped to the same space because they are fine-grained, containing dense information for each frame of video, and these embeddings are then mapped to a coarse space where text is also mapped.
 Their network respects specificity and comparability of different modalities.
@@
 \citet{Nguyen-RSS-20} take a similar approach and perform pairwise cosine similarity to align images and natural language text descriptions to perform object retrieval tasks.
-\citet{triplet_loss_2021_CVPR} take a cross-modal manifold alignment approach to grounded language learning using triplet loss \cite{Chechik:2010:LSO:1756006.1756042}, and perform object retrieval by connecting text descriptions of objects to their corresponding RGB-D images.
+\citet{triplet_loss_2021_CVPR} take a cross-modal manifold alignment approach to grounded language learning using triplet loss \citep{Chechik:2010:LSO:1756006.1756042}, and perform object retrieval by connecting text descriptions of objects to their corresponding RGB-D images.
 
-In general there are two different family of approaches to multimodal learning in the literature.
+In general there are two different families of approaches to multimodal learning in the literature.
 One family is based on classification and cross-entropy losses such as~\citet{NEURIPS2020_supervised_contrastive, chen2020simple} and the other family is geometric based such as~\citet{ICML22GeometricMultimodal,Carvalho-cooking-triplet,salvador2017cooking,triplet_loss_2021_CVPR}.
-In this paper we propose to marry the two families and show that combining geometric approach with cross-entropy method is superior. \todokdinline{Pun intended :-)}
-\todokdinline{Should we move the following two paragraphs to the Approach section?}
-For the classification part of our loss function we employ the supervised contrastive loss proposed by~\citet{NEURIPS2020_supervised_contrastive}.
-The way we formulate our geometric part of the loss function is similar to lifted structured loss~\cite{songCVPR16LiftedStructured}, where they take a geometric approach in a unimodal scenario and consider pairwise distances among all items in a batch, but in this paper we apply our idea to four modalities, and our formulation can be extended to any number of modalities.
+In this paper we propose to marry the two families and show that combining geometric approach with cross-entropy method is superior.
+
+% \todokdinline{Should we move the following two paragraphs to the Approach section?}
+% \todocmi{Actually yes, although I think it's more like 3 paragraphs. Rewriting the subsequent paragraphs so they don't appear out of nowhere would help.}
+% \todokdinline{After moving baselines to the approach section and rewriting it, I think it's better if we leave these here. Let me know if you think otherwise.}
+
+% For the classification part of our loss function we employ the supervised contrastive loss proposed by~\citet{NEURIPS2020_supervised_contrastive}. 
+The way we formulate our geometric loss function is similar to lifted structured loss~\citep{songCVPR16LiftedStructured}, where they take a geometric approach in a unimodal scenario and consider pairwise distances among all items in a batch, but in this paper we apply our idea to four modalities, and our formulation can be extended to any number of modalities.
 This extended geometric formulation results in a loss function that is similar to \citet{ICML22GeometricMultimodal} where they propose a geometric contrastive learning approach in which they %use the same instance of an MLP network (same weights) after extracting embeddings from each modality, while we use different instances of the same MLP network (different weights).
@@
 
-The geometric approach is also known as contrastive learning~\cite{qin2021world}, which is based on the idea of similarity and distance of elements of training data~\cite{Carvalho-cooking-triplet,triplet_loss_2021_CVPR,salvador2017cooking}.
-Triplet loss is also a special case of contrastive loss~\cite{NEURIPS2020_supervised_contrastive}.
-Contrastive loss is mostly used in self-supervised learning~\cite{bui2021self, alayrac2020self,chen2020simple}, with some exceptions such as \citet{NEURIPS2020_supervised_contrastive}, who use contrastive loss for a supervised problem.
+The geometric approach is also known as contrastive learning~\citep{qin2021world}, which is based on the idea of similarity and distance of elements of training data~\citep{Carvalho-cooking-triplet,triplet_loss_2021_CVPR,salvador2017cooking}.
+Triplet loss is also a special case of contrastive loss~\citep{NEURIPS2020_supervised_contrastive}.
+Contrastive loss is mostly used in self-supervised learning~\citep{bui2021self, alayrac2020self,chen2020simple}, with some exceptions such as \citet{NEURIPS2020_supervised_contrastive}, who use contrastive loss for a supervised problem.
 Most of these methods are implemented for a non-multimodal dataset (e.g., RGB images only).
-Our work is a specialization of contrastive loss that is focused on the multimodal input and- language problem.
+Our work is a specialization of contrastive loss that is focused on the multimodal input and language problem.
 
-We note in particular that standard triplet learning approaches often require an expensive \textit{mining} step to find harder negative samples to enable effective learning \cite{Hoffer2015,Schroff2015,DBLP:conf/eccv/2018-9,DBLP:conf/eccv/ZhaoJQLH18,Zhai2018}.
-This is problematic from throughput and scalability, as run time becomes quadratic in batch size.
-Our approach does not require mining to obtain good performance, alleviating these issues along with their complicated history in reproducibility and parameter tuning \cite{Musgrave2020,Raff2020c,Raff2019_quantify_repro}.
+We note in particular that standard triplet learning approaches often require an expensive \textit{mining} step to find harder negative samples to enable effective learning \citep{Hoffer2015,Schroff2015,DBLP:conf/eccv/2018-9,DBLP:conf/eccv/ZhaoJQLH18,Zhai2018}.
+This is problematic for throughput and scalability, as run time becomes quadratic in batch size.
+Our approach does not require mining to obtain good performance, alleviating these issues along with their complicated history in reproducibility and parameter tuning \citep{Musgrave2020,Raff2020c,Raff2019_quantify_repro}.
 
@@
 This work differs from ours in that it becomes computationally complex as more and higher-dimensional modalities are added.
-Our proposed geometric loss function resembles work on quadruplet loss~\cite{chen2017beyond,tursun2021efficient} but is intended to scale to an arbitrary number of modalities.
+Our proposed geometric loss function resembles work on quadruplet loss~\citep{chen2017beyond,tursun2021efficient} but is intended to scale to an arbitrary number of modalities.
 
+To the best of our knowledge, there has been limited work on incorporating more than three modalities into a general learning approach.
 \citet{semihet_three_way_Lei_2020} use inputs from three modalities of image, sketch, and edgemap to perform image retrieval given the sketch.
-They use co-attention mechanism, and their loss function contains alignment loss, cross-entropy loss, and sketch-edgemap contrastive loss.
+They use a co-attention mechanism, and their loss function contains alignment loss, cross-entropy loss, and sketch-edgemap contrastive loss.
 \citet{Mittal2020M3ER} use Canonical Correlational Analysis (CCA) to differentiate between ineffective and effective modalities for the emotion recognition task from multiple input modalities including face, speech, and text.
@@
 % \todo[inline]{talk about CLIP and DALL-E here}
-To the best of our knowledge, there has been limited work on incorporating more than three modalities into a general learning approach.
 
@@
 
-Given a language command, for example either text or speech, that describes an object, we want our model to retrieve the correct object from a set of objects.
+Given a language command (either text or speech) that describes an object, we want our model to retrieve the correct object from a set of objects.
 This problem is an exemplar of tasks found in the area of grounded language learning in the fields of robotics and natural  language processing.
@@
 We demonstrate on a domain containing four modalities, all of which refer to objects in the environment: spoken language, written text, RGB (image) inputs, and depth camera inputs.
-\Cref{fig:experimental-setup} illustrates a small visualization of our object retrieval task: the spoken query ``A white textbook titled algorithms'' is provided to our contrastive model, which identifies the item (outlined in red, and represented by up to four different modalities) as the most likely item the query is referring to.
+\Cref{fig:experimental-setup} illustrates a small visualization of our object retrieval task: the spoken query ``A white textbook titled algorithms'' is provided to our contrastive model, which identifies the item
+(outlined in red in \cref{fig:experimental-setup}) 
+as the most likely item the query is referring to.
 
@@
 \centering
-\includegraphics[width=.5\columnwidth]{Figures/experiment-setup.png}
+\includegraphics[width=.75\columnwidth]{Figures/experiment-setup-rearranged.png}
 \caption{
@@
 Given a query in some modality, our approach seeks to select the object that is the best fit, per a trained model.
-This approach, detailed in \cref{sec:Method}, is able to rank objects as to appropriateness even when one or more modalities is ablated at test time (e.g., depth inputs are missing), and outperforms state-of-the-art contrastive learning approaches on this task.
+This approach, detailed in \cref{sec:Method}, is able to rank objects as to appropriateness even when one or more modalities is ablated at test time (e.g., depth inputs are missing), and outperforms state-of-the-art contrastive learning approaches.
 }
@@
 
-More formally, we consider a spoken language command $x_s$, a textual language command $x_t$, a set of RGB images $X_r = \{x_r^{(1..n)}\}$, and a set of depth images $X_d = \{x_d^{(1..n)}\}$, the task is to retrieve the correct object by choosing the index that has the minimum distance from either of the language commands across all modalities.
-Depending on which modalities are or are not ablated, we consider up-to four distances: ``sr,'' a vector of distances between $x_s$ and all RGB images in $X_r$; ``sd,'' a vector of distances between $x_s$ and all depth images in $X_d$; ``tr,'' a vector of distances between $x_t$ and all RGB images in $X_r$; and ``td,'' a vector of distances between $x_t$ and all depth images in $X_d$.
+More formally, given a spoken language command $x_s$, a textual language command $x_t$, a set of RGB images $X_r = \{x_r^{(1..n)}\}$, and a set of depth images $X_d = \{x_d^{(1..n)}\}$, the task is to retrieve the correct object by choosing the index that has the minimum distance from either of the language commands across all modalities.
+Depending on which modalities are or are not ablated, we consider up to four distances: \textit{sr}, a vector of distances between $x_s$ and all RGB images in $X_r$; \textit{sd}, a vector of distances between $x_s$ and all depth images in $X_d$; \textit{tr}, a vector of distances between $x_t$ and all RGB images in $X_r$; and \textit{td}, a vector of distances between $x_t$ and all depth images in $X_d$.
 In order to select the correct object, we first perform a component-wise average of the relevant modality pair distances for the available modalities.
@@
 For example, if no written instructions are available at test time,\footnote{This setting is particularly salient.
-While large bodies of text are frequently available at training time, a person interacting directly with a physical agent may well prefer to use spoken instructions.} we compute the component-wise average of $sr$ and $sd$, and then select the object whose coordinate resulted in the lowest average distance.
+While large bodies of text are frequently available at training time, a person interacting directly with a physical agent may well prefer to use only spoken instructions.} we compute the component-wise average of $sr$ and $sd$, and then select the object whose coordinate resulted in the lowest average distance.
 This method allows us to extend our model to support arbitrary modalities while remaining robust to test cases in which some modalities are missing or incomplete.
@@
 In keeping with previous work on the closely related problem of image retrieval, we focus on contrastive loss approaches, in which the goal is to learn an embedding of data where similar samples---in our case, samples belonging to the same class of object---are embedded `close' to one another, while dissimilar samples are farther apart.
-We develop a novel geometric loss function that simultaneously minimizes intra-class distances and maximizes inter-class distances across each pair of modalities yielding a model that is both effective at the retrieval task defined above and robust to modality dropouts at test time.
-We take an additional step and combine this geometric loss with a classification-based (cross-entropy) loss function which results in a superior model compared to either of geometric or cross-entropy losses alone.
+We develop a novel geometric loss function, \geom{}, that simultaneously minimizes intra-class distances and maximizes inter-class distances across each pair of modalities, yielding a model that is both effective at the retrieval task defined above and robust to modality dropouts at test time.
+We take an additional step and combine this \geom{} loss with a classification-based (cross-entropy) loss function which results in a superior model compared to either of geometric or cross-entropy losses alone; we refer to this combination as Extended Multi-Modal Alignment, or \ours{}.
 
-\subsection{Geometric Alignment Loss}
-\label{ssec:distanceloss}
+\subsection{Baselines}
+We compare both \ours{} and \geom{} against contrastive learning~\citep{chen2020simple} and supervised contrastive learning~\citep{NEURIPS2020_supervised_contrastive}, which for conciseness we refer to as \supcon{}.
+We consider \supcon{} as the main baseline since it is a general version of multiple contrastive loss functions including triplet loss, the traditional version of contrastive loss usually used in self-supervised settings~\citep{chen2020simple}, and N-pair loss~\citep{NIPS2016N-PairLoss}.
+% and do not include N-pair and triplet loss in our comparisons.
+%contrastive learning methods including triplet loss, the traditional version of contrastive loss usually used in self-supervised settings~\citep{chen2020simple}, and supervised contrastive learning~\citep{NEURIPS2020_supervised_contrastive}.
+
+\begin{comment}
+\subsubsection{Triplet Loss}
+\label{sub:baseline-triplet}
+Triplet loss has two major disadvantages. First, it cannot be used for more than two modalities.
+Second, a batch size of greater than 1 is hard to implement if the anchor, positive, and negative come from different modalities for each batch.
+% To address these issues in EMMA, we sample more than one positive and negative data points.
+%Intuitively, we sample one positive data point and one negative data point, each of which includes $M$ modalities, resulting in 2M data points.
+Each of the $M$ positives modalities become an anchor once.
+% \Cref{table:quantitative} shows that our MMA method perform as good as this baseline while our model can be trained 10x faster since we can use batch sizes larger than 1. Also,
+
+% When we used a fixed anchor modality, a batch size of 64, and applied semantic negative sampling~\citep{Pillai_Matuszek_2018}, our method outperformed this baseline. However, 
+% Since this method cannot be applied to four modalities, we do not include it in the main analysis. 
+\end{comment}
+
+\subsubsection{Contrastive Loss}
+\label{sub:baseline-contrastive}
+
+We compare our model against the contrastive learning method presented by~\citet{chen2020simple} where they use the normalized
+temperature-scaled cross entropy loss (NT-Xent).
+In order to implement this loss function, we use cosine similarity as suggested in the SimCLR paper~\citep{chen2020simple}.
+Another possibility is to use an inner dot product~\citep{NEURIPS2020_supervised_contrastive}; if not normalized, this can lead to numerical instabilities and overflow/underflow since the dot product is not bounded, but the result is the same whether we use normalized inner dot product or cosine similarity.
+Contrastive loss function is formulated in \cref{eq:contrastive-loss}:
+% \todokdinline{try contrastive loss with cross-entropy, dot product, and the supervised contrastive method itself (without triplets).}
+
+% Original Formula
+\begin{equation}\label{eq:contrastive-loss}
+    -\sum_{i \in I} \log \frac{\exp (sim(z_i , z_{j(i)}) / \tau) }{\sum_{a \in A(i)} \exp (sim(z_i, z_a) / \tau)}
+\end{equation}
+% \begin{equation}\label{eq:contrastive-loss}
+%     -\sum_{i \in I} \log \frac{\exp (sim(f(x_i) , z_{j(i)}) / \tau) }{\sum_{a \in A(i)} \exp (sim(f(x_i), f(x_a)) / \tau)}
+% \end{equation}
+where $i$ is the index of anchor, $j(i)$ is the index of positive item with respect to the anchor $z_i$ and is not the same as anchor, $A(i)$ is the set of all negatives and the one positive indices excluding anchor, and $z = f(x)$.
+
+We can treat different modalities of the same instance as additional input for that instance that augment the available information, and consider them as the positive points for the anchor.
+\Cref{eq:contrastive-loss} can be rewritten with the sum over more than one positive item as formulated in \cref{eq:explicit-contrastive}:
+
+%, which is equivalent to ``augmented views'' in the contrastive loss terminology.
+
+% \todo[inline]{Here we use $\sum_{p=1}^{P(i)}$, but later use $\sum_{p \in P(i)}$, we should be consistent.
+IMO I'd go with the latter, as its more consisnt with all other equations and $\sum_{p=1}^{P(i)}$ is a littel unusual looking -Ed I corrected it.
+-Kasra}
+\begin{equation}\label{eq:explicit-contrastive}
+    - \sum_{i \in I} \sum_{p \in P(i)} \log \frac{\exp (sim(z_i , z_{p})/ \tau) }{
+    \sum_{a \in A(i)} \exp (sim(z_i, z_a) / \tau)}
+    %{ \exp (sim(z_i^+ , z_{m}^{+}) / \tau) + \exp (sim(z_i^+, z_{m}^{-}) / \tau)}
+\end{equation}
+where $I$ is a batch consisting of one or more instances each with a set of all its modalities, and $P(i)$ is the set of modalities/augmentations of the anchor $i$ excluding itself (e.g.
+RGB image, depth image, speech, text) and $z = f(x)$.
+Therefore, if we have 4 modalities and the batch size is 64, the size of $I$ is $256$, the size of $P(i)$ is $M-1 = 3$ where $M$ is the number of modalities, and the size of $A(i)$ is $256 - 1 = 255$.
+
+\subsubsection{Supervised Contrastive Learning}
+\label{sub:baseline-supcon}
+
+\citet{NEURIPS2020_supervised_contrastive} extend the contrastive learning method (NT-Xent) and propose a supervised way of performing contrastive learning to treat not only augmentations of the anchor, but also every item that shares the same label with anchor as positives.
+This loss function is shown in \cref{eq:supervised-contrastive}.
+
+\begin{equation}\label{eq:supervised-contrastive}
+    \sum_{i \in I} \frac{-1}{|P(i)|} \sum_{p \in P(i)} \log \frac{\exp (z_i \cdot z_p / \tau) }{\sum_{a \in A(i)} \exp (z_i \cdot z_a / \tau)}
+\end{equation}
+
+% \begin{equation}\label{eq:supervised-contrastive-my-notation}
+%     \sum_{i \in I} \frac{-1}{|P(i)|} \sum_{p \in P(i)} \log \frac{\exp (f_i(x_i) \cdot f_p(x_p) / \tau) }{\sum_{a \in A(i)} \exp (f_i(x_i) \cdot f_a(x_a) / \tau)}
+% \end{equation}
+
+Although this loss function does not use cosine similarity, embeddings are normalized before performing dot product, which is equivalent to cosine similarity.
+
+The main difference between the contrastive loss baseline in \cref{sub:baseline-contrastive} and \supcon{} is that there is no notion of meaningful negative points in contrastive loss, and everything in the batch that is not the anchor or one of the positive views is considered to be negative.
+In \supcon{}, however, all elements in the batch that have same label as the anchor are also considered positives, in addition to different views of the same instance.
+While the denominators of \cref{eq:explicit-contrastive,eq:supervised-contrastive} stay the same, this subtle difference affects the numerator and includes more positive examples, which prevents the unintended use of actual positives as negative examples.
+
+While this model is a strong baseline, the authors applied it to a unimodal dataset.
+In this paper we extend this baseline to work with a multimodal dataset and show that it is slower than \ours{} to learn and its performance is also lower when all modalities are available during test time.
+% if modalities are ablated, this method does not perform as well as our proposed model (we show and discuss this later, in \cref{fig:epochs-mrr.srd}).
+This is especially true when the text modality is dropped, as is likely in physical agent scenarios, where speech is a natural interaction mechanism.
+
+Since \supcon{} considers all pairwise distances in each batch, with $M$ modalities and a batch of size $B$ each batch contains $B$x$M$ items, and \supcon{} computation involves $(BM)^2$ pairwise distance terms which is dependent on batch size.
+However, the computations of our \geom{} approach is agnostic with respect to batch size which makes it scalable.
+
+\subsection{EMMA: Extended Multimodal Alignment}
+\label{sec:emma}
+
+Our proposed multimodal method is composed of two complementary parts.
+The first part is a geometric loss based on distances in the latent space, and the second part is a contrastive loss based on cross-entropy.
+The geometric loss is faster to learn while the cross-entropy method is more aligned with the downstream task of object retrieval.
+Hence we propose to combine them.
+
+\paragraph{Geometric Alignment Loss}
+\label{subsec:geom-loss}
 
 % For $M$ modalities, we 
-We define a distance-based loss function which can be used for arbitrary number of modalities.
-Our proposed method is inspired by the well-known similarity-based triplet loss~\cite{Carvalho-cooking-triplet,triplet_loss_2021_CVPR}, and is similar to contrastive loss~\cite{chen2020simple,NEURIPS2020_supervised_contrastive} under some settings.
+We define a distance-based loss function which can be used for an arbitrary number of modalities.
+Our proposed method is inspired by the well-known similarity-based triplet loss~\citep{Carvalho-cooking-triplet,triplet_loss_2021_CVPR}, and is similar to contrastive loss~\citep{chen2020simple,NEURIPS2020_supervised_contrastive} under some settings.
 Triplet loss-based learning works by forcing similar concepts from different domains `together' in some shared embedding space, while forcing dissimilar concepts `apart.' It is so named because it relies on three data points from the training set: a positive, a negative, and an anchor point.
 However, standard triplet loss cannot be used for more than two modalities.
-Therefore, we modify this concept as follows.
-During training, we sample two different instances and their corresponding representations from all modalities into two sets---one positive set (referring to a specific object) and one negative set (referring to some different object) as shown in \cref{fig:emma-loss}.
-Unlike some prior triplet loss methods~\cite{GoLD_UMBC,triplet_loss_2021_CVPR}, the anchor is not randomly chosen from different modalities in each batch.
+
+To address this issue, we use pairwise distance optimization for all data points. %
+Our method can be used for an arbitrary number of modalities. %
+In contrast to our work, previous works that use triplet loss~\citep{GoLD_UMBC,triplet_loss_2021_CVPR} concatenate RGB and depth to form a single ``vision'' vector to handle three modalities, so they cannot robustly handle RGB or depth sensor ablation during test.
+Moreover, our method has the advantage that it does not require providing hard negative examples.
+Therefore, we modify the concept of triplet loss as follows.
+During training, we sample two different instances and their corresponding representations from all modalities into two sets---one positive set (referring to a specific object) and one negative set (referring to some different object) as shown in \cref{fig:emma-loss-trimmed}.
+Unlike some prior triplet loss methods~\citep{GoLD_UMBC,triplet_loss_2021_CVPR}, the anchor is not randomly chosen from different modalities in each batch.
 Instead, in our setting, every item in the positive set becomes anchor once, and we minimize the distance between that item and other items in the positive set, while minimizing the distance between that item and all items in the negative set.
@@
     % \item data point: a row in the csv/tsv file.
-    \item Positive (Instance): A set of embeddings of one data point (e.g., an RGB image of an apple, corresponding depth image, text description, and speech description of the same apple)
-    \item Negative (Instance): A set of embeddings of another data point of a different object (e.g., an RGB image of a book, corresponding depth image, text description, and speech description of the same book)
+    \item Positive (Instance): A set of embeddings of one data point (e.g., an RGB image of an apple, corresponding depth image, text description, and speech description of the same apple) as shown with green rectangles in \cref{fig:emma-loss-trimmed}
+    \item Negative (Instance): A set of embeddings of another data point of a different object (e.g., an RGB image of a mug, corresponding depth image, text description, and speech description of the same mug) as shown with orange rectangles in \cref{fig:emma-loss-trimmed}
     \item Anchor (Modality): Every modality of the positive set is chosen as the learning anchor once.
 In our formulation, anchor is a ``concept'' referring to the points in the positive sets rather than a single instance in itself.
+In \cref{fig:emma-loss-trimmed}, each of the 4 modalities is selected as an anchor once.
     %For our experiments, text is used. 
@@
 % When we minimize the distance between each pair of positives, and maximize the distance between each pair of modalities one from the positive set and one from the negative set, the model converges faster than \supcon{} while maintaining the same performance when we have all modalities and a little bit less than Supcon when we drop text.
-We refer to this approach as \geom{}, for extended multimodal alignment which is formulated in \cref{eq:full-emma}. An illustration of this loss function is provided in \cref{fig:emma-loss}.
+We refer to this approach as \geom{} which is formulated in \cref{eq:full-emma}. An illustration of this loss function is provided in \cref{fig:emma-loss-trimmed}.
 
@@
 \begin{equation}\label{eq:full-emma}
-    L = \sum_{m_1=1}^{M} \left[ \sum_{m_2=1}^{M} \left[ \max (\cos(z_{m_1}^{+}, z_{m_2}^{-}) -1 + \alpha,0) \right] + \sum_{m_3=m_1+1}^{M} \left[ \max(1 - \cos(z_{m_1}^{+},z_{m_3}^{+}),0) \right] \right]
+\begin{split}
+    L &= \sum_{m_1=1}^{M} \left[ \sum_{m_2=1}^{M} \left[ -\max (dist(z_{m_1}^{+}, z_{m_2}^{-}) + \alpha,0) \right] + \sum_{m_3=m_1+1}^{M} \left[ \max( dist(z_{m_1}^{+},z_{m_3}^{+}),0) \right] \right]
+\end{split}
 \end{equation}
@@
 In \cref{eq:full-emma}, $M$ is the number of modalities, the superscripts $+$ and $-$ represent positive and negative objects, $\alpha$ represents enforced margin between each positive and negative points which we set to 0.4 for all modalities without tuning, and $z$ is the embedding we get by applying a mapping function $f$, which in our case is a neural network on our input data.
-In other words, $z_m = f_m(x_m)$, where each modality $m$ has a specific model $f_m$ that is different from the models for other modalities.
-These models do not share their weights.
-The $\cos(\cdot)$ function is a measure of similarity, not distance, and that is why the signs are reversed.
+In other words, $z_m = f_m(x_m)$, where each modality $m$ has a specific model $f_m$ that is different from the models for other modalities. These models do not share their weights.
 
-To measure distance in embedded space, we use cosine similarity between pairs of embeddings, i.e.
+Cosine similarity is the opposite of distance, and we need to reverse the logic for maximization and minimization. 
+% 
+There are different options to measure distance in embedded space.
+We use cosine similarity between pairs of embeddings, i.e.
 we measure the cosine of the angle between embeddings.
@@
 Other distance metrics, such as Euclidean distance, grow in value with respect to their dimensionality, resulting in a very large distances for data points.
-Cosine similarity is the opposite of distance, and we need to reverse the logic for maximization and minimization.
+% The $\cos(\cdot)$ function is a measure of similarity, not distance, and that is why the signs are reversed.
+
+\begin{figure*}[tb]
+\centering
+% \includegraphics[width=.9\columnwidth]{Figures/4way-MMA.png}
+% \includegraphics[width=2.1\columnwidth]{Figures/e-MMA.png}
+\includegraphics[width=1\columnwidth]{Figures/emma-loss-trimmed.pdf}
+\caption{A high-level prototype of our approach and the distances used in the \geom{} loss with four modalities.
+Gray arrows indicate pre-trained models that are frozen (i.e.
+the parameters are fixed and are not trained).
+The black arrows show 3 fully connected layers with a ReLU activation function~\citep{relu2010} after the first two layers.
+These networks are trained.
+Orange rectangles are negative embeddings and green rectangles are positive embeddings.
+Dashed lines indicate distances to be maximized while dotted lines indicate distances to be minimized.%, and solid lines show the enforced margin.
+}
+\label{fig:emma-loss-trimmed}
+\end{figure*}
+
+Here, the generic \textit{dist} function is replaced with the specific $\cos(\cdot)$, and we omit the max notation for clarity by defining \cref{eq:aux-equation}:
+
+\begin{equation}\label{eq:aux-equation}
+    \begin{split}
+        & g(x,y) = \max(\cos(x,y)-1 + \alpha, 0 ) \\
+        & h(x,y) = \max(1- \cos(x,y), 0 ).
+    \end{split}
+\end{equation}
+
+The first portion of the following equation maximizes all unique pairwise distances between modalities of positive and negative instances.
+The second portion minimizes the unique pairwise distances among the modalities of positive instances.
+
+\begin{equation}\label{eq:objective-geom}
+\begin{split}
+    \mathcal{L} & = \sum_{m_1=1}^{M}  \sum_{m_2=1}^{M} g(z_{m_1}^{+}, z_{m_2}^{-})  + \sum_{m_1=1}^{M} \sum_{m_3=m_1+1}^{M} h(z_{m_1}^{+},z_{m_3}^{+})
+    %  &= \sum_{i=1}^{M-1} \sum_{j=i+1}^{M} dist( z_{i}^{+} , z_{j}^{+}) - dist( z_{i}^{+} , z_{j}^{-}) - dist( z_{i}^{-} , z_{j}^{+}) - \sum_{i=1}^{M} dist( z_{i}^{+} , z_{i}^{-} ) \\
+\end{split}    
+\end{equation}
+Our proposed \geom{} loss function in \cref{eq:objective-geom} can be rewritten as shown in \cref{eq:objective-geom-rewritten} by fully specifying the summations to better understand how our objective function can be reduced to well-known losses such as triplet loss and pairwise loss.
+
+\begin{equation}\label{eq:objective-geom-rewritten}
+\begin{split}
+     \mathcal{L} &= \sum_{i=1}^{M-1} \sum_{j=i+1}^{M} h( z_{i}^{+} , z_{j}^{+})
+     + g( z_{i}^{+} , z_{j}^{-}) + g( z_{i}^{-} , z_{j}^{+}) + \sum_{i=1}^{M} g( z_{i}^{+} , z_{i}^{-} ).
+\end{split}
+\end{equation}
+
+If $M=2$ which means the number of modalities is 2, and we ignore the last two terms in the derived objective function, it results in the triplet loss method. 
+If $M=2$, then our objective function reduces to the quadruplet loss method~\citep{tursun2021efficient,chen2017beyond} if we multiply the first term by 2, ignore the third term, and change the last summation to be up to $M-1$ (which results a single term).
+If $M=1$, only the last term remains in the loss function which is exactly the pairwise distance-based loss function.
+This loss function can be seen as a contrastive loss usually used in the domain of self-supervised learning~\citep{chen2020simple}.
+However, our proposed loss function has two advantages over the traditional contrastive loss expressed in \cref{eq:contrastive-loss}.
+The first advantage is that our loss function does not loop over multiple positives and negatives in a large batch.
+Instead we sample only two objects (positive and negative) each of which have $M$ modalities which gives us $2M$ datapoints (or embeddings).
+Hence, our model can be trained using smaller batch sizes and reduces the number of negative samples we need.
+The second advantage is that this loss function can be used in a multimodal setting with an arbitrary number of modalities, and is not limited to a single data type (e.g.
+RGB images) which is the most common usage of contrastive loss.
+% ~\citep{chen2020simple, NEURIPS2020_supervised_contrastive}.
+Although our \geom{} is technically quadratic in terms of number of modalities, we observe that experimentally, training time increases only by 10 minutes with each additional modality.
 
@@
 It is noteworthy that our training procedure does not perform any stochastic dropout of modalities to obtain test-time robustness to missing modalities.
-Moreover, our approach does not need to compute the distance between all items in the batch as opposed to \supcon{}~\cite{NEURIPS2020_supervised_contrastive}.
+Moreover, our approach does not need to compute the distance between all items in the batch, as opposed to \supcon{}.
+% ~\citep{NEURIPS2020_supervised_contrastive}.
 
-\begin{figure*}[tb]
-\centering
-% \includegraphics[width=.9\columnwidth]{Figures/4way-MMA.png}
-% \includegraphics[width=2.1\columnwidth]{Figures/e-MMA.png}
-\includegraphics[width=1\columnwidth]{Figures/emma-loss.pdf}
-\caption{A high-level prototype of our approach and the distances used in the \geom{} loss with four modalities.
-Gray arrows indicate pre-trained models that are frozen (i.e.
-the parameters are fixed and are not trained).
-The black arrows show 3 fully connected layers with a ReLU activitation function~\cite{relu2010} after the first two layers.
-These networks are trained.
-Orange rectangles are negative embeddings and green rectangles are positive embeddings.
-Dashed lines indicate distances to be maximized while dotted lines indicate distances to be minimized.%, and solid lines show the enforced margin.
-}
-\label{fig:emma-loss}
-\end{figure*}
-
-\subsection{Loss Development Story}
-\todokdinline{Should I move this whole section to supplementary or should I talk about the high-level takeaways in the main paper?}
-
-Our proposed loss function is a modification of the triplet loss function idea where we fix one modality as anchor and do not sample a positive instance from the same modality. 
-Na\"ively, to extend triplet loss to work with arbitrary number of modalities, we can use two anchors (e.g., text and speech), which results in $2(M-2)$ triplet losses as formulated in~\cref{eq:objective-two-anchors}.
-\todoffinline{in this equation block, you're writing things like $|| z_{t}^{+} - z_{m}^{+} ||$. Do you actually mean norm, or would any old scoring function work?}
-\todokdinline{This whole section will be moved to supplementary, but anyway, would it be better if I just get rid of the $||.||$ and use $\cos(x)$ instead?}
-
-\begin{equation}
-\label{eq:objective-two-anchors}
-\begin{split}
-    \mathcal{L}  &= \sum_{m=1}^{M-2} || z_{t}^{+} - z_{m}^{+} || - || z_{t}^{+} - z_{m}^{-} || + || z_{s}^{+} - z_{m}^{+} || - || z_{s}^{+} - z_{m}^{-} || \\
-    &= \sum_{m=1}^{M-2} \cos(z_{t}^{+} ,z_{m}^{-}) - \cos(z_{t}^{+}, z_{m}^{+}) + \cos(z_{s}^{+} ,z_{m}^{-}) - \cos(z_{s}^{+}, z_{m}^{+})
-\end{split}
-\end{equation}
-
-However, this approach performs poorly when one of the anchor modalities is ablated; since there is no explicit minimization between anchors, they do not necessarily map closely to each other, such that only one of the two anchors is actually learned.
-
-An alternative is to apply triplet loss $M-1$ times with a single anchor and $M-1$ `target' modalities.
-In this case, the negative anchor point is disregarded.
-The positive anchor is used as an anchor for all $M-1$ triplet losses, and for each of those triplet losses the positive and negative points are simply selected from the corresponding modalities.
-\Cref{eq:objective-simple-mma} formulates this idea.
-
-\begin{equation}
-\label{eq:objective-simple-mma}
-\begin{split}
-    \mathcal{L}  &= \sum_{m=1}^{M-1} || z_{a}^{+} - z_{m}^{+} || - || z_{a}^{+} - z_{m}^{-} || \\
-    &= \sum_{m=1}^{M-1} \cos(z_{a}^{+} ,z_{m}^{-}) - \cos(z_{a}^{+}, z_{m}^{+})
-\end{split}
-\end{equation}
-
-However, this approach fails to take into account the dissimilarity of the positive and negative anchor points.
-This is an important piece of information that \cref{eq:objective-simple-mma} ignores.
-Although this information is implicitly captured when we maximize the distance between positive anchor and other negative modalities, when the negative data point becomes a positive example itself, the anchor and those points are forced to be closer to each other which can happen in three ways: either moving anchor closer to those points, moving those points closer to anchor, or move both somewhere in between.
-Therefore, in the last two cases we do not have explicit direct control over the distance between these points and previous positive points.
-Our experimental results confirm this hypothesis.
-
-Therefore, we add an extra term which is responsible for maximizing the distance between positive and negative instances of the anchor modality, because a margin between positive and negative points from other modalities (excluding anchor) is enforced.
-% The first dashed line from the top in \cref{fig:emma-loss} is this extra term that captures the dissimilarity between positive and negative text datapoints.
-We formulate this method in \cref{eq:objective-explicit-fixed-anchor}, which is generalized to an arbitrary number of modalities.
-
-\begin{equation}
-\label{eq:objective-explicit-fixed-anchor}
-\begin{split}
-    \mathcal{L}  %&= %\sum_{m=1}^{M-1} || z_{a}^{+} - z_{m}^{+} || - || z_{a}^{+} - z_{m}^{-} ||  - || z_{a}^{+} - z_{a}^{-} || \\
-    &= \max(\cos(z_{a}^{+}, z_{a}^{-}) + \alpha_a, 0) + \sum_{m=1}^{M-1} \max \left(\cos(z_{a}^{+} ,z_{m}^{-}) - \cos(z_{a}^{+}, z_{m}^{+}) + \alpha_m, 0 \right) 
-\end{split}
-\end{equation}
-
-In \cref{eq:objective-explicit-fixed-anchor}, $a$ represents the \textit{anchor} modality, $M$ is the number of modalities, the superscripts $+$ and $-$ represent positive and negative objects, $\alpha_m$ represents enforced margin for each modality which we set to 0.4 for all modalities without tuning, and $z$ is the embedding we get by applying a mapping function $f$, which in our case is a neural network on our input data.
-In other words, $z_m = f_m(x_m)$, where each modality $m$ has a specific model $f_m$ that is different from the models for other modalities.
-These models do not share their weights.
-The $\cos(\cdot)$ function is a measure of similarity, not distance, and that is why the signs are reversed.
-
-Since the downstream task we are considering in the grounded language learning domain is to predict/retrieve a desired object among multiple objects given a natural language description, it makes sense to train the model using language as an anchor, and in fact anchoring learning around language outperforms anchoring on RGB.
-However, if use all modalities as anchor once (looping over modalities and compute distances between each pair), it results in a better performance.
-
-% Triplet loss cannot be used for more than 2 modalities.
-Some previous work has concatenated RGB and depth embeddings to create a single ``vision'' embedding for learning~\cite{triplet_loss_2021_CVPR}, but they cannot handle RGB or depth sensor ablation during test.
-Since our method can handle any number of modalities, we can handle such a case by treating depth as a separate modality.
-% Moreover, when using triplet loss with different anchors per batch, the batch size has to be 1 which makes training very slow.
-
-% This approach is most similar to that of supervised contrastive learning~\cite{NEURIPS2020_supervised_contrastive}, but outperforms both that and regular contrastive learning~\cite{chen2020simple} methods when modalities are ablated, most notably written language.
-
-% We find that the use of instance level losses grounded to one key modality (text) is sufficient to encourage a representation that satisfies our goals.
-
-\begin{comment}
-\subsubsection{Extended Triplet}
-Using an extended version of triplet loss to add all connections, which is formulated in \cref{eq:extended-triplet}, results in a worse performance.
-
-\begin{equation}\label{eq:extended-triplet}
-    L = \sum_{m_1=1}^{M} \left[ \sum_{m_2=m_1+1}^{M} \left[ \text{triplet}(z_{m_1}^{+}, z_{m_2}^{+}, z_{m_2}^{-}) + \text{triplet}(z_{m_1}^{-}, z_{m_2}^{-}, z_{m_2}^{+}) \right] + \cos(z_{m_1}^{+}, z_{m_1}^{-}) -1 + margin \right]
-\end{equation}
-Didn't change much when we drop text: Full eMMA when we have similar objects
-
-Did not work: Run the version with all connections (including among negatives) without triplet and just by using cosine distance between pair of embeddings
-\end{comment}
-
-\begin{comment}
-\subsubsection{Smaller Batch Sizes}
-
-Reducing the batch size from 64 to 16 (for both eMMA and supcon) increases the gap between our model and the baseline in cases where our model beats the baseline (all modalities, dropping speech), but baseline is still better than ours when we drop the text modality.
-
-Batch size of 8 (for both eMMA and supcon) had a similar effect.
-
-In batch size of 2 the trend is still the same, but when we drop the text modality, the MRR graph is flat on 0.4 which is still better than worse case and random case, but it seems model doesn't learn anything with regards to speech.
-To remedy this, we add pulling connections among the modalities of the negative instance as formulated in \cref{eq:full-emma-pull-negatives} and as shown in \cref{fig:full-emma-pull-negatives}, and that solves the flatline problem when we drop the text modality, however, when we have all modalities it performs worse than formulation in \cref{eq:full-emma}.
-
-\begin{equation}\label{eq:full-emma-pull-negatives}
-\begin{split}
-    L = & \sum_{m_1=1}^{M} \left[ \sum_{m_2=1}^{M} \left[ \max (\cos(z_{m_1}^{+}, z_{m_2}^{-}) -1 + margin,0) \right] + \sum_{m_3=m_1+1}^{M} \left[ \max(1 - \cos(z_{m_1}^{+},z_{m_3}^{+}),0) \right] \right.
-\\
-    & \left. + \sum_{m_3=m_1+1}^{M} \left[ \max(1 - \cos(z_{m_1}^{-},z_{m_3}^{-}),0) \right] \right]
-\end{split}
-\end{equation}
-
-\begin{figure*}[tb]
-\centering
-\includegraphics[width=1\columnwidth]{Figures/full-emma-pull-negatives.pdf}
-\caption{A high-level prototype of our approach and the distances used when we have full EMMA and pulling among negatives}
-\label{fig:full-emma-pull-negatives}
-\end{figure*}
-\end{comment}
-
-\begin{comment}
-\subsubsection{Margin}
-% \todokdinline{Try different margins of 0.6, 1.0, and if it improved the results we can try to learn it as a parameter}
-\todokdinline{Should we remove this section?}
-I also tried two different margins so far. A margin of 0.6 and a margin of 1.0. Margin of 0.6 does not change the performance much. However, it converges even faster.
-Margin of 1.0 (this is essentially just cosine similarity because we have: cos -1 +margin) results in a worse performance when we have all modalities.
-The MRR is 0.8995 when margin is 1, while with a margin of 0.6 the MRR is 0.9176, and with a margin of 0.4 the MRR is 0.9198.
-If we drop the text modality, the MRR is 0.7923, 0.7913, and 0.6918 for margins of 0.4, 0.6, and 1.0 respectively.
-\end{comment}
-
-\subsection{EMMA: Extended Multimodal Alignment}
-\label{sub:emma}
-
-\todokdinline{We introduce \supcon{} in Baselines section which is after this section while we use SupCon as a part of our loss function here.
-Should we move the introduction of SupCon to here?
-Or maybe we can discuss the baselines first, and then Approach section follows it?}
-\todocmi{I suggest a section 4.3, Baseline Approach: Supervised Contrastive Learning (pushing EMMA down)}
+\paragraph{Combining Geometric and Cross-Entropy Losses}
+\label{subsec:emma}
 
@@
 Both methods are valid and have some advantages that the other method does not offer.
-In \geom{}, the advantages are intuitive learning objective in terms of distance, interpretability of the learned embedding space, and faster convergence.
-The advantage of \supcon{} is that it uses classification objective which is aligned with the downstream task.
+In \geom{}, the advantages are an intuitive learning objective in terms of distance, interpretability of the learned embedding space, and faster convergence.
+The advantage of \supcon{} is that it uses a classification objective which is aligned with the downstream task.
 
-We propose to combine the \geom{} and \supcon{} and we refer to this approach as \ours{}, for extended multimodal alignment. \ours{} loss function is formulated in \cref{eq:emma}.
+We propose to combine the \geom{} defined in \cref{eq:objective-geom} with \supcon{} defined in \cref{eq:supervised-contrastive}, and we refer to this approach as \ours{}, for extended multimodal alignment.
+The combination is done by taking the sum of the equations for \geom{} and \supcon{} which is shown in \cref{eq:emma}.
+When
 
 \begin{equation}\label{eq:emma}
-    % \mathcal{L} = \geom{} + \supcon{}
-    \ours{} = \geom{} + \supcon{}
+\begin{split}
+    \mathcal{L} =  \sum_{i \in I} \Biggr[ & \Bigr[ \sum_{j=1}^{M-1} \sum_{k=j+1}^{M} h( z_{i,j}^{+} , z_{i,k}^{+})
+     + g( z_{i,j}^{+} , z_{i,k}^{-}) + g( z_{i,j}^{-} , z_{i,k}^{+}) + \sum_{j=1}^{M} g( z_{i,j}^{+} , z_{i,j}^{-} )\Bigr] + \\
+     & \Bigr[ \sum_{m=1}^{M} \frac{-1}{|P(i,m)|}  \sum_{\beta \in P(i,m)} \log \frac{\exp (z_{i,m} \cdot z_{\beta} / \tau) }{\sum_{\gamma \in A(i,m)} \exp (z_{i,m} \cdot z_{\gamma} / \tau)} \Bigr] \Biggr],
+     \end{split}
 \end{equation}
+where $A(i,m)$ includes all items in the batch except for the $z_{i,m}$ itself and $P(i,m)$ includes all the modalities of all instances that have the same label as current instance excluding $z_{i,m}$ itself.
+In other words, the collection of embeddings indexed by $P(i,m)$ is $\{ \bigcup\limits_{r \neq m \in M} z_{i,r}, \bigcup\limits_{l \neq i \in I , y_i = y_l} \bigcup\limits_{m=1}^{M} z_{l,m} \} $.
 
-Combining these two losses results in faster convergence and better performance when we have all modalities available, and maintaining a good performance when modalities are missing.
+% \todo[inline]{Which one do you prefer?
+The first one is my attempt at using almost the same terminology for both SupCon and Geometric method, while the second one is simply copying and pasting the two equations next to each other.}
 
-\subsection{High-Level Findings/Observations}
-We find that using each modality as anchor results in a better performance compared to when we fix one modality as anchor. This difference is more significant when we test with all modalities.
+% \begin{equation}\label{eq:emma-append}
+%     \mathcal{L} = \Bigr[ \sum_{i=1}^{M-1} \sum_{j=i+1}^{M} -\cos( z_{i}^{+} , z_{j}^{+})
+%      + \cos( z_{i}^{+} , z_{j}^{-}) + \cos( z_{i}^{-} , z_{j}^{+}) + \sum_{i=1}^{M} \cos( z_{i}^{+} , z_{i}^{-} ) \Bigr] + 
+%      \Bigr[ \sum_{i \in I} \frac{-1}{|P(i)|} \sum_{p \in P(i)} \log \frac{\exp (z_i \cdot z_p / \tau) }{\sum_{a \in A(i)} \exp (z_i \cdot z_a / \tau)} \Bigr]
+% \end{equation}
 
-While the performance of SupCon deteriorates as the batch size is reduced, our \geom{} stays the same when we have all modalities or when we drop speech.
-However, when we drop text, EMMA deteriorate when we decrease batch size, and with a batch size of 2 without pulling negatives, it flatlines.
-In this case (when batch size is 2 and text is dropped), pulling among negatives is important and results in a better MRR.
+% 
+% \begin{comment}
+% 
+% \begin{equation}\label{eq:emma}
+%     % \mathcal{L} = \geom{} + \supcon{}
+%     \ours{} = \geom{} + \supcon{}
+% \end{equation}
+% \end{comment}
+% 
+Compared to pure \supcon{}, combining these loss functions results in faster convergence and slightly improved performance when all modalities are available, and maintains improved performance when modalities are missing.
+Experimental results are presented in detail in \cref{sec:results}.
 
-We discuss the geometric interpretation of our method versus the classification interpretation of SupCon method and advantages of each.
+% We discuss the geometric interpretation of our method versus the classification interpretation of SupCon method and advantages of each.
 Finally, we show that combining \geom{} with \supcon{} is superior to either the \geom{} approach or \supcon{} alone.
@@
 
-Transformers are de facto architectures in natural language processing community and have shown great success across different tasks.
-Similar to~\citet{GoLD_UMBC}, we use BERT~\cite{devlin-etal-2019-bert} embeddings contained in the FLAIR library ~\cite{akbik2019flair,akbik-etal-2019-pooled} to featurize the textual input, and wav2vec2~\cite{wav2vec2} to extract audio embeddings from speech.
+Transformers have become \textit{de facto} architectures in the natural language processing community and have shown great success across different tasks.
+Similar to~\citet{GoLD_UMBC}, we use BERT~\citep{devlin-etal-2019-bert} embeddings contained in the FLAIR library~\citep{akbik2019flair,akbik-etal-2019-pooled} to featurize textual input, and wav2vec2~\citep{wav2vec2} to extract audio embeddings from speech.
 Both of these encoders output a 3072-dimensional embedding vector which is generated by concatenating the last four hidden layers of their corresponding networks.
-Both BERT~\cite{devlin-etal-2019-bert} and wav2vec2~\cite{wav2vec2} are self-supervised language models using transformers~\cite{vaswani2017attention}.
-% \todo[inline]{A couple of sentences about FLAIR and wav2vec2 and what they do}
+FLAIR has historically been used for different natural language processing tasks such as named entity recognition (NER) and part-of-speech tagging (PoS), and wav2veq2 has supported a a number of audio processing tasks, most notably automated speech recognition.
+Both BERT~\citep{devlin-etal-2019-bert} and wav2vec2~\citep{wav2vec2} are self-supervised language models using transformers~\citep{vaswani2017attention}.
+% 
+To process images, we use ResNet152~\citep{He_resnet_2016} for both RGB and depth images which gives us a 2048-dimensional embedding vector. Depth images are colorized before passing to the ResNet152.
 
-To process images, we use ResNet152~\cite{He_resnet_2016} for both RGB and depth images which gives us a 2048-dimensional embedding vector. Depth images are colorized before passing to the ResNet152.
-
-We then use different instantiations of the same multi-layer perceptron (MLP) consisting of 3 fully connected layers with ReLU activation~\cite{relu2010} to map each of these embeddings to a shared 1024-dimensional space where we can compute the distances between all embeddings.
-We note that these MLP networks are different and do not share any weights.
+We then use different instantiations of the same multi-layer perceptron (MLP) consisting of 3 fully connected layers with ReLU activation~\citep{relu2010} to map each of these embeddings to a shared 1024-dimensional space where we can compute the distances between all embeddings.
+We note that these MLP networks are distinct and do not share any weights.
 
@@
 In this section we evaluate the quality of object retrieval models learned using the EMMA loss function.
-We first describe the dataset we use, give the metrics by which we evaluate the performance, the setup of the experiments, the baselines we compare against, and we end the section by a presenting results and analyzing them.
+We first describe the dataset we use, then describe the metrics by which we evaluate performance, the setup of the experiments, and the baselines against which we compare.
+We end by presenting and analyzing results.
 
@@
 
-We demonstrate the effectiveness of our approach on a recent publicly available multimodal dataset called GoLD~\cite{GoLD_UMBC}, which contains RGB image, depth image, written text descriptions, speech descriptions, and transcribed speech descriptions for 207 object instances across 47 object classes (see \cref{fig:emma-loss}).
-There are a total of 16,500 speech and 16,500 textual descriptions.
-The original GoLD paper uses raw RGB and raw depth images in which other objects are present in the background.
-We use a masked version of the images where the background is deleted (we observe that a masked version converges faster, however, they both converge to the same performance).
-Speech is converted to 16 Hz to match the wav2vec2 speech model~\cite{wav2vec2}.
+We demonstrate the effectiveness of our approach on a recent publicly available multimodal dataset called GoLD~\citep{GoLD_UMBC}, which contains RGB images, depth images, written text descriptions, speech descriptions, and transcribed speech descriptions for 207 object instances across 47 object classes (see \cref{fig:emma-loss-trimmed}).
+There are a total of 16,500 spoken and 16,500 textual descriptions.
+The original GoLD paper uses raw RGB and depth images in which other objects are present in the background.
+We use a masked version of the images where the background is deleted (this masked version converges faster, however, masked and unmasked versions of the GoLD data converge to the same performance).
+Speech is converted to 16 Hz to match the wav2vec2 speech model.
 
-\subsection{Setup and Metrics}
-\label{sec:setupmetrics}
-To evaluate our model we measure different performance metrics on a retrieval task where the model has to select an object among 5 objects given a language description.
-Only one of the objects corresponds to the description and the rest are from different object classes (e.g., one apple among a set including a fork, a mug, a lemon, and a bell pepper).
+\subsection{Setup}
+% \label{sec:setupmetrics}
+\label{sec:setup}
+To evaluate our model we measure different performance metrics on a retrieval task where the model has to select an object from a set of objects given a language description.
+Only one of the objects corresponds to the description and the rest are from different object classes.
+% (e.g., one apple among a set including a fork, a mug, a lemon, and a bell pepper).
+% 
+Similar to~\citet{NEURIPS2020_supervised_contrastive}, we use a stochastic gradient descent (SGD) optimizer with momentum~\citep{ruder2016overviewSGD} with a flexible learning rate starting at 0.05. 
+% For the triplet loss experiments we use the Adam~\citep{kingma_adam_2015} optimizer with scheduling learning rate as suggested by the papers.
+All models are trained for 200 epochs with a batch size of 64 on a Quadro RTX 8000 GPU.
+We used a temperature of $0.1$ for training the contrastive learning method described in \cref{sub:baseline-contrastive}, and a temperature of $0.07$ for training \supcon{} as described in \cref{sub:baseline-supcon}.
 
-\subsubsection{Setup}
-\label{sec:setup}
-Similar to~\citet{NEURIPS2020_supervised_contrastive}, we use a stochastic gradient descent (SGD) optimizer with momentum~\cite{ruder2016overviewSGD} with a flexible learning rate starting at 0.05. 
-% For the triplet loss experiments we use the Adam~\cite{kingma_adam_2015} optimizer with scheduling learning rate as suggested by the papers.
-All models are trained for 200 epochs with a batch size of 64 on a Quadro RTX 8000 GPU.
-
-To evaluate the performance, we compute the distance between the given natural language description and 5 randomly selected images (1 of which corresponds to the description, with the others from different object classes).
-We compute the distance between one language embedding and all candidate RGB embeddings, and we compute the distance between the same language embedding and all candidate depth embeddings corresponding to the RGB embeddings.
+To evaluate the performance, we compute the distance between the given natural language description and 5 randomly selected objects (1 of which corresponds to the description, with the others from different object classes).
+We compute the distance between the language embedding and all available sensory modalities of all candidates as described in \cref{sub:ablations}.
+In case we have RGB and depth, we compute the distance between language embedding and all candidate RGB embeddings, and we compute the distance between the same language embedding and all candidate depth embeddings corresponding to the RGB embeddings.
 We then take average of these two distance matrices.
@@
 
-\subsubsection{Metrics}
+\subsection{Metrics}
 \label{sec:metrics}
@@
 For each query we predict the rank of all objects based on their distance from the language command, and then the inverse rank of the desired objects in all queries are averaged.
-For example, if the model predicts the desired object as the first rank, then MRR = $\frac{1}{1} = 1$ which means a perfect score, and if it predicts the correct object as the last rank among five objects, then MRR = $\frac{1}{5}=0.2$.
+For example, if the model predicts the desired object as the first rank, then MRR = $\frac{1}{1} = 1$ which means a perfect score, and if it predicts the correct object as the fourth rank among five objects, then MRR = $\frac{1}{4}=0.25$.
 
@@
 
+While MRR is more meaningful when it comes to ranking in retrieval tasks, in the real-world scenarios where a robot is asked to hand over an object, if it fails, it does not matter whether the correct object was ranked second or last and the whole system would be considered a failure.
 Accuracy and micro F1 score
@@
 MRR is a more informative metric because it captures the idea that having the correct object as the second choice should be considered better than having it as a last choice, while in accuracy the score is ``all or nothing''---either 0 or 1.
-Because our approach is designed to be robust to missing information across modalities, we also report MRR for different combinations of modality dropouts.
-
-\subsection{Baselines}
-We compare both \ours{} and \geom{} against contrastive learning~\cite{chen2020simple} and supervised contrastive learning~\cite{NEURIPS2020_supervised_contrastive} which we refer to as \supcon{}.
-We consider \supcon{} as the main baseline since it is a general version of multiple contrastive loss functions including triplet loss, the traditional version of contrastive loss usually used in self-supervised settings~\cite{chen2020simple}, and N-pair loss~\cite{NIPS2016N-PairLoss}.
-% and do not include N-pair and triplet loss in our comparisons.
-%contrastive learning methods including triplet loss, the traditional version of contrastive loss usually used in self-supervised settings~\cite{chen2020simple}, and supervised contrastive learning~\cite{NEURIPS2020_supervised_contrastive}.
-
-\begin{table*}[]
-\centering
-\resizebox{.97\textwidth}{!}{
-\begin{tabular}{lc|ll|lll}
-\multicolumn{1}{c}{} &  & \multicolumn{2}{c|}{\textbf{Accuracy}} & \multicolumn{3}{c}{\textbf{Mean Reciprocal Rank}} \\ \hline
-\multicolumn{1}{c|}{\textbf{Method}} & \textbf{Modalities} & \multicolumn{1}{c|}{\textbf{\begin{tabular}[c]{@{}c@{}}Speech/RGB/\\ Depth\end{tabular}}} & \multicolumn{1}{c|}{\textbf{\begin{tabular}[c]{@{}c@{}}Text/RGB/\\ Depth\end{tabular}}} & \multicolumn{1}{c|}{\textbf{\begin{tabular}[c]{@{}c@{}}Speech/RGB/\\ Depth\end{tabular}}} & \multicolumn{1}{c|}{\textbf{\begin{tabular}[c]{@{}c@{}}Text/RGB/\\ Depth\end{tabular}}} & \multicolumn{1}{c}{\textbf{\begin{tabular}[c]{@{}c@{}}Text/Speech/\\ RGB/Depth\end{tabular}}} \\ \hline
-\multicolumn{1}{l|}{EMMA} & TSRD & \multicolumn{1}{l|}{\textbf{0.65}$\pm$0.01} & 0.83$\pm$0.006 & \multicolumn{1}{l|}{\textbf{0.79}$\pm$0.002} & \multicolumn{1}{l|}{0.90$\pm$0.003} & 0.90$\pm$0.004 \\
-\multicolumn{1}{l|}{EMMA} & 3 & \multicolumn{1}{l|}{--} & 0.84$\pm$0.008 & \multicolumn{1}{l|}{--} & \multicolumn{1}{l|}{0.90$\pm$0.004} & -- \\
-\multicolumn{1}{l|}{\supcon{}} & TSRD & \multicolumn{1}{l|}{0.59$\pm$0.007} & 0.84$\pm$0.008 & \multicolumn{1}{l|}{0.75$\pm$0.004} & \multicolumn{1}{l|}{0.91$\pm$0.006} & 0.90$\pm$0.006 \\
-\multicolumn{1}{l|}{\supcon{}} & 3 & \multicolumn{1}{l|}{--} & 0.84$\pm$0.006 & \multicolumn{1}{l|}{--} & \multicolumn{1}{l|}{0.91$\pm$0.004} & -- \\
-\multicolumn{1}{l|}{Contrastive Cosine} & TSRD & \multicolumn{1}{l|}{0.40$\pm$0.153} & 0.83$\pm$0.006 & \multicolumn{1}{l|}{0.62$\pm$0.034} & \multicolumn{1}{l|}{0.90$\pm$0.004} & 0.89$\pm$0.005 \\
-\multicolumn{1}{l|}{Contrastive Cosine} & 3 & \multicolumn{1}{l|}{--} & 0.84$\pm$0.005 & \multicolumn{1}{l|}{--} & \multicolumn{1}{l|}{0.91$\pm$0.003} & -- \\
-\multicolumn{1}{l|}{\ours{}} & TSRD & \multicolumn{1}{l|}{--} & 0.83$\pm$0.008 & \multicolumn{1}{l|}{--} & \multicolumn{1}{l|}{0.90$\pm$0.005} & -- \\ \hline
-\end{tabular}
-}
-\caption{Average and standard deviation of accuracy and MRR over 3 runs with 3 different random seeds on a held-out test set.
-These results represent performance when no modalities are ablated, in which case supervised contrastive loss, contrastive loss, and EMMA perform similarly (``--'' is an ablated modality and can't be evaluated).
-All metrics are from 0 to 1, and higher is better.
-For 5 objects, a random guess would have MRR of 0.33, and the worst case performance would be always ranking the correct item last which gives us 0.2.
-All models are trained using text as anchor.
-The batch size is 64 and optimizer is SGD for all experiments except the triplet loss, where batch size is 1 and optimizer is Adam.
-%In the Triplet experiment, RGB and Depth embeddings are concatenated to form a single vision embedding since triplet loss cannot be used for more than two modalities.
-\todokdinline{Should we keep Acc or is MRR enough since table is out of bounds?
-Also, should we include other training modalities in the table?
-I inserted the results in a file called results-table.tex for now so that we can decide later.}
-}
-\todocmi{Let's include everything, and play tricks to shrink the table horizontally. If necessary it can be split into two tables with the same y headers and different x headers.}
-\todoffinline{I've added a resizebox in order to get this to fit.
-This is not the desired solution (though currently the text isn't that small, so it could be okay) but I needed to see the entire table easily.}
-\label{table:quantitative}
-\end{table*}
-% %Edward moved table above where refernced b/c LaTeX plays games with double-wide objects
-% \begin{table*}[bth]
-% \centering
-% \begin{tabular}{c|c|c|c|c|c|c}
-% \toprule
-% \textbf{Method} & \textbf{Modalities} & \textbf{acc.srd} & \textbf{acc.trd} & \textbf{MRR.srd} & \textbf{MRR.trd} & \textbf{MRR.tsrd} \\ %\hline
-% \midrule
-% EMMA (Text) & 4 & \textbf{0.65}$\pm$0.01 & 0.83$\pm$0.006 & \textbf{0.79}$\pm$0.002 & 0.90$\pm$0.003 & 0.90$\pm$0.004 \\
-% EMMA (Text) & 3 & - & 0.84$\pm$0.008 & - & 0.90$\pm$0.004 & - \\
-% % Simple MMA (Text)  & 3 & ?$\pm$? & ?$\pm$? & ? & ? \\
-% % Simple MMA (Text) & 4 & ?$\pm$? & ?$\pm$? & ? & ? \\
-% % Simple MMA (Speech) & 3 & ?$\pm$? & ?$\pm$? & ? & ?  \\
-% DONE Supervised Contrastive (Text)& 4 & 0.59$\pm$0.007 & 0.84$\pm$0.008 & 0.75$\pm$0.004 & 0.91$\pm$0.006 & 0.90$\pm$0.006  \\
-% TODO Supervised Contrastive (Text)& 3 & - & 0.84$\pm$0.006 & - & 0.91$\pm$0.004 & -\\
-% DONE Contrastive Cosine (Text) & 4 & 0.40$\pm$0.153 & 0.83$\pm$0.006 & 0.62$\pm$0.034  & 0.90$\pm$0.004 & 0.89$\pm$0.005\\
-% TODO Contrastive Cosine (Text) & 3 & - & 0.84$\pm$0.005 & -  & 0.91$\pm$0.003 & -\\
-% PROJ Triplet (Text) & 3 & - & 0.83$\pm$0.008 & - & 0.90$\pm$0.005 & -  \\
-% % DONE EMMA (Speech) & 3 & ?$\pm$? & ?$\pm$? & - & - \\
-% % DONE EMMA (Speech) & 4 & ?$\pm$? & ?$\pm$? & ? & ? \\
-% % DONE Supervised Contrastive (Speech)& 3 & ?$\pm$? & ?$\pm$? & - & -\\
-% % DONE Supervised Contrastive (Speech)& 4 & ?$\pm$? & ?$\pm$? & ? & ? \\
-% % TODO Contrastive Cosine (Speech) & 3 & ?$\pm$? & ?$\pm$?  & - & -\\
-% % TODO Contrastive Cosine (Speech) & 4 & ?$\pm$? & ?$\pm$?  & ? & ?\\
-% % DONE Triplet (Speech) & 3 & ?$\pm$? & ?$\pm$? & - & -  \\
-
-% % Full MMA (Text) w/ neg & 3 & ?$\pm$? & ?$\pm$? \\
-% % Simple MMA (Text) w/ neg Adam & 3  & 0.7949$\pm$? & 0.8778$\pm$? \\
-% % Simple MMA (Text) w/ neg SGD & 3 & 0.7905$\pm$? & 0.8821$\pm$? \\
-% % Triplet (Text) w/ neg & 3 & 0.7497$\pm$? & 0.8522$\pm$? \\
-% % Full MMA (Speech) w/ neg & 3 & ?$\pm$? & ?$\pm$? \\
-% % Simple MMA (Speech) w/ neg & 3 & 0.7949$\pm$? & 0.8778$\pm$? \\
-% % Contrastive Cosine (Text) w/ neg & 3 & 0.7894$\pm$? & 0.8788$\pm$?  & ? & ?\\
-% % Contrastive (Speech) w/ neg & 3 & ?$\pm$? & ?$\pm$? \\
-% % Triplet (Speech) w/ neg & 3 & 0.7497$\pm$? & 0.8522$\pm$? \\
-% \bottomrule
-% \end{tabular}
-% \caption{\label{table:quantitative}Average and standard deviation of accuracy and MRR over 3 runs with 3 different random seeds on a held-out test set.
-These results represent performance when no modalities are ablated, in which case supervised contrastive loss, contrastive loss, and EMMA perform similarly.
-All metrics are from 0 to 1, and higher is better.
-For 5 objects, a random guess would have MRR of 0.33, and the worst case performance would be always ranking the correct item last which gives us 0.2.
-All models are trained using text as anchor.
-The batch size is 64 and optimizer is SGD for all experiments except the triplet loss, where batch size is 1 and optimizer is Adam.
-% % \todokdinline{maybe adding results on cropped and raw dataset as well. Also results using object class instead of negative sampling}
-% }
-% \end{table*}
-
-\subsubsection{Triplet Loss}
-\label{sub:baseline-triplet}
-\todokdinline{Since we do not include triplet in our analysis, it might make more sense to move this section into the introduction of Approach to motivate why we came up with EMMA, thoughts?}
-
-The triplet loss function consists of three data points including anchor, positive, and negative from two modalities.
-The anchor, positive, and negative can be chosen from different modalities in each batch.
-The major disadvantage of the triplet loss method is that it cannot be used for more than two modalities. 
-%Second, a batch size of greater than 1 is hard to implement if the anchor, positive, and negative come from different modalities for each batch.
-% To address these issues in EMMA, we sample more than one positive and negative data points.
-\todokd{If moved to approach section, build on this to introduce our geometric loss function.}
-To address this issue, in \geom{} we use pairwise distance optimization for all data points. 
-%Intuitively, we sample one positive data point and one negative data point, each of which includes $M$ modalities, resulting in 2M data points.
-Each of the $M$ positives modalities become an anchor once.
-% \Cref{table:quantitative} shows that our MMA method perform as good as this baseline while our model can be trained 10x faster since we can use batch sizes larger than 1. Also, 
-Our method can be used for any number of modalities, while previous works that use triplet loss~\cite{GoLD_UMBC,triplet_loss_2021_CVPR} concatenate RGB and depth to form a single ``vision'' vector to handle three modalities, but they cannot handle RGB or depth sensor ablation during test.
-Since our method can handle any number of modalities, we can handle such a case by treating depth as a separate modality.
-Moreover, our method has the advantage that does not require triplet mining or providing hard negative examples.
-% When we used a fixed anchor modality, a batch size of 64, and applied semantic negative sampling~\cite{Pillai_Matuszek_2018}, our method outperformed this baseline. However, 
-Since this method cannot be applied to four modalities, we do not include it in the main analysis.
-
-\subsubsection{Contrastive Loss}
-\label{sub:baseline-contrastive}
-
-We compare our model against contrastive learning method presented by~\citet{chen2020simple} where they use the normalized
-temperature-scaled cross entropy loss (NT-Xent).
-In order to implement this loss function, we use cosine similarity as suggested in the SimCLR paper~\cite{chen2020simple}.
-Another possibility is to use an inner dot product~\cite{NEURIPS2020_supervised_contrastive}; if not normalized, this can lead to numerical instabilities and overflow/underflow since the dot product is not bounded, but the result is the same whether we use normalized inner dot product or cosine similarity.
-Contrastive loss function is formulated in \cref{eq:contrastive-loss}:
-% \todokdinline{try contrastive loss with cross-entropy, dot product, and the supervised contrastive method itself (without triplets).}
-
-% Original Formula
-\begin{equation}\label{eq:contrastive-loss}
-    -\sum_{i \in I} \log \frac{\exp (sim(z_i , z_{j(i)}) / \tau) }{\sum_{a \in A(i)} \exp (sim(z_i, z_a) / \tau)}
-\end{equation}
-% \begin{equation}\label{eq:contrastive-loss}
-%     -\sum_{i \in I} \log \frac{\exp (sim(f(x_i) , z_{j(i)}) / \tau) }{\sum_{a \in A(i)} \exp (sim(f(x_i), f(x_a)) / \tau)}
-% \end{equation}
-where $i$ is anchor, $j(i)$ is the set of positives excluding anchor, $a$ is the set of all positives and negatives excluding anchor, and $z = f(x)$.
-
-We can treat different modalities of the same instance as ``augmentations'' of that instance, and consider them as the positive points for the anchor.
-\Cref{eq:contrastive-loss} can be rewritten with the sum over positives explicitly shown as formulated in \cref{eq:explicit-contrastive}.
-
-%, which is equivalent to ``augmented views'' in the contrastive loss terminology.
-
-\todokdinline{Notation for M - i, it should be something like other modalities of the anchor but it shouldn't be over number of modalities.}
-\begin{equation}\label{eq:explicit-contrastive}
-    - \sum_{i \in I} \sum_{p=1}^{M \setminus \{i\} } \log \frac{\exp (sim(z_i , z_{p})/ \tau) }{
-    \sum_{a \in A(i)} \exp (sim(z_i, z_a) / \tau)}
-    %{ \exp (sim(z_i^+ , z_{m}^{+}) / \tau) + \exp (sim(z_i^+, z_{m}^{-}) / \tau)}
-\end{equation}
-where $M$ is the number of modalities (e.g. RGB image, depth image, speech, text) and $z = f(x)$.
-
-\subsubsection{Supervised Contrastive Learning}
-\label{sub:baseline-supcon}
-
-\citet{NEURIPS2020_supervised_contrastive} extend the contrastive learning method (NT-Xent) and propose a supervised way of performing contrastive learning to treat not only augmentations of the anchor, but also every item that shares the same label with anchor as positives.
-This loss function is shown in \cref{eq:supervised-contrastive}.
-
-\begin{equation}\label{eq:supervised-contrastive}
-    \sum_{i \in I} \frac{-1}{|P(i)|} \sum_{p \in P(i)} \log \frac{\exp (z_i \cdot z_p / \tau) }{\sum_{a \in A(i)} \exp (z_i \cdot z_a / \tau)}
-\end{equation}
-
-% \begin{equation}\label{eq:supervised-contrastive-my-notation}
-%     \sum_{i \in I} \frac{-1}{|P(i)|} \sum_{p \in P(i)} \log \frac{\exp (f_i(x_i) \cdot f_p(x_p) / \tau) }{\sum_{a \in A(i)} \exp (f_i(x_i) \cdot f_a(x_a) / \tau)}
-% \end{equation}
-
-Although, this loss function do not use cosine similarity, they normalize the embeddings before performing dot product which is equivalent to cosine similarity.
-
-The main difference between the contrastive loss baseline in \cref{sub:baseline-contrastive} and \supcon{} is that there is no notion of meaningful negative points in contrastive loss, and everything in the batch that is not the anchor or one of the positive views is considered to be negative.
-In \supcon{}, however, all elements in the batch that have same label as the anchor, are also considered positives in addition to different views of the same instance.
-While the denominators of \cref{eq:explicit-contrastive,eq:supervised-contrastive} stay the same, this subtle difference affects the numerator and includes more positive examples.
-
-While this model is a strong baseline, the authors applied it to a unimodal dataset.
-In this paper we extend this baseline to work with a multimodal dataset and show that it is slower than \ours{} to learn and its performance is also lower when all modalities are available during test time.
-% if modalities are ablated, this method does not perform as well as our proposed model (we show and discuss this later, in \cref{fig:epochs-mrr.srd}).
-This is especially true when the text modality is dropped, as is likely in physical agent scenarios, where speech is a natural interaction mechanism.
+Because our approach is designed to be robust to missing information across modalities, we also report MRR and accuracy for different combinations of modality dropouts.
 
 \subsection{Modality Ablation}
+\label{sub:ablations}
 % \todokdinline{Should I remove this section altogether or should I just mention that our method is equal or better in all cases except when we drop text?
@@
 However, when speech is incorporated as an additional fourth sensory modality, we have three possible choices.
-\textit{First}, we can compute the distance of RGB and depth from text and from speech which gives us 4 distance matrices, and then take average of these four.
-\textit{Second}, we can treat speech in a similar way to RGB and depth: compute the distance of RGB, depth, and speech from text, and then take an average of three of them.
-\textit{Third}, similar to the first method, but adding the distance between language and speech as well and then take the average of 5 distance matrices.
+First, we could compute the distance of RGB and depth from text and from speech which gives us 4 distance matrices, and then take average of these four.
+Second, we could treat speech in a similar way to RGB and depth: compute the distance of RGB, depth, and speech from text, and then take an average of three of them.
+Third, we could compute distances similarly to the first method, but add the distance between language and speech as well and then take the average of 5 distance matrices.
 
@@
 At test time, depending on which modalities are available to the model, we can use speech, text, or both to compute the distance of RGB and depth embeddings from the linguistic query, and then take the average.
-\Cref{eq:distance-matrix} shows all possible cases of modality dropout and the corresponding distance computations, where ``t'' represents \textit{text}, ``s'' represents \textit{speech}, ``r'' represents \textit{RGB}, ``d'' represents \textit{depth}, $K_{ij}$ for each pair of modalities is the distance of one instance in modality $i$ from all candidate instances in modality $j$, and $K$ is the final distance---a matrix if there are multiple queries and a vector if there is one query.
+
+There are total of nine possible cases of modality dropout and the corresponding distance computations.
+In all these cases \textit{t} represents text, \textit{s} represents speech, \textit{r} represents RGB, \textit{d} represents depth, and $K$ is the final distance---a matrix if there are multiple language queries and a vector if there is one query.
+If we only have two modalities, we simply compute the distance between those two modalities; this corresponds to four of the cases; $K_{tr}$ when speech and depth are missing, $K_{sr}$ when text and depth are missing, $K_{td}$ when speech and RGB are missing, and $K_{sd}$ when text and RGB are missing.
+If we have three modalities, we need to take average of two distances.
+There are four cases with three modalities, $K_{trd} = \frac{K_{tr} + K_{td}}{2}$ when speech is missing, $K_{srd} = \frac{K_{sr} + K_{sd}}{2}$ when text is missing,
+and so on.
+% $K_{tsr} = \frac{K_{tr} + K_{sr}}{2}$ when depth is missing, and $K_{tsd} = \frac{K_{td} + K_{sd}}{2}$ when RGB is missing. 
+When we have all modalities available we take average of four distances, $K_{tsrd} = \frac{K_{tr}+ K_{td} +K_{sr} + K_{sd}}{4}$.
+
+% \Cref{eq:distance-matrix} shows all possible cases of modality dropout and the corresponding distance computations, where \textit{t} represents text, \textit{s} represents speech, \textit{r} represents RGB, \textit{d} represents depth, $K_{ij}$ for each pair of modalities is the distance of one instance in modality $i$ from all candidate instances in modality $j$, and $K$ is the final distance---a matrix if there are multiple queries and a vector if there is one query.
+
+\begin{comment}
 \begin{equation}
@@
 \end{equation}
+% \todo[inline]{equation \cref{eq:distance-matrix} takes a lot of space which suggests it's important. Either merge it in text or ...}
+\end{comment}
 
-\Cref{fig:epochs-mrr.lard,fig:epochs-mrr.srd,fig:epochs-mrr.ld,fig:epochs-mrr.lr} show the relative performance of \ours{} and \geom{} against baselines when different modalities are ablated.
+\Cref{fig:result_graphs} shows the relative performance of \ours{} and \geom{} against state-of-the-art methods when different modalities are ablated.
 
-\subsection{Results}
+\section{Results}
 \label{sec:results}
 
-In this section we provide quantitative and qualitative results by comparing our method against supervised contrastive learning~\cite{NEURIPS2020_supervised_contrastive} and na\"ive contrastive loss~\cite{chen2020simple}.
+In this section we provide quantitative and qualitative results by comparing our method against supervised contrastive learning~\citep{NEURIPS2020_supervised_contrastive} and na\"ive contrastive loss~\citep{chen2020simple}.
 % We evaluate our EMMA loss function against the na\"ive triplet loss method.
@@
 % \todo[inline]{Why do we call out triplet loss specifically? We evaluate against several baselines? I don't understand this section particularly}
-To provide a better sense of the performance measure, we can think of a model that always ranks the correct object in the second place which is not very bad.
-Such a model would have MRR of $1/2 = 0.5$.
+To provide a better sense of the performance measure, consider a model that always ranks the correct object in the second place: Such a model would have an MRR of $1/2 = 0.5$.
 
-\Cref{fig:epochs-mrr.lard} Shows that \ours{} learns faster and results in a better performance compared to both \supcon{} and contrastive learning~\cite{chen2020simple} when trained using all modalities and all modalities are available during test.
-We observe that not only does contrastive loss learn more slowly, but that it is prone to overfitting; while this can be addressed with careful tuning of the learning process, it is better not to have to.
+\begin{table*}[tb]
+    \centering
+    \begin{subtable}[t]{1.00\textwidth}
+    \centering
+    \resizebox{1.0\textwidth}{!}{
+    \input{tableMRR}
+    }
+    \caption[]{Average and standard deviation of mean reciprocal rank (MRR) on a held-out test set. MRR is from $\frac{1}{\text{number of objects}}$\% to 100\%.
+    In most cases \ours{} has a better MRR compared to other methods, and all methods have a higher MRR when text modality is present.
+This shows that text modality is a rich source of information compared to speech, and that \ours{} is strongly competitive with state of the art approaches even when relying on speech for the language signal.
+    % \todocmi{Right? Correct}
+    }
+    \label{table:mrr}
+    \end{subtable}
+    \vskip\baselineskip
+    \begin{subtable}[t]{1.00\textwidth}
+    \centering
+    \resizebox{1.0\textwidth}{!}{
+    \input{tableACC}
+    }
+    \caption[]{Average and standard deviation of accuracy (Acc) on a held-out test set. Accuracy is from 0\% to 100\%.
+    Although accuracy is a strictly less forgiving metric than MRR, these results demonstrate that \ours{} still outperforms existing approaches in the majority of cases.
+When we drop the text modality, the accuracy of the contrastive method is slightly better than random, while \ours{} is about 10\% more accurate than the contrastive method.
+    }
+    \label{table:acc}
+    \end{subtable}
+    \caption{Average and standard deviation of mean reciprocal rank (MRR) and accuracy (Acc) over 5 runs with 5 different random seeds on a held-out test set with different modalities ablated during testing.
+\Cref{table:mrr} shows MRR scores and \cref{table:acc} shows accuracy.
+    % . These results represent performance when no modalities are ablated during training. 
+     Higher is better for both metrics.
+For 5 objects, a random guess would have MRR of 0.33 and accuracy of 0.5, and the worst case performance would have MRR of 0.2 and accuracy of 0.0.
+The batch size is 64 and optimizer is SGD for all experiments.
+Column headers refer to modalities that are present at query time.
+    % ---e.g., MRR \textit{sd} contains only speech and depth. 
+    Bold numbers represent the best-performing method.
+\ours{} is either better than or very close to both state-of-the-art methods for most of the cases.
+We observe the same pattern in both MRR and accuracy except the fact that accuracy scores are lower than MRR.
+    }
+    \label{table:quantitative}
+\end{table*}
 
-\begin{figure}[tbh]
+\begin{comment}
+\begin{table*}[tb]
     \centering
-\includegraphics[width=.99\columnwidth]{Figures/average-seeds-epochs-mrr_lard}
-\caption{Mean Reciprocal Rank (MRR) on the held-out test set when all modalities are available, averaged over 5 runs.
-Green is \ours{}, orange is \supcon{}, blue represents our \geom{} loss function, and red is the contrastive loss baseline.
+    \begin{subtable}[t]{1.00\textwidth}
+    \centering
+    \resizebox{.99\textwidth}{!}{
+    \csvautotabular{Figures/unused/resultsMRR.csv}
+    }
+    \end{subtable}
+    \vskip\baselineskip
+    % \hfill
+    \begin{subtable}[t]{1.00\textwidth}
+    \centering
+    \resizebox{.99\textwidth}{!}{
+    \csvautotabular{Figures/unused/resultsACC.csv}
+    }
+    \end{subtable}
+\end{table*}
+\end{comment}
+
+\begin{figure}
+    \centering
+    \begin{subfigure}[b]{0.49\textwidth}
+        \centering
+        \includegraphics[width=\textwidth]{Figures/average-seeds-epochs-mrr_lard-trimmed.pdf}
+        \caption[]{Mean Reciprocal Rank (MRR) on the held-out test set when all modalities are available.}    
+        \label{fig:epochs-mrr.lard}
+    \end{subfigure}
+    \hfill
+    \begin{subfigure}[b]{0.49\textwidth}  
+        \centering 
+        \includegraphics[width=\textwidth]{Figures/average-seeds-epochs-mrr_ard-trimmed.pdf}
+        \caption[]{Mean Reciprocal Rank (MRR) on the held-out test set when the text modality is ablated.}
+        \label{fig:epochs-mrr.srd}
+    \end{subfigure}
+    \vskip\baselineskip
+    \begin{subfigure}[b]{0.49\textwidth}   
+        \centering 
+        \includegraphics[width=\textwidth]{Figures/average-seeds-epochs-mrr_lr-trimmed.pdf}
+        \caption[]{Mean Reciprocal Rank (MRR) on the held-out test set when speech and depth modalities are ablated.}    
+        \label{fig:epochs-mrr.lr}
+    \end{subfigure}
+    \hfill
+    \begin{subfigure}[b]{0.49\textwidth}   
+        \centering 
+        \includegraphics[width=\textwidth]{Figures/average-seeds-epochs-mrr_ld-trimmed.pdf}
+        \caption[]{Mean Reciprocal Rank (MRR) on the held-out test set when speech and RGB modalities are dropped.}    
+        \label{fig:epochs-mrr.ld}
+    \end{subfigure}
+    \caption[]
+    {Mean Reciprocal Rank (MRR, \cref{eq:mrr}) on the held-out test set with the ablation of selected modalities, all averaged over 5 runs for the downstream task of object retrieval.
+Red is self-supervised contrastive learning which is prone to overfitting, orange is supervised contrastive learning, blue represents our proposed \geom{}, and green is our proposed \ours{} loss function.
 Higher is better.
+(a) shows the MRR when all modalities are available, (b) shows the MRR when text is removed at test time, (c) shows MRR when speech and depth are ablated, and (d) shows MRR when speech and RGB are removed.
+We train all models for 200 epochs.
     } 
-\label{fig:epochs-mrr.lard}
+    \label{fig:result_graphs}
 \end{figure}
+%     
+% 
+\Cref{fig:epochs-mrr.lard} Shows that \ours{} learns faster and results in a better performance compared to both \supcon{} and contrastive learning~\citep{chen2020simple} when trained using all modalities and with all modalities available during test.
+We observe that not only does contrastive loss learn more slowly, but that it is prone to overfitting; while this can be addressed with careful tuning of the learning process, an approach that is innately robust to overfitting without tuning is preferable.
 
-% Comparing \cref{fig:epochs-mrr.trd} and 
-When we drop the text modality in \cref{fig:epochs-mrr.srd}, we can see that the performance decreases from about 0.93 to about 0.82, showing that speech cannot completely replace the text.
-\Cref{fig:3d-tsne} depicts that speech modality is not aligned as good as the text modality and that is why when we drop text and use speech as the main query, the performance decreases.
+When we drop the text modality (\cref{fig:epochs-mrr.srd}), we can see that the performance decreases from about 0.93 to about 0.82, showing that speech cannot completely replace text.
+In \cref{fig:3d-tsne} (see \cref{sec:Qualitative}), the alignment of shared embeddings for a randomly sampled set of classes is visualized for all four modalities under consideration, suggesting that the speech modality is not aligned as well as the text modality.
+For this reason, when we drop text and use speech as the main query, the performance decreases.
 This supports our hypothesis that a geometric alignment of the latent space is crucial to a good performance for object retrieval and multimodal understanding.
-In \cref{fig:epochs-mrr.srd}, we observe that when speech is used as the query and text modality is ablated, \supcon{} baseline works slightly better than \ours{}, although \ours{} still learns faster.
-The reason is that \supcon{} optimizes for the classification task, and since the speech modality is not well aligned based on \cref{fig:3d-tsne}, using \geom{} makes the downstream task more difficult by trying to pull and push similar and dissimilar data points, respectively.
-More research needs to be done to come up with strategies to align more chaotic modalities.
+
+In \cref{fig:epochs-mrr.srd}, we observe that when speech is used as the query and the text modality is ablated, the \supcon{} baseline works slightly better than \ours{}, although \ours{} still learns faster.
+The reason is that \supcon{} optimizes for the classification task, and since the speech modality is less well aligned, using \geom{} makes the downstream task more difficult by trying to pull and push similar and dissimilar data points, respectively.
+Future research will consider strategies to align more chaotic modalities.
 % We also observe that speech has a higher variance compared to text, and a na\"ive contrastive learning method cannot take advantage of text modality to reduce this variance.
@@
 
-There is very little gap in performance when depth or RGB are dropped in \cref{fig:epochs-mrr.lr,fig:epochs-mrr.ld} compared to when we have all modalities in \cref{fig:epochs-mrr.lard} which shows that our model is robust enough when rgb or depth sensors fail in real-world deployment.
-Also, when depth is dropped in \cref{fig:epochs-mrr.lr} performance decreases less compared to when rgb is dropped in \cref{fig:epochs-mrr.ld} which suggests that depth is harder to work with compared to RGB.
+There is very little gap in performance when depth or RGB are dropped in \cref{fig:epochs-mrr.lr,fig:epochs-mrr.ld} compared to when we have all modalities in \cref{fig:epochs-mrr.lard}, showing that our model is robust when RGB or depth sensors fail.
+Also, when depth is dropped in \cref{fig:epochs-mrr.lr}, performance decreases less compared to when RGB is dropped in \cref{fig:epochs-mrr.ld}.
+This suggests that depth is less informative when compared to RGB, which is consistent with existing vision research results.
 
-\begin{figure}[tbh]
-\centering
-\includegraphics[width=.99\columnwidth]{Figures/average-seeds-epochs-mrr_ard.pdf}
-\caption{Mean Reciprocal Rank (MRR) on the held-out test set when the text modality is ablated, averaged over 5 runs.
-Green is self-supervised contrastive learning, orange is supervised contrastive learning, and blue represents our proposed EMMA loss function.
-Higher is better.
-}
-\label{fig:epochs-mrr.srd}
-\end{figure}
+% \todocmi{Main remaining task is to sort out figures. They should all have meaningful y axis labels and essentially no whitespace surrounding them.}
 
-\begin{comment}
-\begin{figure}[tbh]
-\centering
-\includegraphics[width=.99\columnwidth]{Figures/average-seeds-epochs-mrr_lrd.pdf}
-\caption{Mean Reciprocal Rank (MRR) on the held-out test set when the speech modality is ablated, averaged over 5 runs, for the downstream task of object retrieval when all objects are unique.
-Green is self-supervised contrastive learning, orange is supervised contrastive learning, and blue represents our proposed EMMA loss function.
-Higher is better.
-When speech as a modality is ablated, all tested methods perform similarly.
-}
-\label{fig:epochs-mrr.trd}
-\end{figure}
-\end{comment}
-
-\begin{figure}[tbh]
-\centering
-\includegraphics[width=.99\columnwidth]{Figures/average-seeds-epochs-mrr_lr.pdf}
-\caption{Mean Reciprocal Rank (MRR) on the held-out test set when speech and depth modalities are ablated, averaged over 5 runs for the downstream task of object retrieval when all objects are unique.
-Green is \ours{}, orange is \supcon{}, and blue represents our proposed \geom{} loss function.
-Higher is better.
-}
-% \todocmi{Maybe MRR is wrong and we should be looking at top-ranked result?}
-\label{fig:epochs-mrr.lr}
-\end{figure}
-
-\begin{figure}[tbh]
-\centering
-\includegraphics[width=.99\columnwidth]{Figures/average-seeds-epochs-mrr_ld.pdf}
-\caption{Mean Reciprocal Rank (MRR) on the held-out test set when speech and RGB modalities are dropped averaged over 5 runs for the downstream task of object retrieval when all objects are unique.
-Green is \ours{}, orange is \supcon{}, and blue represents our proposed \geom{} loss function.
-Higher is better.
-}
-\label{fig:epochs-mrr.ld}
-\end{figure}
-
-\begin{figure*}[tbh]
-\centering
-% \includegraphics[width=1.99\columnwidth]{Figures/3D-tsne-4M-eMMA-cosine-submodalities-text-anchor-gold-no_neg_sampling-1024-sbs.png}
-% \includegraphics[width=1\columnwidth]{Figures/3D-tsne-4M-eMMA-cosine-submodalities-text-anchor-gold-no_neg_sampling-1024-sbs.png}
-\includegraphics[width=1\columnwidth]{Figures/3D-tsne-exp-supcon-emma-lard-64-relu-SGD-0.001-unique_objects-gold-no_neg_sampling-1024-sbs.png}
-\caption{3D T-SNE~\cite{van2008tsne} projection of test embeddings of 10 randomly selected classes of objects using our proposed \ours{}.
-Each modality is separately projected into a three-dimensional space.
-In a perfect embedding, all instances of a class would be clustered in identical areas of the embedding space across all modalities.
-We can see that \ours{} successfully encourages all four modalities to live in a common manifold, allowing accurate retrieval even when modalities are missing.
-}
-\label{fig:3d-tsne}
-\end{figure*}
-
-% \subsection{Qualitative Results}
-% \label{sec:Qualitative}
-We look into the projection of the high-dimensional learned embeddings into a 
-%2-dimensional and 
-3-dimensional space using t-SNE~\cite{van2008tsne}, which is a dimensionality reduction technique to visualize high-dimensional data.
+\paragraph{Qualitative Results}
+\label{sec:Qualitative}
+In order to help visualize the performance of learned embeddings, we consider projections of a randomly selected subset of classes of the high-dimensional learned embeddings into a 3-dimensional space using t-SNE~\citep{van2008tsne}, a dimensionality reduction technique to visualize high-dimensional data.
 T-SNE creates a probability distribution over pairs of high-dimensional data where similar pairs have a higher probability and dissimilar pairs have a lower probability.
@@
 Although these projections are not perfect, combined with the quantitative results, they demonstrate that our model is learning to map instances of the same class closer to each other regardless of their modalities.
+Interestingly, toothbrush and toothpaste are mapped almost on top of each other in the text modality showing semantic and syntax similarity.
+However, in the RGB and depth modality they are close but not on top of each other since they do not look the same.
+Also, we can see that apple and lemon are mapped close to each other in all modalities which suggests that our proposed \ours{} learns some notion of the concept of fruits.
+These qualitative results show that our propose \geom{} and \ours{} have an interpretable latent space.
+
+\begin{figure*}[h]
+\centering
+% \includegraphics[width=1.99\columnwidth]{Figures/3D-tsne-4M-eMMA-cosine-submodalities-text-anchor-gold-no_neg_sampling-1024-sbs.png}
+% \includegraphics[width=1\columnwidth]{Figures/3D-tsne-4M-eMMA-cosine-submodalities-text-anchor-gold-no_neg_sampling-1024-sbs.png}
+% \includegraphics[width=.99\columnwidth]{Figures/3D-tsne-exp-supcon-emma-lard-64-relu-SGD-0.001-unique_objects-gold-no_neg_sampling-1024-sbs-trimmed.png}
+\includegraphics[width=.99\columnwidth]{Figures/3D-tsne-exp-supcon-emma-lard-64-relu-SGD-0.001-unique_objects-gold-no_neg_sampling-1024-trimmed.pdf}
+\caption{3D T-SNE~\citep{van2008tsne} projection of test embeddings of 10 randomly selected classes of objects using \ours{}.
+Each modality is separately projected into a three-dimensional space.
+Each RGB and depth image is associated with several language descriptions, leading to denser plots for text and speech.
+In a perfect embedding, all instances of a class would be clustered in identical areas of the embedding space across all modalities.
+We can see that \ours{} successfully encourages all four modalities to live in a common manifold, allowing accurate retrieval even when modalities are missing.
+}
+% \todo[inline]{get rid of markers for modalities, and change the colors to be brighter. Try combining marker and color for each class.}
+
+\label{fig:3d-tsne}
+\end{figure*}
+
+An example of the need to consider multiple modalities jointly is shown in \cref{fig:rankings}, showing how EMMA is able to correctly select an object instance from several similarly shaped and describable objects.
+
+\begin{figure}[h!]
+\centering
+\includegraphics[width=.99\columnwidth]{Figures/example-rankings-rearranged}
+\caption{\ours{} more accurately ranks objects and handles ambiguities with respect to a retrieval queries.
+Note that the phrase ``light blue'' in the query is very similar to ``light bulb'' and while \supcon{} confuses this and predicts light bulb as the correct object, \ours{} correctly identifies the can opener as the intended object, ranking the light bulb second.}
+\label{fig:rankings}
+\end{figure}
 
@@
 % \includegraphics[width=.99\columnwidth]{Figures/2D-tsne-4M-eMMA-cosine-submodalities-text-anchor-gold-no_neg_sampling-1024.pdf}
-% \caption{2D T-SNE~\cite{van2008tsne} projection of test embeddings of 10 randomly selected classes of objects using our proposed method.
+% \caption{2D T-SNE~\citep{van2008tsne} projection of test embeddings of 10 randomly selected classes of objects using our proposed method.
 Each modality is separately projected into a two-dimensional space.
@@
 Our proposed model performs well and learns fast, has been demonstrated to handle four modalities of shared information effectively, and is robust to test-time situations where information from one or more modalities is missing.
-
-However, there is still room for improvement.
+There remains room for improvement.
 Specifically, the speech modality is harder to handle.
@@
 Text seems to be the best clustered modality, and that makes sense because the variation in written text is much smaller than the other three modalities.
-Variation in speech is higher because of different accents, native language, gender, age, and etc., variation in RGB and depth is higher than text because of lightning conditions, object's texture and shape, the angle of camera, and other factors.
-
-\begin{figure}[h!]
-\centering
-\includegraphics[width=1\columnwidth]{Figures/example-rankings.pdf}
-\caption{\ours{} more accurately ranks objects and handles ambiguities with respect to a retrieval query compared to \supcon{}.
-Note that the phrase ``light blue'' in the query is very similar to ``light bulb'' and while \supcon{} confuses this and predicts light bulb as the correct object, \ours{} correctly identifies can opener as the intended object, and light bulb is ranked as the second object.
-}
-\label{fig:rankings}
-\end{figure}
-
-An example of the need to consider multiple modalities jointly is shown in \cref{fig:rankings}, showing how EMMA is able to correctly select an object instance from several similarly shaped and describable objects.
+Variation in speech is higher because there are a number of factors affecting speech understanding, including different accents, native language, gender, and age~\citep{KebeAAAI2022}.
+Variation in RGB and depth is higher than in text due to variations in lighting conditions, an object's texture and shape, the angle of the camera, and other factors.
 
@@
 Our proposed method is general enough to be applied to a variety of multimodal retrieval problems, and is not limited to purely language-based image retrieval.
-\todokdinline{Don't we need evidence for that?}
+% \todokdinline{Don't we need evidence for that?}
 
@@
 \appendix
-\section{Appendix}
 \label{sec:appendix}
+% \todocmi{should we get rid of this appendix? I am thinking so.}
+% \input{appendix}
 

### 2022-10-06 Paper/tableMRR.tex
Paper submitted to TMLR

\begin{tabular}{|p{{2cm}}|p{{2cm}}|p{{2cm}}|p{{2cm}}|p{{2cm}}|p{{2cm}}|p{{2cm}}|p{{2cm}}|p{{2cm}}|p{{2cm}}|} 
 \hline 
Methods & speech/depth & speech/RGB & text/depth & text/RGB & text/speech/ \newline depth & text/speech/ \newline RGB & speech/RGB/ \newline depth & text/RGB/ \newline depth & all \\ 
 \hline 
Geometric & 76.82$\pm$0.34 & 78.34$\pm$0.29 & 89.64$\pm$0.38 & 91.13$\pm$0.73 & 89.21$\pm$0.45 & 90.95$\pm$0.83 & 79.37$\pm$0.29 & 92.29$\pm$0.51 & 92.14$\pm$0.45 \\ 
SupCon & \textbf{78.18}$\pm$0.58 & \textbf{79.69}$\pm$0.54 & 89.04$\pm$0.88 & 90.56$\pm$0.74 & 88.75$\pm$0.66 & 90.5$\pm$0.69 & \textbf{81.2}$\pm$0.39 & 91.96$\pm$0.42 & 92.03$\pm$0.7 \\ 
EMMA & 77.63$\pm$0.29 & 78.66$\pm$0.64 & \textbf{89.87}$\pm$0.5 & \textbf{91.26}$\pm$0.86 & \textbf{89.66}$\pm$0.36 & \textbf{90.97}$\pm$0.66 & 80.32$\pm$0.45 & \textbf{92.71}$\pm$0.5 & \textbf{92.72}$\pm$0.47 \\ 
Contrastive & 71.74$\pm$0.73 & 73.37$\pm$0.39 & 89.72$\pm$0.54 & 90.82$\pm$0.37 & 89.13$\pm$0.61 & 90.26$\pm$0.58 & 74.96$\pm$0.44 & 91.92$\pm$0.41 & 91.72$\pm$0.53 \\ 
\hline \end{tabular}

### 2023-01-17 Paper/main.tex
Before converting TMLR to SIGIR format

@@
 \newcommand{\OK}[1]{{\color{green}{\textbf{DONE}}} #1}
+\newcommand{\DO}[1]{{\color{blue}{\textbf{PROG}}} #1}
 \newcommand{\ours}{\textsc{EMMA}}
@@
 
-    We propose \textit{extended multimodal alignment (EMMA)}, a generalized geometric method combined with a cross-entropy loss function that can be used to learn retrieval models that incorporate an arbitrary number of views of a particular piece of data, compounded by the challenge of retrieval when a modality becomes unavailable.
     Our study is motivated by needs in robotics and computer-human interaction, where an agent has many sensors and thus modalities with which a human may interact, both to communicate a desired goal and for the agent to recognize a desired target object.
 For such problems, there has been little research on integrating more than two modalities.
-While there have been widely popular works on self-supervised contrastive learning based on cross-entropy, there is an entirely separate family of approaches based on explicit geometric alignment.
+While there have been widely popular works based on cross-entropy, there is an entirely separate family of approaches based on explicit geometric alignment known as contrastive learning.
 However, to the best of our knowledge there has been no work on combining the two approaches for multimodal learning.
 We propose to combine both families of approaches and argue that the two are complementary.
+We introduce \textit{extended multimodal alignment (EMMA)}, a geometric (distance-based) method combined with a contrastive loss function.
+Our method can be used to learn retrieval models that incorporate an arbitrary number of views of a particular piece of data, compounded by the challenge of retrieval when a modality becomes unavailable.
 We demonstrate the usability of our model on a grounded language object retrieval scenario, where an intelligent agent has to select an object given an unconstrained language command.
@@
     %when modalities are ablated. 
+    %TODO: another baseline where we use cross-entropy loss only is missing I guess.
     The code is publicly available on GitHub and will be included for the camera-ready version (it is redacted for anonymity). 
 \end{abstract}
+
+\section{TODO}
+\begin{itemize}
+
+    \item \TD modify it for SIGIR
+    \item \TD EMMA as a contstrained optimization for Geometric portion to not satisfy it.
+    \item \TD constrained SupCon per instance modality constraints: for this particular anchor and corresponding neighbors, we want this constraint to be satisfied.
+    \item \OK Do you control for in batch negatives for supcon portion? Yes.
+    \item \TD alternate training of supcon and emma. E.g. train supcon for 50 epochs and emma for the next 50 epochs
+    \item \TD removing objects that have less than a number of instances? yes, but it's not a good story, since we want emma to work in low-data regime.
+    \item \TD ablating number of data points and have a graph based on different number of training data --> maybe EMMA degrades less than SupCon and that would be good
+    \item \TD use supcon portion to reweight the geometric values or use a method to have different scores for different pairs in the geometric loss. Targeting specific pair of embeddings.
+    \item \TD it seems to be a tension between geometric portion and the supcon loss
+    \item \OK another column for number of NAs in test and train and then correlation between them
+
+    \item \OK add another column for the number of instances and then compute correlation between that and the differential scores.
+Take the correlation between number of instances and each column (metric) and then the correlation and the average of columns.
+
+    \item listen to the speeches for NA descriptions and the ones where there is a significant difference in performance
+
+    \item \OK listen to the speeches for empty transcriptions.
+
+    \item \TD take a look at the depth images
+
+    \item going forward we have two options: Improve the model, or do a deep analysis why our results are like this
+
+    \item We could do something that addresses the "open questions" (which questions remain?)
+
+    I guess the main open question is the mixed results and the fact that they are within the std dev of each other. Basically, I think they are unsure whether being faster is a good result or not.
+    What the reader would take away?
+
+    \item \DO What the rich analysis would consist of?
+
+    \item Why are we getting what we are getting?
+
+    Visualize individual data points and what are the failure modes.
+    You can do the same for type by type (apples vs books)
+    The percent that EMMA got right and SupCon got wrong and vice versa.
+
+    have pictures, positive set and negative set,
+    and all modalities and then show scores for all of them.
+
+    \item \TD Different datasets or tasks
+    \item \TD How can we tell when modalities are complementary?
+
+    I think they meant when loss functions are complementary, am I wrong?
+
+    \item \TD How can we tell what losses to use in what cases?
+
+    \item  \TD Why didn't our approach work better, and what can we do to improve it?
+
+    One reason could be that \supcon{} considers the distance between all pairs of data points in a batch.
+
+    \item What's the next thing to try or add?
+
+    We could add a weighting term to balance the effect of \supcon{} and \geom{}.
+
+\end{itemize}
 
@@
 The ultimate aim of this work is to take arbitrary input modalities about novel objects, including both spoken and written language, and build a model that allows a system to correctly retrieve that object given a subset of those input modalities.
-In this work, we propose our new method \ours{}, a generalized geometric method combined with a cross-entropy loss function that can be used to learn retrieval models that incorporate an arbitrary number of views of a particular piece of data.
+In this work, we propose our new method called \textit{extended multimodal alignment (\ours{})}, a generalized geometric method combined with a cross-entropy based supervised contrastive loss function~\cite{NEURIPS2020_supervised_contrastive}, which we abbreviate \supcon{}.
+Our method can be used to learn retrieval models that incorporate an arbitrary number of views of a particular piece of data.
 We demonstrate steps towards this ultimate goal with a mechanism that learns quickly from a variety of input signals and is robust to ablation of certain inputs.
@@
 One family is based on classification and cross-entropy losses such as~\citet{NEURIPS2020_supervised_contrastive, chen2020simple} and the other family is geometric based such as~\citet{ICML22GeometricMultimodal,Carvalho-cooking-triplet,salvador2017cooking,triplet_loss_2021_CVPR}.
-In this paper we propose to marry the two families and show that combining geometric approach with cross-entropy method is superior.
+In this paper we propose to marry the two families and show that combining geometric approach with cross-entropy based \supcon{} loss function is superior.
 
@@
 
+\paragraph{Core Concepts}
+\label{sec:core-concepts}
+There is a shared terminology across the methods we describe in this section, but they differ in what they incorporate.
+We explain the high-level concepts here and clarify their use cases for each method in the corresponding sections.
+The three terms that are commonly used are \textit{anchor}, \textit{positive}, and \textit{negative}.
+Anchor refers to a data point that acts as a point of reference against which other data points are compared.
+Positive and negative refer to a data point or a set of data points that are similar and dissimilar to the anchor, respectively.
+% 
+For example, to learn the meaning of the concept ``book,'' the anchor might be an RGB image of a book; the corresponding text description and depth image of that book are included in the positive set; and the negative set includes the text description and RGB image of an apple.
+% 
+The methods we discuss here differ both in how they choose these three sets and the objective function they use.
+
 \subsection{Baselines}
@@
 
+As originally proposed, the \supcon{} approach which we choose as our baseline was originally applied to unimodal datasets such as ImageNet~\cite{imagenet2009}, CIFAR-10~\cite{krizhevsky2009learning}, and CIFAR-100~\cite{krizhevsky2009learning}.
+We both demonstrate that it is possible to use on multimodal datasets and extend it with the addition of geometric approaches, which gives benefits on training speed and performance when modalities are dropped.
+
 \subsection{EMMA: Extended Multimodal Alignment}
@@
 Our proposed multimodal method is composed of two complementary parts.
-The first part is a geometric loss based on distances in the latent space, and the second part is a contrastive loss based on cross-entropy.
-The geometric loss is faster to learn while the cross-entropy method is more aligned with the downstream task of object retrieval.
+The first part is a geometric loss based on distances in the latent space, and the second part is a supervised contrastive loss based on cross-entropy (\supcon{}).
+The geometric loss is faster to learn while the cross-entropy based \supcon{} loss is more aligned with the downstream task of object retrieval.
 Hence we propose to combine them.
@@
 % the anchor and positive items are from the same object class, but in different modalities and each of the positives modalities becomes anchor once. 
-We consider the following terminology:
+To make the distinction between triplet loss and our proposed \geom{} more clear, we provide an example for the terminology we use.
 % \todokdinline{I feel like using the terminology of triplet loss especially \textbf{anchor} is confusing here or maybe it needs to be rephrased.
@@
 
-\paragraph{Combining Geometric and Cross-Entropy Losses}
+\paragraph{Combining Geometric Loss and Cross-Entropy based \supcon{} Loss}
 \label{subsec:emma}
@@
 % 
+Both the geometric alignment and cross-entropy portions of EMMA are designed to not anchor on any specific modality, but instead consider all available modalities.
+This geometric portion is in contrast to previous triplet-based approaches.
+To show this, we consider how the geometric portion of EMMA’s loss would apply to the example in Figure 2.
+Considering the apple (on the left) as instance i, the dotted lines connecting the embeddings of the apple’s modalities reflect minimizing the pairwise distances between the different modalities of i: in equation 8, we do this via the $h(x, y)$ function.
+The dashed lines connecting the representations of the apple to a not-apple (the coffee mug) reflect maximizing the distances between these embeddings (done via $g(x,y)$).
+Notice how in both cases we consider all relevant pairs.
+
+Although both \supcon{} and the \geom{} approaches are trying to bring target objects together (and push dissimilar ones away), there is a trade-off in \supcon{} where there is a normalized ranking across them.
+In contrast, in \geom{} distances can be arbitrarily near or far.
+Our approach tries to respect distinct modalities of each instance.
+While our formulation is similar to \supcon{} notationally, conceptually \supcon{} uses different augmentations of RGB images as positive examples (inherently coming from the same underlying empirical distribution), while we use different modalities drawn from different distributions.
+To the best of our knowledge our work is the first to use supervised contrastive learning for such a multimodal scenario, where the different language and sensor inputs provide more robust learning than can be accomplished with single sensor inputs, even augmented.
+
 Compared to pure \supcon{}, combining these loss functions results in faster convergence and slightly improved performance when all modalities are available, and maintains improved performance when modalities are missing.
@@
 
+Our time analysis shows that EMMA takes almost 8 epochs to converge and each epoch takes roughly 0.7 minutes, which makes it 5.6 minutes until convergence, while \supcon{} takes about 36 epochs to converge and each epoch takes 0.52 minutes which amounts to 18.72 minutes.
+That is when we use all four modalities for training.
+When we ablate one or two modalities, the training takes less time.
+
 \paragraph{Qualitative Results}
@@
 Our proposed model performs well and learns fast, has been demonstrated to handle four modalities of shared information effectively, and is robust to test-time situations where information from one or more modalities is missing.
+The bottleneck for agents in different settings may differ, and training speed may not be critical in offline learning scenarios.
+However, since we usually need to finetune models for different tasks when it comes to transfer learning, the training speed becomes relevant.
+
 There remains room for improvement.

### 2023-02-08 Code/analysis.py
analysis for SIGIR

@@ -7,32 +7,32 @@ import matplotlib.pyplot as plt
 import tikzplotlib
 
 
-# exp_ugly_names = {
-#     'Geometric': 'exp-full-emma-lard-64-relu-SGD-0.001-unique_objects-gold-no_neg_sampling-1024',
-#     'SupCon': 'exp-supcon-lard-64-relu-SGD-0.001-unique_objects-gold-no_neg_sampling-1024',
-#     'EMMA': 'exp-supcon-emma-lard-64-relu-SGD-0.001-unique_objects-gold-no_neg_sampling-1024',
-#     'Contrastive': 'exp-contrastive-org-lard-64-relu-SGD-0.001-unique_objects-gold-no_neg_sampling-1024',
-# }
-
 exp_ugly_names = {
-    'SupCon-25': 'exp-0.75-train-supcon-lard-64-relu-SGD-0.001-unique_objects-gold-no_neg_sampling-1024',
-    # 'SupCon-30': 'exp-0.70-train-supcon-lard-64-relu-SGD-0.001-unique_objects-gold-no_neg_sampling-1024',
-    # 'SupCon-35': 'exp-0.65-train-supcon-lard-64-relu-SGD-0.001-unique_objects-gold-no_neg_sampling-1024',
-    # 'SupCon-45': 'exp-0.55-train-supcon-lard-64-relu-SGD-0.001-unique_objects-gold-no_neg_sampling-1024',
-    # 'SupCon-50': 'exp-0.5-train-supcon-lard-64-relu-SGD-0.001-unique_objects-gold-no_neg_sampling-1024',
-    # 'SupCon-75': 'exp-0.25-train-supcon-lard-64-relu-SGD-0.001-unique_objects-gold-no_neg_sampling-1024',
-    # 'SupCon-100': 'exp-supcon-lard-64-relu-SGD-0.001-unique_objects-gold-no_neg_sampling-1024',
-
     'EMMA-25': 'exp-0.75-train-supcon-emma-lard-64-relu-SGD-0.001-unique_objects-gold-no_neg_sampling-1024',
-    # 'EMMA-30': 'exp-0.70-train-supcon-emma-lard-64-relu-SGD-0.001-unique_objects-gold-no_neg_sampling-1024',
-    # 'EMMA-35': 'exp-0.65-train-supcon-emma-lard-64-relu-SGD-0.001-unique_objects-gold-no_neg_sampling-1024',
[…]

### 2023-02-08 Code/explicit-latex-table.py

@@ -9,10 +9,31 @@ import tikzplotlib
 
 
 exp_ugly_names = {
-    'Geometric': 'exp-full-emma-lard-64-relu-SGD-0.001-unique_objects-gold-no_neg_sampling-1024',
-    'SupCon': 'exp-supcon-lard-64-relu-SGD-0.001-unique_objects-gold-no_neg_sampling-1024',
-    'EMMA': 'exp-supcon-emma-lard-64-relu-SGD-0.001-unique_objects-gold-no_neg_sampling-1024',
-    'Contrastive': 'exp-contrastive-org-lard-64-relu-SGD-0.001-unique_objects-gold-no_neg_sampling-1024',
+    # 'EMMA-25': 'exp-0.75-train-supcon-emma-lard-64-relu-SGD-0.001-unique_objects-gold-no_neg_sampling-1024',
+    'EMMA-30': 'exp-0.70-train-supcon-emma-lard-64-relu-SGD-0.001-unique_objects-gold-no_neg_sampling-1024',
+    # 'EMMA-35': 'exp-0.65-train-supcon-emma-lard-64-relu-SGD-0.001-unique_objects-gold-no_neg_sampling-1024',
+    # 'EMMA-45': 'exp-0.55-train-supcon-emma-lard-64-relu-SGD-0.001-unique_objects-gold-no_neg_sampling-1024',
+    # 'EMMA-50': 'exp-0.5-train-supcon-emma-lard-64-relu-SGD-0.001-unique_objects-gold-no_neg_sampling-1024',
+    # 'EMMA-75': 'exp-0.25-train-supcon-emma-lard-64-relu-SGD-0.001-unique_objects-gold-no_neg_sampling-1024',
+    # 'EMMA-100': 'exp-supcon-emma-lard-64-relu-SGD-0.001-unique_objects-gold-no_neg_sampling-1024',
+
+    # 'SupCon-25': 'exp-0.75-train-supcon-lard-64-relu-SGD-0.001-unique_objects-gold-no_neg_sampling-1024',
+    'SupCon-30': 'exp-0.70-train-supcon-lard-64-relu-SGD-0.001-unique_objects-gold-no_neg_sampling-1024',
+    # 'SupCon-35': 'exp-0.65-train-supcon-lard-64-relu-SGD-0.001-unique_objects-gold-no_neg_sampling-1024',
+    # 'SupCon-45': 'exp-0.55-train-supcon-lard-64-relu-SGD-0.001-unique_objects-gold-no_neg_sampling-1024',
[…]

### 2023-02-08 Paper/Inbox.tex
Paper submitted to SIGIR 23

--- 2022-10-06 Paper/main.tex
@@
-\documentclass[10pt]{article} % For LaTeX2e
-\usepackage{tmlr}
-%%\usepackage{showframe}
-% If accepted, instead use the following line for the camera-ready submission:
-% \usepackage[accepted]{tmlr}
-% To de-anonymize and remove mentions to TMLR (for example for posting to preprint servers), instead use the following:
-% \usepackage[preprint]{tmlr}
+\section{TODO}
+\begin{itemize}
 
-% Optional math commands from https://github.com/goodfeli/dlbook_notation.
-\input{math_commands.tex}
+    \item \OK what to cut?
+    I think we can cut table 1.b since we don't really need to report accuracy when we have MRR, and ACC is always lower than MRR.
+Maybe we can report acc in the body of the paper with 1 or 2 sentences.
+    \item \TD look at training performance for SupCon 0.75
+    \item \TD make sure intro is SIGIR focused by comparing it with previous SIGIR
+    \item \OK run Geom with dropped training set 
+    \item \OK graph of converged performance (5 data points? more datapoints from 0.5 and 0.75: 0.55 and 0.65)
+    \item \TD You can also compute AUC of performance vs. epoch. You can also compute AUC of performance vs. epoch. Nisha's AL paper as an ex.
+    \item \OK modify it for SIGIR
+    \item \TD EMMA as a contstrained optimization for Geometric portion to not satisfy it.
+    \item \TD constrained SupCon per instance modality constraints: for this particular anchor and corresponding neighbors, we want this constraint to be satisfied.
+    \item \OK Do you control for in batch negatives for supcon portion? Yes.
+    \item \TD alternate training of supcon and emma. E.g. train supcon for 50 epochs and emma for the next 50 epochs
+    \item \OK removing objects that have less than a number of instances?
+yes, but it's not a good story, since we want emma to work in low-data regime.
+I removed items that has more than a number of NA descriptions and/or less than a number of instances, and the results are not that different.
+    \item \DO ablating number of data points and have a graph based on different number of training data --> maybe EMMA degrades less than SupCon and that would be good
+    \item \TD use supcon portion to reweight the geometric values or use a method to have different scores for different pairs in the geometric loss. Targeting specific pair of embeddings.
+    \item \TD it seems to be a tension between geometric portion and the supcon loss
+    \item \OK another column for number of NAs in test and train and then correlation between them
 
-\usepackage{hyperref}
-\usepackage{url}
+    \item \OK add another column for the number of instances and then compute correlation between that and the differential scores.
+Take the correlation between number of instances and each column (metric) and then the correlation and the average of columns.
 
-% My Packages
-\usepackage{cleveref}
-\usepackage{comment}
-\usepackage{csvsimple}
-\usepackage{datatool}
-\usepackage{todonotes}[disable]
-\usepackage{graphicx}
-\usepackage{subcaption}
-\newcommand{\todokdinline}[1]{\todo[color=red!20,inline]{{KD: \small #1}}}
-\newcommand{\todokd}[1]{\todo[color=red!20]{{\small #1 -- Kasra}}}
-% \newcommand{\todocainline}[1]{\todo[color=yellow!20,inline]{{CA: \small #1}}}
-\newcommand{\todocm}[1]{\todo[color=green!40]{\small #1 -- Cynthia}}
-\newcommand{\todocmi}[1]{\todo[inline,color=green!40]{\small #1 -- Cynthia}}
-\newcommand{\todoff}[1]{\todo[color=blue!20]{\small #1 -- Frank}}
-\newcommand{\todoffinline}[1]{\todo[inline,color=blue!20]{\small #1 -- Frank}}
-\newcommand{\todoed}[1]{\todo[color=cyan!20]{\small #1 -- Ed}}
+    \item listen to the speeches for NA descriptions and the ones where there is a significant difference in performance
 
-\newcommand{\CitationNeeded}[1]{{\textbf {\color{red}Cite #1}}}
-\newcommand{\ProofRead}[1]{{\textbf {\color{red}Proofread please #1}}}
-\newcommand{\Complete}[0]{{\textbf {\color{red}Complete it }}}
-\newcommand{\Rephrase}[1]{{\textbf {\color{red}Rephrase please #1}}}
-\newcommand{\TD}[1]{{\color{red}{\textbf{TODO}}} #1}
-\newcommand{\OK}[1]{{\color{green}{\textbf{DONE}}} #1}
-\newcommand{\ours}{\textsc{EMMA}}
-% \newcommand{\ours}{\textsc{EMMA}}
-\newcommand{\geom}{\textsc{Geometric Alignment}}
-\newcommand{\supcon}{\textsc{SupCon}}
+    \item \OK listen to the speeches for empty transcriptions.
 
-%%%%%%%%%%%%%%%%%%%%
-% EDITING LINK: 
-% https://www.overleaf.com/3575751412sjsdmszkxsyh
-%%%%%%%%%%%%%%%%%%%%
+    \item \TD take a look at the depth images
 
-\title{Multimodal Language Learning in the Face of Missing Modalities}
-% \\ Extended Multimodal Alignment \\ Combining Geometric and classification ... }
+    \item going forward we have two options: Improve the model, or do a deep analysis why our results are like this
 
-% Authors must not appear in the submitted version. They should be hidden
-% as long as the tmlr package is used without the [accepted] or [preprint] options.
-% Non-anonymous submissions will be rejected without review.
+    \item We could do something that addresses the "open questions" (which questions remain?)
 
-\author{\name Kasra Darvish \email kasradarvish@umbc.edu \\
-      \addr Department of Computer Science\\
-      University of Maryland Baltimore County
-      \AND
-      \name Edward Raff \email Raff\_Edward@bah.com \\
-      \addr University of Maryland Baltimore County \\
-      Booz Allen Hamilton
-      \AND
-      \name Francis Ferraro \email ferraro@umbc.edu \\
-      \addr University of Maryland Baltimore County
-      \AND
-      \name Cynthia Matuszek \email cmat@umbc.edu\\
-      \addr University of Maryland Baltimore County \\
-      CIFAR Fellow}
+    I guess the main open question is the mixed results and the fact that they are within the std dev of each other. Basically, I think they are unsure whether being faster is a good result or not.
+    What the reader would take away?
 
-% The \author macro works with any number of authors. Use \AND 
-% to separate the names and addresses of multiple authors.
+    \item \DO What the rich analysis would consist of?
 
-\newcommand{\fix}{\marginpar{FIX}}
-\newcommand{\new}{\marginpar{NEW}}
+    \item Why are we getting what we are getting?
 
-\def\month{MM}  % Insert correct month for camera-ready version
-\def\year{YYYY} % Insert correct year for camera-ready version
-\def\openreview{\url{https://openreview.net/forum?id=XXXX}} % Insert correct link to OpenReview for camera-ready version
+    Visualize individual data points and what are the failure modes.
+    You can do the same for type by type (apples vs books)
+    The percent that EMMA got right and SupCon got wrong and vice versa.
 
-\begin{document}
+    have pictures, positive set and negative set,
+    and all modalities and then show scores for all of them.
 
-\maketitle
+    \item \TD Different datasets or tasks
+    \item \TD How can we tell when modalities are complementary?
 
-\begin{abstract}
-    % Grounded language understanding, in which natural language is used as a query against objects in a physical environment, allows a real-world, intuitive mechanism by which users can instruct physical agents to engage in tasks such as object retrieval.
-Visuolinguistic approaches to such object inference tasks typically involve training on large pools of image/text pairs and then using language to subselect elements of the sensed environment.
-However, physical agents such as robots typically have access to sensory and interactive modalities beyond vision, and learning from multiple modalities can improve performance on downstream tasks.
-In order to fully leverage multimodal training data while being robust to missing information, we propose a generalized distance-based loss function that can be extended to learn retrieval models that incorporate an arbitrary number of modalities.
-We demonstrate the usability of our model on a grounded language object retrieval scenario, where an intelligent agent has to select an object given an unconstrained language command.
-We leverage four modalities including vision, depth sensing, text, and speech, and we show that our model can outperform state-of-the-art contrastive models when modalities are ablated.
+    I think they meant when loss functions are complementary, am I wrong?
 
-    % shorter version
+    \item \TD How can we tell what losses to use in what cases?
 
-    We propose \textit{extended multimodal alignment (EMMA)}, a generalized geometric method combined with a cross-entropy loss function that can be used to learn retrieval models that incorporate an arbitrary number of views of a particular piece of data, compounded by the challenge of retrieval when a modality becomes unavailable.
-Our study is motivated by needs in robotics and computer-human interaction, where an agent has many sensors and thus modalities with which a human may interact, both to communicate a desired goal and for the agent to recognize a desired target object.
-For such problems, there has been little research on integrating more than two modalities.
-While there have been widely popular works on self-supervised contrastive learning based on cross-entropy, there is an entirely separate family of approaches based on explicit geometric alignment.
-However, to the best of our knowledge there has been no work on combining the two approaches for multimodal learning.
-We propose to combine both families of approaches and argue that the two are complementary.
-We demonstrate the usability of our model on a grounded language object retrieval scenario, where an intelligent agent has to select an object given an unconstrained language command.
-We leverage four modalities including vision, depth sensing, text, and speech, and we show that our model converges approximately 5 times faster than previous strong baselines, and out-performs or is strongly competitive with state-of-the-art contrastive learning.
-% (in 80\% less time).
-    %when modalities are ablated. 
-    The code is publicly available on GitHub and will be included for the camera-ready version (it is redacted for anonymity). 
-\end{abstract}
+    \item  \TD Why didn't our approach work better, and what can we do to improve it?
 
-\section{Introduction}
-\label{intro}
+    One reason could be that \supcon{} considers the distance between all pairs of data points in a batch.
 
-Inspired by the multimodal nature of human interaction with the world, it is intuitive that agents learning about the world, upon encountering new concepts and new objects, should form a model that incorporates information from all available sensors and data sources.
-The benefits of integrating multiple modalities are twofold; first, complementary information can be extracted from different modalities that can help with understanding the world, and second, additional modalities can help in the cases when one or more sources of data about the world become unavailable.
-% 
-Grounded language understanding, in which natural language is used as a query against objects in a physical environment, allows a real-world, intuitive mechanism by which users can instruct physical agents to engage in tasks such as object retrieval.
-Visuolinguistic approaches to such object inference tasks typically involve training on large pools of image/text pairs and then using language to subselect elements of the sensed environment~\citep{hong2021gilbert,Zhuge_2021_CVPR_VLP}.
+    \item What's the next thing to try or add?
 
-Although physical agents such as robots typically have access to sensory and interactive modalities beyond vision, and learning from multiple modalities can improve performance on downstream tasks, most approaches use at most two sensory inputs (e.g., visual data such as RGB plus depth images) with single labels, such as those provided by textual natural language.
-Simultaneously using additional inputs from different modalities is an underexplored area, in part due to the domain-specific nature of such \textit{n}-ary learning approaches.
-With the modern proliferation of audio and text based communication and home agents (e.g., Alexa/Google Home), there is a growing need to handle more modalities, and simultaneously their potential failures.
+    We could add a weighting term to balance the effect of \supcon{} and \geom{}.
 
-One difficulty with working with complex multimodal data is the increased likelihood that one or more modalities may have missing information.
-Hardware can become damaged or defective, sensors can get blocked or obstructed, and various adverse but not uncommon conditions can remove a modality from use.
-Current multimodal approaches are typically not robust to the loss of one or more modalities at test time, as may happen if, for example, a physical agent fails to retrieve data from a particular sensor.
-In order to fully leverage multimodal training data while being robust to missing information, we propose a generalized distance-based loss function that can be extended to learn retrieval models that incorporate an arbitrary number of modalities.
-
-We consider the domain of grounded language-based object retrieval~\citep{hu2016natural, triplet_loss_2021_CVPR}, in which objects in an environment must be identified based on linguistic instructions.
-This can be considered a special case of image retrieval~\citep{huang2017deep, ma2020large, novak2015large, vo2019composing} in which objects are identified using visual inputs in combination with other sensor modalities.
-Approaches to acquiring grounded language have explored various combinations of sensor inputs such as depth and RGB with labels provided by textual language or  speech~\citep{RichardsDarvishMatuszekCategoryFree20}.
-However, despite the multisensory nature of object retrieval, much of the existing work has not previously been extended to include an arbitrary number of modalities.
-
-% \todo[inline]{Whats the name of our method?
-When I skim the paper its hard to tell what ``the new thing is'', and then I have to read.
-I think being explicit here, with something like ``\textit{we propose our new method BLA}'' will help -Ed}
-
-The ultimate aim of this work is to take arbitrary input modalities about novel objects, including both spoken and written language, and build a model that allows a system to correctly retrieve that object given a subset of those input modalities.
-In this work, we propose our new method \ours{}, a generalized geometric method combined with a cross-entropy loss function that can be used to learn retrieval models that incorporate an arbitrary number of views of a particular piece of data.
-We demonstrate steps towards this ultimate goal with a mechanism that learns quickly from a variety of input signals and is robust to ablation of certain inputs.
-This is a generalization of approaches to grounded language learning in which specific input modalities are labeled with language to allow for future identification, but explicitly seeks to be agnostic about the nature of the individual sensory and linguistic inputs.
-
-Our contributions are as follows:
-\begin{enumerate}
-    \item Proposing a new approach to multimodal learning by combining geometric and cross-entropy methods and showing that it outperforms and converges faster compared to each method separately.
-    % \item Outperforming state-of-the-art models for grounded language learning by differentiating visual input modalities such as RGB and depth, rather than combining them into a single modality.
-
-    \item Our method is designed to be applicable to additional modalities as they become available.
-
-    \item Our method converges approximately 5 times faster than previous strong baselines, and out-performs or is strongly competitive with state-of-the-art contrastive learning~\citep{chen2020simple} and supervised contrastive learning~\citep{NEURIPS2020_supervised_contrastive} models when all modalities are available.
-
-    \item Demonstrating robustness in the face of missing input signals when one or more modalities are ablated at test time, and proposing a simple averaging method to quantify performance for multimodal object retrieval scenarios.
-
-    \item Demonstrating the separate utility of speech and text as sources of label information by treating them as sensory input modalities, rather than explicitly as labels.
-
-    % \item Stable model that outperforms state-of-the-art contrastive learning models when batch size decreases.\todoff{add motivation in the intro about why robustness to a smaller batch size is important}
-
-\end{enumerate}
-
-% \todoffinline{Super pet peeve of mine: except for books or theses, don't take up space with this roadmap.
-Instead, in the above contributions, add pointers to the sections (or even subsections) where we provide evidence for each contribution}
-% \todokdinline{I think Cynthia added this part for the last submission since we were short a few sentences to reach the page limit. We can remove this.}
-% The remainder of this paper is organized as follows.
-In \cref{sec:Related-Work}, we describe existing work on visuolinguistic retrieval, contrastive learning, and multimodal learning.
-In \cref{sec:Problem-Description}, we formalize the object retrieval problem and give examples of the data under consideration.
-In \cref{sec:Method}, we describe our novel approach to learning models of linguistically-driven queries against a multimodal set of objects.
-Finally, in \cref{sec:Experiments}, we demonstrate the effectiveness of our approach against traditional contrastive learning and supervised contrastive learning baselines, both for the general learning problem and in the case of missing information at test time.
-
-% \todocmi{NOTES: Some things are probably less informative or more ambiguous (e.g., depth for differentiating round things).
-So instead of predicting a single vector output, predict a vector and a covariance that captures uncertainty; could potentially use that to do more accurate search, or to handle really uncertain modalities (e.g., just drop speech when it's uninformative).}
-
-%===================================================================
-
-\section{Related Work}
-\label{sec:Related-Work}
-
-% \todoffinline{I'm not convinced this section shouldn't appear toward the end of the paper. Maybe adding that narrative structure will help?}
-
-% \todocmi{I'm not sure what narrative structure is being referred to here. I have a preference for having related work sections towards the beginning.}
-
-There is extensive work on the task of image retrieval, of which physical object retrieval can be considered a special case.
-In this task, language is used to formulate queries against datasets of images, for example in text-and-image matching tasks for fashion data~\citep{gao2020fashionbert, wen2021comprehensive}, sketch retrieval~\citep{huang2017deep}, and general photographs of objects~\citep{ma2020large, novak2015large, hong2021gilbert}.
-Prior work has focused solely on language and visual modalities, using text or a combination of language and vision to perform visual retrieval~\citep{vo2019composing}.
-While recent works have focused on grounding with respect to a knowledge set \citep{10.1145/3357384.3357889,10.1145/3397271.3401097}, our work focuses on robust multimodal learning to perform grounding over an arbitrary number of input modalities.
-
-The growing number of datasets that contain modalities beyond images and text demonstrates the importance and the challenge of this task.
-\citet{bagher-zadeh-etal-2018-multimodal} introduce a large dataset of multimodal opinion sentiment and emotion intensity (CMU-MOSEI) for sentiment analysis and emotion recognition.
-They also introduce a novel multimodal fusion technique called the Dynamic Fusion Graph (DFG).~\citet{GoLD_UMBC} present a multimodal dataset of household objects containing RGB images, depth images, written language, and spoken language, which has been used to support learning grounded language directly from speech given small datasets~\citep{KebeAAAI2022}.
-\citet{baltrusaitisMultimodalMachineLearning2019} propose a new taxonomy of multimodal machine learning by introducing five technical challenges in addition to typical early and late fusion categorization.
-They cover different approaches to multimodal machine learning such as \textit{representation}, \textit{translation}, \textit{alignment}, \textit{fusion}, and \textit{co-learning}.
-In this work, we focus on alignment, the process of identifying the direct relationship between elements across modalities.
-
-We highlight a key difference with recent work like \citet{10.1145/3397271.3401232} and \citet{10.1145/3331184.3331213} who develop multi-modal retrieval models.
-These retrieval models are class based, where the model is trained to recognize any object of the same retrieval class as being the `same.' Our method and data is instance based, where different objects are not considered the same even if they have the same class, due to the grounding goals of our work (i.e., the agent should identify the specified object, not equivalent objects).
-These works also do not consider performance in the case of failure of an input modality.
-
-A number of different approaches have been proposed for linguistic-based alignment learning.
-\citet{alayrac2020self} use a self-supervised contrastive learning method to learn and combine representations from three modalities of visual, audio, and language.
-Audio and visual domains are mapped to the same space because they are fine-grained, containing dense information for each frame of video, and these embeddings are then mapped to a coarse space where text is also mapped.
-Their network respects specificity and comparability of different modalities.
-The loss function they define consists of two distance-based terms for the corresponding spaces, differing from ours in that they do not handle arbitrarily many modalities and do not focus on robustness in the face of modality drop-outs.
-\citet{Nguyen-RSS-20} take a similar approach and perform pairwise cosine similarity to align images and natural language text descriptions to perform object retrieval tasks.
-\citet{triplet_loss_2021_CVPR} take a cross-modal manifold alignment approach to grounded language learning using triplet loss \citep{Chechik:2010:LSO:1756006.1756042}, and perform object retrieval by connecting text descriptions of objects to their corresponding RGB-D images.
-
-In general there are two different families of approaches to multimodal learning in the literature.
-One family is based on classification and cross-entropy losses such as~\citet{NEURIPS2020_supervised_contrastive, chen2020simple} and the other family is geometric based such as~\citet{ICML22GeometricMultimodal,Carvalho-cooking-triplet,salvador2017cooking,triplet_loss_2021_CVPR}.
-In this paper we propose to marry the two families and show that combining geometric approach with cross-entropy method is superior.
-
-% \todokdinline{Should we move the following two paragraphs to the Approach section?}
-% \todocmi{Actually yes, although I think it's more like 3 paragraphs. Rewriting the subsequent paragraphs so they don't appear out of nowhere would help.}
-% \todokdinline{After moving baselines to the approach section and rewriting it, I think it's better if we leave these here. Let me know if you think otherwise.}
-
-% For the classification part of our loss function we employ the supervised contrastive loss proposed by~\citet{NEURIPS2020_supervised_contrastive}. 
-The way we formulate our geometric loss function is similar to lifted structured loss~\citep{songCVPR16LiftedStructured}, where they take a geometric approach in a unimodal scenario and consider pairwise distances among all items in a batch, but in this paper we apply our idea to four modalities, and our formulation can be extended to any number of modalities.
-This extended geometric formulation results in a loss function that is similar to \citet{ICML22GeometricMultimodal} where they propose a geometric contrastive learning approach in which they %use the same instance of an MLP network (same weights) after extracting embeddings from each modality, while we use different instances of the same MLP network (different weights).
-They also
-fuse all embeddings as a central embedding and then minimize the distance between each embedding and that central embedding.
-However, their method does not take into account any notion of classification.
-
-The geometric approach is also known as contrastive learning~\citep{qin2021world}, which is based on the idea of similarity and distance of elements of training data~\citep{Carvalho-cooking-triplet,triplet_loss_2021_CVPR,salvador2017cooking}.
-Triplet loss is also a special case of contrastive loss~\citep{NEURIPS2020_supervised_contrastive}.
-Contrastive loss is mostly used in self-supervised learning~\citep{bui2021self, alayrac2020self,chen2020simple}, with some exceptions such as \citet{NEURIPS2020_supervised_contrastive}, who use contrastive loss for a supervised problem.
-Most of these methods are implemented for a non-multimodal dataset (e.g., RGB images only).
-Our work is a specialization of contrastive loss that is focused on the multimodal input and language problem.
-
-We note in particular that standard triplet learning approaches often require an expensive \textit{mining} step to find harder negative samples to enable effective learning \citep{Hoffer2015,Schroff2015,DBLP:conf/eccv/2018-9,DBLP:conf/eccv/ZhaoJQLH18,Zhai2018}.
-This is problematic for throughput and scalability, as run time becomes quadratic in batch size.
-Our approach does not require mining to obtain good performance, alleviating these issues along with their complicated history in reproducibility and parameter tuning \citep{Musgrave2020,Raff2020c,Raff2019_quantify_repro}.
-
-Some of the most closely related work focuses on learning based on more than two input modalities.
-\citet{Veit_2018_CVPR} train a joint model of images, hashtags, and users to perform image retrieval.
-Their task is similar to ours---given a description, we want to find the image/object in the scene that best matches the description.
-They tackle this problem by forming a three-way tensor product model.
-They use a ranking loss to train the model where the score of an observed triplet is higher than an unobserved triplet.
-They sample six negative triplets per positive sample triplet, and use each of them as a negative in the loss.
-The downstream retrieval task is then simply done by taking the argmax of the tensor product for a given user.
-This work differs from ours in that it becomes computationally complex as more and higher-dimensional modalities are added.
-Our proposed geometric loss function resembles work on quadruplet loss~\citep{chen2017beyond,tursun2021efficient} but is intended to scale to an arbitrary number of modalities.
-
-To the best of our knowledge, there has been limited work on incorporating more than three modalities into a general learning approach.
-\citet{semihet_three_way_Lei_2020} use inputs from three modalities of image, sketch, and edgemap to perform image retrieval given the sketch.
-They use a co-attention mechanism, and their loss function contains alignment loss, cross-entropy loss, and sketch-edgemap contrastive loss.
-\citet{Mittal2020M3ER} use Canonical Correlational Analysis (CCA) to differentiate between ineffective and effective modalities for the emotion recognition task from multiple input modalities including face, speech, and text.
-Multimodal learning has been used for different applications including but not limited to object retrieval.
-\citet{Deception_ICMI_2014} use language, physiological response, and thermal sensing to detect deceit.
-\citet{het_data_fusion_liu_IEEE_2017} take two modalities as input and output predictions in a third modality.
-\citet{tursun2021efficient} uses quadruplet loss for two modalities only; image and sketch.
-A quadruplet is composed of a sketch picture as an anchor, a negative example from sketch domain, a negative example from picture domain, and a positive example from picture domain.
-\citet{chen2017beyond} also uses quadruplet loss, but does not support multiple domains.
-In contrast, our method is not restricted to two modalities.
-% \todo[inline]{talk about CLIP and DALL-E here}
-
-%===================================================================
-
-\section{Problem Description}
-\label{sec:Problem-Description}
-
-Given a language command (either text or speech) that describes an object, we want our model to retrieve the correct object from a set of objects.
-This problem is an exemplar of tasks found in the area of grounded language learning in the fields of robotics and natural  language processing.
-Intuitively, the goal is to take unconstrained natural language queries and select the appropriate object based on the complete set of sensor inputs available to the agent.
-We demonstrate on a domain containing four modalities, all of which refer to objects in the environment: spoken language, written text, RGB (image) inputs, and depth camera inputs.
-\Cref{fig:experimental-setup} illustrates a small visualization of our object retrieval task: the spoken query ``A white textbook titled algorithms'' is provided to our contrastive model, which identifies the item
-(outlined in red in \cref{fig:experimental-setup}) 
-as the most likely item the query is referring to.
-
-\begin{figure}[tb]
-\centering
-\includegraphics[width=.75\columnwidth]{Figures/experiment-setup-rearranged.png}
-\caption{
-The experimental object retrieval setup, in which objects are represented by four modalities: an RGB image, a depth image, spoken descriptions, and textual descriptions.
-Given a query in some modality, our approach seeks to select the object that is the best fit, per a trained model.
-This approach, detailed in \cref{sec:Method}, is able to rank objects as to appropriateness even when one or more modalities is ablated at test time (e.g., depth inputs are missing), and outperforms state-of-the-art contrastive learning approaches.
-}
-\label{fig:experimental-setup}
-\end{figure}
-
-More formally, given a spoken language command $x_s$, a textual language command $x_t$, a set of RGB images $X_r = \{x_r^{(1..n)}\}$, and a set of depth images $X_d = \{x_d^{(1..n)}\}$, the task is to retrieve the correct object by choosing the index that has the minimum distance from either of the language commands across all modalities.
-Depending on which modalities are or are not ablated, we consider up to four distances: \textit{sr}, a vector of distances between $x_s$ and all RGB images in $X_r$; \textit{sd}, a vector of distances between $x_s$ and all depth images in $X_d$; \textit{tr}, a vector of distances between $x_t$ and all RGB images in $X_r$; and \textit{td}, a vector of distances between $x_t$ and all depth images in $X_d$.
-In order to select the correct object, we first perform a component-wise average of the relevant modality pair distances for the available modalities.
-Then, we select the object which had the minimum distance, i.e., we perform an argmin on this average vector of multiple-modality distances.
-Depending on the missing/available sensors during test time, we might have any combination of these four distances.
-For example, if no written instructions are available at test time,\footnote{This setting is particularly salient.
-While large bodies of text are frequently available at training time, a person interacting directly with a physical agent may well prefer to use only spoken instructions.} we compute the component-wise average of $sr$ and $sd$, and then select the object whose coordinate resulted in the lowest average distance.
-This method allows us to extend our model to support arbitrary modalities while remaining robust to test cases in which some modalities are missing or incomplete.
-
-%===================================================================
-
-\section{Approach}
-\label{sec:Method}
-
-In keeping with previous work on the closely related problem of image retrieval, we focus on contrastive loss approaches, in which the goal is to learn an embedding of data where similar samples---in our case, samples belonging to the same class of object---are embedded `close' to one another, while dissimilar samples are farther apart.
-We develop a novel geometric loss function, \geom{}, that simultaneously minimizes intra-class distances and maximizes inter-class distances across each pair of modalities, yielding a model that is both effective at the retrieval task defined above and robust to modality dropouts at test time.
-We take an additional step and combine this \geom{} loss with a classification-based (cross-entropy) loss function which results in a superior model compared to either of geometric or cross-entropy losses alone; we refer to this combination as Extended Multi-Modal Alignment, or \ours{}.
-
-\subsection{Baselines}
-We compare both \ours{} and \geom{} against contrastive learning~\citep{chen2020simple} and supervised contrastive learning~\citep{NEURIPS2020_supervised_contrastive}, which for conciseness we refer to as \supcon{}.
-We consider \supcon{} as the main baseline since it is a general version of multiple contrastive loss functions including triplet loss, the traditional version of contrastive loss usually used in self-supervised settings~\citep{chen2020simple}, and N-pair loss~\citep{NIPS2016N-PairLoss}.
-% and do not include N-pair and triplet loss in our comparisons.
-%contrastive learning methods including triplet loss, the traditional version of contrastive loss usually used in self-supervised settings~\citep{chen2020simple}, and supervised contrastive learning~\citep{NEURIPS2020_supervised_contrastive}.
-
-\begin{comment}
-\subsubsection{Triplet Loss}
-\label{sub:baseline-triplet}
-Triplet loss has two major disadvantages. First, it cannot be used for more than two modalities.
-Second, a batch size of greater than 1 is hard to implement if the anchor, positive, and negative come from different modalities for each batch.
-% To address these issues in EMMA, we sample more than one positive and negative data points.
-%Intuitively, we sample one positive data point and one negative data point, each of which includes $M$ modalities, resulting in 2M data points.
-Each of the $M$ positives modalities become an anchor once.
-% \Cref{table:quantitative} shows that our MMA method perform as good as this baseline while our model can be trained 10x faster since we can use batch sizes larger than 1. Also,
-
-% When we used a fixed anchor modality, a batch size of 64, and applied semantic negative sampling~\citep{Pillai_Matuszek_2018}, our method outperformed this baseline. However, 
-% Since this method cannot be applied to four modalities, we do not include it in the main analysis. 
-\end{comment}
-
-\subsubsection{Contrastive Loss}
-\label{sub:baseline-contrastive}
-
-We compare our model against the contrastive learning method presented by~\citet{chen2020simple} where they use the normalized
-temperature-scaled cross entropy loss (NT-Xent).
-In order to implement this loss function, we use cosine similarity as suggested in the SimCLR paper~\citep{chen2020simple}.
-Another possibility is to use an inner dot product~\citep{NEURIPS2020_supervised_contrastive}; if not normalized, this can lead to numerical instabilities and overflow/underflow since the dot product is not bounded, but the result is the same whether we use normalized inner dot product or cosine similarity.
-Contrastive loss function is formulated in \cref{eq:contrastive-loss}:
-% \todokdinline{try contrastive loss with cross-entropy, dot product, and the supervised contrastive method itself (without triplets).}
-
-% Original Formula
-\begin{equation}\label{eq:contrastive-loss}
-    -\sum_{i \in I} \log \frac{\exp (sim(z_i , z_{j(i)}) / \tau) }{\sum_{a \in A(i)} \exp (sim(z_i, z_a) / \tau)}
-\end{equation}
-% \begin{equation}\label{eq:contrastive-loss}
-%     -\sum_{i \in I} \log \frac{\exp (sim(f(x_i) , z_{j(i)}) / \tau) }{\sum_{a \in A(i)} \exp (sim(f(x_i), f(x_a)) / \tau)}
-% \end{equation}
-where $i$ is the index of anchor, $j(i)$ is the index of positive item with respect to the anchor $z_i$ and is not the same as anchor, $A(i)$ is the set of all negatives and the one positive indices excluding anchor, and $z = f(x)$.
-
-We can treat different modalities of the same instance as additional input for that instance that augment the available information, and consider them as the positive points for the anchor.
-\Cref{eq:contrastive-loss} can be rewritten with the sum over more than one positive item as formulated in \cref{eq:explicit-contrastive}:
-
-%, which is equivalent to ``augmented views'' in the contrastive loss terminology.
-
-% \todo[inline]{Here we use $\sum_{p=1}^{P(i)}$, but later use $\sum_{p \in P(i)}$, we should be consistent.
-IMO I'd go with the latter, as its more consisnt with all other equations and $\sum_{p=1}^{P(i)}$ is a littel unusual looking -Ed I corrected it.
--Kasra}
-\begin{equation}\label{eq:explicit-contrastive}
-    - \sum_{i \in I} \sum_{p \in P(i)} \log \frac{\exp (sim(z_i , z_{p})/ \tau) }{
-    \sum_{a \in A(i)} \exp (sim(z_i, z_a) / \tau)}
-    %{ \exp (sim(z_i^+ , z_{m}^{+}) / \tau) + \exp (sim(z_i^+, z_{m}^{-}) / \tau)}
-\end{equation}
-where $I$ is a batch consisting of one or more instances each with a set of all its modalities, and $P(i)$ is the set of modalities/augmentations of the anchor $i$ excluding itself (e.g.
-RGB image, depth image, speech, text) and $z = f(x)$.
-Therefore, if we have 4 modalities and the batch size is 64, the size of $I$ is $256$, the size of $P(i)$ is $M-1 = 3$ where $M$ is the number of modalities, and the size of $A(i)$ is $256 - 1 = 255$.
-
-\subsubsection{Supervised Contrastive Learning}
-\label{sub:baseline-supcon}
-
-\citet{NEURIPS2020_supervised_contrastive} extend the contrastive learning method (NT-Xent) and propose a supervised way of performing contrastive learning to treat not only augmentations of the anchor, but also every item that shares the same label with anchor as positives.
-This loss function is shown in \cref{eq:supervised-contrastive}.
-
-\begin{equation}\label{eq:supervised-contrastive}
-    \sum_{i \in I} \frac{-1}{|P(i)|} \sum_{p \in P(i)} \log \frac{\exp (z_i \cdot z_p / \tau) }{\sum_{a \in A(i)} \exp (z_i \cdot z_a / \tau)}
-\end{equation}
-
-% \begin{equation}\label{eq:supervised-contrastive-my-notation}
-%     \sum_{i \in I} \frac{-1}{|P(i)|} \sum_{p \in P(i)} \log \frac{\exp (f_i(x_i) \cdot f_p(x_p) / \tau) }{\sum_{a \in A(i)} \exp (f_i(x_i) \cdot f_a(x_a) / \tau)}
-% \end{equation}
-
-Although this loss function does not use cosine similarity, embeddings are normalized before performing dot product, which is equivalent to cosine similarity.
-
-The main difference between the contrastive loss baseline in \cref{sub:baseline-contrastive} and \supcon{} is that there is no notion of meaningful negative points in contrastive loss, and everything in the batch that is not the anchor or one of the positive views is considered to be negative.
-In \supcon{}, however, all elements in the batch that have same label as the anchor are also considered positives, in addition to different views of the same instance.
-While the denominators of \cref{eq:explicit-contrastive,eq:supervised-contrastive} stay the same, this subtle difference affects the numerator and includes more positive examples, which prevents the unintended use of actual positives as negative examples.
-
-While this model is a strong baseline, the authors applied it to a unimodal dataset.
-In this paper we extend this baseline to work with a multimodal dataset and show that it is slower than \ours{} to learn and its performance is also lower when all modalities are available during test time.
-% if modalities are ablated, this method does not perform as well as our proposed model (we show and discuss this later, in \cref{fig:epochs-mrr.srd}).
-This is especially true when the text modality is dropped, as is likely in physical agent scenarios, where speech is a natural interaction mechanism.
-
-Since \supcon{} considers all pairwise distances in each batch, with $M$ modalities and a batch of size $B$ each batch contains $B$x$M$ items, and \supcon{} computation involves $(BM)^2$ pairwise distance terms which is dependent on batch size.
-However, the computations of our \geom{} approach is agnostic with respect to batch size which makes it scalable.
-
-\subsection{EMMA: Extended Multimodal Alignment}
-\label{sec:emma}
-
-Our proposed multimodal method is composed of two complementary parts.
-The first part is a geometric loss based on distances in the latent space, and the second part is a contrastive loss based on cross-entropy.
-The geometric loss is faster to learn while the cross-entropy method is more aligned with the downstream task of object retrieval.
-Hence we propose to combine them.
-
-\paragraph{Geometric Alignment Loss}
-\label{subsec:geom-loss}
-
-% For $M$ modalities, we 
-We define a distance-based loss function which can be used for an arbitrary number of modalities.
-Our proposed method is inspired by the well-known similarity-based triplet loss~\citep{Carvalho-cooking-triplet,triplet_loss_2021_CVPR}, and is similar to contrastive loss~\citep{chen2020simple,NEURIPS2020_supervised_contrastive} under some settings.
-Triplet loss-based learning works by forcing similar concepts from different domains `together' in some shared embedding space, while forcing dissimilar concepts `apart.' It is so named because it relies on three data points from the training set: a positive, a negative, and an anchor point.
-However, standard triplet loss cannot be used for more than two modalities.
-
-To address this issue, we use pairwise distance optimization for all data points. %
-Our method can be used for an arbitrary number of modalities. %
-In contrast to our work, previous works that use triplet loss~\citep{GoLD_UMBC,triplet_loss_2021_CVPR} concatenate RGB and depth to form a single ``vision'' vector to handle three modalities, so they cannot robustly handle RGB or depth sensor ablation during test.
-Moreover, our method has the advantage that it does not require providing hard negative examples.
-Therefore, we modify the concept of triplet loss as follows.
-During training, we sample two different instances and their corresponding representations from all modalities into two sets---one positive set (referring to a specific object) and one negative set (referring to some different object) as shown in \cref{fig:emma-loss-trimmed}.
-Unlike some prior triplet loss methods~\citep{GoLD_UMBC,triplet_loss_2021_CVPR}, the anchor is not randomly chosen from different modalities in each batch.
-Instead, in our setting, every item in the positive set becomes anchor once, and we minimize the distance between that item and other items in the positive set, while minimizing the distance between that item and all items in the negative set.
-It can be seen as an one-to-many relationship instead of an one-to-two relationship in the triplet loss formulation.
-% the anchor and positive items are from the same object class, but in different modalities and each of the positives modalities becomes anchor once. 
-We consider the following terminology:
-% \todokdinline{I feel like using the terminology of triplet loss especially \textbf{anchor} is confusing here or maybe it needs to be rephrased.
-We can alternatively say that we consider all pairwise distances.}
-% \todoffinline{I don't see the problem with anchor, but if you'd rather use a different term, what about query?
-As a note, while you do consider all pairwise distances but you treat some differently than others.}
-\begin{itemize}
-    % \item data point: a row in the csv/tsv file.
-    \item Positive (Instance): A set of embeddings of one data point (e.g., an RGB image of an apple, corresponding depth image, text description, and speech description of the same apple) as shown with green rectangles in \cref{fig:emma-loss-trimmed}
-    \item Negative (Instance): A set of embeddings of another data point of a different object (e.g., an RGB image of a mug, corresponding depth image, text description, and speech description of the same mug) as shown with orange rectangles in \cref{fig:emma-loss-trimmed}
-    \item Anchor (Modality): Every modality of the positive set is chosen as the learning anchor once.
-In our formulation, anchor is a ``concept'' referring to the points in the positive sets rather than a single instance in itself.
-In \cref{fig:emma-loss-trimmed}, each of the 4 modalities is selected as an anchor once.
-    %For our experiments, text is used. 
-    The anchor is used as the basis for learning distances between positive and negative samples.
-    % variance of text is lower and clusters of different objects are more distinct compared to other modalities, therefore, it can train a better model.
-    % where we don't minimize the distance between the pair of embeddings from all modalities of the same data point, and we don't maximize the distance between positive and negative datapoints of the same modality.
-    % \todo{explain this better}
-    % \todokd{I rephrased it. Good?}
 \end{itemize}
-
-The objective is then to first minimize the distance between each pair of positive points from heterogeneous modalities, and second, maximize the distance between each positive and negative points from all modalities.
-%and finally minimize the distance between each pair of negative points from heterogeneous modalities.
-% The objective is then to first minimize the distance between the anchor and each of the other positive points from heterogeneous modalities, and second, maximize the distance between the anchor and each of the negative points from all modalities.
-%and finally, maximize the distance between positive and negative anchor points from the same modality. 
-% When we minimize the distance between each pair of positives, and maximize the distance between each pair of modalities one from the positive set and one from the negative set, the model converges faster than \supcon{} while maintaining the same performance when we have all modalities and a little bit less than Supcon when we drop text.
-We refer to this approach as \geom{} which is formulated in \cref{eq:full-emma}. An illustration of this loss function is provided in \cref{fig:emma-loss-trimmed}.
-
-% \begin{equation}\label{eq:full-emma}
-%     L = \sum_{m_1=1}^{M} \left[ \sum_{m_2=1}^{M} \left[ \max (\cos(z_{m_1}^{+}, z_{m_2}^{-}),0) \right] + \sum_{m_3=m_1+1}^{M} \left[ 1 - \max(\cos(z_{m_1}^{+},z_{m_3}^{+}),0) \right] \right]
-% \end{equation}
-\begin{equation}\label{eq:full-emma}
-\begin{split}
-    L &= \sum_{m_1=1}^{M} \left[ \sum_{m_2=1}^{M} \left[ -\max (dist(z_{m_1}^{+}, z_{m_2}^{-}) + \alpha,0) \right] + \sum_{m_3=m_1+1}^{M} \left[ \max( dist(z_{m_1}^{+},z_{m_3}^{+}),0) \right] \right]
-\end{split}
-\end{equation}
-
-In \cref{eq:full-emma}, $M$ is the number of modalities, the superscripts $+$ and $-$ represent positive and negative objects, $\alpha$ represents enforced margin between each positive and negative points which we set to 0.4 for all modalities without tuning, and $z$ is the embedding we get by applying a mapping function $f$, which in our case is a neural network on our input data.
-In other words, $z_m = f_m(x_m)$, where each modality $m$ has a specific model $f_m$ that is different from the models for other modalities. These models do not share their weights.
-
-Cosine similarity is the opposite of distance, and we need to reverse the logic for maximization and minimization. 
-% 
-There are different options to measure distance in embedded space.
-We use cosine similarity between pairs of embeddings, i.e.
-we measure the cosine of the angle between embeddings.
-Cosine similarity is a good choice for high-dimensional data as it is bounded between -1 and 1.
-Other distance metrics, such as Euclidean distance, grow in value with respect to their dimensionality, resulting in a very large distances for data points.
-% The $\cos(\cdot)$ function is a measure of similarity, not distance, and that is why the signs are reversed.
-
-\begin{figure*}[tb]
-\centering
-% \includegraphics[width=.9\columnwidth]{Figures/4way-MMA.png}
-% \includegraphics[width=2.1\columnwidth]{Figures/e-MMA.png}
-\includegraphics[width=1\columnwidth]{Figures/emma-loss-trimmed.pdf}
-\caption{A high-level prototype of our approach and the distances used in the \geom{} loss with four modalities.
-Gray arrows indicate pre-trained models that are frozen (i.e.
-the parameters are fixed and are not trained).
-The black arrows show 3 fully connected layers with a ReLU activation function~\citep{relu2010} after the first two layers.
-These networks are trained.
-Orange rectangles are negative embeddings and green rectangles are positive embeddings.
-Dashed lines indicate distances to be maximized while dotted lines indicate distances to be minimized.%, and solid lines show the enforced margin.
-}
-\label{fig:emma-loss-trimmed}
-\end{figure*}
-
-Here, the generic \textit{dist} function is replaced with the specific $\cos(\cdot)$, and we omit the max notation for clarity by defining \cref{eq:aux-equation}:
-
-\begin{equation}\label{eq:aux-equation}
-    \begin{split}
-        & g(x,y) = \max(\cos(x,y)-1 + \alpha, 0 ) \\
-        & h(x,y) = \max(1- \cos(x,y), 0 ).
-    \end{split}
-\end{equation}
-
-The first portion of the following equation maximizes all unique pairwise distances between modalities of positive and negative instances.
-The second portion minimizes the unique pairwise distances among the modalities of positive instances.
-
-\begin{equation}\label{eq:objective-geom}
-\begin{split}
-    \mathcal{L} & = \sum_{m_1=1}^{M}  \sum_{m_2=1}^{M} g(z_{m_1}^{+}, z_{m_2}^{-})  + \sum_{m_1=1}^{M} \sum_{m_3=m_1+1}^{M} h(z_{m_1}^{+},z_{m_3}^{+})
-    %  &= \sum_{i=1}^{M-1} \sum_{j=i+1}^{M} dist( z_{i}^{+} , z_{j}^{+}) - dist( z_{i}^{+} , z_{j}^{-}) - dist( z_{i}^{-} , z_{j}^{+}) - \sum_{i=1}^{M} dist( z_{i}^{+} , z_{i}^{-} ) \\
-\end{split}    
-\end{equation}
-Our proposed \geom{} loss function in \cref{eq:objective-geom} can be rewritten as shown in \cref{eq:objective-geom-rewritten} by fully specifying the summations to better understand how our objective function can be reduced to well-known losses such as triplet loss and pairwise loss.
-
-\begin{equation}\label{eq:objective-geom-rewritten}
-\begin{split}
-     \mathcal{L} &= \sum_{i=1}^{M-1} \sum_{j=i+1}^{M} h( z_{i}^{+} , z_{j}^{+})
-     + g( z_{i}^{+} , z_{j}^{-}) + g( z_{i}^{-} , z_{j}^{+}) + \sum_{i=1}^{M} g( z_{i}^{+} , z_{i}^{-} ).
-\end{split}
-\end{equation}
-
-If $M=2$ which means the number of modalities is 2, and we ignore the last two terms in the derived objective function, it results in the triplet loss method. 
-If $M=2$, then our objective function reduces to the quadruplet loss method~\citep{tursun2021efficient,chen2017beyond} if we multiply the first term by 2, ignore the third term, and change the last summation to be up to $M-1$ (which results a single term).
-If $M=1$, only the last term remains in the loss function which is exactly the pairwise distance-based loss function.
-This loss function can be seen as a contrastive loss usually used in the domain of self-supervised learning~\citep{chen2020simple}.
-However, our proposed loss function has two advantages over the traditional contrastive loss expressed in \cref{eq:contrastive-loss}.
-The first advantage is that our loss function does not loop over multiple positives and negatives in a large batch.
-Instead we sample only two objects (positive and negative) each of which have $M$ modalities which gives us $2M$ datapoints (or embeddings).
-Hence, our model can be trained using smaller batch sizes and reduces the number of negative samples we need.
-The second advantage is that this loss function can be used in a multimodal setting with an arbitrary number of modalities, and is not limited to a single data type (e.g.
-RGB images) which is the most common usage of contrastive loss.
-% ~\citep{chen2020simple, NEURIPS2020_supervised_contrastive}.
-Although our \geom{} is technically quadratic in terms of number of modalities, we observe that experimentally, training time increases only by 10 minutes with each additional modality.
-
-Altogether, our proposed \geom{} function contains 
-$3M^2-M/2$ terms: 
-$M(M-1)/2$ anchor-to-positive distance minimizations and $M^2$ anchor-to-negative distance maximizations.
-%, and $M(M-1)/2$ \textcolor{red}{optional} negative-to-negative distance minimizations.
-%1 positive-to-negative anchor distance maximization.
-It is noteworthy that our training procedure does not perform any stochastic dropout of modalities to obtain test-time robustness to missing modalities.
-Moreover, our approach does not need to compute the distance between all items in the batch, as opposed to \supcon{}.
-% ~\citep{NEURIPS2020_supervised_contrastive}.
-
-\paragraph{Combining Geometric and Cross-Entropy Losses}
-\label{subsec:emma}
-
-The main difference between \geom{} and \supcon{} is that in \geom{} we focus on the geometric interpretation of similarity using cosine distances, while in \supcon{}, cosine distances are used to compute a classification-based loss function similar to cross-entropy.
-Both methods are valid and have some advantages that the other method does not offer.
-In \geom{}, the advantages are an intuitive learning objective in terms of distance, interpretability of the learned embedding space, and faster convergence.
-The advantage of \supcon{} is that it uses a classification objective which is aligned with the downstream task.
-
-We propose to combine the \geom{} defined in \cref{eq:objective-geom} with \supcon{} defined in \cref{eq:supervised-contrastive}, and we refer to this approach as \ours{}, for extended multimodal alignment.
-The combination is done by taking the sum of the equations for \geom{} and \supcon{} which is shown in \cref{eq:emma}.
-When
-
-\begin{equation}\label{eq:emma}
-\begin{split}
-    \mathcal{L} =  \sum_{i \in I} \Biggr[ & \Bigr[ \sum_{j=1}^{M-1} \sum_{k=j+1}^{M} h( z_{i,j}^{+} , z_{i,k}^{+})
-     + g( z_{i,j}^{+} , z_{i,k}^{-}) + g( z_{i,j}^{-} , z_{i,k}^{+}) + \sum_{j=1}^{M} g( z_{i,j}^{+} , z_{i,j}^{-} )\Bigr] + \\
-     & \Bigr[ \sum_{m=1}^{M} \frac{-1}{|P(i,m)|}  \sum_{\beta \in P(i,m)} \log \frac{\exp (z_{i,m} \cdot z_{\beta} / \tau) }{\sum_{\gamma \in A(i,m)} \exp (z_{i,m} \cdot z_{\gamma} / \tau)} \Bigr] \Biggr],
-     \end{split}
-\end{equation}
-where $A(i,m)$ includes all items in the batch except for the $z_{i,m}$ itself and $P(i,m)$ includes all the modalities of all instances that have the same label as current instance excluding $z_{i,m}$ itself.
-In other words, the collection of embeddings indexed by $P(i,m)$ is $\{ \bigcup\limits_{r \neq m \in M} z_{i,r}, \bigcup\limits_{l \neq i \in I , y_i = y_l} \bigcup\limits_{m=1}^{M} z_{l,m} \} $.
-
-% \todo[inline]{Which one do you prefer?
-The first one is my attempt at using almost the same terminology for both SupCon and Geometric method, while the second one is simply copying and pasting the two equations next to each other.}
-
-% \begin{equation}\label{eq:emma-append}
-%     \mathcal{L} = \Bigr[ \sum_{i=1}^{M-1} \sum_{j=i+1}^{M} -\cos( z_{i}^{+} , z_{j}^{+})
-%      + \cos( z_{i}^{+} , z_{j}^{-}) + \cos( z_{i}^{-} , z_{j}^{+}) + \sum_{i=1}^{M} \cos( z_{i}^{+} , z_{i}^{-} ) \Bigr] + 
-%      \Bigr[ \sum_{i \in I} \frac{-1}{|P(i)|} \sum_{p \in P(i)} \log \frac{\exp (z_i \cdot z_p / \tau) }{\sum_{a \in A(i)} \exp (z_i \cdot z_a / \tau)} \Bigr]
-% \end{equation}
-
-% 
-% \begin{comment}
-% 
-% \begin{equation}\label{eq:emma}
-%     % \mathcal{L} = \geom{} + \supcon{}
-%     \ours{} = \geom{} + \supcon{}
-% \end{equation}
-% \end{comment}
-% 
-Compared to pure \supcon{}, combining these loss functions results in faster convergence and slightly improved performance when all modalities are available, and maintains improved performance when modalities are missing.
-Experimental results are presented in detail in \cref{sec:results}.
-
-% We discuss the geometric interpretation of our method versus the classification interpretation of SupCon method and advantages of each.
-Finally, we show that combining \geom{} with \supcon{} is superior to either the \geom{} approach or \supcon{} alone.
-
-\begin{comment}
-\subsubsection{EMMA + Binary Classification}
-\todokdinline{Should we remove this section?}
-
-When batch size is 64, it performs worse than baseline (SupCon) and EMMA itself. The experiment for the batch size of 2 also performs worse than other versions of EMMA, except for when we drop text.
-
-There is a slight chance that I have made a mistake in my implementation of binary cross-entropy loss. My guesses (in decreasing order of likelihood) are:
-
-1.
-The inputs to the BCE loss are $sigmoid(\cos_{sim}(z_i,z_j))$ for all $i \neq j$ where i is over the positive modalities only, while j is over both positive and negative modalities.
-Maybe I should've used $1-\cos(pos_i, pos_j)$ for similar embeddings and $cos(pos_i, neg_j) -1$ for dissimilar embeddings instead; similar to how EMMA's geometric loss is structured.
-
-2.
-Let's assume $\text{predicts} = \text{sigmoid}(\cos_{sim}(z_i,z_j))$.
-Their corresponding labels/targets are 1 if i and j are from positive modalities, and 0 if j is from the negative modalities.
-The way I used PyTorch's BCE is that I passed the predicts whose targets are 1 to BCE first, passed the predicts whose targets are 0 to BCE in the next step, and finally added the two together.
-In other words:
-
-\begin{equation}
-\begin{aligned}    
-loss &= 0.0 \\
-loss &= loss + BCE(\text{similar predicts, targets}=1) \\
-loss &= loss + BCE(\text{dissimilar predicts, targets}=0)
-\end{aligned}
-\end{equation}
-
-Adding temperature gives mixed results; when we have all the modalities during the test, adding temperature improves performance over not having a temperature, but is still slightly worse than other methods.
-When we drop the text modality, however, the performance goes down.
-\end{comment}
-
-\subsection{Network Architecture}
-\label{sec:Model}
-
-Transformers have become \textit{de facto} architectures in the natural language processing community and have shown great success across different tasks.
-Similar to~\citet{GoLD_UMBC}, we use BERT~\citep{devlin-etal-2019-bert} embeddings contained in the FLAIR library~\citep{akbik2019flair,akbik-etal-2019-pooled} to featurize textual input, and wav2vec2~\citep{wav2vec2} to extract audio embeddings from speech.
-Both of these encoders output a 3072-dimensional embedding vector which is generated by concatenating the last four hidden layers of their corresponding networks.
-FLAIR has historically been used for different natural language processing tasks such as named entity recognition (NER) and part-of-speech tagging (PoS), and wav2veq2 has supported a a number of audio processing tasks, most notably automated speech recognition.
-Both BERT~\citep{devlin-etal-2019-bert} and wav2vec2~\citep{wav2vec2} are self-supervised language models using transformers~\citep{vaswani2017attention}.
-% 
-To process images, we use ResNet152~\citep{He_resnet_2016} for both RGB and depth images which gives us a 2048-dimensional embedding vector. Depth images are colorized before passing to the ResNet152.
-
-We then use different instantiations of the same multi-layer perceptron (MLP) consisting of 3 fully connected layers with ReLU activation~\citep{relu2010} to map each of these embeddings to a shared 1024-dimensional space where we can compute the distances between all embeddings.
-We note that these MLP networks are distinct and do not share any weights.
-
-%===================================================================
-
-\section{Experiments}
-\label{sec:Experiments}
-
-In this section we evaluate the quality of object retrieval models learned using the EMMA loss function.
-We first describe the dataset we use, then describe the metrics by which we evaluate performance, the setup of the experiments, and the baselines against which we compare.
-We end by presenting and analyzing results.
-
-\subsection{Data}
-\label{sec:Data}
-
-We demonstrate the effectiveness of our approach on a recent publicly available multimodal dataset called GoLD~\citep{GoLD_UMBC}, which contains RGB images, depth images, written text descriptions, speech descriptions, and transcribed speech descriptions for 207 object instances across 47 object classes (see \cref{fig:emma-loss-trimmed}).
-There are a total of 16,500 spoken and 16,500 textual descriptions.
-The original GoLD paper uses raw RGB and depth images in which other objects are present in the background.
-We use a masked version of the images where the background is deleted (this masked version converges faster, however, masked and unmasked versions of the GoLD data converge to the same performance).
-Speech is converted to 16 Hz to match the wav2vec2 speech model.
-
-\subsection{Setup}
-% \label{sec:setupmetrics}
-\label{sec:setup}
-To evaluate our model we measure different performance metrics on a retrieval task where the model has to select an object from a set of objects given a language description.
-Only one of the objects corresponds to the description and the rest are from different object classes.
-% (e.g., one apple among a set including a fork, a mug, a lemon, and a bell pepper).
-% 
-Similar to~\citet{NEURIPS2020_supervised_contrastive}, we use a stochastic gradient descent (SGD) optimizer with momentum~\citep{ruder2016overviewSGD} with a flexible learning rate starting at 0.05. 
-% For the triplet loss experiments we use the Adam~\citep{kingma_adam_2015} optimizer with scheduling learning rate as suggested by the papers.
-All models are trained for 200 epochs with a batch size of 64 on a Quadro RTX 8000 GPU.
-We used a temperature of $0.1$ for training the contrastive learning method described in \cref{sub:baseline-contrastive}, and a temperature of $0.07$ for training \supcon{} as described in \cref{sub:baseline-supcon}.
-
-To evaluate the performance, we compute the distance between the given natural language description and 5 randomly selected objects (1 of which corresponds to the description, with the others from different object classes).
-We compute the distance between the language embedding and all available sensory modalities of all candidates as described in \cref{sub:ablations}.
-In case we have RGB and depth, we compute the distance between language embedding and all candidate RGB embeddings, and we compute the distance between the same language embedding and all candidate depth embeddings corresponding to the RGB embeddings.
-We then take average of these two distance matrices.
-Instead of choosing an empirical threshold beyond which objects are considered to be `referred to,' we choose the closest image embedding (average distance of RGB and/or depth from language) as the prediction.
-In order to use cosine \textit{distance}, we have to subtract the cosine of the \textit{angle} between two embeddings (which represents similarity) from 1: that is, we compute $1 - \cos(e_1, e_2)$.
-
-\subsection{Metrics}
-\label{sec:metrics}
-The best metric to capture the performance in such a scenario is mean reciprocal rank (MRR, \cref{eq:mrr} for $Q$ queries).
-For each query we predict the rank of all objects based on their distance from the language command, and then the inverse rank of the desired objects in all queries are averaged.
-For example, if the model predicts the desired object as the first rank, then MRR = $\frac{1}{1} = 1$ which means a perfect score, and if it predicts the correct object as the fourth rank among five objects, then MRR = $\frac{1}{4}=0.25$.
-
-\begin{equation} \label{eq:mrr}
-    \mathrm{MRR}=\frac{1}{|Q|} \sum_{i=1}^{|Q|} \frac{1}{\operatorname{rank}_{i}}
-\end{equation}
-
-While MRR is more meaningful when it comes to ranking in retrieval tasks, in the real-world scenarios where a robot is asked to hand over an object, if it fails, it does not matter whether the correct object was ranked second or last and the whole system would be considered a failure.
-Accuracy and micro F1 score
-% (sklearn), flattened binary f1 score (sklearn), and F1 score 
-are the same in this task, since for each prediction we either have a true positive and no false positives and no false negatives, or we have no true positives, one false positive and one false negative.
-MRR is a more informative metric because it captures the idea that having the correct object as the second choice should be considered better than having it as a last choice, while in accuracy the score is ``all or nothing''---either 0 or 1.
-Because our approach is designed to be robust to missing information across modalities, we also report MRR and accuracy for different combinations of modality dropouts.
-
-\subsection{Modality Ablation}
-\label{sub:ablations}
-% \todokdinline{Should I remove this section altogether or should I just mention that our method is equal or better in all cases except when we drop text?
-Or should I spend more time to come up with something so that we can perform better than baseline even in this case?}
-We consider an experiment in which we incorporate RGB, depth, speech, and written language to train the model.
-The loss function requires no changes beyond adjusting the value of $M$ in \cref{eq:full-emma} according to the number of modalities available during training.
-Our goal is the non-trivial downstream prediction task: determining what objects are being referred to by arbitrary language from a small number of examples.
-When we consider only text, RGB, and depth, written language is used as the query modality, and we compute the distance of RGB and depth modalities from it and then average them.
-However, when speech is incorporated as an additional fourth sensory modality, we have three possible choices.
-First, we could compute the distance of RGB and depth from text and from speech which gives us 4 distance matrices, and then take average of these four.
-Second, we could treat speech in a similar way to RGB and depth: compute the distance of RGB, depth, and speech from text, and then take an average of three of them.
-Third, we could compute distances similarly to the first method, but add the distance between language and speech as well and then take the average of 5 distance matrices.
-
-Of these, the first method is the most appropriate choice for a robust multimodal alignment approach.
-The second and third options are possible during training, but in real-world object retrieval scenarios, having only one form of language instructions is a reasonable scenario---people are not likely to \textit{both} speak about \textit{and} type in instructions for an agent.
-At test time, depending on which modalities are available to the model, we can use speech, text, or both to compute the distance of RGB and depth embeddings from the linguistic query, and then take the average.
-
-There are total of nine possible cases of modality dropout and the corresponding distance computations.
-In all these cases \textit{t} represents text, \textit{s} represents speech, \textit{r} represents RGB, \textit{d} represents depth, and $K$ is the final distance---a matrix if there are multiple language queries and a vector if there is one query.
-If we only have two modalities, we simply compute the distance between those two modalities; this corresponds to four of the cases; $K_{tr}$ when speech and depth are missing, $K_{sr}$ when text and depth are missing, $K_{td}$ when speech and RGB are missing, and $K_{sd}$ when text and RGB are missing.
-If we have three modalities, we need to take average of two distances.
-There are four cases with three modalities, $K_{trd} = \frac{K_{tr} + K_{td}}{2}$ when speech is missing, $K_{srd} = \frac{K_{sr} + K_{sd}}{2}$ when text is missing,
-and so on.
-% $K_{tsr} = \frac{K_{tr} + K_{sr}}{2}$ when depth is missing, and $K_{tsd} = \frac{K_{td} + K_{sd}}{2}$ when RGB is missing. 
-When we have all modalities available we take average of four distances, $K_{tsrd} = \frac{K_{tr}+ K_{td} +K_{sr} + K_{sd}}{4}$.
-
-% \Cref{eq:distance-matrix} shows all possible cases of modality dropout and the corresponding distance computations, where \textit{t} represents text, \textit{s} represents speech, \textit{r} represents RGB, \textit{d} represents depth, $K_{ij}$ for each pair of modalities is the distance of one instance in modality $i$ from all candidate instances in modality $j$, and $K$ is the final distance---a matrix if there are multiple queries and a vector if there is one query.
-
-\begin{comment}
-\begin{equation}
-\label{eq:distance-matrix}
-K = 
-    \begin{cases} 
-        K_{tr} & \text{speech and depth are missing} \\
-        K_{sr} & \text{text and depth are missing} \\
-        K_{td} & \text{speech and RGB are missing} \\
-        K_{sd} & \text{text and RGB are missing} \\
-        K_{trd} = \frac{K_{tr} + K_{td} }{2} & \text{speech is missing} \\
-        K_{srd} = \frac{K_{sr} + K_{sd} }{2} & \text{text is missing} \\
-        K_{tsr} = \frac{K_{tr} + K_{sr} }{2} & \text{depth is missing} \\
-        K_{tsd} = \frac{K_{td} + K_{sd} }{2} & \text{RGB is missing} \\
-        K_{tsrd} = \frac{K_{tr} + K_{td} + K_{sr} + K_{sd} }{4} & \text{o.w.} \\
-    \end{cases}
-\end{equation}
-% \todo[inline]{equation \cref{eq:distance-matrix} takes a lot of space which suggests it's important. Either merge it in text or ...}
-\end{comment}
-
-\Cref{fig:result_graphs} shows the relative performance of \ours{} and \geom{} against state-of-the-art methods when different modalities are ablated.
-
-\section{Results}
-\label{sec:results}
-
-In this section we provide quantitative and qualitative results by comparing our method against supervised contrastive learning~\citep{NEURIPS2020_supervised_contrastive} and na\"ive contrastive loss~\citep{chen2020simple}.
-% We evaluate our EMMA loss function against the na\"ive triplet loss method.
-Our model significantly outperforms the triplet method across all epochs while it converges faster as shown in \cref{fig:epochs-mrr.srd}.
-\Cref{table:quantitative} summarizes the performance of all models with different ablation metrics.
-% \todo[inline]{Why do we call out triplet loss specifically? We evaluate against several baselines? I don't understand this section particularly}
-To provide a better sense of the performance measure, consider a model that always ranks the correct object in the second place: Such a model would have an MRR of $1/2 = 0.5$.
-
-\begin{table*}[tb]
-    \centering
-    \begin{subtable}[t]{1.00\textwidth}
-    \centering
-    \resizebox{1.0\textwidth}{!}{
-    \input{tableMRR}
-    }
-    \caption[]{Average and standard deviation of mean reciprocal rank (MRR) on a held-out test set. MRR is from $\frac{1}{\text{number of objects}}$\% to 100\%.
-    In most cases \ours{} has a better MRR compared to other methods, and all methods have a higher MRR when text modality is present.
-This shows that text modality is a rich source of information compared to speech, and that \ours{} is strongly competitive with state of the art approaches even when relying on speech for the language signal.
-    % \todocmi{Right? Correct}
-    }
-    \label{table:mrr}
-    \end{subtable}
-    \vskip\baselineskip
-    \begin{subtable}[t]{1.00\textwidth}
-    \centering
-    \resizebox{1.0\textwidth}{!}{
-    \input{tableACC}
-    }
-    \caption[]{Average and standard deviation of accuracy (Acc) on a held-out test set. Accuracy is from 0\% to 100\%.
-    Although accuracy is a strictly less forgiving metric than MRR, these results demonstrate that \ours{} still outperforms existing approaches in the majority of cases.
-When we drop the text modality, the accuracy of the contrastive method is slightly better than random, while \ours{} is about 10\% more accurate than the contrastive method.
-    }
-    \label{table:acc}
-    \end{subtable}
-    \caption{Average and standard deviation of mean reciprocal rank (MRR) and accuracy (Acc) over 5 runs with 5 different random seeds on a held-out test set with different modalities ablated during testing.
-\Cref{table:mrr} shows MRR scores and \cref{table:acc} shows accuracy.
-    % . These results represent performance when no modalities are ablated during training. 
-     Higher is better for both metrics.
-For 5 objects, a random guess would have MRR of 0.33 and accuracy of 0.5, and the worst case performance would have MRR of 0.2 and accuracy of 0.0.
-The batch size is 64 and optimizer is SGD for all experiments.
-Column headers refer to modalities that are present at query time.
-    % ---e.g., MRR \textit{sd} contains only speech and depth. 
-    Bold numbers represent the best-performing method.
-\ours{} is either better than or very close to both state-of-the-art methods for most of the cases.
-We observe the same pattern in both MRR and accuracy except the fact that accuracy scores are lower than MRR.
-    }
-    \label{table:quantitative}
-\end{table*}
-
-\begin{comment}
-\begin{table*}[tb]
-    \centering
-    \begin{subtable}[t]{1.00\textwidth}
-    \centering
-    \resizebox{.99\textwidth}{!}{
-    \csvautotabular{Figures/unused/resultsMRR.csv}
-    }
-    \end{subtable}
-    \vskip\baselineskip
-    % \hfill
-    \begin{subtable}[t]{1.00\textwidth}
-    \centering
-    \resizebox{.99\textwidth}{!}{
-    \csvautotabular{Figures/unused/resultsACC.csv}
-    }
-    \end{subtable}
-\end{table*}
-\end{comment}
-
-\begin{figure}
-    \centering
-    \begin{subfigure}[b]{0.49\textwidth}
-        \centering
-        \includegraphics[width=\textwidth]{Figures/average-seeds-epochs-mrr_lard-trimmed.pdf}
-        \caption[]{Mean Reciprocal Rank (MRR) on the held-out test set when all modalities are available.}    
-        \label{fig:epochs-mrr.lard}
-    \end{subfigure}
-    \hfill
-    \begin{subfigure}[b]{0.49\textwidth}  
-        \centering 
-        \includegraphics[width=\textwidth]{Figures/average-seeds-epochs-mrr_ard-trimmed.pdf}
-        \caption[]{Mean Reciprocal Rank (MRR) on the held-out test set when the text modality is ablated.}
-        \label{fig:epochs-mrr.srd}
-    \end{subfigure}
-    \vskip\baselineskip
-    \begin{subfigure}[b]{0.49\textwidth}   
-        \centering 
-        \includegraphics[width=\textwidth]{Figures/average-seeds-epochs-mrr_lr-trimmed.pdf}
-        \caption[]{Mean Reciprocal Rank (MRR) on the held-out test set when speech and depth modalities are ablated.}    
-        \label{fig:epochs-mrr.lr}
-    \end{subfigure}
-    \hfill
-    \begin{subfigure}[b]{0.49\textwidth}   
-        \centering 
-        \includegraphics[width=\textwidth]{Figures/average-seeds-epochs-mrr_ld-trimmed.pdf}
-        \caption[]{Mean Reciprocal Rank (MRR) on the held-out test set when speech and RGB modalities are dropped.}    
-        \label{fig:epochs-mrr.ld}
-    \end{subfigure}
-    \caption[]
-    {Mean Reciprocal Rank (MRR, \cref{eq:mrr}) on the held-out test set with the ablation of selected modalities, all averaged over 5 runs for the downstream task of object retrieval.
-Red is self-supervised contrastive learning which is prone to overfitting, orange is supervised contrastive learning, blue represents our proposed \geom{}, and green is our proposed \ours{} loss function.
-Higher is better.
-(a) shows the MRR when all modalities are available, (b) shows the MRR when text is removed at test time, (c) shows MRR when speech and depth are ablated, and (d) shows MRR when speech and RGB are removed.
-We train all models for 200 epochs.
-    } 
-    \label{fig:result_graphs}
-\end{figure}
-%     
-% 
-\Cref{fig:epochs-mrr.lard} Shows that \ours{} learns faster and results in a better performance compared to both \supcon{} and contrastive learning~\citep{chen2020simple} when trained using all modalities and with all modalities available during test.
-We observe that not only does contrastive loss learn more slowly, but that it is prone to overfitting; while this can be addressed with careful tuning of the learning process, an approach that is innately robust to overfitting without tuning is preferable.
-
-When we drop the text modality (\cref{fig:epochs-mrr.srd}), we can see that the performance decreases from about 0.93 to about 0.82, showing that speech cannot completely replace text.
-In \cref{fig:3d-tsne} (see \cref{sec:Qualitative}), the alignment of shared embeddings for a randomly sampled set of classes is visualized for all four modalities under consideration, suggesting that the speech modality is not aligned as well as the text modality.
-For this reason, when we drop text and use speech as the main query, the performance decreases.
-This supports our hypothesis that a geometric alignment of the latent space is crucial to a good performance for object retrieval and multimodal understanding.
-
-In \cref{fig:epochs-mrr.srd}, we observe that when speech is used as the query and the text modality is ablated, the \supcon{} baseline works slightly better than \ours{}, although \ours{} still learns faster.
-The reason is that \supcon{} optimizes for the classification task, and since the speech modality is less well aligned, using \geom{} makes the downstream task more difficult by trying to pull and push similar and dissimilar data points, respectively.
-Future research will consider strategies to align more chaotic modalities.
-% We also observe that speech has a higher variance compared to text, and a na\"ive contrastive learning method cannot take advantage of text modality to reduce this variance.
-While the other two methods leverage text modality to align speech better.
-
-There is very little gap in performance when depth or RGB are dropped in \cref{fig:epochs-mrr.lr,fig:epochs-mrr.ld} compared to when we have all modalities in \cref{fig:epochs-mrr.lard}, showing that our model is robust when RGB or depth sensors fail.
-Also, when depth is dropped in \cref{fig:epochs-mrr.lr}, performance decreases less compared to when RGB is dropped in \cref{fig:epochs-mrr.ld}.
-This suggests that depth is less informative when compared to RGB, which is consistent with existing vision research results.
-
-% \todocmi{Main remaining task is to sort out figures. They should all have meaningful y axis labels and essentially no whitespace surrounding them.}
-
-\paragraph{Qualitative Results}
-\label{sec:Qualitative}
-In order to help visualize the performance of learned embeddings, we consider projections of a randomly selected subset of classes of the high-dimensional learned embeddings into a 3-dimensional space using t-SNE~\citep{van2008tsne}, a dimensionality reduction technique to visualize high-dimensional data.
-T-SNE creates a probability distribution over pairs of high-dimensional data where similar pairs have a higher probability and dissimilar pairs have a lower probability.
-A similar probability distribution is also defined over pairs of data in the lower dimension (either 2D or 3D), and T-SNE minimizes the KL divergence between these two probability distributions.
-% \textcolor{red}{\dots}\todo{explanation}. 
-% \Cref{fig:2d-tsne} shows the projection of all test embeddings from all different modalities in a 2D space.
-As shown in the graph, instances of the same class but different modality are mapped to almost the same locations and are closer to each other.
-\Cref{fig:3d-tsne} shows the projection onto 3D space to give a better view of the location of embeddings.
-Although these projections are not perfect, combined with the quantitative results, they demonstrate that our model is learning to map instances of the same class closer to each other regardless of their modalities.
-Interestingly, toothbrush and toothpaste are mapped almost on top of each other in the text modality showing semantic and syntax similarity.
-However, in the RGB and depth modality they are close but not on top of each other since they do not look the same.
-Also, we can see that apple and lemon are mapped close to each other in all modalities which suggests that our proposed \ours{} learns some notion of the concept of fruits.
-These qualitative results show that our propose \geom{} and \ours{} have an interpretable latent space.
-
-\begin{figure*}[h]
-\centering
-% \includegraphics[width=1.99\columnwidth]{Figures/3D-tsne-4M-eMMA-cosine-submodalities-text-anchor-gold-no_neg_sampling-1024-sbs.png}
-% \includegraphics[width=1\columnwidth]{Figures/3D-tsne-4M-eMMA-cosine-submodalities-text-anchor-gold-no_neg_sampling-1024-sbs.png}
-% \includegraphics[width=.99\columnwidth]{Figures/3D-tsne-exp-supcon-emma-lard-64-relu-SGD-0.001-unique_objects-gold-no_neg_sampling-1024-sbs-trimmed.png}
-\includegraphics[width=.99\columnwidth]{Figures/3D-tsne-exp-supcon-emma-lard-64-relu-SGD-0.001-unique_objects-gold-no_neg_sampling-1024-trimmed.pdf}
-\caption{3D T-SNE~\citep{van2008tsne} projection of test embeddings of 10 randomly selected classes of objects using \ours{}.
-Each modality is separately projected into a three-dimensional space.
-Each RGB and depth image is associated with several language descriptions, leading to denser plots for text and speech.
-In a perfect embedding, all instances of a class would be clustered in identical areas of the embedding space across all modalities.
-We can see that \ours{} successfully encourages all four modalities to live in a common manifold, allowing accurate retrieval even when modalities are missing.
-}
-% \todo[inline]{get rid of markers for modalities, and change the colors to be brighter. Try combining marker and color for each class.}
-
-\label{fig:3d-tsne}
-\end{figure*}
-
-An example of the need to consider multiple modalities jointly is shown in \cref{fig:rankings}, showing how EMMA is able to correctly select an object instance from several similarly shaped and describable objects.
-
-\begin{figure}[h!]
-\centering
-\includegraphics[width=.99\columnwidth]{Figures/example-rankings-rearranged}
-\caption{\ours{} more accurately ranks objects and handles ambiguities with respect to a retrieval queries.
-Note that the phrase ``light blue'' in the query is very similar to ``light bulb'' and while \supcon{} confuses this and predicts light bulb as the correct object, \ours{} correctly identifies the can opener as the intended object, ranking the light bulb second.}
-\label{fig:rankings}
-\end{figure}
-
-% \begin{figure}[tbh]
-% \centering
-% \includegraphics[width=.99\columnwidth]{Figures/2D-tsne-4M-eMMA-cosine-submodalities-text-anchor-gold-no_neg_sampling-1024.pdf}
-% \caption{2D T-SNE~\citep{van2008tsne} projection of test embeddings of 10 randomly selected classes of objects using our proposed method.
-Each modality is separately projected into a two-dimensional space.
-\Rephrase{} \textcolor{red}{We may only want 3D}
-% }
-% \label{fig:2d-tsne}
-% \end{figure}
-
-% \todokdinline{Add the T-SNE projection for supervised contrastive loss as well.}
-
-% \subsection{Failed Experiments}
-% \todokdinline{I suppose I should remove this subsection, right?}
-% We tried a lot of different approaches and they did not work better than our model or the baselines, but we list them here to save time for readers who want to try them.
-
-% \subsubsection{Simple MMA}
-% Copy text from \ref{sub:simple-mma} if we decide to keep this section.
-
-% \subsubsection{Two Anchors}
-% Since speech is similar to text in its usage when it comes to the downstream task of object retrieval (given a single language command, find the correct object among multiple objects), we decided to have text and speech as two anchors instead of having one anchor only.
-When there is one anchor, the distance between positive pairs text and speech is minimized and the distance between negative pairs are minimized.
-However, this is not related to the downstream task since we never want to find the correct speech for a given text or vice versa.
-The second term added to the last function is exactly similar to the loss function when speech is the anchor, and the first term is the same as when language is the anchor.
-Therefore, the performance is somewhere in the middle of those two approaches but more towards the speech when the metrics are computed based and speech matrices.
-
-% \subsection{Example Prediction}
-
-\subsection{Discussion}
-Our proposed model performs well and learns fast, has been demonstrated to handle four modalities of shared information effectively, and is robust to test-time situations where information from one or more modalities is missing.
-There remains room for improvement.
-Specifically, the speech modality is harder to handle.
-\Cref{fig:3d-tsne} shows that although the relative position of instances are correct in the speech space, the distinction and clustering of different objects are not as good as the other three modalities.
-
-Text seems to be the best clustered modality, and that makes sense because the variation in written text is much smaller than the other three modalities.
-Variation in speech is higher because there are a number of factors affecting speech understanding, including different accents, native language, gender, and age~\citep{KebeAAAI2022}.
-Variation in RGB and depth is higher than in text due to variations in lighting conditions, an object's texture and shape, the angle of the camera, and other factors.
-
-%===================================================================
-
-\section{Conclusion}
-\label{sec:conclusion}
-
-In this work, we have demonstrated the effectiveness of a novel approach to learning from high-dimensional multimodal information even when one or more modalities is unavailable at test time.
-Our approach performs well on an object retrieval task from a testbed that contains four separate modalities, consistent with information that might be available to a physical agent, and outperforms state of the art contrastive learning approaches.
-Our proposed method is general enough to be applied to a variety of multimodal retrieval problems, and is not limited to purely language-based image retrieval.
-% \todokdinline{Don't we need evidence for that?}
-
-In future, this work will be extended to solve less clearly delineated problems; for example, differentiating among members of a class, as well as across classes.
-However, this work represents a significant step towards handling such retrieval problems, while not arbitrarily limiting the number of sensor and other modalities that can be incorporated.
-
-% Another future work is to improve the speech modality so it can contribute more to the training.
-Right now when text modality is dropped, the performance decreases by about 0.12.
-Ideally a better model should have the same performance in case of dropout when it is trained using all modalities.
-
-%===================================================================
-
-% ***************** The END of the paper *****************
-
-\bibliography{tmlr}
-\bibliographystyle{tmlr}
-
-\appendix
-\label{sec:appendix}
-% \todocmi{should we get rid of this appendix? I am thinking so.}
-% \input{appendix}
-
-\end{document}

### 2023-02-08 Paper/abstract-submission.tex
Paper submitted to SIGIR 23

Dealing with robots and other physical systems often leads to large amounts of complex, multimodal data, which may not be reliably available throughout the span of an interaction. Our study is motivated by needs in robotics, where an agent may have many sensors and thus many modalities by which a human may interact. Tying these various modalities together relates to a task in grounded language understanding in which natural language is used as a retrieval query against objects in a physical environment. However, current work has neglected the real-world constraint that sensors fail due to obstructions or adverse conditions. We propose a generalized distance-based loss function that can be used to learn retrieval models that incorporate an arbitrary number of views of a particular piece of data, compounded by the challenge of retrieval when a modality becomes unavailable. We demonstrate the usability of our model on a grounded language object retrieval scenario. We leverage four modalities including vision, depth sensing, text, and speech, and demonstrate that our model learns faster than state-of-the-art baselines, while outperforming them when less training data is available. The code is publicly available on GitHub and will be included for the camera-ready version.

### 2023-02-08 Paper/main-SIGIR.tex
Paper submitted to SIGIR 23

--- 2023-01-17 Paper/main.tex
@@
-\documentclass[10pt]{article} % For LaTeX2e
-\usepackage{tmlr}
-%%\usepackage{showframe}
-% If accepted, instead use the following line for the camera-ready submission:
-% \usepackage[accepted]{tmlr}
-% To de-anonymize and remove mentions to TMLR (for example for posting to preprint servers), instead use the following:
-% \usepackage[preprint]{tmlr}
-
-% Optional math commands from https://github.com/goodfeli/dlbook_notation.
-\input{math_commands.tex}
-
-\usepackage{hyperref}
-\usepackage{url}
-
+%%
+%% This is file `sample-sigconf.tex',
+%% generated with the docstrip utility.
+%%
+%% The original source files were:
+%%
+%% samples.dtx  (with options: `sigconf')
+%% 
+%% IMPORTANT NOTICE:
+%% 
+%% For the copyright see the source file.
+%% 
+%% Any modified versions of this file must be renamed
+%% with new filenames distinct from sample-sigconf.tex.
+%% 
+%% For distribution of the original source see the terms
+%% for copying and modification in the file samples.dtx.
+%% 
+%% This generated file may be distributed as long as the
+%% original source files, as listed above, are part of the
+%% same distribution. (The sources need not necessarily be
+%% in the same archive or directory.)
+%%
+%% Commands for TeXCount
+%TC:macro \cite [option:text,text]
+%TC:macro \citep [option:text,text]
+%TC:macro \citet [option:text,text]
+%TC:envir table 0 1
+%TC:envir table* 0 1
+%TC:envir tabular [ignore] word
+%TC:envir displaymath 0 word
+%TC:envir math 0 word
+%TC:envir comment 0 0
+%%
+%%
+%% The first command in your LaTeX source must be the \documentclass command.
+% \documentclass[sigconf,review,natbib=true]{acmart}
+\documentclass[sigconf,natbib=true,anonymous=true]{acmart}
+%% NOTE that a single column version may be required for 
+%% submission and peer review. This can be done by changing
+%% the \doucmentclass[...]{acmart} in this template to 
+%% \documentclass[manuscript,screen]{acmart}
+%% 
+%% To ensure 100% compatibility, please check the white list of
+%% approved LaTeX packages to be used with the Master Article Template at
+%% https://www.acm.org/publications/taps/whitelist-of-latex-packages 
+%% before creating your document. The white list page provides 
+%% information on how to submit additional LaTeX packages for 
+%% review and adoption.
+%% Fonts used in the template cannot be substituted; margin 
+%% adjustments are not allowed.
+
+%% Rights management information.  This information is sent to you
+%% when you complete the rights form.  These commands have SAMPLE
+%% values in them; it is your responsibility as an author to replace
+%% the commands and values with those provided to you when you
+%% complete the rights form.
+\setcopyright{acmcopyright}
+\copyrightyear{2023}
+\acmYear{2023}
+\acmDOI{XXXXXXX.XXXXXXX}
+
+%% These commands are for a PROCEEDINGS abstract or paper.
+\acmConference[SIGIR '23]{}{2023}{Taipei, Taiwan}
+%
+%  Uncomment \acmBooktitle if th title of the proceedings is different
+%  from ``Proceedings of ...''!
+%
+%\acmBooktitle{Woodstock '18: ACM Symposium on Neural Gaze Detection,
+%  June 03--05, 2018, Woodstock, NY} 
+\acmPrice{15.00}
+\acmISBN{978-1-4503-XXXX-X/18/06}
+
+%%
+%% Submission ID.
+%% Use this when submitting an article to a sponsored event. You'll
+%% receive a unique submission ID from the organizers
+%% of the event, and this ID should be used as the parameter to this command.
+%%\acmSubmissionID{123-A56-BU3}
+
+%%
+%% For managing citations, it is recommended to use bibliography
+%% files in BibTeX format.
+%%
+%% You can then either use BibTeX with the ACM-Reference-Format style,
+%% or BibLaTeX with the acmnumeric or acmauthoryear sytles, that include
+%% support for advanced citation of software artefact from the
+%% biblatex-software package, also separately available on CTAN.
+%%
+%% Look at the sample-*-biblatex.tex files for templates showcasing
+%% the biblatex styles.
+%%
+
+%%
+%% The majority of ACM publications use numbered citations and
+%% references.  The command \citestyle{authoryear} switches to the
+%% "author year" style.
+%%
+%% If you are preparing content for an event
+%% sponsored by ACM SIGGRAPH, you must use the "author year" style of
+%% citations and references.
+%% Uncommenting
+%% the next command will enable that style.
+%%\citestyle{acmauthoryear}
+
+%%
+%% end of the preamble, start of the body of the document source.
 % My Packages
@@
 
-\title{Multimodal Language Learning in the Face of Missing Modalities}
+% Remove for submission!!
+% Remove for submission!!
+% Remove for submission!!
+% Remove for submission!!
+\pagenumbering{arabic} 
+\settopmatter{printfolios=true}
+
+\begin{document}
+
+%%
+%% The "title" command has an optional parameter,
+%% allowing the author to define a "short title" to be used in page headers.
+\title{Multimodal Language Learning for Object Retrieval in the Face of Missing Modalities}
 % \\ Extended Multimodal Alignment \\ Combining Geometric and classification ... }
 
-% Authors must not appear in the submitted version. They should be hidden
-% as long as the tmlr package is used without the [accepted] or [preprint] options.
-% Non-anonymous submissions will be rejected without review.
-
-\author{\name Kasra Darvish \email kasradarvish@umbc.edu \\
-      \addr Department of Computer Science\\
-      University of Maryland Baltimore County
-      \AND
-      \name Edward Raff \email Raff\_Edward@bah.com \\
-      \addr University of Maryland Baltimore County \\
-      Booz Allen Hamilton
-      \AND
-      \name Francis Ferraro \email ferraro@umbc.edu \\
-      \addr University of Maryland Baltimore County
-      \AND
-      \name Cynthia Matuszek \email cmat@umbc.edu\\
-      \addr University of Maryland Baltimore County \\
-      CIFAR Fellow}
-
-% The \author macro works with any number of authors. Use \AND 
-% to separate the names and addresses of multiple authors.
-
-\newcommand{\fix}{\marginpar{FIX}}
-\newcommand{\new}{\marginpar{NEW}}
-
-\def\month{MM}  % Insert correct month for camera-ready version
-\def\year{YYYY} % Insert correct year for camera-ready version
-\def\openreview{\url{https://openreview.net/forum?id=XXXX}} % Insert correct link to OpenReview for camera-ready version
-
-\begin{document}
-
-\maketitle
+%%
+%% The "author" command and its associated commands are used to define
+%% the authors and their affiliations.
+%% Of note is the shared affiliation of the first two authors, and the
+%% "authornote" and "authornotemark" commands
+%% used to denote shared contribution to the research.
+\author{Kasra Darvish}
+% \authornote{Both authors contributed equally to this research.}
+% \email{kasradarvish@umbc.edu}
+% \orcid{1234-5678-9012}
+% \author{G.K.M. Tobin}
+% \authornotemark[1]
+% \email{webmaster@marysville-ohio.com}
+\affiliation{%
+  \institution{University of Maryland Baltimore County}
+  % \streetaddress{P.O. Box 1212}
+  \city{Baltimore}
+  \state{MD}
+  \country{USA}
+  % \postcode{43017-6221}
+}
+\email{kasradarvish@umbc.edu}
+
+\author{Edward Raff}
+\affiliation{%
+  \institution{University of Maryland Baltimore County \\
+      Booz Allen Hamilton}
+  % \streetaddress{1 Th{\o}rv{\"a}ld Circle}
+  \city{Baltimore}
+  \country{USA}}
+\email{Raff\_Edward@bah.com}
+
+\author{Francis Ferraro}
+\affiliation{%
+  \institution{University of Maryland Baltimore County}
+  % \streetaddress{1 Th{\o}rv{\"a}ld Circle}
+  \city{Baltimore}
+  \country{USA}}
+\email{ferraro@umbc.edu}
+
+\author{Cynthia Matuszek}
+\affiliation{%
+  \institution{University of Maryland Baltimore County}
+  \city{Baltimore}
+  \country{USA}}
+\email{cmat@umbc.edu}
+
+%%
+%% By default, the full list of authors will be used in the page
+%% headers. Often, this list is too long, and will overlap
+%% other information printed in the page headers. This command allows
+%% the author to define a more concise list
+%% of authors' names for this purpose.
+% \renewcommand{\shortauthors}{Trovato and Tobin, et al.}
+
+%%
+%% The abstract is a short summary of the work to be presented in the
+%% article.
 
 \begin{abstract}
-    % Grounded language understanding, in which natural language is used as a query against objects in a physical environment, allows a real-world, intuitive mechanism by which users can instruct physical agents to engage in tasks such as object retrieval.
-Visuolinguistic approaches to such object inference tasks typically involve training on large pools of image/text pairs and then using language to subselect elements of the sensed environment.
-However, physical agents such as robots typically have access to sensory and interactive modalities beyond vision, and learning from multiple modalities can improve performance on downstream tasks.
-In order to fully leverage multimodal training data while being robust to missing information, we propose a generalized distance-based loss function that can be extended to learn retrieval models that incorporate an arbitrary number of modalities.
-We demonstrate the usability of our model on a grounded language object retrieval scenario, where an intelligent agent has to select an object given an unconstrained language command.
-We leverage four modalities including vision, depth sensing, text, and speech, and we show that our model can outperform state-of-the-art contrastive models when modalities are ablated.
-
-    % shorter version
-
-    Our study is motivated by needs in robotics and computer-human interaction, where an agent has many sensors and thus modalities with which a human may interact, both to communicate a desired goal and for the agent to recognize a desired target object.
+    Dealing with robots and other physical systems often leads to large amounts of complex, multimodal data, which may not be reliably available throughout the span of an interaction.
+    Our study is motivated by needs in robotics, where an agent may have many sensors and thus many modalities by which a human may interact.
+    % , such as voice, text, images, and depth.
+    Tying these various modalities together relates to a task in grounded language understanding in which natural language is used as a retrieval query against objects in a physical environment.
+    However, current work has neglected the real-world constraint that sensors fail due to obstructions or adverse conditions.
+    We propose a generalized distance-based loss function that can be used to learn retrieval models that incorporate an arbitrary number of views of a particular piece of data, compounded by the challenge of retrieval when a modality becomes unavailable.
+    We demonstrate the usability of our model on a grounded language object retrieval scenario.
+We leverage four modalities including vision, depth sensing, text, and speech, and demonstrate that our model learns faster than state-of-the-art baselines, while outperforming them when less training data is available.
+    The code is publicly available on GitHub and will be included for the camera-ready version.
+%%% END of previous abstract
+
+    % TMLR abstract
+    % Our study is motivated by needs in robotics and computer-human interaction, where an agent has many sensors and thus modalities with which a human may interact, both to communicate a desired goal and for the agent to recognize a desired target object.
 For such problems, there has been little research on integrating more than two modalities.
@@
 We demonstrate the usability of our model on a grounded language object retrieval scenario, where an intelligent agent has to select an object given an unconstrained language command.
-We leverage four modalities including vision, depth sensing, text, and speech, and we show that our model converges approximately 5 times faster than previous strong baselines, and out-performs or is strongly competitive with state-of-the-art contrastive learning.
+We leverage four modalities including vision, depth sensing, text, and speech, and we show that our model converges approximately 2.3 times faster than previous strong baselines, and out-performs or is strongly competitive with state-of-the-art contrastive learning.
 % (in 80\% less time).
-    %when modalities are ablated. 
-    %TODO: another baseline where we use cross-entropy loss only is missing I guess.
-    The code is publicly available on GitHub and will be included for the camera-ready version (it is redacted for anonymity). 
+    % END of TMLR abstract
+
 \end{abstract}
 
-\section{TODO}
-\begin{itemize}
-
-    \item \TD modify it for SIGIR
-    \item \TD EMMA as a contstrained optimization for Geometric portion to not satisfy it.
-    \item \TD constrained SupCon per instance modality constraints: for this particular anchor and corresponding neighbors, we want this constraint to be satisfied.
-    \item \OK Do you control for in batch negatives for supcon portion? Yes.
-    \item \TD alternate training of supcon and emma. E.g. train supcon for 50 epochs and emma for the next 50 epochs
-    \item \TD removing objects that have less than a number of instances? yes, but it's not a good story, since we want emma to work in low-data regime.
-    \item \TD ablating number of data points and have a graph based on different number of training data --> maybe EMMA degrades less than SupCon and that would be good
-    \item \TD use supcon portion to reweight the geometric values or use a method to have different scores for different pairs in the geometric loss. Targeting specific pair of embeddings.
-    \item \TD it seems to be a tension between geometric portion and the supcon loss
-    \item \OK another column for number of NAs in test and train and then correlation between them
-
-    \item \OK add another column for the number of instances and then compute correlation between that and the differential scores.
-Take the correlation between number of instances and each column (metric) and then the correlation and the average of columns.
-
-    \item listen to the speeches for NA descriptions and the ones where there is a significant difference in performance
-
-    \item \OK listen to the speeches for empty transcriptions.
-
-    \item \TD take a look at the depth images
-
-    \item going forward we have two options: Improve the model, or do a deep analysis why our results are like this
-
-    \item We could do something that addresses the "open questions" (which questions remain?)
-
-    I guess the main open question is the mixed results and the fact that they are within the std dev of each other. Basically, I think they are unsure whether being faster is a good result or not.
-    What the reader would take away?
-
-    \item \DO What the rich analysis would consist of?
-
-    \item Why are we getting what we are getting?
-
-    Visualize individual data points and what are the failure modes.
-    You can do the same for type by type (apples vs books)
-    The percent that EMMA got right and SupCon got wrong and vice versa.
-
-    have pictures, positive set and negative set,
-    and all modalities and then show scores for all of them.
-
-    \item \TD Different datasets or tasks
-    \item \TD How can we tell when modalities are complementary?
-
-    I think they meant when loss functions are complementary, am I wrong?
-
-    \item \TD How can we tell what losses to use in what cases?
-
-    \item  \TD Why didn't our approach work better, and what can we do to improve it?
-
-    One reason could be that \supcon{} considers the distance between all pairs of data points in a batch.
-
-    \item What's the next thing to try or add?
-
-    We could add a weighting term to balance the effect of \supcon{} and \geom{}.
-
-\end{itemize}
+%%
+%% The code below is generated by the tool at http://dl.acm.org/ccs.cfm.
+%% Please copy and paste the code instead of the example below.
+%%
+\begin{CCSXML}
+<ccs2012>
+ <concept>
+  <concept_id>10010520.10010553.10010562</concept_id>
+  <concept_desc>Machine Learning~Computer Vision</concept_desc>
+  <concept_significance>500</concept_significance>
+ </concept>
+ <concept>
+  <concept_id>10010520.10010575.10010755</concept_id>
+  <concept_desc>Machine Learning~Natural Language Processing</concept_desc>
+  <concept_significance>300</concept_significance>
+ </concept>
+ <concept>
+  <concept_id>10010520.10010553.10010554</concept_id>
+  <concept_desc>Information Retrieval~Object Detection</concept_desc>
+  <concept_significance>100</concept_significance>
+ </concept>
+ <concept>
+  <concept_id>10003033.10003083.10003095</concept_id>
+  <concept_desc>Deep Learning~Neural Networks</concept_desc>
+  <concept_significance>100</concept_significance>
+ </concept>
+</ccs2012>
+\end{CCSXML}
+
+% \ccsdesc[500]{Machine Learning~Computer Vision}
+\ccsdesc{Machine Learning~Computer Vision}
+\ccsdesc{Machine Learning~Natural Language Processing}
+\ccsdesc{Information Retrieval~Object Detection}
+% \ccsdesc[100]{Deep Learning~Neural Networks}
+
+%%
+%% Keywords. The author(s) should pick words that accurately describe
+%% the work being presented. Separate the keywords with commas.
+\keywords{deep learning, neural networks, multimodal learning}
+
+%% A "teaser" image appears between the author and affiliation
+%% information and the body of the document, and typically spans the
+%% page.
+\begin{comment}
+
+\begin{teaserfigure}
+  \includegraphics[width=\textwidth]{sampleteaser}
+  \caption{Seattle Mariners at Spring Training, 2010.}
+  \Description{Enjoying the baseball game from the third-base
+  seats. Ichiro Suzuki preparing to bat.}
+  \label{fig:teaser}
+\end{teaserfigure}
+\end{comment}
+
+% \received{\todokdinline{Should we remove these?}20 February 2007}
+% \received[revised]{12 March 2009}
+% \received[accepted]{5 June 2009}
+
+%%
+%% This command processes the author and affiliation and title
+%% information and builds the first part of the formatted document.
+\maketitle
+
+% \input{Inbox.tex}
 
@@
 \label{intro}
-
-Inspired by the multimodal nature of human interaction with the world, it is intuitive that agents learning about the world, upon encountering new concepts and new objects, should form a model that incorporates information from all available sensors and data sources.
-The benefits of integrating multiple modalities are twofold; first, complementary information can be extracted from different modalities that can help with understanding the world, and second, additional modalities can help in the cases when one or more sources of data about the world become unavailable.
+Humans use multiple sensory inputs to understand, interact with, and retrieve information from the world around them.
+It is intuitive that agents learning about the world, upon encountering new concepts and new objects, should form a model that incorporates information from all available sensors and data sources.
+The benefits of integrating multiple modalities are twofold: first, complementary information from different modalities can help with understanding the world, and second, additional modalities can help when one or more sources of data become unavailable.
 % 
-Grounded language understanding, in which natural language is used as a query against objects in a physical environment, allows a real-world, intuitive mechanism by which users can instruct physical agents to engage in tasks such as object retrieval.
+Grounded language understanding-based object retrieval, in which natural language is used as a query against objects in a physical environment, allows an intuitive mechanism by which users can instruct agents to engage in tasks.
 Visuolinguistic approaches to such object inference tasks typically involve training on large pools of image/text pairs and then using language to subselect elements of the sensed environment~\citep{hong2021gilbert,Zhuge_2021_CVPR_VLP}.
 
-Although physical agents such as robots typically have access to sensory and interactive modalities beyond vision, and learning from multiple modalities can improve performance on downstream tasks, most approaches use at most two sensory inputs (e.g., visual data such as RGB plus depth images) with single labels, such as those provided by textual natural language.
-Simultaneously using additional inputs from different modalities is an underexplored area, in part due to the domain-specific nature of such \textit{n}-ary learning approaches.
-With the modern proliferation of audio and text based communication and home agents (e.g., Alexa/Google Home), there is a growing need to handle more modalities, and simultaneously their potential failures.
-
-One difficulty with working with complex multimodal data is the increased likelihood that one or more modalities may have missing information.
-Hardware can become damaged or defective, sensors can get blocked or obstructed, and various adverse but not uncommon conditions can remove a modality from use.
-Current multimodal approaches are typically not robust to the loss of one or more modalities at test time, as may happen if, for example, a physical agent fails to retrieve data from a particular sensor.
-In order to fully leverage multimodal training data while being robust to missing information, we propose a generalized distance-based loss function that can be extended to learn retrieval models that incorporate an arbitrary number of modalities.
-
+Although physical agents such as robots typically have access to multiple sensory and interactive modalities, most approaches use at most two sensory inputs (e.g., RGB plus depth images) with single labels.
+Simultaneously using additional modalities and labels is an underexplored area.
+% , in part due to the domain-specific nature of such \textit{n}-ary learning approaches.
+With the modern proliferation of audio and text based communication and home agents (e.g., Alexa/Google Home),
+% there is a growing need to handle more modalities and their potential failures. 
+However, one difficulty of working with complex multimodal data is the increased likelihood that one or more modalities may have missing information.
+Hardware can become damaged or defective, sensors can get blocked or obstructed, and various adverse conditions can remove a modality from use.
+Current multimodal approaches are typically not robust to the loss of one or more modalities at test time.
+
+In order to fully leverage multimodal training data, we propose a generalized distance-based loss function that can be extended to learn retrieval models that incorporate an arbitrary number of modalities, while being robust to missing information.
 We consider the domain of grounded language-based object retrieval~\citep{hu2016natural, triplet_loss_2021_CVPR}, in which objects in an environment must be identified based on linguistic instructions.
-This can be considered a special case of image retrieval~\citep{huang2017deep, ma2020large, novak2015large, vo2019composing} in which objects are identified using visual inputs in combination with other sensor modalities.
-Approaches to acquiring grounded language have explored various combinations of sensor inputs such as depth and RGB with labels provided by textual language or  speech~\citep{RichardsDarvishMatuszekCategoryFree20}.
+This can be considered a special case of image retrieval~\citep{huang2017deep, ma2020large, novak2015large, vo2019composing}.
+Previous approaches have explored combinations of sensors such as depth and RGB with labels provided by language~\citep{RichardsDarvishMatuszekCategoryFree20}.
 However, despite the multisensory nature of object retrieval, much of the existing work has not previously been extended to include an arbitrary number of modalities.
-
-% \todo[inline]{Whats the name of our method?
-When I skim the paper its hard to tell what ``the new thing is'', and then I have to read.
-I think being explicit here, with something like ``\textit{we propose our new method BLA}'' will help -Ed}
-
 The ultimate aim of this work is to take arbitrary input modalities about novel objects, including both spoken and written language, and build a model that allows a system to correctly retrieve that object given a subset of those input modalities.
+
 In this work, we propose our new method called \textit{extended multimodal alignment (\ours{})}, a generalized geometric method combined with a cross-entropy based supervised contrastive loss function~\cite{NEURIPS2020_supervised_contrastive}, which we abbreviate \supcon{}.
 Our method can be used to learn retrieval models that incorporate an arbitrary number of views of a particular piece of data.
-We demonstrate steps towards this ultimate goal with a mechanism that learns quickly from a variety of input signals and is robust to ablation of certain inputs.
-This is a generalization of approaches to grounded language learning in which specific input modalities are labeled with language to allow for future identification, but explicitly seeks to be agnostic about the nature of the individual sensory and linguistic inputs.
+We demonstrate steps towards this goal with a mechanism that learns quickly from a variety of input signals and is robust to ablation of inputs.
+% This is a generalization of approaches to grounded language learning in which specific input modalities are labeled with language to allow for future identification, but explicitly seeks to be agnostic about the nature of the individual sensory and linguistic inputs.
 
 Our contributions are as follows:
-\begin{enumerate}
-    \item Proposing a new approach to multimodal learning by combining geometric and cross-entropy methods and showing that it outperforms and converges faster compared to each method separately.
-    % \item Outperforming state-of-the-art models for grounded language learning by differentiating visual input modalities such as RGB and depth, rather than combining them into a single modality.
-
-    \item Our method is designed to be applicable to additional modalities as they become available.
-
-    \item Our method converges approximately 5 times faster than previous strong baselines, and out-performs or is strongly competitive with state-of-the-art contrastive learning~\citep{chen2020simple} and supervised contrastive learning~\citep{NEURIPS2020_supervised_contrastive} models when all modalities are available.
-
-    \item Demonstrating robustness in the face of missing input signals when one or more modalities are ablated at test time, and proposing a simple averaging method to quantify performance for multimodal object retrieval scenarios.
-
-    \item Demonstrating the separate utility of speech and text as sources of label information by treating them as sensory input modalities, rather than explicitly as labels.
-
-    % \item Stable model that outperforms state-of-the-art contrastive learning models when batch size decreases.\todoff{add motivation in the intro about why robustness to a smaller batch size is important}
-
-\end{enumerate}
+% \begin{enumerate}
+
+(1) We propose a new approach to multimodal learning by combining geometric and cross-entropy methods. Our method is designed to be applicable to additional modalities as they become available.
+
+(2) Our method converges approximately 2.3 times faster than previous baselines, and out-performs state-of-the-art contrastive learning~\citep{chen2020simple} and supervised contrastive learning~\citep{NEURIPS2020_supervised_contrastive} when less training data is available.
+    % and/or modalities are ablated.
+
+(3) We demonstrate robustness in the face of missing input signals when one or more modalities are ablated at test time, and propose a simple averaging method to quantify performance for multimodal object retrieval scenarios.
+
+(4) We demonstrate the separate utility of speech and text as sources of label information by treating them as sensory input modalities, rather than explicitly as labels.
+
+% \end{enumerate}
 
@@
 
-There is extensive work on the task of image retrieval, of which physical object retrieval can be considered a special case.
-In this task, language is used to formulate queries against datasets of images, for example in text-and-image matching tasks for fashion data~\citep{gao2020fashionbert, wen2021comprehensive}, sketch retrieval~\citep{huang2017deep}, and general photographs of objects~\citep{ma2020large, novak2015large, hong2021gilbert}.
+There is extensive work on the task of image retrieval, of which physical object retrieval can be considered a special case, for example in
+% . In this task, language is used to formulate queries against datasets of images, 
+% for example in text-and-image matching tasks for fashion data~\citep{gao2020fashionbert, wen2021comprehensive}, sketch retrieval~\citep{huang2017deep}, and 
+general photographs of objects~\citep{ma2020large, novak2015large, hong2021gilbert}.
 Prior work has focused solely on language and visual modalities, using text or a combination of language and vision to perform visual retrieval~\citep{vo2019composing}.
@@
 \citet{bagher-zadeh-etal-2018-multimodal} introduce a large dataset of multimodal opinion sentiment and emotion intensity (CMU-MOSEI) for sentiment analysis and emotion recognition.
-They also introduce a novel multimodal fusion technique called the Dynamic Fusion Graph (DFG).~\citet{GoLD_UMBC} present a multimodal dataset of household objects containing RGB images, depth images, written language, and spoken language, which has been used to support learning grounded language directly from speech given small datasets~\citep{KebeAAAI2022}.
+% They also introduce a novel multimodal fusion technique called the Dynamic Fusion Graph (DFG).
+~\citet{GoLD_UMBC} present a multimodal dataset of household objects containing RGB images, depth images, written language, and spoken language, which has been used to support learning grounded language directly from speech given small datasets~\citep{KebeAAAI2022}.
 \citet{baltrusaitisMultimodalMachineLearning2019} propose a new taxonomy of multimodal machine learning by introducing five technical challenges in addition to typical early and late fusion categorization.
@@
 
-We highlight a key difference with recent work like \citet{10.1145/3397271.3401232} and \citet{10.1145/3331184.3331213} who develop multi-modal retrieval models.
-These retrieval models are class based, where the model is trained to recognize any object of the same retrieval class as being the `same.' Our method and data is instance based, where different objects are not considered the same even if they have the same class, due to the grounding goals of our work (i.e., the agent should identify the specified object, not equivalent objects).
+We highlight a key difference with recent work like on multi-modal retrieval models~\cite{10.1145/3397271.3401232, 10.1145/3331184.3331213}.
+These retrieval models are class based, whereas our method is instance based, where different objects are not considered the same even if they have the same class.
 These works also do not consider performance in the case of failure of an input modality.
@@
 \citet{alayrac2020self} use a self-supervised contrastive learning method to learn and combine representations from three modalities of visual, audio, and language.
-Audio and visual domains are mapped to the same space because they are fine-grained, containing dense information for each frame of video, and these embeddings are then mapped to a coarse space where text is also mapped.
+% Audio and visual domains are mapped to the same space because they are fine-grained, containing dense information for each frame of video, and these embeddings are then mapped to a coarse space where text is also mapped.
 Their network respects specificity and comparability of different modalities.
-The loss function they define consists of two distance-based terms for the corresponding spaces, differing from ours in that they do not handle arbitrarily many modalities and do not focus on robustness in the face of modality drop-outs.
+The loss function they define consists of two distance-based terms for dense and coarse spaces, differing from ours in that they do not handle arbitrarily many modalities and do not focus on robustness in the face of modality drop-outs.
 \citet{Nguyen-RSS-20} take a similar approach and perform pairwise cosine similarity to align images and natural language text descriptions to perform object retrieval tasks.
@@
 
-In general there are two different families of approaches to multimodal learning in the literature.
-One family is based on classification and cross-entropy losses such as~\citet{NEURIPS2020_supervised_contrastive, chen2020simple} and the other family is geometric based such as~\citet{ICML22GeometricMultimodal,Carvalho-cooking-triplet,salvador2017cooking,triplet_loss_2021_CVPR}.
-In this paper we propose to marry the two families and show that combining geometric approach with cross-entropy based \supcon{} loss function is superior.
-
-% \todokdinline{Should we move the following two paragraphs to the Approach section?}
-% \todocmi{Actually yes, although I think it's more like 3 paragraphs. Rewriting the subsequent paragraphs so they don't appear out of nowhere would help.}
-% \todokdinline{After moving baselines to the approach section and rewriting it, I think it's better if we leave these here. Let me know if you think otherwise.}
+There are two broad families of approaches to multimodal learning in the literature.
+One family is based on classification and cross-entropy losses~\cite{NEURIPS2020_supervised_contrastive, chen2020simple}, and the other is geometric~\cite{ICML22GeometricMultimodal,Carvalho-cooking-triplet,salvador2017cooking,triplet_loss_2021_CVPR}.
+In this paper we marry the two families and show that combining a geometric approach with cross-entropy based \supcon{} loss function is superior.
 
 % For the classification part of our loss function we employ the supervised contrastive loss proposed by~\citet{NEURIPS2020_supervised_contrastive}. 
-The way we formulate our geometric loss function is similar to lifted structured loss~\citep{songCVPR16LiftedStructured}, where they take a geometric approach in a unimodal scenario and consider pairwise distances among all items in a batch, but in this paper we apply our idea to four modalities, and our formulation can be extended to any number of modalities.
-This extended geometric formulation results in a loss function that is similar to \citet{ICML22GeometricMultimodal} where they propose a geometric contrastive learning approach in which they %use the same instance of an MLP network (same weights) after extracting embeddings from each modality, while we use different instances of the same MLP network (different weights).
-They also
+Our geometric loss function formulation is similar to lifted structured loss~\citep{songCVPR16LiftedStructured}.
+They take a geometric approach in a unimodal scenario and consider pairwise distances among all items in a batch, but we apply our idea to four modalities (and could incorporate more).
+This extended geometric formulation results in a loss function that is similar to \citet{ICML22GeometricMultimodal}, where they 
+% propose a geometric contrastive learning approach in which they 
 fuse all embeddings as a central embedding and then minimize the distance between each embedding and that central embedding.
@@
 Triplet loss is also a special case of contrastive loss~\citep{NEURIPS2020_supervised_contrastive}.
-Contrastive loss is mostly used in self-supervised learning~\citep{bui2021self, alayrac2020self,chen2020simple}, with some exceptions such as \citet{NEURIPS2020_supervised_contrastive}, who use contrastive loss for a supervised problem.
+Contrastive loss is mostly used in self-supervised learning~\citep{bui2021self, alayrac2020self,chen2020simple}, with some exceptions such as \citet{NEURIPS2020_supervised_contrastive}.
 Most of these methods are implemented for a non-multimodal dataset (e.g., RGB images only).
@@
 
-We note in particular that standard triplet learning approaches often require an expensive \textit{mining} step to find harder negative samples to enable effective learning \citep{Hoffer2015,Schroff2015,DBLP:conf/eccv/2018-9,DBLP:conf/eccv/ZhaoJQLH18,Zhai2018}.
+We note that standard triplet learning approaches often require an expensive \textit{mining} step to find harder negative samples to enable effective learning \citep{Hoffer2015,Schroff2015,DBLP:conf/eccv/2018-9,DBLP:conf/eccv/ZhaoJQLH18,Zhai2018}.
 This is problematic for throughput and scalability, as run time becomes quadratic in batch size.
@@
 \citet{Veit_2018_CVPR} train a joint model of images, hashtags, and users to perform image retrieval.
-Their task is similar to ours---given a description, we want to find the image/object in the scene that best matches the description.
-They tackle this problem by forming a three-way tensor product model.
-They use a ranking loss to train the model where the score of an observed triplet is higher than an unobserved triplet.
-They sample six negative triplets per positive sample triplet, and use each of them as a negative in the loss.
-The downstream retrieval task is then simply done by taking the argmax of the tensor product for a given user.
+Their task is similar to ours---given a description, find the image/object in the scene that best matches the description.
+They tackle this problem by forming a three-way tensor product model, using a ranking loss to train the model where the score of an observed triplet is higher than an unobserved triplet.
+% They sample six negative triplets per positive sample triplet, and use each of them as a negative in the loss.
+The downstream retrieval task is then done by taking the argmax of the tensor product for a given user.
 This work differs from ours in that it becomes computationally complex as more and higher-dimensional modalities are added.
-Our proposed geometric loss function resembles work on quadruplet loss~\citep{chen2017beyond,tursun2021efficient} but is intended to scale to an arbitrary number of modalities.
+Our proposed geometric loss function resembles quadruplet loss~\citep{chen2017beyond,tursun2021efficient} but is intended to scale to an arbitrary number of modalities.
 
@@
 \citet{chen2017beyond} also uses quadruplet loss, but does not support multiple domains.
-In contrast, our method is not restricted to two modalities.
-% \todo[inline]{talk about CLIP and DALL-E here}
+% In contrast, our method is not restricted to two modalities.
 
@@
 We demonstrate on a domain containing four modalities, all of which refer to objects in the environment: spoken language, written text, RGB (image) inputs, and depth camera inputs.
-\Cref{fig:experimental-setup} illustrates a small visualization of our object retrieval task: the spoken query ``A white textbook titled algorithms'' is provided to our contrastive model, which identifies the item
-(outlined in red in \cref{fig:experimental-setup}) 
-as the most likely item the query is referring to.
-
-\begin{figure}[tb]
-\centering
-\includegraphics[width=.75\columnwidth]{Figures/experiment-setup-rearranged.png}
-\caption{
-The experimental object retrieval setup, in which objects are represented by four modalities: an RGB image, a depth image, spoken descriptions, and textual descriptions.
+%\Cref{fig:experimental-setup} illustrates a small visualization of our object retrieval task: the spoken query ``A white textbook titled algorithms'' is provided to our contrastive model, which identifies the item  (outlined in red in \cref{fig:experimental-setup}) as the most likely item the query is referring to.
+
+% \begin{figure}[tb]
+% \centering
+% \includegraphics[width=.75\columnwidth]{Figures/experiment-setup-rearranged.png}
+% \caption{
+% The experimental object retrieval setup, in which objects are represented by four modalities: an RGB image, a depth image, spoken descriptions, and textual descriptions.
 Given a query in some modality, our approach seeks to select the object that is the best fit, per a trained model.
 This approach, detailed in \cref{sec:Method}, is able to rank objects as to appropriateness even when one or more modalities is ablated at test time (e.g., depth inputs are missing), and outperforms state-of-the-art contrastive learning approaches.
-}
-\label{fig:experimental-setup}
-\end{figure}
+% }
+% \label{fig:experimental-setup}
+% \end{figure}
 
@@
 In keeping with previous work on the closely related problem of image retrieval, we focus on contrastive loss approaches, in which the goal is to learn an embedding of data where similar samples---in our case, samples belonging to the same class of object---are embedded `close' to one another, while dissimilar samples are farther apart.
-We develop a novel geometric loss function, \geom{}, that simultaneously minimizes intra-class distances and maximizes inter-class distances across each pair of modalities, yielding a model that is both effective at the retrieval task defined above and robust to modality dropouts at test time.
-We take an additional step and combine this \geom{} loss with a classification-based (cross-entropy) loss function which results in a superior model compared to either of geometric or cross-entropy losses alone; we refer to this combination as Extended Multi-Modal Alignment, or \ours{}.
-
-\paragraph{Core Concepts}
-\label{sec:core-concepts}
+We describe a novel geometric loss function, \geom{}, that simultaneously minimizes intra-class distances and maximizes inter-class distances across each pair of modalities, yielding a model that is both effective at the retrieval task defined above and robust to modality dropouts at test time.
+We additionally combine this \geom{} loss with a classification-based (cross-entropy) loss function which results in a superior model compared to either of geometric or cross-entropy losses alone; we refer to this combination as Extended Multi-Modal Alignment, or \ours{}.
+
+% \paragraph{Core Concepts}
+% \label{sec:core-concepts}
+\textit{Core Concepts. }
 There is a shared terminology across the methods we describe in this section, but they differ in what they incorporate.
@@
 While this model is a strong baseline, the authors applied it to a unimodal dataset.
-In this paper we extend this baseline to work with a multimodal dataset and show that it is slower than \ours{} to learn and its performance is also lower when all modalities are available during test time.
+In this paper we extend this baseline to work with a multimodal dataset and show that it is slower than \ours{} to learn and requires more data to converge.
 % if modalities are ablated, this method does not perform as well as our proposed model (we show and discuss this later, in \cref{fig:epochs-mrr.srd}).
@@
 
-As originally proposed, the \supcon{} approach which we choose as our baseline was originally applied to unimodal datasets such as ImageNet~\cite{imagenet2009}, CIFAR-10~\cite{krizhevsky2009learning}, and CIFAR-100~\cite{krizhevsky2009learning}.
+As originally proposed, the \supcon{} approach which we choose as our baseline was applied to unimodal datasets such as ImageNet~\cite{imagenet2009}, CIFAR-10~\cite{krizhevsky2009learning}, and CIFAR-100~\cite{krizhevsky2009learning}.
 We both demonstrate that it is possible to use on multimodal datasets and extend it with the addition of geometric approaches, which gives benefits on training speed and performance when modalities are dropped.
@@
 \label{sec:emma}
-
 Our proposed multimodal method is composed of two complementary parts.
@@
 
-\paragraph{Geometric Alignment Loss}
-\label{subsec:geom-loss}
-
+% \paragraph{Geometric Alignment Loss.}
+% \label{subsec:geom-loss}
+
+\textit{Geometric Alignment Loss.}
 % For $M$ modalities, we 
@@
 Therefore, we modify the concept of triplet loss as follows.
-During training, we sample two different instances and their corresponding representations from all modalities into two sets---one positive set (referring to a specific object) and one negative set (referring to some different object) as shown in \cref{fig:emma-loss-trimmed}.
+During training, we sample two different instances and their corresponding representations from all modalities into two sets---one positive set (referring to a specific object) and one negative set (referring to some different object) as shown in \Cref{fig:emma-loss-trimmed}.
 Unlike some prior triplet loss methods~\citep{GoLD_UMBC,triplet_loss_2021_CVPR}, the anchor is not randomly chosen from different modalities in each batch.
@@
 As a note, while you do consider all pairwise distances but you treat some differently than others.}
-\begin{itemize}
+% \begin{itemize}
     % \item data point: a row in the csv/tsv file.
-    \item Positive (Instance): A set of embeddings of one data point (e.g., an RGB image of an apple, corresponding depth image, text description, and speech description of the same apple) as shown with green rectangles in \cref{fig:emma-loss-trimmed}
-    \item Negative (Instance): A set of embeddings of another data point of a different object (e.g., an RGB image of a mug, corresponding depth image, text description, and speech description of the same mug) as shown with orange rectangles in \cref{fig:emma-loss-trimmed}
-    \item Anchor (Modality): Every modality of the positive set is chosen as the learning anchor once.
+    % \item
+
+    \textit{Positive (Instance)}: A set of embeddings of one data point (e.g., an RGB image of an apple, corresponding depth image, text description, and speech description of the same apple) as shown with green rectangles in \Cref{fig:emma-loss-trimmed};
+
+    % \item 
+    \textit{Negative (Instance)}: A set of embeddings of another data point of a different object (e.g., an RGB image of a mug, corresponding depth image, text description, and speech description of the same mug) as shown with orange rectangles in \Cref{fig:emma-loss-trimmed};
+    % \item
+
+    \textit{Anchor (Modality)}: Every modality of the positive set is chosen as the learning anchor once.
 In our formulation, anchor is a ``concept'' referring to the points in the positive sets rather than a single instance in itself.
-In \cref{fig:emma-loss-trimmed}, each of the 4 modalities is selected as an anchor once.
-    %For our experiments, text is used. 
+In \Cref{fig:emma-loss-trimmed}, each of the 4 modalities is selected as an anchor once.
     The anchor is used as the basis for learning distances between positive and negative samples.
@@
     % \todokd{I rephrased it. Good?}
-\end{itemize}
+% \end{itemize}
 
@@
 % When we minimize the distance between each pair of positives, and maximize the distance between each pair of modalities one from the positive set and one from the negative set, the model converges faster than \supcon{} while maintaining the same performance when we have all modalities and a little bit less than Supcon when we drop text.
-We refer to this approach as \geom{} which is formulated in \cref{eq:full-emma}. An illustration of this loss function is provided in \cref{fig:emma-loss-trimmed}.
+We refer to this approach as \geom{} which is formulated in \Cref{eq:full-emma}. An illustration of this loss function is provided in \Cref{fig:emma-loss-trimmed}.
 
@@
 % \end{equation}
+
 \begin{equation}\label{eq:full-emma}
 \begin{split}
-    L &= \sum_{m_1=1}^{M} \left[ \sum_{m_2=1}^{M} \left[ -\max (dist(z_{m_1}^{+}, z_{m_2}^{-}) + \alpha,0) \right] + \sum_{m_3=m_1+1}^{M} \left[ \max( dist(z_{m_1}^{+},z_{m_3}^{+}),0) \right] \right]
+    L &= \sum_{m_1=1}^{M} \Biggl[ \sum_{m_2=1}^{M} \left[ -\max (dist(z_{m_1}^{+}, z_{m_2}^{-}) + \alpha,0) \right]  \\
+    & + \sum_{m_3=m_1+1}^{M} \left[ \max( dist(z_{m_1}^{+},z_{m_3}^{+}),0) \right] \Biggr]
 \end{split}
@@
 
-In \cref{eq:full-emma}, $M$ is the number of modalities, the superscripts $+$ and $-$ represent positive and negative objects, $\alpha$ represents enforced margin between each positive and negative points which we set to 0.4 for all modalities without tuning, and $z$ is the embedding we get by applying a mapping function $f$, which in our case is a neural network on our input data.
-In other words, $z_m = f_m(x_m)$, where each modality $m$ has a specific model $f_m$ that is different from the models for other modalities. These models do not share their weights.
+In \cref{eq:full-emma}, $M$ is the number of modalities, the superscripts $+$ and $-$ represent positive and negative objects, $\alpha$ represents an enforced margin between each positive and negative points which (set to 0.4 for all modalities without tuning), and $z$ is the embedding from applying a mapping function $f$, which is a neural network on our input data.
+In other words, $z_m = f_m(x_m)$, where each modality $m$ has a specific model $f_m$ that is different from the models for other modalities.
+These models do not share their weights.
 
@@
 Other distance metrics, such as Euclidean distance, grow in value with respect to their dimensionality, resulting in a very large distances for data points.
-% The $\cos(\cdot)$ function is a measure of similarity, not distance, and that is why the signs are reversed.
-
-\begin{figure*}[tb]
+The $\cos(\cdot)$ function is a measure of similarity, not distance, and hence the reversed signs.
+
+\begin{figure}[bt]
 \centering
@@
 % \includegraphics[width=2.1\columnwidth]{Figures/e-MMA.png}
-\includegraphics[width=1\columnwidth]{Figures/emma-loss-trimmed.pdf}
+% \includegraphics[width=1\columnwidth]{Figures/emma-loss-trimmed.pdf}
+\includegraphics[width=1\columnwidth]{Figures/EMMA-Arch-Diagram.png}
 \caption{A high-level prototype of our approach and the distances used in the \geom{} loss with four modalities.
@@
 Dashed lines indicate distances to be maximized while dotted lines indicate distances to be minimized.%, and solid lines show the enforced margin.
+% \todocmi{Fix text in all graphics to be readable}
 }
 \label{fig:emma-loss-trimmed}
-\end{figure*}
+\end{figure}
 
@@
 
-If $M=2$ which means the number of modalities is 2, and we ignore the last two terms in the derived objective function, it results in the triplet loss method. 
+If $M=2$ (the number of modalities is 2), and we ignore the last two terms in the derived objective function, it results in the triplet loss method. 
 If $M=2$, then our objective function reduces to the quadruplet loss method~\citep{tursun2021efficient,chen2017beyond} if we multiply the first term by 2, ignore the third term, and change the last summation to be up to $M-1$ (which results a single term).
 If $M=1$, only the last term remains in the loss function which is exactly the pairwise distance-based loss function.
-This loss function can be seen as a contrastive loss usually used in the domain of self-supervised learning~\citep{chen2020simple}.
-However, our proposed loss function has two advantages over the traditional contrastive loss expressed in \cref{eq:contrastive-loss}.
-The first advantage is that our loss function does not loop over multiple positives and negatives in a large batch.
-Instead we sample only two objects (positive and negative) each of which have $M$ modalities which gives us $2M$ datapoints (or embeddings).
+This loss function can be seen as a contrastive loss, usually used in the domain of self-supervised learning~\citep{chen2020simple}.
+
+However, our proposed loss function has two advantages over the traditional loss expressed in \cref{eq:contrastive-loss}.
+The first is that our loss function does not loop over multiple positives and negatives in a large batch.
+Instead we sample only two objects (positive and negative), each of which has $M$ modalities, giving $2M$ datapoints (or embeddings).
 Hence, our model can be trained using smaller batch sizes and reduces the number of negative samples we need.
-The second advantage is that this loss function can be used in a multimodal setting with an arbitrary number of modalities, and is not limited to a single data type (e.g.
-RGB images) which is the most common usage of contrastive loss.
+The second advantage is that this loss function can be used in a multimodal setting with an arbitrary number of modalities, and is not limited to a single data type.
 % ~\citep{chen2020simple, NEURIPS2020_supervised_contrastive}.
@@
 It is noteworthy that our training procedure does not perform any stochastic dropout of modalities to obtain test-time robustness to missing modalities.
-Moreover, our approach does not need to compute the distance between all items in the batch, as opposed to \supcon{}.
-% ~\citep{NEURIPS2020_supervised_contrastive}.
-
-\paragraph{Combining Geometric Loss and Cross-Entropy based \supcon{} Loss}
-\label{subsec:emma}
-
+Moreover, our approach does not need to compute the distance between all items in the batch, as opposed to \supcon{}~\citep{NEURIPS2020_supervised_contrastive}.
+
+% \paragraph{Combining Geometric Loss and Cross-Entropy based \supcon{} Loss}
+% \label{subsec:emma}
+
+\textit{Combining Geometric Loss and Cross-Entropy based \supcon{} Loss.}
 The main difference between \geom{} and \supcon{} is that in \geom{} we focus on the geometric interpretation of similarity using cosine distances, while in \supcon{}, cosine distances are used to compute a classification-based loss function similar to cross-entropy.
@@
 \begin{split}
-    \mathcal{L} =  \sum_{i \in I} \Biggr[ & \Bigr[ \sum_{j=1}^{M-1} \sum_{k=j+1}^{M} h( z_{i,j}^{+} , z_{i,k}^{+})
-     + g( z_{i,j}^{+} , z_{i,k}^{-}) + g( z_{i,j}^{-} , z_{i,k}^{+}) + \sum_{j=1}^{M} g( z_{i,j}^{+} , z_{i,j}^{-} )\Bigr] + \\
-     & \Bigr[ \sum_{m=1}^{M} \frac{-1}{|P(i,m)|}  \sum_{\beta \in P(i,m)} \log \frac{\exp (z_{i,m} \cdot z_{\beta} / \tau) }{\sum_{\gamma \in A(i,m)} \exp (z_{i,m} \cdot z_{\gamma} / \tau)} \Bigr] \Biggr],
+    \mathcal{L} =  \sum_{i \in I} \Biggl[ & \sum_{j=1}^{M-1} \sum_{k=j+1}^{M} h( z_{i,j}^{+} , z_{i,k}^{+})
+     + g( z_{i,j}^{+} , z_{i,k}^{-}) + g( z_{i,j}^{-} , z_{i,k}^{+}) \\ 
+     + & \sum_{j=1}^{M} g( z_{i,j}^{+} , z_{i,j}^{-} ) \\
+    + & \sum_{m=1}^{M} \frac{-1}{|P(i,m)|}  \sum_{\beta \in P(i,m)} \log \frac{\exp (z_{i,m} \cdot z_{\beta} / \tau) }{\sum_{\gamma \in A(i,m)} \exp (z_{i,m} \cdot z_{\gamma} / \tau)} \Biggr]
      \end{split}
 \end{equation}
+% \todokdinline{I don't like how this equation is shown. Any ideas what to do?}
 where $A(i,m)$ includes all items in the batch except for the $z_{i,m}$ itself and $P(i,m)$ includes all the modalities of all instances that have the same label as current instance excluding $z_{i,m}$ itself.
@@
 
-We demonstrate the effectiveness of our approach on a recent publicly available multimodal dataset called GoLD~\citep{GoLD_UMBC}, which contains RGB images, depth images, written text descriptions, speech descriptions, and transcribed speech descriptions for 207 object instances across 47 object classes (see \cref{fig:emma-loss-trimmed}).
+We demonstrate the effectiveness of our approach on a recent publicly available multimodal dataset called GoLD~\citep{GoLD_UMBC}, which contains RGB images, depth images, written text descriptions, speech descriptions, and transcribed speech descriptions for 207 object instances across 47 object classes (see \Cref{fig:emma-loss-trimmed}).
 There are a total of 16,500 spoken and 16,500 textual descriptions.
+The training set, validation set, and test set contain 7380, 4160, and 4960 items (each consisting of four modalities), respectively.
 The original GoLD paper uses raw RGB and depth images in which other objects are present in the background.
@@
 \label{sec:metrics}
-The best metric to capture the performance in such a scenario is mean reciprocal rank (MRR, \cref{eq:mrr} for $Q$ queries).
+The best metric to capture the performance in such a scenario is mean reciprocal rank (MRR)
+% , \cref{eq:mrr} 
+for $Q$ queries. 
 For each query we predict the rank of all objects based on their distance from the language command, and then the inverse rank of the desired objects in all queries are averaged.
+We compute MRR by $\frac{1}{|Q|} \sum_{i=1}^{|Q|} \frac{1}{\operatorname{rank}_{i}}$ which results in a float number between $\frac{1}{\text{number of objects}}$ and 1.0.
 For example, if the model predicts the desired object as the first rank, then MRR = $\frac{1}{1} = 1$ which means a perfect score, and if it predicts the correct object as the fourth rank among five objects, then MRR = $\frac{1}{4}=0.25$.
 
-\begin{equation} \label{eq:mrr}
-    \mathrm{MRR}=\frac{1}{|Q|} \sum_{i=1}^{|Q|} \frac{1}{\operatorname{rank}_{i}}
-\end{equation}
+% \begin{equation} \label{eq:mrr}
+%     \mathrm{MRR}=\frac{1}{|Q|} \sum_{i=1}^{|Q|} \frac{1}{\operatorname{rank}_{i}}
+% \end{equation}
 
@@
 Accuracy and micro F1 score
-% (sklearn), flattened binary f1 score (sklearn), and F1 score 
 are the same in this task, since for each prediction we either have a true positive and no false positives and no false negatives, or we have no true positives, one false positive and one false negative.
 MRR is a more informative metric because it captures the idea that having the correct object as the second choice should be considered better than having it as a last choice, while in accuracy the score is ``all or nothing''---either 0 or 1.
-Because our approach is designed to be robust to missing information across modalities, we also report MRR and accuracy for different combinations of modality dropouts.
+Because our approach is designed to be robust to missing information across modalities, we report MRR %and accuracy
+for different combinations of modality dropouts.
 
@@
 
-In this section we provide quantitative and qualitative results by comparing our method against supervised contrastive learning~\citep{NEURIPS2020_supervised_contrastive} and na\"ive contrastive loss~\citep{chen2020simple}.
+In this section we provide quantitative and qualitative results by comparing our methods against supervised contrastive learning~\citep{NEURIPS2020_supervised_contrastive} and na\"ive contrastive loss~\citep{chen2020simple}.
 % We evaluate our EMMA loss function against the na\"ive triplet loss method.
 Our model significantly outperforms the triplet method across all epochs while it converges faster as shown in \cref{fig:epochs-mrr.srd}.
-\Cref{table:quantitative} summarizes the performance of all models with different ablation metrics.
+\Cref{table:quantitative} summarizes the performance of all models with different ablation metrics and different percentages of training data available.
 % \todo[inline]{Why do we call out triplet loss specifically? We evaluate against several baselines? I don't understand this section particularly}
-To provide a better sense of the performance measure, consider a model that always ranks the correct object in the second place: Such a model would have an MRR of $1/2 = 0.5$.
-
+To provide a better sense of the metric, consider a model that always ranks the correct object in the second place: Such a model would have an MRR of 0.5.
+
+\begin{table*}[tb]
+    \centering
+    % \begin{subtable}[t]{1.00\textwidth}
+    % \centering
+    \resizebox{1.0\textwidth}{!}{
+    \input{tableMRR-all}
+    }
+    % \caption[]{Average and standard deviation of mean reciprocal rank (MRR) on a held-out test set. MRR is from $\frac{1}{\text{number of objects}}$\% to 100\%.
+    % In most cases \ours{} has a better MRR compared to other methods, and all methods have a higher MRR when text modality is present.
+This shows that text modality is a rich source of information compared to speech, and that \ours{} is strongly competitive with state of the art approaches even when relying on speech for the language signal.
+    % }
+    \label{table:mrr}
+    % \end{subtable}
+    % \vskip\baselineskip
+    % \begin{subtable}[t]{1.00\textwidth}
+    % \centering
+    % \resizebox{1.0\textwidth}{!}{
+    % \input{tableACC}
+    % }
+    % \caption[]{Average and standard deviation of accuracy (Acc) on a held-out test set. Accuracy is from 0\% to 100\%.
+    % Although accuracy is a strictly less forgiving metric than MRR, these results demonstrate that \ours{} still outperforms existing approaches in the majority of cases.
+When we drop the text modality, the accuracy of the contrastive method is slightly better than random, while \ours{} is about 10\% more accurate than the contrastive method.
+    % }
+    % \label{table:acc}
+    % \end{subtable}
+    \caption{Average and standard deviation of mean reciprocal rank (MRR) 
+    % and accuracy (Acc) 
+    over 5 runs on a held-out test set with different modalities ablated during testing.
+    % \Cref{table:mrr} shows MRR scores and \cref{table:acc} shows accuracy.
+    % . These results represent performance when no modalities are ablated during training. 
+     For 5 objects, MRR is from 20\% to 100\% and higher is better. A random guess would have MRR of 0.46 
+     % and accuracy of 0.5, 
+     and the worst case performance would have MRR of 0.2.
+     % and accuracy of 0.0. 
+     The batch size is 64 and optimizer is SGD for all experiments.
+Column headers refer to modalities that are present at query time.
+The numbers in the name of each method indicate the percentage of training data we used to train that model.
+    % ---e.g., MRR \textit{sd} contains only speech and depth. 
+    Bold numbers represent the best-performing method.
+\ours{} is either better than or very close to both state-of-the-art methods for most of the cases.
+All methods have a higher MRR when the text modality is present.
+With 25\% of the training data present, \ours{} outperforms other methods in all cases.
+    % We observe the same pattern in both MRR and accuracy except the fact that accuracy scores are lower than MRR.
+    }
+    \label{table:quantitative}
+\end{table*}
+
+\begin{comment}    
 \begin{table*}[tb]
@@
     }
-    \caption[]{Average and standard deviation of mean reciprocal rank (MRR) on a held-out test set. MRR is from $\frac{1}{\text{number of objects}}$\% to 100\%.
-    In most cases \ours{} has a better MRR compared to other methods, and all methods have a higher MRR when text modality is present.
+    \caption[]{In most cases \ours{} has a better MRR compared to other methods, and all methods have a higher MRR when text modality is present.
 This shows that text modality is a rich source of information compared to speech, and that \ours{} is strongly competitive with state of the art approaches even when relying on speech for the language signal.
-    % \todocmi{Right? Correct}
     }
@@
     \resizebox{1.0\textwidth}{!}{
-    \input{tableACC}
-    }
-    \caption[]{Average and standard deviation of accuracy (Acc) on a held-out test set. Accuracy is from 0\% to 100\%.
-    Although accuracy is a strictly less forgiving metric than MRR, these results demonstrate that \ours{} still outperforms existing approaches in the majority of cases.
-When we drop the text modality, the accuracy of the contrastive method is slightly better than random, while \ours{} is about 10\% more accurate than the contrastive method.
+    \input{tableMRR25}
+    }
+    \caption[]{Write the caption.
     }
@@
     \end{subtable}
-    \caption{Average and standard deviation of mean reciprocal rank (MRR) and accuracy (Acc) over 5 runs with 5 different random seeds on a held-out test set with different modalities ablated during testing.
-\Cref{table:mrr} shows MRR scores and \cref{table:acc} shows accuracy.
-    % . These results represent performance when no modalities are ablated during training. 
-     Higher is better for both metrics.
-For 5 objects, a random guess would have MRR of 0.33 and accuracy of 0.5, and the worst case performance would have MRR of 0.2 and accuracy of 0.0.
-The batch size is 64 and optimizer is SGD for all experiments.
-Column headers refer to modalities that are present at query time.
-    % ---e.g., MRR \textit{sd} contains only speech and depth. 
-    Bold numbers represent the best-performing method.
-\ours{} is either better than or very close to both state-of-the-art methods for most of the cases.
-We observe the same pattern in both MRR and accuracy except the fact that accuracy scores are lower than MRR.
+    \caption{Average and standard deviation of mean reciprocal rank (MRR) 
+    over 5 runs on a held-out test set with different modalities ablated during testing.
+MRR is from $\frac{1}{\text{number of objects}}$\% to 100\%.
+Higher is better.
+For 5 objects, a random guess would have MRR of 0.46
+    % and accuracy of 0.5, 
+    and the worst case performance would have MRR of 0.2 
+    % and accuracy of 0.0. 
+    The batch size is 64 and optimizer is SGD for all experiments. Column headers refer to modalities that are present at query time.
+    Bold numbers represent the best-performing method. \ours{} is either better than or very close to both state-of-the-art methods for most of the cases.
+    \todokdinline{Shall we combine the two tables and have a horizontal line to separate different percentages?}
     }
@@
 \end{table*}
+\end{comment}
 
@@
     \centering
-    \begin{subfigure}[b]{0.49\textwidth}
-        \centering
-        \includegraphics[width=\textwidth]{Figures/average-seeds-epochs-mrr_lard-trimmed.pdf}
-        \caption[]{Mean Reciprocal Rank (MRR) on the held-out test set when all modalities are available.}    
+    \begin{subfigure}[b]{0.49\columnwidth}
+        \centering
+        \includegraphics[width=\columnwidth]{Figures/average-seeds-converged-partial-train-mrr_lard-trimmed.pdf}
+        \caption[]{All modalities are available.}    
+        \label{fig:partial-mrr.lard}
+    \end{subfigure}
+    \hfill
+    \begin{subfigure}[b]{0.49\columnwidth}  
+        \centering 
+        \includegraphics[width=\columnwidth]{Figures/average-seeds-converged-partial-train-mrr_ard-trimmed.pdf}
+        \caption[]{The text modality is ablated.}
+        \label{fig:partial-mrr.srd}
+    \end{subfigure}
+    \vskip\baselineskip
+    \begin{subfigure}[b]{0.49\columnwidth}   
+        \centering 
+        \includegraphics[width=\columnwidth]{Figures/average-seeds-converged-partial-train-mrr_lr-trimmed.pdf}
+        \caption[]{Speech and depth are ablated.}    
+        \label{fig:partial-mrr.lr}
+    \end{subfigure}
+    \hfill
+    \begin{subfigure}[b]{0.49\columnwidth}   
+        \centering 
+        \includegraphics[width=\columnwidth]{Figures/average-seeds-converged-partial-train-mrr_ld-trimmed.pdf}
+        \caption[]{Speech and RGB are dropped.}    
+        \label{fig:partial-mrr.ld}
+    \end{subfigure}
+    \caption[]
+    {Mean Reciprocal Rank for selected modalities of the held-out test set, shown as a function of amount of training data.
+All averaged over 5 runs for the downstream task of object retrieval.
+Higher is better.
+    Orange is \supcon{}, green represents our proposed \geom{}, and blue is our proposed \ours{} loss function. We train all models for 200 epochs.
+    In all cases, \ours{} outperforms other models when relatively little training data is available.
+    % \todokdinline{I feel like the explanations about (a) through (c) in this main caption are extra since we mention it in subcaptions.}
+    % \todokdinline{I have uploaded the same graphs wihtout Geometric if we think that's better. We just need to trim them.}
+    } 
+    \label{fig:partial-data-result-graphs}
+\end{figure}
+%
+
+%
+
+In~\Cref{fig:partial-data-result-graphs} we show that in all cases of ablations, \ours{} outperforms \supcon{} in the low data scenarios where different percentages of the training set are made available to each model, but the same held-out test set is used to evaluate all models.
+When we use $25\%$ of the training data and ablate text, \supcon{} struggles to learn and the converged MRR is at 0.46, while \ours{} achieves an MRR of 0.73.
+Note that a random model retrieving an object among five candidates would have an average MRR score of $0.46$.
+
+Moreover, \Cref{fig:epochs-mrr.lard} shows that \ours{} learns faster and results in a better performance compared to both \supcon{} and contrastive learning~\citep{chen2020simple} when trained using all modalities and with all modalities available during test.
+We observe that not only does contrastive loss learn more slowly, but that it is prone to overfitting; while this can be addressed with careful tuning, an approach that is innately robust to overfitting without tuning is preferable.
+
+When we drop the text modality (\Cref{fig:epochs-mrr.srd}), we can see that the performance decreases from about 0.93 to about 0.82, showing that speech cannot completely replace text in our current effort.
+In \Cref{fig:3d-tsne} (see \cref{sec:results}), the alignment of shared embeddings for a randomly sampled set of classes is visualized for all four modalities under consideration, suggesting that the speech modality is not aligned as well as the text modality.
+For this reason, when we drop text and use speech as the main query, the performance decreases.
+This supports our hypothesis that a geometric alignment of the latent space is crucial to a good performance for object retrieval and multimodal understanding.
+
+In \Cref{fig:epochs-mrr.srd,fig:partial-mrr.srd}, we observe that when speech is used as the query and the text modality is ablated, the \supcon{} baseline works slightly better than \ours{} as more training data becomes available, although \ours{} still learns faster.
+The reason is that \supcon{} optimizes for the classification task, and since the speech modality is less well aligned, using \geom{} makes the downstream task more difficult by trying to pull and push similar and dissimilar data points, respectively.
+However, both \geom{} and \ours{} work significantly better than \supcon{} as shown in~\Cref{fig:partial-mrr.srd} when less training data is available suggesting that the classifier cannot be trained well with less data, while distance-based alignment is reliable.
+Future research will consider strategies to align more chaotic modalities.
+% We also observe that speech has a higher variance compared to text, and a na\"ive contrastive learning method cannot take advantage of text modality to reduce this variance.
+While the other two methods leverage text modality to align speech better.
+
+There is very little gap in performance when depth or RGB are dropped (\Cref{fig:epochs-mrr.lr,fig:epochs-mrr.ld,fig:partial-mrr.lr,fig:partial-mrr.ld}) compared to when we have all modalities in \Cref{fig:epochs-mrr.lard,fig:partial-mrr.lard}, showing that our model is robust when RGB or depth sensors fail.
+Also, when depth is dropped (\Cref{fig:epochs-mrr.lr,fig:partial-mrr.lr}), performance decreases than when RGB is dropped (\Cref{fig:epochs-mrr.ld,fig:partial-mrr.ld}).
+This suggests that depth is less informative compared to RGB, which is consistent with existing vision research.
+
+% \todocmi{Main remaining task is to sort out figures. They should all have meaningful y axis labels and essentially no whitespace surrounding them.}
+
+A time analysis shows that \ours{} takes almost 8 epochs to converge and each epoch takes roughly 0.7 minutes, which makes it ~5.6 minutes until convergence, while \supcon{} takes about 36 epochs to converge and each epoch takes 0.52 minutes, summing to ~18.72 minutes.
+That is when we use the entire training set and all four modalities for training.
+When we ablate modalities subset the training data, the training takes less time.
+
+\begin{figure}
+    \centering
+    \begin{subfigure}[b]{0.49\columnwidth}
+        \centering
+        \includegraphics[width=\columnwidth]{Figures/average-seeds-epochs-mrr_lard-trimmed.pdf}
+        \caption[]{All modalities are available.}    
         \label{fig:epochs-mrr.lard}
@@
     \hfill
-    \begin{subfigure}[b]{0.49\textwidth}  
-        \centering 
-        \includegraphics[width=\textwidth]{Figures/average-seeds-epochs-mrr_ard-trimmed.pdf}
-        \caption[]{Mean Reciprocal Rank (MRR) on the held-out test set when the text modality is ablated.}
+    \begin{subfigure}[b]{0.49\columnwidth}  
+        \centering 
+        \includegraphics[width=\columnwidth]{Figures/average-seeds-epochs-mrr_ard-trimmed.pdf}
+        \caption[]{The text modality is ablated.}
         \label{fig:epochs-mrr.srd}
@@
     \vskip\baselineskip
-    \begin{subfigure}[b]{0.49\textwidth}   
-        \centering 
-        \includegraphics[width=\textwidth]{Figures/average-seeds-epochs-mrr_lr-trimmed.pdf}
-        \caption[]{Mean Reciprocal Rank (MRR) on the held-out test set when speech and depth modalities are ablated.}    
+    \begin{subfigure}[b]{0.49\columnwidth}   
+        \centering 
+        \includegraphics[width=\columnwidth]{Figures/average-seeds-epochs-mrr_lr-trimmed.pdf}
+        \caption[]{Speech and depth are ablated.}    
         \label{fig:epochs-mrr.lr}
@@
     \hfill
-    \begin{subfigure}[b]{0.49\textwidth}   
-        \centering 
-        \includegraphics[width=\textwidth]{Figures/average-seeds-epochs-mrr_ld-trimmed.pdf}
-        \caption[]{Mean Reciprocal Rank (MRR) on the held-out test set when speech and RGB modalities are dropped.}    
+    \begin{subfigure}[b]{0.49\columnwidth}   
+        \centering 
+        \includegraphics[width=\columnwidth]{Figures/average-seeds-epochs-mrr_ld-trimmed.pdf}
+        \caption[]{Speech and RGB are dropped.}    
         \label{fig:epochs-mrr.ld}
@@
     \caption[]
-    {Mean Reciprocal Rank (MRR, \cref{eq:mrr}) on the held-out test set with the ablation of selected modalities, all averaged over 5 runs for the downstream task of object retrieval.
-Red is self-supervised contrastive learning which is prone to overfitting, orange is supervised contrastive learning, blue represents our proposed \geom{}, and green is our proposed \ours{} loss function.
-Higher is better.
-(a) shows the MRR when all modalities are available, (b) shows the MRR when text is removed at test time, (c) shows MRR when speech and depth are ablated, and (d) shows MRR when speech and RGB are removed.
-We train all models for 200 epochs.
+    {Mean Reciprocal Rank (MRR, \cref{sec:metrics}) on the held-out test set with the ablation of selected modalities, after each epoch of training, all averaged over 5 runs for the downstream task of object retrieval.
+Higher is better and all available training data is used.
+Red is self-supervised contrastive learning which is prone to overfitting, orange is \supcon{}, green is our \geom{}, and blue is \ours{}.
+In all cases except text ablation, \ours{} learns faster (in fewer epochs) than other models.
+    % (a) through (d) show the performance on different modality ablations of the test set after each epoch of training. We train all models for 200 epochs using the entire training set.
     } 
@@
 %
-% 
-\Cref{fig:epochs-mrr.lard} Shows that \ours{} learns faster and results in a better performance compared to both \supcon{} and contrastive learning~\citep{chen2020simple} when trained using all modalities and with all modalities available during test.
-We observe that not only does contrastive loss learn more slowly, but that it is prone to overfitting; while this can be addressed with careful tuning of the learning process, an approach that is innately robust to overfitting without tuning is preferable.
-
-When we drop the text modality (\cref{fig:epochs-mrr.srd}), we can see that the performance decreases from about 0.93 to about 0.82, showing that speech cannot completely replace text.
-In \cref{fig:3d-tsne} (see \cref{sec:Qualitative}), the alignment of shared embeddings for a randomly sampled set of classes is visualized for all four modalities under consideration, suggesting that the speech modality is not aligned as well as the text modality.
-For this reason, when we drop text and use speech as the main query, the performance decreases.
-This supports our hypothesis that a geometric alignment of the latent space is crucial to a good performance for object retrieval and multimodal understanding.
-
-In \cref{fig:epochs-mrr.srd}, we observe that when speech is used as the query and the text modality is ablated, the \supcon{} baseline works slightly better than \ours{}, although \ours{} still learns faster.
-The reason is that \supcon{} optimizes for the classification task, and since the speech modality is less well aligned, using \geom{} makes the downstream task more difficult by trying to pull and push similar and dissimilar data points, respectively.
-Future research will consider strategies to align more chaotic modalities.
-% We also observe that speech has a higher variance compared to text, and a na\"ive contrastive learning method cannot take advantage of text modality to reduce this variance.
-While the other two methods leverage text modality to align speech better.
-
-There is very little gap in performance when depth or RGB are dropped in \cref{fig:epochs-mrr.lr,fig:epochs-mrr.ld} compared to when we have all modalities in \cref{fig:epochs-mrr.lard}, showing that our model is robust when RGB or depth sensors fail.
-Also, when depth is dropped in \cref{fig:epochs-mrr.lr}, performance decreases less compared to when RGB is dropped in \cref{fig:epochs-mrr.ld}.
-This suggests that depth is less informative when compared to RGB, which is consistent with existing vision research results.
-
-% \todocmi{Main remaining task is to sort out figures. They should all have meaningful y axis labels and essentially no whitespace surrounding them.}
-
-Our time analysis shows that EMMA takes almost 8 epochs to converge and each epoch takes roughly 0.7 minutes, which makes it 5.6 minutes until convergence, while \supcon{} takes about 36 epochs to converge and each epoch takes 0.52 minutes which amounts to 18.72 minutes.
-That is when we use all four modalities for training.
-When we ablate one or two modalities, the training takes less time.
-
-\paragraph{Qualitative Results}
-\label{sec:Qualitative}
+
+% \paragraph{Qualitative Results}
+% \label{sec:Qualitative}
+\textit{Qualitative Results.}
 In order to help visualize the performance of learned embeddings, we consider projections of a randomly selected subset of classes of the high-dimensional learned embeddings into a 3-dimensional space using t-SNE~\citep{van2008tsne}, a dimensionality reduction technique to visualize high-dimensional data.
@@
 Although these projections are not perfect, combined with the quantitative results, they demonstrate that our model is learning to map instances of the same class closer to each other regardless of their modalities.
-Interestingly, toothbrush and toothpaste are mapped almost on top of each other in the text modality showing semantic and syntax similarity.
-However, in the RGB and depth modality they are close but not on top of each other since they do not look the same.
-Also, we can see that apple and lemon are mapped close to each other in all modalities which suggests that our proposed \ours{} learns some notion of the concept of fruits.
-These qualitative results show that our propose \geom{} and \ours{} have an interpretable latent space.
-
-\begin{figure*}[h]
+Qualitatively, we note that toothbrush and toothpaste are mapped almost on top of each other in the text modality, showing semantic and syntactic similarity, but not in the visual modalities, since they do not look the same.
+Also, we note that apple and lemon are mapped close to each other in all modalities.
+These qualitative results suggest that our proposed \geom{} and \ours{} may have an interpretable latent space.
+
+\begin{figure}[h]
 \centering
@@
 % \includegraphics[width=.99\columnwidth]{Figures/3D-tsne-exp-supcon-emma-lard-64-relu-SGD-0.001-unique_objects-gold-no_neg_sampling-1024-sbs-trimmed.png}
-\includegraphics[width=.99\columnwidth]{Figures/3D-tsne-exp-supcon-emma-lard-64-relu-SGD-0.001-unique_objects-gold-no_neg_sampling-1024-trimmed.pdf}
+\includegraphics[width=1\columnwidth]{Figures/3D-tsne-exp-supcon-emma-lard-64-relu-SGD-0.001-unique_objects-gold-no_neg_sampling-1024-trimmed.pdf}
 \caption{3D T-SNE~\citep{van2008tsne} projection of test embeddings of 10 randomly selected classes of objects using \ours{}.
@@
 }
+% \todokdinline{We can put this graph in one column as well and save some space.}
+
 % \todo[inline]{get rid of markers for modalities, and change the colors to be brighter. Try combining marker and color for each class.}
-
 \label{fig:3d-tsne}
-\end{figure*}
-
-An example of the need to consider multiple modalities jointly is shown in \cref{fig:rankings}, showing how EMMA is able to correctly select an object instance from several similarly shaped and describable objects.
-
+\end{figure}
+
+% Uncomment the following if you want to have figure of rankings
+% An example of the need to consider multiple modalities jointly is shown in \cref{fig:rankings}, showing how EMMA is able to correctly select an object instance from several similarly shaped and describable objects.
+EMMA more accurately ranks objects and handles ambiguities with respect to a retrieval queries.
+In a retrieval case where the query is "This is a can opener.
+It is light blue in color." and  the candidate objects are can opener, potato, soda bottle, book, and light bulb, \supcon{} is confused and predicts light bulb as the correct object since the phrase “light blue” in the query is very similar to “light bulb” which is also among the candidates, while EMMA correctly identifies the can opener as the intended object, ranking the light bulb second.
+% This is a   can opener. It is light blue in color. can_opener_3_4	0 [0 2 4 3 1]	lightbulb_4_3	['can_opener_3_4' 'potato_5_1' 'soda_bottle_3_2' 'book_3_1' 'lightbulb_4_3']
+
+\begin{comment}
 \begin{figure}[h!]
@@
 \caption{\ours{} more accurately ranks objects and handles ambiguities with respect to a retrieval queries.
-Note that the phrase ``light blue'' in the query is very similar to ``light bulb'' and while \supcon{} confuses this and predicts light bulb as the correct object, \ours{} correctly identifies the can opener as the intended object, ranking the light bulb second.}
+Note that the phrase ``light blue'' in the query is very similar to ``light bulb'' and while \supcon{} confuses this and predicts light bulb as the correct object, \ours{} correctly identifies the can opener as the intended object, ranking the light bulb second.
+% \todocmi{Make sure this is discussed in text, and get rid of the figure.}
+}
 \label{fig:rankings}
 \end{figure}
+\end{comment}
 
@@
 \Cref{fig:3d-tsne} shows that although the relative position of instances are correct in the speech space, the distinction and clustering of different objects are not as good as the other three modalities.
-
 Text seems to be the best clustered modality, and that makes sense because the variation in written text is much smaller than the other three modalities.
@@
 In this work, we have demonstrated the effectiveness of a novel approach to learning from high-dimensional multimodal information even when one or more modalities is unavailable at test time.
-Our approach performs well on an object retrieval task from a testbed that contains four separate modalities, consistent with information that might be available to a physical agent, and outperforms state of the art contrastive learning approaches.
+On an object retrieval task from a testbed that contains four separate modalities, consistent with information that might be available to a physical agent, our model achieves an MRR score of 0.91 when trained on 25\% of training data and all modalities are available during test, while the previous state-of-the-art baseline achieves an MRR of 0.83.
+This gap is even larger when the text modality is missing during test and speech is used as the query.
+Moreover, \ours{}, our model, is about 2.3 times faster than \supcon{} and requires about 30\% of the time that \supcon{} needs to converge to the same performance when we train both models using all of the training data.
 Our proposed method is general enough to be applied to a variety of multimodal retrieval problems, and is not limited to purely language-based image retrieval.
@@
 
-In future, this work will be extended to solve less clearly delineated problems; for example, differentiating among members of a class, as well as across classes.
+% \todokdinline{Shall we remove the future work paragraph below? Removed for the sake of space}
+% In future, this work will be extended to solve less clearly delineated problems; for example, differentiating among members of a class, as well as across classes.
 However, this work represents a significant step towards handling such retrieval problems, while not arbitrarily limiting the number of sensor and other modalities that can be incorporated.
@@
 
-\bibliography{tmlr}
-\bibliographystyle{tmlr}
-
+%%
+%% The acknowledgments section is defined using the "acks" environment
+%% (and NOT an unnumbered section). This ensures the proper
+%% identification of the section in the article metadata, and the
+%% consistent spelling of the heading.
+% \begin{acks}
+% To Robert, for the bagels and explaining CMYK and color spaces.
+% \end{acks}
+
+%%
+%% The next two lines define the bibliography style to be used, and
+%% the bibliography file.
+\bibliographystyle{ACM-Reference-Format}
+\bibliography{references}
+
+%%
+%% If your work has an appendix, this is the place to put it.
 \appendix
@@
 % \input{appendix}
-
 \end{document}
+\endinput
+%%
+%% End of file `sample-sigconf.tex'.

### 2023-02-09 Code/ML.py
different language models for text embeddings

@@ -347,6 +347,7 @@ def load_configs():
     parser.add_argument('--noise', default='clean', type=str)
     parser.add_argument('--drop_train_data', default=0, type=float)
     parser.add_argument('--drop_train_data_objects', default='objects', type=str, help='inputs are: objects, randomly')
+    parser.add_argument('--lm', default='bert', type=str, help='language models for text embeddings: bert, bart-base, bart-large, t5')
     
     args = parser.parse_args()
     if args.method == 'supervised-contrastive':
@@ -370,6 +371,12 @@ if __name__ == "__main__":
     if 'd' in config.data_type:
         config.modalities.append('depth')
     
+    if config.lm == 'bert':
+        config.text_dim = 3072
+    elif config.lm == 'bart-large':
+        config.text_dim = 1024
+    elif config.lm in ['bart-base', 't5', 't5-large']:
+        config.text_dim = 768
     
     set_seeds(config)
     device = setup_device(config.gpu_num)
@@ -382,7 +389,7 @@ if __name__ == "__main__":
 
     rgb_model = TheModel(config, feature_size=2048, device=device, embed_dim=config.embed_dim)
     depth_model = TheModel(config, feature_size=2048, device=device, embed_dim=config.embed_dim)
-    language_model = TheModel(config, feature_size=3072, device=device, embed_dim=config.embed_dim)
+    language_model = TheModel(config, feature_size=config.text_dim, device=device, embed_dim=config.embed_dim)
     audio_model = TheModel(config, feature_size=3072, device=device, embed_dim=config.embed_dim)
     models = {'rgb': rgb_model, 'depth':depth_model, 'language': language_model, 'audio': audio_model}
     # TODO: for modality in modalities: 

### 2023-02-09 Code/analysis.py

@@ -8,29 +8,45 @@ import tikzplotlib
 
 
 exp_ugly_names = {
-    'EMMA-25': 'exp-0.75-train-supcon-emma-lard-64-relu-SGD-0.001-unique_objects-gold-no_neg_sampling-1024',
-    'EMMA-30': 'exp-0.70-train-supcon-emma-lard-64-relu-SGD-0.001-unique_objects-gold-no_neg_sampling-1024',
-    'EMMA-35': 'exp-0.65-train-supcon-emma-lard-64-relu-SGD-0.001-unique_objects-gold-no_neg_sampling-1024',
-    'EMMA-45': 'exp-0.55-train-supcon-emma-lard-64-relu-SGD-0.001-unique_objects-gold-no_neg_sampling-1024',
-    'EMMA-50': 'exp-0.5-train-supcon-emma-lard-64-relu-SGD-0.001-unique_objects-gold-no_neg_sampling-1024',
-    'EMMA-75': 'exp-0.25-train-supcon-emma-lard-64-relu-SGD-0.001-unique_objects-gold-no_neg_sampling-1024',
-    'EMMA-100': 'exp-supcon-emma-lard-64-relu-SGD-0.001-unique_objects-gold-no_neg_sampling-1024',
+    'EMMAbert-25': 'exp-0.75-train-supcon-emma-lard-64-relu-SGD-0.001-unique_objects-gold-no_neg_sampling-1024',
+    'EMMAt5base-25': 'exp-0.75-train-t5-base-supcon-emma-lard-64-relu-SGD-0.001-unique_objects-gold-no_neg_sampling-1024',
+    'EMMAbartbase-25': 'exp-0.75-train-bart-supcon-emma-lard-64-relu-SGD-0.001-unique_objects-gold-no_neg_sampling-1024',
+    'EMMAbartlarge-25': 'exp-0.75-train-bart-large-supcon-emma-lard-64-relu-SGD-0.001-unique_objects-gold-no_neg_sampling-1024',
 
-    'SupCon-25': 'exp-0.75-train-supcon-lard-64-relu-SGD-0.001-unique_objects-gold-no_neg_sampling-1024',
-    'SupCon-30': 'exp-0.70-train-supcon-lard-64-relu-SGD-0.001-unique_objects-gold-no_neg_sampling-1024',
-    'SupCon-35': 'exp-0.65-train-supcon-lard-64-relu-SGD-0.001-unique_objects-gold-no_neg_sampling-1024',
[…]

### 2023-02-09 Code/datasets.py

@@ -8,6 +8,7 @@ from torch.utils.data import Dataset, DataLoader
 import torchvision
 from sklearn.model_selection import train_test_split
 import flair
+from sentence_transformers import SentenceTransformer
 import re
 from skimage import io
 import skimage
@@ -95,17 +96,25 @@ class GoLD_Dataset(Dataset):
             self.rgb_embeddings, self.depth_embeddings = self.image_embeddings[0], self.image_embeddings[1]
         
         if computed_lang_embeds is None:
-            if os.path.exists('../data/'+task+'_language_embeddings.pkl'):
-                print(f'loading {task} language embeddings previously saved')
-                self.language_embeddings = pickle.load(open('../data/'+task+'_language_embeddings.pkl', 'rb'))
+            if os.path.exists('../data/'+task+'_'+self.config.lm+'_language_embeddings.pkl'):
+                print(f'loading {self.config.lm} language embeddings previously saved for {task}')
+                self.language_embeddings = pickle.load(open('../data/'+task+'_'+self.config.lm+'_language_embeddings.pkl', 'rb'))
             else:
-                print('computing language embeddings once ...')
+                print(f'computing {self.config.lm} language embeddings once ...')
                 self.language_embeddings = []
-                self.language_model = flair.embeddings.DocumentPoolEmbeddings([flair.embeddings.BertEmbeddings()])
+                if self.config.lm == 'bert':
+                    self.language_model = flair.embeddings.DocumentPoolEmbeddings([flair.embeddings.BertEmbeddings()])
+                elif self.config.lm == 'bart-base':
+                    self.language_model = flair.embeddings.TransformerDocumentEmbeddings('facebook/bart-base')
+                elif self.config.lm == 'bart-large':
+                    self.language_model = flair.embeddings.TransformerDocumentEmbeddings('facebook/bart-large')
+                elif self.config.lm == 't5':
+                    self.language_model = SentenceTransformer('sentence-transformers/sentence-t5-base')
+                    # self.language_model = flair.embeddings.TransformerDocumentEmbeddings('t5-base') # Not working
                 for i in range(len(self.descriptions)):
                     self.language_embeddings.append(self.proc_sentence(self.descriptions[i]).detach().cpu().numpy())
-                print('Done computing language embeddings once!')
-                pickle.dump(self.language_embeddings, open('../data/'+task+'_language_embeddings.pkl', 'wb'))
[…]
